### 2017-02-09 paper/main.tex
re-add paper as the actual files

\documentclass[modern, letterpaper]{aastex61}

% to-do list
% ----------
% -

% style notes
% -----------
% - This file generates by Makefile; don't be typing ``pdflatex'' or some bullshit.
% - Line break between sentences to make the git diffs readable.
% - Simple Monte Carlo gets a capital S to indicate that it is a defined thing.
% - Use \, as a multiply operator.
% - Reserve () for function arguments; use [] or {} for outer shit.
% - Always prior pdf or posterior pdf, never prior or posterior (that's your arse).
% - Use \sectionname not Section, \figname not Figure, \documentname not Article or Paper or paper.

\include{gitstuff}
\include{preamble}

% define macros for text
\newcommand{\apogee}{\project{\acronym{APOGEE}}}
\newcommand{\sdssiii}{\project{\acronym{SDSS-III}}}
\newcommand{\thejoker}{\project{The~Joker}}
\newcommand{\DR}{\acronym{DR13}}

% for response to referee
% \renewcommand{\resp}[1]{#1}

\shortauthors{Price-Whelan et al.}

\begin{document}\sloppy\sloppypar\raggedbottom\frenchspacing % trust me

\title{Population statistics of companions to APOGEE red clump stars}

\author{Adrian~M.~Price-Whelan}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}
\email{adrn@astro.princeton.edu}
\correspondingauthor{Adrian M. Price-Whelan}

\author{David~W.~Hogg}
\affiliation{Center for Cosmology and Particle Physics,
             Department of Physics,
             New York University, 726 Broadway,
             New York, NY 10003, USA}
\affiliation{Max-Planck-Institut f\"ur Astronomie,
             K\"onigstuhl 17, D-69117 Heidelberg, Germany}
\affiliation{Flatiron Institute,
             Simons Foundation,
             162 Fifth Avenue,
             New York, NY 10010, USA}

\author{Hans-Walter~Rix}
\affiliation{Max-Planck-Institut f\"ur Astronomie,
             K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\begin{abstract}
% Context
% Aims
% Methods
% Results
\end{abstract}

\keywords{
  binaries: spectroscopic
  ---
  methods: data analysis
  ---
  methods: statistical
  ---
  planets and satellites: fundamental parameters
  ---
  surveys
  ---
  techniques: radial velocities
}

\section{Introduction} \label{sec:intro}

Foo.

\end{document}

### 2017-02-09 ci/build-paper.sh
build paper

@@ -1,0 +1,15 @@
+#!/bin/bash -x
+
+# Build the paper
+cd paper
+make
+
+# Push to GitHub
+if [ -n "$GITHUB_API_KEY" ]; then
+  cd $TRAVIS_BUILD_DIR
+  git checkout --orphan $TRAVIS_BRANCH-pdf
+  git rm -rf .
+  git add -f paper/main.pdf
+  git -c user.name='travis' -c user.email='travis' commit -m "making a paper"
+  git push -q -f https://$GITHUB_USER:$GITHUB_API_KEY@github.com/$TRAVIS_REPO_SLUG $TRAVIS_BRANCH-pdf
+fi

### 2017-02-09 ci/setup-texlive.sh

@@ -1,0 +1,12 @@
+#!/bin/bash -x
+
+# Download and extract texlive installer
+wget http://mirror.ctan.org/systems/texlive/tlnet/install-tl-unx.tar.gz -O install-tl.tar.gz
+mkdir -p install-tl
+tar -xf install-tl.tar.gz -C install-tl --strip-components 1
+
+# Install only the necessary packages
+mkdir -p texlive
+TEXLIVE_INSTALL_PREFIX=`pwd`/texlive install-tl/install-tl -profile ci/texlive.profile
+export PATH=`pwd`/texlive/bin/x86_64-linux:$PATH
+tlmgr install lineno revtex textcase epsf xcolor ulem microtype multirow helvetic

### 2017-02-09 ci/setup-paper.sh
setup

@@ -1,0 +1,12 @@
+#!/bin/bash -x
+
+# If building the paper, do that here
+if [[ $BUILD_PAPER == paper ]]; then
+  if git diff --name-only $TRAVIS_COMMIT_RANGE | grep 'paper/'
+  then
+    echo "Building the paper..."
+    source "$( dirname "${BASH_SOURCE[0]}" )"/setup-texlive.sh
+    return
+  fi
+  return
+fi

### 2017-02-09 ci/setup-paper.sh
shuffling

@@ -1,12 +1,9 @@
 #!/bin/bash -x
 
-# If building the paper, do that here
-if [[ $BUILD_PAPER == paper ]]; then
-  if git diff --name-only $TRAVIS_COMMIT_RANGE | grep 'paper/'
-  then
-    echo "Building the paper..."
-    source "$( dirname "${BASH_SOURCE[0]}" )"/setup-texlive.sh
-    return
-  fi
+if git diff --name-only $TRAVIS_COMMIT_RANGE | grep 'paper/'
+then
+  echo "Building the paper..."
+  source "$( dirname "${BASH_SOURCE[0]}" )"/setup-texlive.sh
   return
 fi
+return

### 2017-02-09 ci/setup-paper.sh
wtf

@@ -6,4 +6,5 @@ then
   source "$( dirname "${BASH_SOURCE[0]}" )"/setup-texlive.sh
   return
 fi
+export BUILD_PAPER='true'
 return

### 2017-02-09 ci/build-paper.sh
please execute this

@@ -1,5 +1,9 @@
 #!/bin/bash -x
 
+if [[ $BUILDING_PAPER != true ]]; then
+  return
+fi
+
 # Build the paper
 cd paper
 make

### 2017-02-09 ci/setup-paper.sh

@@ -1,10 +1,5 @@
 #!/bin/bash -x
 
-if git diff --name-only $TRAVIS_COMMIT_RANGE | grep 'paper/'
-then
-  echo "Building the paper..."
-  source "$( dirname "${BASH_SOURCE[0]}" )"/setup-texlive.sh
-  return
-fi
-export BUILD_PAPER='true'
-return
+echo "Building the paper..."
+export BUILDING_PAPER=true
+source "$( dirname "${BASH_SOURCE[0]}" )"/setup-texlive.sh

### 2017-02-09 ci/before_script.sh
split into scripts

@@ -1,0 +1,11 @@
+#!/bin/bash -x
+
+set -ev
+
+if [[ $BUILD_PAPER != true ]]; then
+   export TWOFACE_CACHE_PATH=~/twoface_test/
+
+   if [[ $SETUP_CMD != *"egg_info"* ]]; then
+       python scripts/initdb.py --test -v
+   fi
+fi

### 2017-02-09 ci/install.sh

@@ -1,0 +1,11 @@
+#!/bin/bash -x
+
+set -ev
+
+if [[ $BUILD_PAPER == true ]]; then
+   source ci/setup-paper.sh
+else
+   git clone git://github.com/astropy/ci-helpers.git
+   source ci-helpers/travis/setup_conda_$TRAVIS_OS_NAME.sh
+   if [[ $SETUP_CMD != *"egg_info"* ]]; then python setup.py install; fi
+fi

### 2017-02-09 ci/script.sh

@@ -1,0 +1,13 @@
+#!/bin/bash -x
+
+set -ev
+
+echo $BUILDING_PAPER
+
+if [[ $BUILD_PAPER == true ]]; then
+   source ci/build-paper.sh
+   echo "YES3"
+else
+   $MAIN_CMD $SETUP_CMD
+   echo "NO3"
+fi

### 2017-02-28 paper/main.tex
abstract attempt

@@
 \newcommand{\thejoker}{\project{The~Joker}}
+\newcommand{\thecannon}{\project{The~Cannon}}
 \newcommand{\DR}{\acronym{DR13}}
+\newcommand{\RC}{\acronym{RC}}
 
@@
 % Context
+Radial-velocity measurements can be used to find stellar, sub-stellar,
+and planetary companions of stars.
+In the \apogee\ survey, we have multi-epoch radial-velocity measurements
+for many red-giant stars, including many red-clump (\RC) stars.
+Spectral analysis of these stars can be used to deliver masses (with
+XXX precision) and 20-ish chemical abundance measurements (with YYY
+precision).
 % Aims
+Here we perform a search for secondary companions of [number] \RC\ stars.
+Because the primary-star masses are known, the mass-function degeneracy
+is broken and the secondary-companion $m\,\sin i$ values can be inferred.
+Also, if the primary and companion are assumed to be co-eval, the
+abundances of the \RC\ star can be assumed to represent accurately the
+abundances of the secondary.
 % Methods
+We use a data-driven model (\thecannon) of stellar spectra to deliver
+each \RC\ mass and chemical abundances, and a custom-built Monte Carlo
+sampler (\thejoker) to deliver (often highly multi-modal) posterior
+beliefs about companion $m\,\sin i$, period, and other orbital parameters.
+Our sampler permits us to accurately sample even the badly behaved
+posterior pdfs obtained when there are as few an NNN observation
+epochs per star.
+We put these posterior samples---made under an interim prior---into a
+hierarchical inference to obtain individual-companion-marginalized
+posterior beliefs about the properties of the full secondary-companion
+distribution in mass and period.
 % Results
+We deliver a catalog of QQQ companions [that meet some criteria], and
+posterior samplings for all [number] stars in the parent sample.
+We find that the period and mass distributions [have some properties].
+We look for trends with chemical abundances, and we find tentative evidence
+for ZZZ and WWW.
 \end{abstract}

### 2017-03-02 paper/main.tex
title, abstract tweaks

@@
 
-\title{Population statistics of companions to APOGEE red clump stars}
+\title{Binary companions of red clump stars}
 
@@
 for many red-giant stars, including many red-clump (\RC) stars.
-Spectral analysis of these stars can be used to deliver masses (with
+Spectral analysis of \RC\ stars can be used to deliver masses (with
 XXX precision) and 20-ish chemical abundance measurements (with YYY
@@
 distribution in mass and period.
+In this analysis, even stars with only a few observations (and therefore
+badly multimodal posterior pdfs) contribute valuable information to the
+population inference.
 % Results
@@
 \keywords{
-  binaries: spectroscopic
+  binaries:~spectroscopic
   ---
-  methods: data analysis
+  methods:~data~analysis
   ---
-  methods: statistical
+  methods:~statistical
   ---
-  planets and satellites: fundamental parameters
+  planets~and~satellites:~fundamental~parameters
   ---
@@
   ---
-  techniques: radial velocities
+  techniques:~radial~velocities
 }

### 2017-03-02 paper/main.tex
intro content

@@
 \newcommand{\RC}{\acronym{RC}}
+\newcommand{\RGB}{\acronym{RGB}}
 
@@
 
-Foo.
+If we want to use a spectroscopic survey to responsibly measure the
+binary-star population---that is, if we want to get (say) the mass,
+mass ratio, and period distribution of binary star systems---we face
+(at least) two challenges.
+The first is how, given a small number of observations of a primary
+star (in the case, say, of SB1 stars), do we reliably obtain posterior
+information about the binary-system properties?
+After all, if there are only a few radial-velocity measurements made
+per star, any observed radial-velocity values will be consistent with
+many different combinations of period and amplitude (not to mention
+eccentricity and argument of perihelion) for the primary orbit.
+In general, the likelihood function, and the posterior pdf under any
+reasonable prior pdf, will be highly multimodal in these kinds of
+problems, and many of the modes will have comparable integrated
+probability density.
+We have solved this problem previously, though with limitations (to be
+discussed more below), with \thejoker\ (CITE).
+\thejoker\ is a Simple Monte Carlo rejection sampler that is
+computationally expensive but probabilistically righeous: It delivers
+independent (zero-auto-correlations) posterior pdf samples for
+single-companion binary model parameters, given any number of
+radial-velocity measurements.
+
+The second challenge is how, given all these noisy, multi-modal
+posterior pdfs overy binary-system properties, many of which will
+permit a wide range of qualitatively different system properties, and
+many of which will be consistent with no companion at all, can we
+constrain parameters of a model of the full binary-system population?
+No histogram of best-fit values, nor stack of multi-modal posteriors,
+would be anything like a good estimate of the true population
+distribution.
+We have also solved this problem previously, though in the exoplanet
+domain, with hierarchical probabilistic models based on importance
+sampling (CITE DFM, and also HOGG).
+These hierarchical models derive posterior estimates of the population
+parameters that accurately describe the population, given the noisy
+information about the individual systems, provided that the noise
+model is accurate---that is, provided that the likelihood functions
+used in the individual-system inferences are themselves accurate.
+
+Here we mash up these two solutions to these two challenges.
+We deliver a catalog of binary-star systems with \RC\ primaries, and also
+posterior information about the full population statistics.
+
+HOGG: Why are \RC\ or \RGB\ stars ideal candidates? It has to do with the
+unlikelihood that they will be SB2. Also: abundances! Also: masses!
+Also: distances! and so on.!
+
+HOGG: Why and how does \apogee\ rock it?
+
+HOGG: What is up with delivering a catalog (it requires decision-making).
 

### 2017-03-02 paper/main.tex
typos and explanation

@@
 The first is how, given a small number of observations of a primary
-star (in the case, say, of SB1 stars), do we reliably obtain posterior
+star (a single-lined spectroscopic binary star), do we reliably obtain posterior
 information about the binary-system properties?
@@
 \thejoker\ is a Simple Monte Carlo rejection sampler that is
-computationally expensive but probabilistically righeous: It delivers
+computationally expensive but probabilistically righteous: It delivers
 independent (zero-auto-correlations) posterior pdf samples for
@@
 The second challenge is how, given all these noisy, multi-modal
-posterior pdfs overy binary-system properties, many of which will
+posterior pdfs over binary-system properties, many of which will
 permit a wide range of qualitatively different system properties, and
@@
 We deliver a catalog of binary-star systems with \RC\ primaries, and also
-posterior information about the full population statistics.
+posterior information about the full population statistics, for at least
+a toy model of that full population.
+We use all the data---even the non-detections, marginal detections, and
+badly multi-modal posterior pdfs---when we infer the population model;
+that is, we don't arbitrarily cut to a catalog of ``clean'' binaries.
 

### 2017-03-05 paper/main.tex
intro: why RC stars

@@
 % ----------
-% -
+% - write a first draft of the introduction
+% - list our assumptions for detection & characterization (TheJoker)
+% - list our assumptions for hierarchical inference
 
@@
 
-HOGG: Why are \RC\ or \RGB\ stars ideal candidates? It has to do with the
-unlikelihood that they will be SB2. Also: abundances! Also: masses!
-Also: distances! and so on.!
+For the study of single-line binaries, red-giant stars---and even
+better red-clump stars---are ideal.
+For one, because they are so luminous, they are unlikely (in general)
+to have equally-bright companions, and therefore are well approximated
+or fit as single-line objects.
+For two, they are standard candles, or can be distance-calibrated.
+For three, they have masses that can be estimated spectrally (making
+use of dredged-up elements; CITE MARTIG and CITE NESS).
+With primary-star mass estimates, the binary-orbit fitting will return
+$m\,\sin i$ estimates for the secondary, and not just mass-function
+estimates.
+For four, the \apogee\ pipelines, and also \thecannon\, produce
+detailed abundance estimates for red giants and \RC\ stars.
+If there are causal relationships between chemical abundances and
+binary companions---and we expect there will be---these should become
+visible here.
 
@@
 
+\section{Method: Detection and orbit fitting}
+
+\section{A catalog of red-clump binary systems}
+
+\section{Method: Population inferences}
+
+\section{The population of red-clump binary companions}
+
+\section{Discussion}
+
 \end{document}

### 2017-03-06 scripts/run_redclump.py
untested shit

@@ -10,7 +10,7 @@ import numpy as np
 from schwimmbad import choose_pool
 from sqlalchemy import func
 from thejoker.log import log as joker_logger
-from thejoker.sampler import JokerParams, TheJoker, save_prior_samples
+from thejoker.sampler import JokerParams, TheJoker
 from thejoker.utils import quantity_to_hdf5
 import yaml
 
@@ -70,7 +70,7 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
 
         else:
             # no jitter specified, set to extra suggested systemic velocity
-            run.jitter = 130 * u.m/u.s
+            run.jitter = 130. * u.m/u.s
 
         # TODO: need a way to pass in velocity trend info
 
@@ -101,10 +101,7 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
     if not os.path.exists(run.prior_samples_file):
         logger.debug("Prior samples file not found - generating now...")
 
-        # TODO: also need to get value of the prior pdf and save
-        prior_samples = joker.sample_prior(config['prior']['num_cache'])
-        prior_units = save_prior_samples(run.prior_samples_file, prior_samples, u.km/u.s) # data in km/s
-        del prior_samples
+        # TODO: prior samples
         # TODO: for large prior samples libraries, may want a way to write
         #   without storing all samples in memory
 

### 2017-03-06 twoface/sample_prior.py

@@ -1,0 +1,44 @@
+import astropy.units as u
+import h5py
+import numpy as np
+from thejoker.sampler import pack_prior_samples
+
+def make_prior_cache(filename, joker, N, max_batch_size=2**24):
+    """
+
+    """
+
+    max_batch_size = int(max_batch_size)
+    N = int(N)
+
+    # first just make an empty file
+    with h5py.File(filename, 'w') as f:
+        pass
+
+    num_added = 0
+    for i in range(10000): # HACK: magic number, maximum num. iterations
+        samples, ln_probs = joker.sample_prior(max_batch_size, return_logprobs=True)
+        packed_samples, units = pack_prior_samples(samples, u.km/u.s) # TODO: make rv_unit configurable?
+
+        batch_size,K = packed_samples.shape
+
+        if (num_added + batch_size) > N:
+            packed_samples = packed_samples[:N - (num_added + batch_size)+1]
+            batch_size,K = packed_samples.shape
+            if batch_size <= 0:
+                break
+
+        with h5py.File(filename, 'r+') as f:
+            if 'samples' not in f:
+                # make the HDF5 file with placeholder datasets
+                f.create_dataset('samples', shape=(N, K), dtype=np.float32)
+                f.create_dataset('ln_prior_probs', shape=(N,), dtype=np.float32)
+                f.attrs['units'] = np.array([str(x) for x in units]).astype('|S6')
+
+            i1 = num_added
+            i2 = num_added+batch_size
+
+            f['samples'][i1:i2,:] = packed_samples
+            f['ln_prior_probs'][i1:i2] = ln_probs
+
+        num_added += batch_size

### 2017-03-06 twoface/tests/test_sample_prior.py

@@ -1,0 +1,19 @@
+# Standard library
+import os
+import sys
+
+# Third-party
+import astropy.units as u
+import numpy as np
+from thejoker.sampler import TheJoker, JokerParams
+
+# Project
+from ..sample_prior import make_prior_cache
+
+def test_make_prior_cache(tmpdir):
+
+    filename = str(tmpdir / 'prior_samples.h5')
+    params = JokerParams(P_min=8*u.day, P_max=8192*u.day)
+    joker = TheJoker(params)
+
+    make_prior_cache(filename, joker, N=2**16, max_batch_size=2**14)

### 2017-03-06 twoface/sample_prior.py
test and docstring

@@ -6,6 +6,17 @@ from thejoker.sampler import pack_prior_samples
 def make_prior_cache(filename, joker, N, max_batch_size=2**24):
     """
 
+    Parameters
+    ----------
+    filename : str
+        The HDF5 file name to cache to.
+    joker : `~thejoker.sampler.TheJoker`
+        An instance of ``TheJoker``.
+    N : int
+        Number of samples to generate in the cache.
+    max_batch_size : int (optional)
+        The batch size to generate each iteration.
+
     """
 
     max_batch_size = int(max_batch_size)

### 2017-03-06 twoface/tests/test_sample_prior.py

@@ -1,19 +1,19 @@
-# Standard library
-import os
-import sys
-
 # Third-party
 import astropy.units as u
-import numpy as np
+import h5py
 from thejoker.sampler import TheJoker, JokerParams
 
 # Project
 from ..sample_prior import make_prior_cache
 
 def test_make_prior_cache(tmpdir):
+    N = 2**16
 
     filename = str(tmpdir / 'prior_samples.h5')
     params = JokerParams(P_min=8*u.day, P_max=8192*u.day)
     joker = TheJoker(params)
 
-    make_prior_cache(filename, joker, N=2**16, max_batch_size=2**14)
+    make_prior_cache(filename, joker, N=N, max_batch_size=2**14)
+
+    with h5py.File(filename, 'r') as f:
+        assert f['samples'].shape == (N, 5)

### 2017-03-06 scripts/run_redclump.py
use make_prior_cache in script

@@ -20,6 +20,7 @@ from twoface.log import log as logger
 from twoface.db import Session, db_connect
 from twoface.db import JokerRun, AllStar, AllVisit, StarResult, Status, AllVisitToAllStar, RedClump
 from twoface.config import TWOFACE_CACHE_PATH
+from twoface.sample_prior import make_prior_cache
 
 def main(config_file, pool, seed, overwrite=False, _continue=False):
     config_file = abspath(expanduser(config_file))
@@ -100,11 +101,8 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
     # create prior samples cache, store to file and store filename in DB
     if not os.path.exists(run.prior_samples_file):
         logger.debug("Prior samples file not found - generating now...")
-
-        # TODO: prior samples
-        # TODO: for large prior samples libraries, may want a way to write
-        #   without storing all samples in memory
-
+        make_prior_cache(run.prior_samples_file, joker, N=run.max_prior_samples,
+                         max_batch_size=2**22)
         logger.debug("...done")
 
     else:

### 2017-03-06 twoface/tests/test_sample_prior.py

@@ -7,7 +7,7 @@ from thejoker.sampler import TheJoker, JokerParams
 from ..sample_prior import make_prior_cache
 
 def test_make_prior_cache(tmpdir):
-    N = 2**16
+    N = 2**18
 
     filename = str(tmpdir / 'prior_samples.h5')
     params = JokerParams(P_min=8*u.day, P_max=8192*u.day)

### 2017-03-06 twoface/db/model.py
test cascade

@@ -7,7 +7,7 @@ import datetime
 import astropy.units as u
 from sqlalchemy import Table, Column, types
 from sqlalchemy.schema import ForeignKey
-from sqlalchemy.orm import relationship
+from sqlalchemy.orm import relationship, backref
 
 # Project
 from .core import Base
@@ -33,7 +33,8 @@ class StarResult(Base):
                        default=0)
 
     star = relationship("AllStar")
-    jokerrun = relationship("JokerRun")
+    jokerrun = relationship("JokerRun",
+                            backref=backref("results", cascade="all, delete-orphan"))
     status = relationship("Status")
 
     def __repr__(self):
@@ -561,6 +562,9 @@ class JokerRun(Base):
 
     # The Joker parameters:
     jitter = Column('jitter', JitterType, default=float('nan')) # nan = not fixed
+    jitter_mean = Column('jitter_mean', types.Numeric, default=float('nan'))
+    jitter_stddev = Column('jitter_stddev', types.Numeric, default=float('nan'))
+    jitter_unit = Column('jitter_unit', types.String, default="")
 
     P_min = Column('P_min', PeriodType, nullable=False)
     P_max = Column('P_max', PeriodType, nullable=False)

### 2017-03-06 twoface/db/tests/test_database.py

@@ -1,9 +1,8 @@
-from __future__ import division, print_function
-
 # Standard library
 import os
 
 # Third-party
+import astropy.units as u
 from astropy.utils.data import get_pkg_data_filename
 import pytest
 import yaml
@@ -41,6 +40,40 @@ class TestDB(object):
         # get a visit and check that it has one star
         visit = session.query(AllVisit).filter(AllVisit.target_id == test_target_ID).limit(1).one()
         assert len(visit.stars) == 2
+
+        session.close()
+
+    def test_jokerrun_cascade(self):
+        # make sure the Results and Statuses are deleted when a JokerRun is deleted
+
+        NAME = 'test-cascade'
+        session = Session()
+
+        # first set up:
+        stars = session.query(AllStar).all()
+
+        run = JokerRun()
+        run.config_file = ''
+        run.name = NAME
+        run.P_min = 1.*u.day
+        run.P_max = 100.*u.day
+        run.requested_samples_per_star = 128
+        run.max_prior_samples = 1024
+        run.prior_samples_file = ''
+        run.stars = stars
+        session.add(run)
+        session.commit()
+
+        assert session.query(JokerRun).count() == 1
+        assert session.query(AllStar).count() == session.query(StarResult).count()
+
+        for run in session.query(JokerRun).filter(JokerRun.name == NAME).all():
+            session.delete(run)
+        session.commit()
+
+        assert session.query(JokerRun).count() == 0
+        assert session.query(StarResult).count() == 0
+        assert session.query(Status).count() > 0 # just to be safe
 
         session.close()
 

### 2017-03-06 twoface/sample_prior.py
fix logic

@@ -28,13 +28,13 @@ def make_prior_cache(filename, joker, N, max_batch_size=2**24):
 
     num_added = 0
     for i in range(10000): # HACK: magic number, maximum num. iterations
-        samples, ln_probs = joker.sample_prior(max_batch_size, return_logprobs=True)
+        samples, ln_probs = joker.sample_prior(min(max_batch_size, N), return_logprobs=True)
         packed_samples, units = pack_prior_samples(samples, u.km/u.s) # TODO: make rv_unit configurable?
 
         batch_size,K = packed_samples.shape
 
         if (num_added + batch_size) > N:
-            packed_samples = packed_samples[:N - (num_added + batch_size)+1]
+            packed_samples = packed_samples[:N - (num_added + batch_size)]
             batch_size,K = packed_samples.shape
             if batch_size <= 0:
                 break
@@ -49,7 +49,10 @@ def make_prior_cache(filename, joker, N, max_batch_size=2**24):
             i1 = num_added
             i2 = num_added+batch_size
 
-            f['samples'][i1:i2,:] = packed_samples
-            f['ln_prior_probs'][i1:i2] = ln_probs
+            f['samples'][i1:i2,:] = packed_samples[i1:i2]
+            f['ln_prior_probs'][i1:i2] = ln_probs[i1:i2]
 
         num_added += batch_size
+
+        if num_added >= N:
+            break

### 2017-03-06 twoface/data.py
remove apogee systematic error

@@ -20,9 +20,8 @@ class APOGEERVData(RVData):
         jd = [float(v.jd) for v in visits]
         rv = [float(v.vhelio) for v in visits]
         rv_rand_err = [float(v.vrelerr) for v in visits]
-        noise_floor = 100 * u.m/u.s # from APOGEE team
 
         t = Time(jd, format='jd', scale='utc')
         rv = rv * u.km/u.s
-        rv_err = np.sqrt(noise_floor**2 + (rv_rand_err*u.km/u.s)**2)
+        rv_err = rv_rand_err*u.km/u.s
         return cls(t=t, rv=rv, stddev=rv_err)

### 2017-03-06 scripts/run_redclump.py
modify redclump run

@@ -67,13 +67,15 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
             run.jitter = u.Quantity(*config['hyperparams']['jitter'].split())
 
         elif 'jitter_prior_mean' in config['hyperparams']:
-            raise NotImplementedError("NOT YET SORRY")
+            run.jitter_mean = config['hyperparams']['jitter_prior_mean']
+            run.jitter_stddev = config['hyperparams']['jitter_prior_stddev']
+            run.jitter_unit = config['hyperparams']['jitter_prior_unit']
 
         else:
-            # no jitter specified, set to extra suggested systemic velocity
-            run.jitter = 130. * u.m/u.s
+            # no jitter specified
+            run.jitter = 0. * u.m/u.s
 
-        # TODO: need a way to pass in velocity trend info
+        # TODO: need a way to pass in velocity trend specification - right now, assumed flat
 
         stars = session.query(AllStar).join(AllVisitToAllStar, AllVisit, RedClump)\
                                       .group_by(AllStar.apstar_id)\
@@ -91,23 +93,30 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
     else:
         raise ValueError("Multiple JokerRun rows found for name '{}'".format(config['name']))
 
+    if run.jitter is None or np.isnan(run.jitter):
+        jitter_kwargs = dict(jitter=(float(run.jitter_mean), float(run.jitter_stddev)),
+                             jitter_unit=u.Unit(run.jitter_unit))
+
+    else:
+        jitter_kwargs = dict(jitter=run.jitter)
+
     # create TheJoker sampler instance
     rnd = np.random.RandomState(seed=seed)
     logger.debug("Creating TheJoker instance with {}, {}".format(rnd, pool))
-    params = JokerParams(P_min=run.P_min, P_max=run.P_max, jitter=run.jitter,
-                         anomaly_tol=1E-11)
+    params = JokerParams(P_min=run.P_min, P_max=run.P_max, anomaly_tol=1E-11,
+                         **jitter_kwargs)
     joker = TheJoker(params, random_state=rnd, pool=pool)
 
     # create prior samples cache, store to file and store filename in DB
     if not os.path.exists(run.prior_samples_file):
         logger.debug("Prior samples file not found - generating now...")
-        make_prior_cache(run.prior_samples_file, joker, N=run.max_prior_samples,
[…]

### 2017-07-29 paper/main.tex
outline text

@@
 % - Use \sectionname not Section, \figname not Figure, \documentname not Article or Paper or paper.
+% - Hyphenate binary-star when it is an adjective, not when it is a noun!
 
@@
 \include{preamble}
+
+% adjust ASS-TEX shit
+\setlength{\parindent}{1.\baselineskip}
 
@@
 
-\begin{abstract}
+\begin{abstract}\noindent % trust me
 % Context
@@
 binary-star population---that is, if we want to get (say) the mass,
-mass ratio, and period distribution of binary star systems---we face
+mass ratio, and period distribution of binary-star systems---we face
 (at least) two challenges.
@@
 
-HOGG: What is up with delivering a catalog (it requires decision-making).
+HOGG: What is up with delivering a catalog (it requires decision-making)?
 
-\section{Method: Detection and orbit fitting}
+\section{Data}
+
+Overview of what \apogee\ is, and what data we are using.
+
+How we updated the individual-visit radial velocity measurements.
+
+What we did (if anything) with the missing visits.
+
+\section{Method: Detection and orbit fitting}\label{sec:fitting}
+
+APW: Should we start this section with a set of named assumptions, as we
+did in Anderson et al?
+
+Summary of \thejoker\ and what it does.
+
+Any modifications or knob-settings for \thejoker?
 
@@
 
-\section{Method: Population inferences}
+\section{Method: Population inferences}\label{sec:popinference}
+
+Now we want to take the individual-star fits---or samplings---that
+we made in \sectionname~\ref{sec:fitting} and, from these, produce
+an inference of the properties of the whole \RC\ binary-star population.
+Important necessary properties of this inference are the following:
+\begin{description}
+\item[foo] whatevs
+\item[bar] and other
+\end{description}
+We can meet these requirements with a hierarchical Bayesian inference
+(hierarchical probablistic model) of the binary-star population.
+
+In addition to the above requirements, we make the following assumptions
+that restrict our attention to a well-defined model space:
+\begin{description}
+\item[foo] whatevs
+\item[bar] and other
+\end{description}
+
+The binary-system parameters for each individual \RC\ star were
+inferred (in \sectionname~\ref{sec:fitting}) with an interim prior that
+is very simple and doesn't represent our true beliefs about the
+binary-star population...
+
+In a hierarchical Bayesian inference, we replace the binary-star
+system priors used in \sectionname~\ref{sec:fitting} with a parameterized
+model for the binary-star population...
+
+Because all we have is a sampling of each system under the interim prior,
+we have to re-weight or importance-sample the individual-star samplings
+in this hierarchical inference... Cite Hogg et al; Cite DFM et al.
 

### 2017-07-29 twoface/db/tests/test_database.py
add astropy to deps

@@ -4,7 +4,6 @@ import os
 # Third-party
 import astropy.units as u
 from astropy.utils.data import get_pkg_data_filename
-import pytest
 import yaml
 
 # Project

### 2017-07-29 paper/main.tex
description lists

@@
 \begin{description}
-\item[foo] whatevs
-\item[bar] and other
+\item[utilize non-detections] Even stars which are consistent with
+  zero radial-velocity variation (no binary companion) are relevant to
+  a populations inference. This is for two reasons: The first is that
+  the small-mass and long-period ends of the binary population will
+  imprint only low-amplitude signals in the data. These signals are
+  present at low signal-to-noise, but not if we remove them because
+  they don't make some catalog or threshold cut. The second reason
+  these stars are important is that they constrain the fraction of
+  stars with no companions. Any populations-inference method we employ
+  must responsibly harvest the information in these non-detection
+  stars.
+\item[marginalize out individual-star parameters] Every star---even
+  any star with a high-confidence companion detection---will have a
+  highly multi-modal likelihood function, with many qualitatively
+  different companion models that are nonetheless locally
+  optimal. This means that most posterior samplings (created in
+  \sectionname~\ref{sec:fitting}) under our interim prior will show
+  multiple, qualitatively distinct modes in binary-system parameter
+  space.  Any correct population inference will marginalize out these
+  extremely non-trivial distributions over individual-star
+  binary-system parameters.
 \end{description}
-We can meet these requirements with a hierarchical Bayesian inference
-(hierarchical probablistic model) of the binary-star population.
+We can meet these requirements---at least in principle---with a
+hierarchical Bayesian inference (hierarchical probablistic model) of
+the binary-star population.
 
@@
 \begin{description}
-\item[foo] whatevs
-\item[bar] and other
+\item[correct samplings] accurate representations of posterior pdf
+  given prior and likelihood; correct likelihood function.
+\item[no trinaries] related to the above: no trinaries, quads, or anything
+  crazier.
+\item[sufficent samplings] density and support of samplings
+\item[independent stars] no relation between one star and another; each
+  drawn from the same population; no multiple populations.
+\item[population parameterization] whatevs
 \end{description}

### 2017-07-29 paper/main.tex
add some bullshit at the end (acks, software, facility)

--- 2017-07-29 paper/main.tex
@@
 
+\acknowledgements
+
+It is a pleasure to thank XX.
+
+The authors are pleased to acknowledge that the work reported on in this
+paper was substantially performed at the TIGRESS high performance computer
+center at Princeton University which is jointly supported by the Princeton
+Institute for Computational Science and Engineering and the Princeton
+University Office of Information Technology's Research Computing department.
+
+\software{
+The code used in this project is available from
+\url{https://github.com/adrn/TwoFace} under the MIT open-source
+software license.
+This research utilized the following open-source \python\ packages:
+    \package{Astropy} (\citealt{Astropy-Collaboration:2013}),
+    % \package{corner} (\citealt{Foreman-Mackey:2016}),
+    % \package{emcee} (\citealt{Foreman-Mackey:2013ascl}),
+    \package{IPython} (\citealt{Perez:2007}),
+    \package{matplotlib} (\citealt{Hunter:2007}),
+    \package{numpy} (\citealt{Van-der-Walt:2011}),
+    \package{scipy} (\url{https://www.scipy.org/}),
+    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}).
+}
+
+\facility{\sdssiii, \apogee}
+
+\bibliographystyle{aasjournal}
+\bibliography{refs}
+
 \end{document}

### 2017-07-29 scripts/run_redclump.py
cleaning up pep8

@@ -18,7 +18,8 @@ import yaml
 from twoface.data import APOGEERVData
 from twoface.log import log as logger
 from twoface.db import Session, db_connect
-from twoface.db import JokerRun, AllStar, AllVisit, StarResult, Status, AllVisitToAllStar, RedClump
+from twoface.db import (JokerRun, AllStar, AllVisit, StarResult, Status,
+                        AllVisitToAllStar, RedClump)
 from twoface.config import TWOFACE_CACHE_PATH
 from twoface.sample_prior import make_prior_cache
 
@@ -39,8 +40,8 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
     db_path = join(TWOFACE_CACHE_PATH, database_file)
 
     if not os.path.exists(db_path):
-        raise IOError("sqlite database not found at '{}'\n Did you run scripts/initdb.py yet?"
-                      .format(db_path))
+        raise IOError("sqlite database not found at '{}'\n Did you run "
+                      "scripts/initdb.py yet?".format(db_path))
 
     logger.debug("Connecting to sqlite database at '{}'".format(db_path))
     engine = db_connect(database_path=db_path, ensure_db_exists=False)
@@ -50,7 +51,7 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
     # see if this run (by name) is in db already, if so, just grab
     _run = session.query(JokerRun).filter(JokerRun.name == config['name']).all()
     if len(_run) == 0:
-        logger.info("JokerRun '{}' not found in database - creating new entry..."
+        logger.info("JokerRun '{}' not found in database, creating new entry..."
                     .format(config['name']))
 
         # create a JokerRun for this run
@@ -59,9 +60,11 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
         run.name = config['name']
         run.P_min = u.Quantity(*config['hyperparams']['P_min'].split())
         run.P_max = u.Quantity(*config['hyperparams']['P_max'].split())
-        run.requested_samples_per_star = int(config['hyperparams']['requested_samples_per_star'])
+        run.requested_samples_per_star = int(
+            config['hyperparams']['requested_samples_per_star'])
         run.max_prior_samples = int(config['prior']['max_samples'])
-        run.prior_samples_file = join(TWOFACE_CACHE_PATH, config['prior']['samples_file'])
+        run.prior_samples_file = join(TWOFACE_CACHE_PATH,
+                                      config['prior']['samples_file'])
 
         if 'jitter' in config['hyperparams']:
[…]

### 2017-07-29 twoface/sample_prior.py

@@ -27,7 +27,7 @@ def make_prior_cache(filename, joker, N, max_batch_size=2**24):
         pass
 
     num_added = 0
-    for i in range(10000): # HACK: magic number, maximum num. iterations
+    for i in range(2**16): # HACK: magic number, maximum num. iterations
         samples, ln_probs = joker.sample_prior(min(max_batch_size, N), return_logprobs=True)
         packed_samples, units = pack_prior_samples(samples, u.km/u.s) # TODO: make rv_unit configurable?
 

### 2017-07-29 scripts/initdb.py
change howinitialization is done

@@ -1,148 +1,11 @@
-__author__ = "adrn <adrn@astro.princeton.edu>"
-
-# Standard library
-import os
-from os.path import abspath, expanduser, join, dirname
-
-# Third-party
-from astropy.table import Table
-import numpy as np
-
 # Project
-import twoface
-from twoface.util import Timer
 from twoface.log import log as logger
-from twoface.db import Session, db_connect, AllStar, AllVisit, Status, RedClump
-from twoface.db.core import Base
-from twoface.config import TWOFACE_CACHE_PATH
+from TwoFace.db.init import initialize_db
 
-def main(allVisit_file=None, allStar_file=None, rc_file=None, test=False, **kwargs):
-
-    # if running in test mode, get test files
-    if test:
-        base_path = abspath(dirname(twoface.__file__))
-        allVisit_file = os.path.join(base_path, 'db', 'tests', 'test-allVisit.fits')
-        allStar_file = os.path.join(base_path, 'db', 'tests', 'test-allStar.fits')
-        rc_file = os.path.join(base_path, 'db', 'tests', 'test-rc.fits')
-
-        database_path = join(TWOFACE_CACHE_PATH, 'test.sqlite')
-
-    else:
-        assert allVisit_file is not None
-        assert allStar_file is not None
-        assert rc_file is not None
-
-        database_path = join(TWOFACE_CACHE_PATH, 'apogee.sqlite')
-
-    norm = lambda x: abspath(expanduser(x))
-    allvisit_tbl = Table.read(norm(allVisit_file), format='fits', hdu=1)
-    allvisit_tbl = allvisit_tbl[np.isfinite(allvisit_tbl['VHELIO'])]
-
-    allstar_tbl = Table.read(norm(allStar_file), format='fits', hdu=1)
-    rc_tbl = Table.read(norm(rc_file), format='fits', hdu=1)
-
-    engine = db_connect(database_path)
-    # engine.echo = True
-    logger.debug("Connected to database at '{}'".format(database_path))
-
-    # this is the magic that creates the tables based on the definitions in twoface/db/model.py
-    Base.metadata.drop_all()
-    Base.metadata.create_all()
-
-    session = Session()
-
-    logger.debug("Loading allStar, allVisit tables...")
-
-    allstar_colnames = [str(x).split('.')[1].upper() for x in AllStar.__table__.columns]
-    i = allstar_colnames.index('ID')
[…]

### 2017-07-29 twoface/db/__init__.py

@@ -1,2 +1,3 @@
-from .core import db_connect, Session
+from .connect import db_connect, Session
 from .model import *
+from .init import initialize_db

### 2017-07-29 twoface/db/model.py

@@ -10,7 +10,7 @@ from sqlalchemy.schema import ForeignKey
 from sqlalchemy.orm import relationship, backref
 
 # Project
-from .core import Base
+from .connect import Base
 from .quantity_type import QuantityTypeClassFactory
 from . import numpy_adapt # just need to execute code
 
@@ -361,63 +361,64 @@ class RedClump(Base):
     allstar_id = Column("allstar_id", types.Integer, ForeignKey('allstar.id'))
     star = relationship("AllStar")
 
-    apstar_id = Column("apstar_id", types.String)
-    aspcap_id = Column("aspcap_id", types.String)
-    apogee_id = Column("apogee_id", types.String)
-    telescope = Column("telescope", types.String)
-    location_id = Column("location_id", types.SmallInteger)
-    field = Column("field", types.String)
-    j = Column("j", types.REAL)
-    j_err = Column("j_err", types.REAL)
-    h = Column("h", types.REAL)
-    h_err = Column("h_err", types.REAL)
-    k = Column("k", types.REAL)
-    k_err = Column("k_err", types.REAL)
-    ra = Column("ra", types.Numeric)
-    dec = Column("dec", types.Numeric)
-    glon = Column("glon", types.Numeric)
-    glat = Column("glat", types.Numeric)
-    apogee_target1 = Column("apogee_target1", types.Integer)
-    apogee_target2 = Column("apogee_target2", types.Integer)
-    apogee_target3 = Column("apogee_target3", types.Integer)
-    targflags = Column("targflags", types.String)
-    nvisits = Column("nvisits", types.Integer)
-    commiss = Column("commiss", types.SmallInteger)
-    snr = Column("snr", types.REAL)
-    starflag = Column("starflag", types.Integer)
-    starflags = Column("starflags", types.String)
-    andflag = Column("andflag", types.Integer)
-    andflags = Column("andflags", types.String)
-    vhelio_avg = Column("vhelio_avg", types.REAL)
-    vscatter = Column("vscatter", types.REAL)
-    verr = Column("verr", types.REAL)
-    verr_med = Column("verr_med", types.REAL)
[…]

### 2017-07-29 twoface/db/tests/test_database.py

@@ -1,5 +1,5 @@
 # Standard library
-import os
+from os import path, unlink
 
 # Third-party
 import astropy.units as u
@@ -10,6 +10,7 @@ import yaml
 from ...config import TWOFACE_CACHE_PATH
 from ..core import db_connect, Session
 from ..model import AllStar, AllVisit, StarResult, Status, JokerRun
+from ..init import initialize_db
 
 class TestDB(object):
 
@@ -18,38 +19,43 @@ class TestDB(object):
         with open(get_pkg_data_filename('travis_db.yml')) as f:
             config = yaml.load(f)
 
-        self.db_path = os.path.join(TWOFACE_CACHE_PATH, config['database_file'])
-        if not os.path.exists(self.db_path):
-            cmd = "python scripts/initdb.py --test -v"
-            raise IOError("Test database file doest not exist! Before running the tests, you "
-                          "should run: \n {}".format(cmd))
+        self.db_path = path.join(TWOFACE_CACHE_PATH, config['database_file'])
+
+        # initialize the database
+        initialize_db(allVisit_file=get_pkg_data_filename('test-allVisit.fits'),
+                      allStar_file=get_pkg_data_filename('test-allStar.fits'),
+                      rc_file=get_pkg_data_filename('test-rc.fits'),
+                      database_file=self.db_path,
+                      drop_all=True)
 
         self.engine = db_connect(self.db_path)
+        self.session = Session()
 
     def test_one(self):
+        s = self.session
+
         # a target in both test FITS files included in the repo
         test_target_ID = "4264.2M00000032+5737103"
 
-        session = Session()
-
         # get star entry and check total num of visits
-        star = session.query(AllStar).filter(AllStar.target_id == test_target_ID).one()
+        star = s.query(AllStar).filter(AllStar.target_id == test_target_ID).one()
         assert len(star.visits) == 6
 
         # get a visit and check that it has one star
-        visit = session.query(AllVisit).filter(AllVisit.target_id == test_target_ID).limit(1).one()
+        visit = s.query(AllVisit).filter(AllVisit.target_id == test_target_ID).limit(1).one()
         assert len(visit.stars) == 2
 
-        session.close()
-
     def test_jokerrun_cascade(self):
-        # make sure the Results and Statuses are deleted when a JokerRun is deleted
+        """
+        Make sure the Results and Statuses are deleted when a JokerRun is
+        deleted
+        """
+        s = self.session
 
         NAME = 'test-cascade'
-        session = Session()
 
[…]

### 2017-07-31 paper/main.tex
equations

--- 2017-07-29 paper/main.tex
@@
 \item[correct samplings] accurate representations of posterior pdf
-  given prior and likelihood; correct likelihood function.
+  given prior and likelihood; correct likelihood function. We hereby
+  take on all the assumptions given in \sectionname~\ref{sec:fitting},
+  about noise and kinematics, and so on.
 \item[no trinaries] related to the above: no trinaries, quads, or anything
@@
 
-Because all we have is a sampling of each system under the interim prior,
+Because all we have is a sampling of each system under the interim prior
+$p(\omega\given H_0)$,
 we have to re-weight or importance-sample the individual-star samplings
 in this hierarchical inference... Cite Hogg et al; Cite DFM et al.
+% The following needs checking for symbol conventions
+\begin{eqnarray}
+  \ln p(y\given\alpha,H_1) &=& \sum_{n=1}^N \ln p(y_n\given\alpha,H_1)
+  \\
+  \ln p(y_n\given\alpha,H_1) &\propto& \ln\left[\frac{1}{K_n}\,\sum_{k=1}^{K_n}\frac{p(\omega_{nk}\given\alpha,H_1)}{p(\omega_{nk}\given H_0)}\right]
+  \quad .
+\end{eqnarray}
 

### 2017-07-31 scripts/run_redclump.py
ack

@@ -51,7 +51,7 @@ import yaml
 # Project
 from twoface.data import APOGEERVData
 from twoface.log import log as logger
-from twoface.db import Session, db_connect
+from twoface.db import db_connect
 from twoface.db import (JokerRun, AllStar, AllVisit, StarResult, Status,
                         AllVisitToAllStar, RedClump)
 from twoface.config import TWOFACE_CACHE_PATH
@@ -78,7 +78,8 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
                       .format(db_path))
 
     logger.debug("Connecting to sqlite database at '{0}'".format(db_path))
-    engine = db_connect(database_path=db_path, ensure_db_exists=False)
+    Session, engine = db_connect(database_path=db_path,
+                                 ensure_db_exists=False)
     session = Session()
 
     # See if this run (by name) is in the database already, if so, grab that.
@@ -215,15 +216,16 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
         n_samples = 0 # running total of good samples returned
         start_idx = 0
         all_samples = None
+        all_ln_probs = np.array([])
         for n in range(128): # MAGIC NUMBER: but we should never hit this
             logger.debug("\t Iteration {}, using {} prior samples"
                          .format(n, n_process))
 
             # Process n_process samples from prior cache, pulled from the file
             # starting from the index `start_idx`
-            samples = joker._rejection_sample_from_cache(data, n_process,
-                                                         run.prior_samples_file,
-                                                         start_idx)
+            samples, ln_probs = joker._rejection_sample_from_cache(
+                data, n_process, run.prior_samples_file, start_idx,
+                return_logprobs=True)
 
             # Number of samples that survive the rejection sampling step
             n_survive = samples.shape[0]
@@ -236,6 +238,8 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
                     all_samples = samples
                 else:
                     all_samples = np.vstack((all_samples, samples))
+
+                all_ln_probs = np.concatenate((all_ln_probs, ln_probs))
 
             # End processing for this star if the number of samples we have is
             # larger than or equal to the number requested per star
@@ -258,6 +262,7 @@ def main(config_file, pool, seed, overwrite=False, _continue=False):
 
                 if all_samples is None:
                     all_samples = samples
+                    all_ln_probs = np.array(ln_probs)
 
[…]

### 2017-07-31 twoface/db/init.py

@@ -10,7 +10,8 @@ from ..config import TWOFACE_CACHE_PATH
 from ..util import Timer
 from ..log import log as logger
 from .connect import db_connect, Base
-from .model import AllStar, AllVisit, AllVisitToAllStar, Status, RedClump
+from .model import (AllStar, AllVisit, AllVisitToAllStar, Status, RedClump,
+                    StarResult)
 
 __all__ = ['initialize_db', 'load_red_clump']
 
@@ -202,13 +203,14 @@ def load_red_clump(filename, database_file, overwrite=False, batch_size=4096):
             # Retrieve the parent AllStar record
             try:
                 star = session.query(AllStar).filter(
-                    AllStar.apstar_id == row_data['apstar_id']).one()
+                    AllStar.apstar_id == row['APSTAR_ID']).one()
             except:
+                logger.debug('Red clump star not found in AllStar - skipping')
                 continue
 
-            if row['apstar_id'] in rc_ap_ids:
+            if row['APSTAR_ID'] in rc_ap_ids:
                 q = session.query(RedClump).join(AllStar).filter(
-                    AllStar.apstar_id == row['apstar_id'])
+                    AllStar.apstar_id == row['APSTAR_ID'])
 
                 if overwrite:
                     q.delete()

### 2017-07-31 twoface/db/model.py

@@ -23,6 +23,9 @@ class Status(Base):
     id = Column(types.Integer, primary_key=True, autoincrement=False)
     message = Column('message', types.String)
 
+    def __repr__(self):
+        return "<Status id={0} message='{1}'>".format(self.id, self.message)
+
 class StarResult(Base):
     __tablename__ = 'starresult'
 

### 2017-08-04 paper/main.tex
content

@@
 
-% adjust ASS-TEX shit
+% adjust AAS-TEX shit
 \setlength{\parindent}{1.\baselineskip}
@@
 
+APW: Opening sentence: If we see RV changes, there must be binaries!
+Can we infer binary populations from bad data?
+
 If we want to use a spectroscopic survey to responsibly measure the
@@
 
-APW: Should we start this section with a set of named assumptions, as we
-did in Anderson et al?
+Our approach here is to proceed in two phases.
+In the first phase, we obtain a posterior sampling in binary-system
+parameter space for every individual star, treating it as a
+single-lined (SB1) spectroscopic binary system with a single
+companion.
+This sampling is performed under a relatively uninformative prior pdf,
+which will be called the ``\emph{interim prior}''.
+These samplings are used to discover and characterize individual binary-star
+systems, and generate (among other things) a catalog of binaries.
+In the second phase, we perform a hierarchical inference of the binary
+population, using (responsibly) these samplings under the interim
+prior as inputs.
+In the second phase we use all samplings of all stars, not just those that
+pass our thresholds for discovery as binary systems.
+In this \sectionname, we describe the first phase, that is, the
+individual-system sampling phase for discovery and characterization.
 
-Summary of \thejoker\ and what it does.
+We make our posterior samplings under the following assumptions:
+\begin{description}
+\item[foo] whatevs
+\item[bar] and more
+\end{description}
+
+Summary of \thejoker\ and what it does; that is, that it obeys the
+relevant parts of our assumptions.
 
@@
 
+Examples of outputs and etc.
+
 \section{A catalog of red-clump binary systems}
+
+Thresholding on what now?
+
+Catalog.
+
+Some highlights from this catalog.
 

### 2017-08-04 paper/main.tex
assumptions listed (but not described)

@@
 % - Use \sectionname not Section, \figname not Figure, \documentname not Article or Paper or paper.
+% - Make sure that two assumptions in the two assumption lists only have the same name if they really are the same assumption: These are proper names!
 % - Hyphenate binary-star when it is an adjective, not when it is a noun!
@@
 
-We make our posterior samplings under the following assumptions:
+We perform our individual system fits---that is, make our posterior
+samplings---under the following assumptions:
 \begin{description}
-\item[foo] whatevs
-\item[bar] and more
+\item[no trinaries] whatevs: No trinaries or beyond. Obviously wrong!
+\item[SB1] whatevs: There is no significant light contributed by the secondary.
+\item[Kepler] whatevs: No non-gravitational contributions to RV history.
+\item[noise model] whatevs: Measurements are unbiased and noise estimates are correct up to jitter; everything Gaussian.
+\item[interim prior] We adopt a prior pdf on binary-system parameters
+  with the following properties...
 \end{description}
@@
 
+Return to our various assumptions and make sure that we discuss them
+\emph{by name} here.
+
 \acknowledgements

### 2017-08-04 paper/main.tex
authors
equations of pain

@@
 % adjust AAS-TEX shit
-\setlength{\parindent}{1.\baselineskip}
+\setlength{\parindent}{1.1\baselineskip}
 
@@
 
-\author{David~W.~Hogg}
+\author[0000-0003-2866-9403]{David~W.~Hogg}
+\affiliation{Max-Planck-Institut f\"ur Astronomie,
+             K\"onigstuhl 17, D-69117 Heidelberg, Germany}
 \affiliation{Center for Cosmology and Particle Physics,
@@
              New York, NY 10003, USA}
-\affiliation{Max-Planck-Institut f\"ur Astronomie,
-             K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+\affiliation{Center for Data Science,
+             New York University, 60 Fifth Ave,
+             New York, NY 10011, USA}
 \affiliation{Flatiron Institute,
@@
              K\"onigstuhl 17, D-69117 Heidelberg, Germany}
+
+\author{Jason~Cao}
+\affiliation{Center for Cosmology and Particle Physics,
+             Department of Physics,
+             New York University, 726 Broadway,
+             New York, NY 10003, USA}
 
@@
 
-APW: Opening sentence: If we see RV changes, there must be binaries!
-Can we infer binary populations from bad data?
+APW: Opening sentence: Even with one single star and two epochs, if we
+see an RV change, there must be binaries! What more can we say? Can we
+infer binary populations from bad data?
 
@@
   drawn from the same population; no multiple populations.
-\item[population parameterization] whatevs
+\item[masses known] The mass of each primary \RC\ star is known, with
+  a correctly known posterior pdf, from prior work (CITE NESS).
+\item[population parameterization] We adopt the following
+  parameterization of the population. In what follows, $\theta$ will
+  stand in for the full set of hyper-parameters (population
+  parameters).
+  \begin{itemize}
+  \item Each primary \RC\ star has a probability $F$ (pure probability
+    $0<F<1$) of having a secondary companion. $F$ is a hyper-parameter
+    in $\theta$.
+  \item The log-period distribution $p(\ln P\given\theta)$ is a power
+    law between end-points, or
+    \begin{eqnarray}
+      \ln p(\ln P\given\theta) &=& g_0 + (g_1 - g_0)\,\frac{\ln P - x_0}{x_1 - x_0}
+      \quad \mbox{for $x_0<\ln P<x_1$,}
+    \end{eqnarray}
+    where $g_0, g_1, x_0, x_1$ are hyper-parameters in $\theta$.
+  \item Each primary star has a true mass $M$ (which we only know
+    noisily) and the companion mass is potentially larger or
+    smaller (!). The log-mass-ratio distribution $p(\ln q\given\theta)$ is
+    \begin{eqnarray}
+      \ln p(\ln q\given\theta) &=& h_0 + (h_1 - h_0)\,\frac{\ln q - y_0}{y_1 - y_0}
+      \quad \mbox{for $y_0<\ln q<y_1$,}
+    \end{eqnarray}
+    where $h_0, h_1, y_0, y_1$ are hyper-parameters in $\theta$.
+  \item The eccentricity distribution $p(e\given P,\theta)$ is a beta,
+    and permitted to vary with period:
+    \begin{eqnarray}
+      p(e\given P,\theta) &=& B(e\given \alpha(P),\beta(P))
+      \\
+      \alpha(P) &=& a_0 + (a_1 - a_0)\,\frac{P - x_0}{x_1 - x_0}
+      \\
+      \beta(P) &=& b_0 + (b_1 - b_0)\,\frac{P - x_0}{x_1 - x_0}
+      \quad ,
+    \end{eqnarray}
+    where $B(x\given\alpha,\beta)$ is the beta distribution with
+    control parameters $\alpha,\beta$, and $a_0, a_1, b_0, b_1$ are
+    hyper-parameters in $\theta$.
+  \item The inclination distribution $p(i)$ and the argument-of-perihelion
+    distribution $p(\omega)$ and the phase distribution $p(\phi)$ are all
+    fixed to their isotropic distributions. They bring no hyper-parameters
+    to $\theta$.
+  \end{itemize}
+\item[hyper-priors] whatevs
 \end{description}

### 2017-08-05 paper/main.tex
parameter names and definitions

@@
 % - Hyphenate binary-star when it is an adjective, not when it is a noun!
+% - Where there are defined math symbols (like \pars), use them!
 
@@
 \newcommand{\RGB}{\acronym{RGB}}
+
+% define macros for math
+\newcommand{\hyperpars}{\gamma}
+\newcommand{\pars}{\theta}
 
@@
 \item[population parameterization] We adopt the following
-  parameterization of the population. In what follows, $\theta$ will
+  parameterization of the population. In what follows, $\hyperpars$ will
   stand in for the full set of hyper-parameters (population
@@
     $0<F<1$) of having a secondary companion. $F$ is a hyper-parameter
-    in $\theta$.
-  \item The log-period distribution $p(\ln P\given\theta)$ is a power
-    law between end-points, or
+    in $\hyperpars$.
+  \item For those with secondary companions, the log-period $\ln P$
+    distribution $p(\ln P\given\hyperpars)$ is a power law between
+    end-points, or
     \begin{eqnarray}
-      \ln p(\ln P\given\theta) &=& g_0 + (g_1 - g_0)\,\frac{\ln P - x_0}{x_1 - x_0}
+      \ln p(\ln P\given\hyperpars) &=& g_0 + (g_1 - g_0)\,\frac{\ln P - x_0}{x_1 - x_0}
       \quad \mbox{for $x_0<\ln P<x_1$,}
     \end{eqnarray}
-    where $g_0, g_1, x_0, x_1$ are hyper-parameters in $\theta$.
+    where $g_0, g_1, x_0, x_1$ are hyper-parameters in $\hyperpars$.
+    Parameters $g_0, g_1$ are natural logs of densities in log-period space.
+    Parameters $x_0, x_1$ are log-period limits.
   \item Each primary star has a true mass $M$ (which we only know
     noisily) and the companion mass is potentially larger or
-    smaller (!). The log-mass-ratio distribution $p(\ln q\given\theta)$ is
+    smaller (!). The log-mass-ratio $\ln q$ distribution
+    $p(\ln q\given\hyperpars)$ is
     \begin{eqnarray}
-      \ln p(\ln q\given\theta) &=& h_0 + (h_1 - h_0)\,\frac{\ln q - y_0}{y_1 - y_0}
+      \ln p(\ln q\given\hyperpars) &=& h_0 + (h_1 - h_0)\,\frac{\ln q - y_0}{y_1 - y_0}
       \quad \mbox{for $y_0<\ln q<y_1$,}
     \end{eqnarray}
-    where $h_0, h_1, y_0, y_1$ are hyper-parameters in $\theta$.
-  \item The eccentricity distribution $p(e\given P,\theta)$ is a beta,
+    where $h_0, h_1, y_0, y_1$ are hyper-parameters in $\hyperpars$.
+    Parameters $h_0, h_1$ are natural logs of densities in log-mass-ratio space.
+    Parameters $y_0, y_1$ are log-mass-ratio limits.
+  \item The eccentricity distribution $p(e\given P,\hyperpars)$ is a beta,
     and permitted to vary with period:
     \begin{eqnarray}
-      p(e\given P,\theta) &=& B(e\given \alpha(P),\beta(P))
+      p(e\given P,\hyperpars) &=& B(e\given \alpha(P),\beta(P))
       \\
@@
     control parameters $\alpha,\beta$, and $a_0, a_1, b_0, b_1$ are
-    hyper-parameters in $\theta$.
-  \item The inclination distribution $p(i)$ and the argument-of-perihelion
-    distribution $p(\omega)$ and the phase distribution $p(\phi)$ are all
-    fixed to their isotropic distributions. They bring no hyper-parameters
-    to $\theta$.
+    hyper-parameters in $\hyperpars$.
+    Parameters $a_0, a_1$ are values of the beta-distribution alpha-parameter
+    at the log-period limits, and parameters $b_0, b_1$ are values of the
+    beta-distribution beta-parameter at the limits.
+  \item The inclination $i$ distribution $p(i)$ and the
+    argument-of-perihelion $\omega$ distribution $p(\omega)$ and the
+    orbital-phase $\phi$ (at the fiducial time DEFINE THIS)
+    distribution $p(\phi)$ are all fixed to their isotropic
+    distributions. They bring no hyper-parameters to $\hyperpars$.
   \end{itemize}
@@
 Because all we have is a sampling of each system under the interim prior
-$p(\omega\given H_0)$,
+$p(\pars\given H_0)$,
 we have to re-weight or importance-sample the individual-star samplings
@@
 \begin{eqnarray}
-  \ln p(y\given\alpha,H_1) &=& \sum_{n=1}^N \ln p(y_n\given\alpha,H_1)
+  \ln p(y\given\hyperpars,H_1) &=& \sum_{n=1}^N \ln p(y_n\given\hyperpars,H_1)
   \\
-  \ln p(y_n\given\alpha,H_1) &\propto& \ln\left[\frac{1}{K_n}\,\sum_{k=1}^{K_n}\frac{p(\omega_{nk}\given\alpha,H_1)}{p(\omega_{nk}\given H_0)}\right]
-  \quad .
+  \ln p(y_n\given\hyperpars,H_1) &=& \ln\left[\frac{1}{K_n}\,\sum_{k=1}^{K_n}\frac{p(\pars_{nk}\given\hyperpars,H_1)}{p(\pars_{nk}\given H_0)}\right] + Q_n
+  \quad ,
 \end{eqnarray}
+where $y$ is the full data set, $y_n$ is the data set for star $n$,
+$\hyperpars$ is the vector of hyper-parameters,
+$H_1$ is the hypothesis (or model) that is our
+hyper-parameter-controlled hierarchical model,
+$H_0$ is the hypothesis corresponding to the interim prior of \sectionname~\ref{sec:fitting},
+the $\pars_{nk}$ are, for each star $n$, the $K_n$ samples from
+the posterior under $H_0$ (the interim prior),
+and $Q_n$ is an arbitrary constant that does not depend on any of the
+hyper-parameters.
 

### 2017-08-08 twoface/plot.py
add to __all

@@ -7,7 +7,7 @@ import numpy as np
 from thejoker.plot import plot_rv_curves
 from thejoker.sampler.utils import get_ivar
 
-__all__ = []
+__all__ = ['plot_data_orbits']
 
 def plot_data_orbits(data, samples_dict, n_orbits=128, jitter=None,
                      xlim_choice='wide', n_times=4096, title=None):

### 2017-08-08 scripts/initdb.py
load melissas rg masses into db as well

@@ -1,14 +1,17 @@
 # Project
 from twoface.log import log as logger
-from twoface.db.init import initialize_db, load_red_clump
+from twoface.db.init import initialize_db, load_red_clump, load_nessrg
 
-def main(allVisit_file, allStar_file, rc_file, overwrite=False, **_):
+def main(allVisit_file, allStar_file, rc_file, nessrg_file,
+         overwrite=False, **_):
     database_name = 'apogee.sqlite' # TODO: should this be enforced?
 
     initialize_db(allVisit_file, allStar_file, database_name,
                   overwrite=overwrite)
 
     load_red_clump(rc_file, database_name, overwrite=overwrite)
+
+    load_nessrg(nessrg_file, database_name, overwrite=overwrite)
 
 if __name__ == "__main__":
     from argparse import ArgumentParser
@@ -33,6 +36,8 @@ if __name__ == "__main__":
                         type=str, help="Path to APOGEE allVisit FITS file.")
     parser.add_argument("--redclump", dest="rc_file", required=True, type=str,
                         help="Path to APOGEE Red Clump catalog FITS file.")
+    parser.add_argument("--nessrg", dest="nessrg_file", required=True, type=str,
+                        help="Path to Ness Red Giant masses FITS file.")
 
     args = parser.parse_args()
 

### 2017-08-08 twoface/db/init.py

@@ -11,9 +11,9 @@ from ..util import Timer
 from ..log import log as logger
 from .connect import db_connect, Base
 from .model import (AllStar, AllVisit, AllVisitToAllStar, Status, RedClump,
-                    StarResult, CaoVelocity)
+                    StarResult, CaoVelocity, NessRG)
 
-__all__ = ['initialize_db', 'load_red_clump', 'load_cao']
+__all__ = ['initialize_db', 'load_red_clump', 'load_cao', 'load_nessrg']
 
 def tblrow_to_dbrow(tblrow, colnames, varchar_cols=[]):
     row_data = dict()
@@ -406,6 +406,112 @@ def load_cao(filename, database_file, overwrite=False, batch_size=1024):
 
     if len(cvs) > 0:
         session.add_all(cvs)
+        session.commit()
+
+    logger.debug("tables loaded in {:.2f} seconds".format(t.elapsed()))
+
+    session.close()
+
+# ----------------------------------------------------------------------------
+# Ness RG masses
+
+def ness_tblrow_to_dbrow(tblrow, colnames):
+    row_data = dict()
+    for _c in tblrow.colnames:
+        c = _c.replace('(','').replace(')','').replace('[','').replace(']','')
+        c = c.replace('/', '_')
+
+        if c in colnames:
+            row_data[c] = tblrow[_c]
+
+    return row_data
+
+def load_nessrg(filename, database_file, overwrite=False, batch_size=4096):
+    """Load the Ness red giant mass catalog.
+
+    Parameters
+    ----------
+    filename : str
+        Full path to Ness red giant info file.
+    database_file : str
+        Filename (not path) of database file in cache path.
+    overwrite : bool (optional)
+        Overwrite any data already loaded into the database.
+    batch_size : int (optional)
+        How many rows to create before committing.
+    """
+
+    database_path = join(TWOFACE_CACHE_PATH, database_file)
+
+    norm = lambda x: abspath(expanduser(x))
+    tbl = Table.read(norm(filename), format='fits', hdu=1)
+
+    Session, engine = db_connect(database_path)
+    logger.debug("Connected to database at '{}'".format(database_path))
+    session = Session()
+
+    # What columns do we load?
+    skip = ['ID', 'ALLSTAR_ID']
+    colnames = []
+    for x in NessRG.__table__.columns:
+        col = str(x).split('.')[1]
+        if col in skip:
+            continue
[…]

### 2017-08-08 twoface/db/model.py

@@ -16,7 +16,7 @@ from .quantity_type import QuantityTypeClassFactory
 from . import numpy_adapt # just need to execute code
 
 __all__ = ['AllStar', 'AllVisit', 'RedClump', 'JokerRun', 'StarResult',
-           'Status', 'AllVisitToAllStar', 'CaoVelocity']
+           'Status', 'AllVisitToAllStar', 'CaoVelocity', 'NessRG']
 
 class Status(Base):
     __tablename__ = 'status'
@@ -655,3 +655,33 @@ class JokerRun(Base):
     def __repr__(self):
         return ("<JokerRun(name='{}', date='{}')>"
                 .format(self.name, self.date.isoformat()))
+
+class NessRG(Base):
+    __tablename__ = 'nessrg'
+
+    id = Column(types.Integer, primary_key=True)
+
+    allstar_id = Column('allstar_id', types.Integer,
+                        ForeignKey('allstar.id', ondelete='CASCADE'),
+                        index=True)
+    star = relationship("AllStar")
+
+    # Melissa Ness' masses and ages from:
+    # http://iopscience.iop.org/article/10.3847/0004-637X/823/2/114/meta
+
+    # strip ()[], replace / with _
+    Type = Column('Type', types.Integer)
+    Teff = Column('Teff', types.REAL)
+    logg = Column('logg', types.REAL)
+    Fe_H = Column('Fe_H', types.REAL)
+    a_Fe = Column('a_Fe', types.REAL)
+    lnM = Column('lnM', types.REAL)
+    lnAge = Column('lnAge', types.REAL)
+    chi2 = Column('chi2', types.REAL)
+
+    e_Teff = Column('e_Teff', types.REAL)
+    e_logg = Column('e_logg', types.REAL)
+    e_Fe_H = Column('e_Fe_H', types.REAL)
+    e_a_Fe = Column('e_a_Fe', types.REAL)
+    e_logM = Column('e_logM', types.REAL)
+    e_logAge = Column('e_logAge', types.REAL)

### 2017-08-10 paper/1-catalog/twoface-catalog.tex
titles
author shit
new outline

--- 2017-08-05 paper/main.tex
@@
 % - list our assumptions for detection & characterization (TheJoker)
-% - list our assumptions for hierarchical inference
 
@@
 % -----------
-% - This file generates by Makefile; don't be typing ``pdflatex'' or some bullshit.
+% - This file generates by Makefile; don't be typing ``pdflatex'' or some
+%   bullshit.
 % - Line break between sentences to make the git diffs readable.
@@
 % - Reserve () for function arguments; use [] or {} for outer shit.
-% - Always prior pdf or posterior pdf, never prior or posterior (that's your arse).
-% - Use \sectionname not Section, \figname not Figure, \documentname not Article or Paper or paper.
-% - Make sure that two assumptions in the two assumption lists only have the same name if they really are the same assumption: These are proper names!
+% - Always prior pdf or posterior pdf, never prior or posterior (that's your
+%   arse).
+% - Use \sectionname not Section, \figname not Figure, \documentname not Article
+%   or Paper or paper.
+% - Make sure that two assumptions in the two assumption lists only have the
+%   same name if they really are the same assumption: These are proper names!
 % - Hyphenate binary-star when it is an adjective, not when it is a noun!
 % - Where there are defined math symbols (like \pars), use them!
+
+% other notes
+% - Binary engulfment reference: https://arxiv.org/pdf/1002.2216.pdf
+
+% Outline:
+% - Intro
+% - Data (APOGEE description)
+% - Methods
+%     - Re-measure velocities
+%     - The Joker / orbit sampling
+% - Results
+%     - Catalog of companions (including samplings)
+%     - Differences between RGB / RC
+%     - Other interesting shit
+% - Discussion
 
@@
 
+\newcommand{\nRC}{XX}
+\newcommand{\nRGB}{XX}
+\newcommand{\ntotal}{XX}
+\newcommand{\ncompanions}{XX}
+
 % define macros for math
@@
 
-\title{Binary companions of red clump stars}
+\title{Binary companions of red clump stars I: \\
+       catalog and surviving companions}
 
-\author{Adrian~M.~Price-Whelan}
+\author[0000-0003-0872-7098]{Adrian~M.~Price-Whelan}
 \affiliation{Department of Astrophysical Sciences,
@@
 % Context
-Radial-velocity measurements can be used to find stellar, sub-stellar,
-and planetary companions of stars.
-In the \apogee\ survey, we have multi-epoch radial-velocity measurements
-for many red-giant stars, including many red-clump (\RC) stars.
-Spectral analysis of \RC\ stars can be used to deliver masses (with
-XXX precision) and 20-ish chemical abundance measurements (with YYY
-precision).
+Repeat radial-velocity measurements of stars can be used to find stellar,
+sub-stellar, and planetary companions of stars.
+The \apogee\ survey has measured multi-epoch but sparse radial velocities for
+$> XX$ (mainly red giant) stars.
+Though even few observations are useful for detecting companions, such data are
+difficult to use for constraining the physical properties of companions because
+permitted orbital solutions are highly degenerate.
 % Aims
-Here we perform a search for secondary companions of [number] \RC\ stars.
-Because the primary-star masses are known, the mass-function degeneracy
-is broken and the secondary-companion $m\,\sin i$ values can be inferred.
-Also, if the primary and companion are assumed to be co-eval, the
-abundances of the \RC\ star can be assumed to represent accurately the
-abundances of the secondary.
+Here we perform a search for secondary companions of \nRC\ red clump (\RC) and
+\nRGB\ red giant branch (\RGB) stars that have similar radii to \RC\ stars
+observed as part of the \apogee\ survey.
+Because the primary-star masses are known, the mass-function degeneracy is
+broken and the secondary-companion masses ($m\,\sin i$) can be inferred.
 % Methods
-We use a data-driven model (\thecannon) of stellar spectra to deliver
-each \RC\ mass and chemical abundances, and a custom-built Monte Carlo
-sampler (\thejoker) to deliver (often highly multi-modal) posterior
-beliefs about companion $m\,\sin i$, period, and other orbital parameters.
-Our sampler permits us to accurately sample even the badly behaved
-posterior pdfs obtained when there are as few an NNN observation
-epochs per star.
-We put these posterior samples---made under an interim prior---into a
-hierarchical inference to obtain individual-companion-marginalized
-posterior beliefs about the properties of the full secondary-companion
-distribution in mass and period.
-In this analysis, even stars with only a few observations (and therefore
-badly multimodal posterior pdfs) contribute valuable information to the
-population inference.
+We use a custom-built Monte Carlo sampler (\thejoker) combined with prior
+measurements of primary masses and stellar parameters to deliver (often highly
+multi-modal) posterior beliefs about companion mass, pericenter distance, and
+other orbital parameters for the \RC\ and \RGB\ stars.
 % Results
-We deliver a catalog of QQQ companions [that meet some criteria], and
-posterior samplings for all [number] stars in the parent sample.
-We find that the period and mass distributions [have some properties].
-We look for trends with chemical abundances, and we find tentative evidence
-for ZZZ and WWW.
+We deliver a catalog of \ncompanions\ companions \todo{that meet some criteria},
+and posterior samplings for all \ntotal\ stars in the parent sample.
+We find \todo{describe} differences between the companions around \RC\ and \RGB\
+stars that are \todo{in/consistent} with the theoretical prediction that
+stellar companions should not survive the common-envelope \RGB\ phase.
+We additionally find trends with chemical abundances, and we find tentative
+evidence for ZZZ and WWW.
 \end{abstract}

### 2017-08-11 twoface/db/model.py
fix relationship

@@ -664,7 +664,7 @@ class NessRG(Base):
     allstar_id = Column('allstar_id', types.Integer,
                         ForeignKey('allstar.id', ondelete='CASCADE'),
                         index=True)
-    star = relationship("AllStar")
+    star = relationship("AllStar", backref=backref("ness_rg", uselist=False))
 
     # Melissa Ness' masses and ages from:
     # http://iopscience.iop.org/article/10.3847/0004-637X/823/2/114/meta

### 2017-08-11 twoface/plot.py
remove import and fix slicing

@@ -9,7 +9,7 @@ from thejoker.sampler.utils import get_ivar
 
 __all__ = ['plot_data_orbits']
 
-def plot_data_orbits(data, samples_dict, n_orbits=128, jitter=None,
+def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
                      xlim_choice='wide', n_times=4096, title=None):
     """
     Plot the APOGEE RV data vs. time and orbits computed from The Joker samples.
@@ -18,8 +18,8 @@ def plot_data_orbits(data, samples_dict, n_orbits=128, jitter=None,
     ----------
     data : :class:`~twoface.data.APOGEERVData`
         The radial velocity data.
-    samples_dict : dict
-        Dictionary of posterior samples from The Joker.
+    samples : :class:`~thejoker.samples.JokerSamples`
+        Posterior samples from The Joker.
     n_orbits : int, optional
         Number of orbits to plot over the data.
     jitter : :class:`~astropy.units.Quantity`, optional
@@ -52,7 +52,7 @@ def plot_data_orbits(data, samples_dict, n_orbits=128, jitter=None,
 
     elif xlim_choice == 'wide': # twice the longest period sample
         t_min = data.t.mjd.min()
-        t_max = max(data.t.mjd.min() + 2*samples_dict['P'].max().value,
+        t_max = max(data.t.mjd.min() + 2*samples['P'].max().value,
                     data.t.mjd.max())
         span = t_max - t_min
         t_grid = np.linspace(t_min - 0.05*span, t_max + 0.05*span, 2048)
@@ -61,26 +61,20 @@ def plot_data_orbits(data, samples_dict, n_orbits=128, jitter=None,
         raise ValueError('Invalid xlim_choice {0}. Can be "wide" or "tight".'
                          .format(xlim_choice))
 
-    plot_rv_curves(samples_dict, t_grid, rv_unit=u.km/u.s, data=data, ax=ax,
-                   n_plot=min(len(samples_dict['P']), n_orbits),
+    plot_rv_curves(samples, t_grid, rv_unit=u.km/u.s, data=data, ax=ax,
+                   n_plot=min(len(samples['P']), n_orbits),
                    plot_kwargs=dict(color='#aaaaaa', alpha=0.25, linewidth=0.5),
[…]

### 2017-08-11 paper/1-catalog/twoface-catalog.tex
intro shite

@@
 % -----------
-% - This file generates by Makefile; don't be typing ``pdflatex'' or some
-%   bullshit.
+% - This file generates by Makefile; don't be typing ``pdflatex'' or some BS.
 % - Line break between sentences to make the git diffs readable.
-% - Simple Monte Carlo gets a capital S to indicate that it is a defined thing.
 % - Use \, as a multiply operator.
@@
 % - Use \sectionname not Section, \figname not Figure, \documentname not Article
-%   or Paper or paper.
+%   \tablename not Table, \eqname not Equation.
 % - Make sure that two assumptions in the two assumption lists only have the
@@
 % other notes
+% -----------
 % - Binary engulfment reference: https://arxiv.org/pdf/1002.2216.pdf
@@
 
-\title{Binary companions of red clump stars I: \\
-       catalog and surviving companions}
+\title{Binary companions of red giant stars I: \\
+       catalog and companions that survive the common envelope phase}
 
@@
 Repeat radial-velocity measurements of stars can be used to find stellar,
-sub-stellar, and planetary companions of stars.
-The \apogee\ survey has measured multi-epoch but sparse radial velocities for
-$> XX$ (mainly red giant) stars.
+sub-stellar, and planetary companions.
+The \apogee\ survey (\DR) has measured multi-epoch but sparse radial velocities
+for $> XX$ (mainly red giant) stars.
 Though even few observations are useful for detecting companions, such data are
@@
 \nRGB\ red giant branch (\RGB) stars that have similar radii to \RC\ stars
-observed as part of the \apogee\ survey.
+using data from the \apogee\ survey.
 Because the primary-star masses are known, the mass-function degeneracy is
@@
 
-APW: Opening sentence: Even with one single star and two epochs, if we
-see an RV change, there must be binaries! What more can we say? Can we
-infer binary populations from bad data?
+Time-domain radial velocity measurements of stars fundamentally contain
+information about massive companions: even with two observations of a single
+star, a difference in the measured radial velocities implies the existence of at
+least one companion.
+Of course with just two observations, the orbital properties of the companion(s)
+will be very poorly constrained.
+Most prior searches for companions using radial velocity data have therefore
+restricted their searches to only stars with ``sufficient'' data where the
+orbital solution is unamibiguous under the ... noise model.
+
+Many of the largest (in number of objects) stellar spectroscopic surveys observe
+large samples of stars sparsely, non-uniformly, and imprecisely.
+
+Even if period, generic orbital parameters are poorly known, limits on companion
+mass or pericenter
 

### 2017-08-12 paper/1-catalog/twoface-catalog.tex
change structure, write intro, shit

@@
 % - Binary engulfment reference: https://arxiv.org/pdf/1002.2216.pdf
-
-% Outline:
-% - Intro
-% - Data (APOGEE description)
-% - Methods
-%     - Re-measure velocities
-%     - The Joker / orbit sampling
-% - Results
-%     - Catalog of companions (including samplings)
-%     - Differences between RGB / RC
-%     - Other interesting shit
-% - Discussion
 
@@
 \title{Binary companions of red giant stars I: \\
-       catalog and companions that survive the common envelope phase}
+       catalog and companions that survive the common envelope}
 
@@
 
-Time-domain radial velocity measurements of stars fundamentally contain
-information about massive companions: even with two observations of a single
-star, a difference in the measured radial velocities implies the existence of at
-least one companion.
-Of course with just two observations, the orbital properties of the companion(s)
-will be very poorly constrained.
-Most prior searches for companions using radial velocity data have therefore
-restricted their searches to only stars with ``sufficient'' data where the
-orbital solution is unamibiguous under the ... noise model.
+Time-domain radial-velocity measurements of stars contain information about
+massive companions: even with two successive observations of a single star, a
+difference in the measured radial velocities implies the existence of at least
+one companion.
+However, with few or imprecise radial-velocity measurements, the orbital
+properties of the companion(s) are very poorly constrained.
+Most prior searches for companions using survey RV data have therefore
+restricted their searches to only sources with many epochs, so that the orbital
+solution can be unambiguously determined.
+The vast majority of spectroscopic targets with repeat observations in the
+largest (by number of objects) stellar spectroscopic surveys are in the opposite
+regime: targets are often observed just a few times with sparse, non-uniform
+phase coverage.
 
-Many of the largest (in number of objects) stellar spectroscopic surveys observe
-large samples of stars sparsely, non-uniformly, and imprecisely.
+% With such data, even if the period or other generic orbital parameters
+% are very poorly constrained, limits on companion mass or pericentric distance
+% of the companion may still XX.
 
-Even if period, generic orbital parameters are poorly known, limits on companion
-mass or pericenter
+If there are only a few radial-velocity measurements made per star, and the
+companion spectrum is not observed, any measured radial-velocity values will be
+consistent with many different combinations of (primary) orbital parameters
+(period, amplitude, eccentricity, etc.).
+To identify companions to the typical star observed in a spectroscopic survey,
+we therefore face at least one major challenge: how, given a small number of
+observations of a primary star, do we reliably obtain posterior information
+about the binary-system properties?
+In general, the likelihood function---and the posterior probability distribution
+function (pdf) under any reasonable prior pdf---will be highly multimodal, and
+many of the modes will have comparable integrated probability density.
+For example, with just two radial-velocity measurements, a harmonic series of
+period modes will exist in the likelihood function.
 
-If we want to use a spectroscopic survey to responsibly measure the
-binary-star population---that is, if we want to get (say) the mass,
-mass ratio, and period distribution of binary-star systems---we face
-(at least) two challenges.
-The first is how, given a small number of observations of a primary
-star (a single-lined spectroscopic binary star), do we reliably obtain posterior
-information about the binary-system properties?
-After all, if there are only a few radial-velocity measurements made
-per star, any observed radial-velocity values will be consistent with
-many different combinations of period and amplitude (not to mention
-eccentricity and argument of perihelion) for the primary orbit.
-In general, the likelihood function, and the posterior pdf under any
-reasonable prior pdf, will be highly multimodal in these kinds of
-problems, and many of the modes will have comparable integrated
-probability density.
-We have solved this problem previously, though with limitations (to be
-discussed more below), with \thejoker\ (CITE).
-\thejoker\ is a Simple Monte Carlo rejection sampler that is
-computationally expensive but probabilistically righteous: It delivers
-independent (zero-auto-correlations) posterior pdf samples for
-single-companion binary model parameters, given any number of
-radial-velocity measurements.
+We have solved this problem previously, though with limitations (to be discussed
+more below), with \thejoker\ (\citealt{Price-Whelan:2017}).
+\thejoker\ is a Monte Carlo rejection sampler that is computationally expensive
+but probabilistically righteous:
+It delivers independent posterior pdf samples for single-companion binary
+orbital parameters, given any number of radial-velocity measurements.
+Here we use \thejoker\ to generate posterior pdf samples for \todo{which?} stars
+observed by the \apogee\ surveys (see \sectionname~\ref{sec:data};
+\citealt{Majewski:2015}).
 
-The second challenge is how, given all these noisy, multi-modal
-posterior pdfs over binary-system properties, many of which will
-permit a wide range of qualitatively different system properties, and
-many of which will be consistent with no companion at all, can we
-constrain parameters of a model of the full binary-system population?
-No histogram of best-fit values, nor stack of multi-modal posteriors,
-would be anything like a good estimate of the true population
-distribution.
-We have also solved this problem previously, though in the exoplanet
-domain, with hierarchical probabilistic models based on importance
-sampling (CITE DFM, and also HOGG).
-These hierarchical models derive posterior estimates of the population
-parameters that accurately describe the population, given the noisy
-information about the individual systems, provided that the noise
-model is accurate---that is, provided that the likelihood functions
-used in the individual-system inferences are themselves accurate.
+The \apogee\ surveys primarily target red-giant-branch (\RGB) stars, which are
+ideal for the study of single-line binary systems.
+For one, because they are so luminous, they are unlikely (in general) to have
+equally-bright companions, and their spectra are therefore well-approximated or
+fit as single-line objects.
+The subset of \RGB\ stars in the ``red clump'' (\RC) are even more powerful as
+they are standard candles and have masses that can be estimated using
+spectroscopy (using dredged-up elements; \citealt{Martig:2016,Ness:2016}).
+With primary-star mass estimates, the binary-orbit fitting will return
+$m_2\,\sin i$ estimates for the secondary, and not just estimates of the
+mass-function.
+Additionally, the \apogee\ pipelines (\citealt{Garcia-Perez:2016}) and also
+\thecannon\ (\citealt{Ness:2015}) produce detailed abundance estimates for \RGB\
+and \RC\ stars.
+If there are causal relationships between chemical abundances and binary
+companions---as are expected---these should be measurable here.
 
-Here we mash up these two solutions to these two challenges.
-We deliver a catalog of binary-star systems with \RC\ primaries, and also
-posterior information about the full population statistics, for at least
-a toy model of that full population.
-We use all the data---even the non-detections, marginal detections, and
-badly multi-modal posterior pdfs---when we infer the population model;
-that is, we don't arbitrarily cut to a catalog of ``clean'' binaries.
+By making cuts on this library of posterior pdf samples (described in detail in
+\sectionname~\ref{sec:whatever}), we deliver a catalog of binary-star systems
+from the \apogee\ survey with no cuts or constraints on data quality or volume.
+We additionally highlight the subset of this catalog with \RC\ star primaries,
+for which we can compute companion mass and pericentric distance.
 
-For the study of single-line binaries, red-giant stars---and even
-better red-clump stars---are ideal.
-For one, because they are so luminous, they are unlikely (in general)
-to have equally-bright companions, and therefore are well approximated
-or fit as single-line objects.
-For two, they are standard candles, or can be distance-calibrated.
-For three, they have masses that can be estimated spectrally (making
-use of dredged-up elements; CITE MARTIG and CITE NESS).
-With primary-star mass estimates, the binary-orbit fitting will return
-$m\,\sin i$ estimates for the secondary, and not just mass-function
-estimates.
-For four, the \apogee\ pipelines, and also \thecannon\, produce
-detailed abundance estimates for red giants and \RC\ stars.
-If there are causal relationships between chemical abundances and
-binary companions---and we expect there will be---these should become
-visible here.
+\section{Data} \label{sec:data}
+
+\todo{HOGG: fill in this ish}
 
@@
 
-HOGG: What is up with delivering a catalog (it requires decision-making)?
-
-\section{Data}
+ASPCAP: \citealt{Garcia-Perez:2016}
+SDSSIV: \citealt{Blanton:2017}
 
@@
 
-\section{Method: Detection and orbit fitting}\label{sec:fitting}
+\section{Methods}
+
+\subsection{Re-measured radial velocities}\label{sec:rv-remeasure}
+
+\todo{HOGG: fill in this ish}
+
+\subsection{Detection and orbit fitting}\label{sec:fitting}
 
@@
 
-\section{A catalog of red-clump binary systems}
+\section{Results}
+
+\subsection{Catalog / posterior samples whatever} \label{sec:full-catalog}
+
+\todo{APW}
+
+\subsection{A catalog of red-clump binary systems} \label{sec:rc-catalog}
+
+\todo{APW}
 
@@
 
-\section{Method: Population inferences}\label{sec:popinference}
+\subsection{Differences in companions of RC and RGB stars} \label{sec:rc-rgb}
 
-Now we want to take the individual-star fits---or samplings---that
-we made in \sectionname~\ref{sec:fitting} and, from these, produce
-an inference of the properties of the whole \RC\ binary-star population.
-Important necessary properties of this inference are the following:
-\begin{description}
-\item[utilize non-detections] Even stars which are consistent with
-  zero radial-velocity variation (no binary companion) are relevant to
-  a populations inference. This is for two reasons: The first is that
-  the small-mass and long-period ends of the binary population will
-  imprint only low-amplitude signals in the data. These signals are
-  present at low signal-to-noise, but not if we remove them because
-  they don't make some catalog or threshold cut. The second reason
-  these stars are important is that they constrain the fraction of
-  stars with no companions. Any populations-inference method we employ
-  must responsibly harvest the information in these non-detection
-  stars.
-\item[marginalize out individual-star parameters] Every star---even
-  any star with a high-confidence companion detection---will have a
-  highly multi-modal likelihood function, with many qualitatively
-  different companion models that are nonetheless locally
-  optimal. This means that most posterior samplings (created in
-  \sectionname~\ref{sec:fitting}) under our interim prior will show
-  multiple, qualitatively distinct modes in binary-system parameter
-  space.  Any correct population inference will marginalize out these
-  extremely non-trivial distributions over individual-star
-  binary-system parameters.
-\end{description}
-We can meet these requirements---at least in principle---with a
-hierarchical Bayesian inference (hierarchical probablistic model) of
-the binary-star population.
-
-In addition to the above requirements, we make the following assumptions
-that restrict our attention to a well-defined model space:
-\begin{description}
-\item[correct samplings] accurate representations of posterior pdf
-  given prior and likelihood; correct likelihood function. We hereby
-  take on all the assumptions given in \sectionname~\ref{sec:fitting},
-  about noise and kinematics, and so on.
-\item[no trinaries] related to the above: no trinaries, quads, or anything
-  crazier.
-\item[sufficent samplings] density and support of samplings
-\item[independent stars] no relation between one star and another; each
-  drawn from the same population; no multiple populations.
-\item[masses known] The mass of each primary \RC\ star is known, with
-  a correctly known posterior pdf, from prior work (CITE NESS).
-\item[population parameterization] We adopt the following
-  parameterization of the population. In what follows, $\hyperpars$ will
-  stand in for the full set of hyper-parameters (population
-  parameters).
-  \begin{itemize}
-  \item Each primary \RC\ star has a probability $F$ (pure probability
-    $0<F<1$) of having a secondary companion. $F$ is a hyper-parameter
-    in $\hyperpars$.
-  \item For those with secondary companions, the log-period $\ln P$
-    distribution $p(\ln P\given\hyperpars)$ is a power law between
-    end-points, or
-    \begin{eqnarray}
-      \ln p(\ln P\given\hyperpars) &=& g_0 + (g_1 - g_0)\,\frac{\ln P - x_0}{x_1 - x_0}
-      \quad \mbox{for $x_0<\ln P<x_1$,}
-    \end{eqnarray}
-    where $g_0, g_1, x_0, x_1$ are hyper-parameters in $\hyperpars$.
-    Parameters $g_0, g_1$ are natural logs of densities in log-period space.
-    Parameters $x_0, x_1$ are log-period limits.
-  \item Each primary star has a true mass $M$ (which we only know
-    noisily) and the companion mass is potentially larger or
-    smaller (!). The log-mass-ratio $\ln q$ distribution
-    $p(\ln q\given\hyperpars)$ is
-    \begin{eqnarray}
-      \ln p(\ln q\given\hyperpars) &=& h_0 + (h_1 - h_0)\,\frac{\ln q - y_0}{y_1 - y_0}
-      \quad \mbox{for $y_0<\ln q<y_1$,}
-    \end{eqnarray}
-    where $h_0, h_1, y_0, y_1$ are hyper-parameters in $\hyperpars$.
-    Parameters $h_0, h_1$ are natural logs of densities in log-mass-ratio space.
-    Parameters $y_0, y_1$ are log-mass-ratio limits.
-  \item The eccentricity distribution $p(e\given P,\hyperpars)$ is a beta,
-    and permitted to vary with period:
-    \begin{eqnarray}
-      p(e\given P,\hyperpars) &=& B(e\given \alpha(P),\beta(P))
-      \\
-      \alpha(P) &=& a_0 + (a_1 - a_0)\,\frac{P - x_0}{x_1 - x_0}
-      \\
-      \beta(P) &=& b_0 + (b_1 - b_0)\,\frac{P - x_0}{x_1 - x_0}
-      \quad ,
-    \end{eqnarray}
-    where $B(x\given\alpha,\beta)$ is the beta distribution with
-    control parameters $\alpha,\beta$, and $a_0, a_1, b_0, b_1$ are
-    hyper-parameters in $\hyperpars$.
-    Parameters $a_0, a_1$ are values of the beta-distribution alpha-parameter
-    at the log-period limits, and parameters $b_0, b_1$ are values of the
-    beta-distribution beta-parameter at the limits.
-  \item The inclination $i$ distribution $p(i)$ and the
-    argument-of-perihelion $\omega$ distribution $p(\omega)$ and the
-    orbital-phase $\phi$ (at the fiducial time DEFINE THIS)
-    distribution $p(\phi)$ are all fixed to their isotropic
-    distributions. They bring no hyper-parameters to $\hyperpars$.
-  \end{itemize}
-\item[hyper-priors] whatevs
-\end{description}
-
-The binary-system parameters for each individual \RC\ star were
-inferred (in \sectionname~\ref{sec:fitting}) with an interim prior that
-is very simple and doesn't represent our true beliefs about the
-binary-star population...
-
-In a hierarchical Bayesian inference, we replace the binary-star
-system priors used in \sectionname~\ref{sec:fitting} with a parameterized
-model for the binary-star population...
-
-Because all we have is a sampling of each system under the interim prior
-$p(\pars\given H_0)$,
-we have to re-weight or importance-sample the individual-star samplings
-in this hierarchical inference... Cite Hogg et al; Cite DFM et al.
-% The following needs checking for symbol conventions
-\begin{eqnarray}
-  \ln p(y\given\hyperpars,H_1) &=& \sum_{n=1}^N \ln p(y_n\given\hyperpars,H_1)
-  \\
-  \ln p(y_n\given\hyperpars,H_1) &=& \ln\left[\frac{1}{K_n}\,\sum_{k=1}^{K_n}\frac{p(\pars_{nk}\given\hyperpars,H_1)}{p(\pars_{nk}\given H_0)}\right] + Q_n
-  \quad ,
-\end{eqnarray}
-where $y$ is the full data set, $y_n$ is the data set for star $n$,
-$\hyperpars$ is the vector of hyper-parameters,
-$H_1$ is the hypothesis (or model) that is our
-hyper-parameter-controlled hierarchical model,
-$H_0$ is the hypothesis corresponding to the interim prior of \sectionname~\ref{sec:fitting},
-the $\pars_{nk}$ are, for each star $n$, the $K_n$ samples from
-the posterior under $H_0$ (the interim prior),
-and $Q_n$ is an arbitrary constant that does not depend on any of the
-hyper-parameters.
-
-\section{The population of red-clump binary companions}
+\todo{APW}
 

### 2017-08-14 paper/1-catalog/twoface-catalog.tex
refs and text

@@
 
+\todo{An alternate approach to this intro is to start with an overview of
+      binary stars, radial velocity measurements of binarity in general...}
+
+\todo{As is, the intro is missing many citations}
+
 Time-domain radial-velocity measurements of stars contain information about
@@
 
-\subsection{Detection and orbit fitting}\label{sec:fitting}
+\subsection{Orbit fitting and velocity modeling}\label{sec:fitting}
 
-Our approach here is to proceed in two phases.
-In the first phase, we obtain a posterior sampling in binary-system
-parameter space for every individual star, treating it as a
-single-lined (SB1) spectroscopic binary system with a single
-companion.
-This sampling is performed under a relatively uninformative prior pdf,
-which will be called the ``\emph{interim prior}''.
-These samplings are used to discover and characterize individual binary-star
-systems, and generate (among other things) a catalog of binaries.
-In the second phase, we perform a hierarchical inference of the binary
-population, using (responsibly) these samplings under the interim
-prior as inputs.
-In the second phase we use all samplings of all stars, not just those that
-pass our thresholds for discovery as binary systems.
-In this \sectionname, we describe the first phase, that is, the
-individual-system sampling phase for discovery and characterization.
+For every source in the sample of \apogee\ stars defined in
+\sectionname~\ref{sec:data}, we obtain a posterior sampling in binary-system
+parameter space, treating it as a single-lined (SB1) spectroscopic binary system
+with a single companion.
+This sampling is performed under a relatively uninformative prior pdf, and the
+resulting posterior samplings are used to discover and characterize individual
+binary-star systems and generate a catalog of companions.
+We use \thejoker\ to perform the posterior samplings; we briefly describe the
+algorithm below, but a full description is given in previous work
+(\citealt{Price-Whelan:2017}).
+In this \sectionname, we describe the assumptions and method used to generate
+individual-system samplings.
 
@@
 \begin{description}
-\item[no trinaries] whatevs: No trinaries or beyond. Obviously wrong!
-\item[SB1] whatevs: There is no significant light contributed by the secondary.
-\item[Kepler] whatevs: No non-gravitational contributions to RV history.
-\item[noise model] whatevs: Measurements are unbiased and noise estimates are correct up to jitter; everything Gaussian.
-\item[interim prior] We adopt a prior pdf on binary-system parameters
-  with the following properties...
+\item[no multiplets] All radial-velocity variations of the primary star are
+  induced by a single companion.
+  This is motivated by the idea that triple star systems are usually
+  hierarchical so that the period of the inner binary is typically much shorter
+  than the orbital period of the outer body (\todo{CITATION NEEDED}).
+  At present, we ignore the possibility of higher-order multiple systems.
+  \todo{revisit this - do we want to also sample over a long-term trend?}
+\item[Kepler] Related to the first point, all velocity variations of the primary
+  are gravitational, and we therefore ignore the possibility of intrinsic
+  variation from, e.g., stellar oscillations.
+\item[SB1] All spectra are single-lined; that is, we assume that the secondary
+  is significantly fainter and is thus undetected in the spectra.
+  This assumption is motivated by the fact that we expect \RGB\ stars to be
+  substantially more luminous than their typical companion.
+  However, there are known double-lined binary stars in the \apogee\ catalog
+  (El-Badry et al., in prep.), and an unknown fraction of \RGB--\RGB\ binaries.
+\item[simple noise model] Measurements are unbiased and noise estimates are
+  correct up to an \todo{unknown} extra variance (the ``jitter'').
+  All noise contributions result in Gaussian uncertainties on the individual
+  radial-velocity measurements.
 \end{description}
 
-Summary of \thejoker\ and what it does; that is, that it obeys the
-relevant parts of our assumptions.
+\thejoker\ is a custom-built Monte Carlo sampler designed to produce independent
+posterior samples in orbital parameters, given radial-velocity measurements
+under the assumptions above.
+Our parametrization is similar to the notation in \citet{Murray:2010}:
+The radial velocity $v$ at time $t$ is given by
+\begin{equation}
+  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, \phi_0)\right) +
+    e\,\cos\omega]
+\end{equation}
+where $\bs{\theta} = (P,e,\phi_0,\omega,K,v_0)$---period, eccentricity, initial
+phase, argument of pericenter, velocity semi-amplitude, Barycentric velocity---
+and the true anomaly, $f$, is a function of the specified parameters (see
+\sectionname~2 of \citealt{Price-Whelan:2017} or \eqname~63 in
+\citealt{Murray:2010}).
 
-Any modifications or knob-settings for \thejoker?
+\thejoker\ was designed for the extremely multi-modal pdfs expected when the
+number of radial-velocity measurements of a source is small, or the data have
+low signal-to-noise.
+While other Markov Chain Monte Carlo (MCMC) methods have difficulty producing
+independent samples with such data, \thejoker\ succeeds by brute force.
+After generating an initial (very large) library of prior samples from an
+assumed prior pdf (see below), the (typically multi-modal) likelihood is
+evaluated at each sample and used to rejection sample.
+In practice, given a number of requested samples for each star, the sampling
+proceeds iteratively: since it is easier to accept samples when the data is
+sparse or noisy, far more prior sample draws (and thus likelihood
+evaluations) must occur under very constraining data.
+
+In addition to generating posterior samples for each source, we also compute the
+fully marginalized likelihood (FML) under two other models: (1) a model in which
+the radial velocity of the source is constant, and (2) a model in which the
+radial-velocity variations are purely linear.
+We later use these FML values to define cuts on our full catalog of posterior
+samples to produce a catalog of companions.
+
+\todo{equations?}
 

### 2017-12-01 setup.py
ignore scripts

@@ -65,9 +65,11 @@ generate_version_py(PACKAGENAME, VERSION, RELEASE,
                     get_debug_option(PACKAGENAME))
 
 # Treat everything in scripts except README.rst as a script to be installed
-scripts = [fname for fname in glob.glob(os.path.join('scripts', '*'))
-           if os.path.basename(fname) != 'README.rst']
+# scripts = [fname for fname in glob.glob(os.path.join('scripts', '*'))
+#            if os.path.basename(fname) != 'README.rst']
 
+# APW: don't install any scripts
+scripts = []
 
 # Get configuration information from all of the various subpackages.
 # See the docstring for setup_helpers.update_package_files for more

### 2017-12-01 paper/1-catalog/twoface-catalog.tex

@@
 % - list our assumptions for detection & characterization (TheJoker)
+% - Check abstract numbers (~80,000 RG stars with 3 or more good visits?)
 
@@
 % adjust AAS-TEX shit
-\setlength{\parindent}{1.1\baselineskip}
+% \setlength{\parindent}{1.1\baselineskip}
 
@@
 \newcommand{\thecannon}{\project{The~Cannon}}
-\newcommand{\DR}{\acronym{DR13}}
+\newcommand{\DR}{\acronym{DR14}}
 \newcommand{\RC}{\acronym{RC}}
@@
 
-\title{Binary companions of red giant stars I: \\
+\title{Binary companions of red giant stars with APOGEE I: \\
        catalog and companions that survive the common envelope}
@@
 
-\author{Jason~Cao}
-\affiliation{Center for Cosmology and Particle Physics,
-             Department of Physics,
-             New York University, 726 Broadway,
-             New York, NY 10003, USA}
+% \author{Jason~Cao}
+% \affiliation{Center for Cosmology and Particle Physics,
+%              Department of Physics,
+%              New York University, 726 Broadway,
+%              New York, NY 10003, USA}
 
@@
 % Context
-Repeat radial-velocity measurements of stars can be used to find stellar,
-sub-stellar, and planetary companions.
-The \apogee\ survey (\DR) has measured multi-epoch but sparse radial velocities
-for $> XX$ (mainly red giant) stars.
-Though even few observations are useful for detecting companions, such data are
+Repeat radial-velocity measurements of stars can be used to identify stellar-,
+sub-stellar, and planetary-mass companions.
+Even few observations are useful for detecting companions, though such data are
 difficult to use for constraining the physical properties of companions because
@@
 % Aims
-Here we perform a search for secondary companions of \nRC\ red clump (\RC) and
-\nRGB\ red giant branch (\RGB) stars that have similar radii to \RC\ stars
-using data from the \apogee\ survey.
-Because the primary-star masses are known, the mass-function degeneracy is
-broken and the secondary-companion masses ($m\,\sin i$) can be inferred.
+Here we perform a search for secondary companions of red giant branch (\RGB)
+stars using data from the \apogee\ survey (\DR), which has measured multi-epoch
+($\geq 3$) but sparse radial velocities for $> 80,000$ \RGB\ stars.
+In some cases, where the primary-star masses are known, the mass-function
+degeneracy is broken and the secondary-companion masses ($m_2\,\sin i$) can be
+inferred.
 % Methods
-We use a custom-built Monte Carlo sampler (\thejoker) combined with prior
-measurements of primary masses and stellar parameters to deliver (often highly
-multi-modal) posterior beliefs about companion mass, pericenter distance, and
-other orbital parameters for the \RC\ and \RGB\ stars.
+We use a custom-built Monte Carlo sampler (\thejoker) combined with stellar
+parameters and primary mass estimates to deliver (often highly multi-modal)
+posterior beliefs about companion mass, pericenter distance, and other orbital
+parameters for the stars.
+We compare the companion populations of red clump (\RC) versus \RGB\ stars at
+similar surface gravity and temperature, and study the metallicity dependence
+of companion properties.
 % Results
-We deliver a catalog of \ncompanions\ companions \todo{that meet some criteria},
-and posterior samplings for all \ntotal\ stars in the parent sample.
-We find \todo{describe} differences between the companions around \RC\ and \RGB\
-stars that are \todo{in/consistent} with the theoretical prediction that
-stellar companions should not survive the common-envelope \RGB\ phase.
-We additionally find trends with chemical abundances, and we find tentative
-evidence for ZZZ and WWW.
+We provide a catalog of \ncompanions\ companions for which the companion
+properties can be confidently determined, along with posterior samplings for all
+\ntotal\ stars in the parent sample.
+\todo{We find differences between the companions around \RC\ and \RGB\
+stars that are in/consistent with the theoretical prediction that
+stellar companions should not survive the common-envelope \RGB\ phase.}
+\todo{We additionally find trends in XX with chemical abundances.}
+
 \end{abstract}
@@
 
-\subsection{Re-measured radial velocities}\label{sec:rv-remeasure}
-
-\todo{HOGG: fill in this ish}
-
 \subsection{Orbit fitting and velocity modeling}\label{sec:fitting}
@@
 
-\subsection{A catalog of red-clump binary systems} \label{sec:rc-catalog}
+\subsection{A catalog of confident companions} \label{sec:conf-companions}
 
@@
 
-\subsection{Differences in companions of RC and RGB stars} \label{sec:rc-rgb}
+\subsection{Differences in companions of RC and RGB stars}
+\label{sec:rc-rgb}
+
+\todo{APW}
+
+\subsection{Trends in RGB companions with chemical abundances}
+\label{sec:rgb-chemistry}
 

### 2017-12-01 paper/notes/1-dr13_dr14_diff/dr13_dr14_diffs.tex

\documentclass{article}

\usepackage{graphicx}
\usepackage{subfig}

\begin{document}

\section{}

\begin{figure*}[hp]
\begin{center}
\includegraphics[width=\textwidth]{figures/snr.pdf}
\end{center}
\caption{%
Signal-to-noise ratio (SNR) distribution for DR13 visits that also appear in
DR14.
}
\end{figure*}

\begin{figure*}[hp]
\begin{center}
\includegraphics[width=\textwidth]{figures/visit_vhelio_compare.pdf}
\end{center}
\caption{%
One point per \emph{visit}, comparing the derived barycentric radial velocity.
Only visits that appear in DR13 and DR14 are plotted here, and I've downsampled
so that an equal number of points are plotted for each SNR bin (there are
$\approx10^5$ in each SNR bin).
In case you can't see the legend, blue points are the higher SNR bin, orange are
lower.
I'm somewhat troubled by the cloud of blue points centered on 0; it seems like
many moderate SNR visits are getting large ($\approx$10's of km/s) velocity
shifts between DR13 and DR14.
}
\end{figure*}

\begin{figure*}[hp]
\begin{center}
\includegraphics[width=\textwidth]{figures/visit_vhelio_vs_snr.pdf}
\end{center}
\caption{%
Another way to look at the last point.
Again, one point per visit, plotting the difference in derived RV normalized by
the reported DR14 uncertainty of the measured RV.
}
\end{figure*}

\begin{figure*}[hp]
\begin{center}
\includegraphics[width=\textwidth]{figures/std_vhelio_compare.pdf}
\end{center}
\caption{%
Now I compute aggregate statistics on all visits for unique
\texttt{APOGEE\_ID}'s as a way to see how the RMS or velocity scatter changes
between DR13 and DR14. Each point here is now a unique \emph{source}, and in
this case represents the standard deviation of all visit velocities for a given
source in DR13 vs. DR14 in linear (left) and log (right).
}
\end{figure*}

\begin{figure*}[hp]
\begin{center}
\includegraphics[width=\textwidth]{figures/MAD_vhelio_compare.pdf}
\end{center}
\caption{%
Same as last figure, but using $\sigma = 1.5 \times {\rm MAD}$, using the median
absolute deviation (MAD) as a more robust estimator of the scatter.
}
\end{figure*}

\begin{figure*}[hp]
\begin{center}
\includegraphics[width=\textwidth]{figures/MAD_vhelio_diff.pdf}
\end{center}
\caption{%
Now I'm plotting the difference in the MAD-estimated scatter vs. median SNR for
each source.
The vertical axis has is log, but linear from $-10^{-3}$ to $10^{-3}$.
The fact that this is symmetric is strange to me.
}
\end{figure*}

\begin{figure}
\begin{tabular}{cc}
\subfloat{\includegraphics[width=2.5in]{figures/0.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/1.pdf}} \\
\subfloat{\includegraphics[width=2.5in]{figures/2.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/3.pdf}} \\
\subfloat{\includegraphics[width=2.5in]{figures/4.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/5.pdf}} \\
\subfloat{\includegraphics[width=2.5in]{figures/6.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/7.pdf}} \\
\end{tabular}
\caption{Visit velocities for a random batch of APOGEE stars. Error bars are
plotted for all points, but are sometimes smaller than the marker.}
\end{figure}

\begin{figure}
\begin{tabular}{cc}
\subfloat{\includegraphics[width=2.5in]{figures/8.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/9.pdf}} \\
\subfloat{\includegraphics[width=2.5in]{figures/10.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/11.pdf}} \\
\subfloat{\includegraphics[width=2.5in]{figures/12.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/13.pdf}} \\
\subfloat{\includegraphics[width=2.5in]{figures/14.pdf}} &
\subfloat{\includegraphics[width=2.5in]{figures/15.pdf}} \\
\end{tabular}
\caption{Visit velocities for a random batch of APOGEE stars.}
\end{figure}

\end{document}

### 2017-12-10 paper/1-catalog/twoface-catalog.tex
added importance sampling math

@@
 
+\subsection{Infer the jitter}
+
+For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
+primary orbital parameters $\bs{w} = (P, e, \omega, \phi_0, K, v_0)$ and the
+jitter parameter, $y = \ln s^2$, using {\it The Joker}.
+To obtain this sampling, we use an interim prior on the jitter parameter
+parametrized by a mean and root-variance $\alpha_0 = (\mu_{y,0}, \sigma_{y,0})$.
+
+\begin{eqnarray}
+    \bs{\theta}_{k} &=& (\bs{w}, y)_{k}\\
+    \bs{\theta}_{k} &\sim& p(\bs{\theta}_k \given \{D_n\}, \alpha_0)
+\end{eqnarray}
+
+New hyperprior, parameters to infer: $\alpha = (\mu_y, \sigma_y)$.
+
+We want to obtain posterior samples in
+\begin{eqnarray}
+    p(\alpha \given \{D_n\}) &\propto& p(\{D_n\} \given \alpha) \, p(\alpha)
+\end{eqnarray}
+where
+\begin{eqnarray}
+    p(\{D_n\} \given \alpha) &=& \prod_n^N p(D_n \given \alpha) \quad .
+\end{eqnarray}
+
+\begin{eqnarray}
+    p(D_n \given \alpha) &=& \int \dd \theta_n \, p(D_n \given \theta_n) \,
+      p(\theta_n \given \alpha)\\
+    &=& \int \dd \theta_n \, p(D_n \given \theta_n) \, p(\theta_n \given \alpha) \,
+      \frac{p(\theta_n \given D_n, \alpha_0)}{p(\theta_n \given D_n, \alpha_0)}\\
+    &=& \int \dd \theta_n \, p(D_n \given \theta_n) \, p(\theta_n \given \alpha) \,
+      p(\theta_n \given D_n, \alpha_0) \,
+      \frac{p(D_n \given \alpha_0)}{p(D_n \given \theta_n) \,
+      p(\theta_n \given \alpha_0)}\\
+    &=& p(D_n \given \alpha_0) \, \int \dd \theta_n \,
+      \frac{p(\theta_n \given \alpha)}{p(\theta_n \given \alpha_0)} \,
+      p(\theta_n \given D_n, \alpha_0)\\
+    &\approx& \mathcal{Z}_n \, \frac{1}{K_n} \,
+      \sum_k^{K_n} \frac{p(\theta_n^{(k)} \given \alpha)}{p(\theta_n^{(k)} \given \alpha_0)}\\
+    &=& \frac{\mathcal{Z}_n}{K_n} \,
+      \sum_k^{K_n} \frac{p(y_n^{(k)} \given \alpha)}{p(y_n^{(k)} \given \alpha_0)}
+\end{eqnarray}
+
+\begin{eqnarray}
+    p(\{D_n\} \given \alpha) &\propto& \prod_n^N \frac{1}{K_n} \,
+      \sum_k^{K_n} \frac{p(y_n^{(k)} \given \alpha)}{p(y_n^{(k)} \given \alpha_0)}\\
+    \ln p(\{D_n\} \given \alpha) &\propto& \sum_n^N \left[
+      \ln\left( \sum_k^{K_n} \frac{p(y_n^{(k)} \given \alpha)}{p(y_n^{(k)} \given \alpha_0)} \right)
+      - \ln K_n\right]\\
+    &\propto& \sum_n^N \left[
+      \underset{k}{\rm logsumexp}\left[ \ln{p(y_n^{(k)} \given \alpha)} - \ln{p(y_n^{(k)} \given \alpha_0)} \right]
+      - \ln K_n\right]
+\end{eqnarray}
+
 \section{Results}

### 2017-12-14 twoface/db/model.py
add fparam columns

@@ -133,10 +133,6 @@ class AllStar(Base):
     min_h = Column("min_h", types.REAL)
     max_h = Column("max_h", types.REAL)
     min_jk = Column("min_jk", types.REAL)
-    # param = Column("param", postgresql.ARRAY(types.REAL))
-    # fparam = Column("fparam", postgresql.ARRAY(types.REAL))
-    # param_cov = Column("param_cov", postgresql.ARRAY(types.REAL))
-    # fparam_cov = Column("fparam_cov", postgresql.ARRAY(types.REAL))
     teff = Column("teff", types.REAL)
     teff_err = Column("teff_err", types.REAL)
     logg = Column("logg", types.REAL)
@@ -267,6 +263,18 @@ class AllStar(Base):
     pm_src = Column("pm_src", types.String)
     # fparam_class = Column("fparam_class", postgresql.ARRAY(types.REAL))
     # chi2_class = Column("chi2_class", postgresql.ARRAY(types.REAL))
+
+    # need to split up FPARAM
+    # Teff, logg, vmicro, [M/H], [C/M], [N/M], [alpha/M], vsini, vmacro
+    # param = Column("param", postgresql.ARRAY(types.REAL))
+    # param_cov = Column("param_cov", postgresql.ARRAY(types.REAL))
+    _fp_str = "fparam{i} = Column('fparam{i}', types.REAL)"
+    _cov_str = "fparam_cov{i}{j} = Column('fparam_cov{i}{j}', types.REAL)"
+    for i in range(9):
+        exec(_fp_str.format(i=i))
+        for j in range(9):
+            if j < i: continue
+            exec(_cov_str.format(i=i, j=j))
 
     def __repr__(self):
         return ("<ApogeeStar(id='{0}', apogee_id='{1}', {2} results)>"

### 2017-12-14 twoface/db/tests/test_database.py

@@ -41,7 +41,7 @@ class TestDB(object):
 
         # get star entry and check total num of visits
         star = s.query(AllStar).filter(AllStar.target_id == test_target_ID).one()
-        assert len(star.visits) == 6
+        assert len(star.visits) == 3
 
         # get a visit and check that it has one star
         visit = s.query(AllVisit).filter(AllVisit.target_id == test_target_ID).limit(1).one()

### 2017-12-14 twoface/db/model.py
get fparam, fparam_cov

@@ -5,6 +5,7 @@ import datetime
 
 # Third-party
 import astropy.units as u
+import numpy as np
 from sqlalchemy import Table, Column, types
 from sqlalchemy.schema import ForeignKey
 from sqlalchemy.orm import relationship, backref
@@ -283,6 +284,29 @@ class AllStar(Base):
     def apogeervdata(self, clean=False):
         """Return a `twoface.data.APOGEERVData` instance for this star. """
         return star_to_apogeervdata(self, clean=clean)
+
+    @property
+    def fparam(self):
+        vals = np.zeros(9)
+        for i in range(9):
+            vals[i] = getattr(self, 'fparam{0}'.format(i))
+        return vals
+
+    @property
+    def fparam_cov(self):
+        vals = np.zeros((9,9))
+        for i in range(9):
+            for j in range(9):
+                name = 'fparam{0}{1}'.format(i, j)
+                if hasattr(self, name):
+                    vals[i,j] = getattr(self, name)
+                    vals[j,i] = getattr(self, name)
+
+                name = 'fparam{0}{1}'.format(j, i)
+                if hasattr(self, name):
+                    vals[i,j] = getattr(self, name)
+                    vals[j,i] = getattr(self, name)
+        return vals
 
 class AllVisit(Base):
     __tablename__ = 'allvisit'

### 2017-12-14 paper/1-catalog/twoface-catalog.tex
change subscripting

@@
 
-\title{Binary companions of red giant stars with APOGEE I: \\
+\title{Binary companions of red giant stars with APOGEE: \\
        catalog and companions that survive the common envelope}
@@
 
-\todo{An alternate approach to this intro is to start with an overview of
-      binary stars, radial velocity measurements of binarity in general...}
+\todo{Intro is missing many citations}
 
-\todo{As is, the intro is missing many citations}
+Nearly all stars form with stellar companions in multiplets or clusters, through
+subsequent dynamical evolution that stars may lose their neighbors.
+Main sequence stars in the solar neighborhood more often appear in binary or
+multiple star systems than singles (\citealt{ragavhan,tokovinin,todo}).
+- Related to fragmentation in collapsing protostellar / molecular cloud ->
+  multiple stars (e.g., \citealt{Raskutti:2016})
+- Population of stellar binaries and multiplets -> information about star
+  formation, dynamics
+
+Even without stellar or sub-stellar companions, single stars likely have
+planetary-mass companions (\citealt{todo}.
+- Different formation channel for planets?
+- Sub-stellar companions?
+
+- Almost all stars have companions of some mass
+These systems have a large dynamic range of constituent and orbital
+characteristics.
+For example, for binary main-sequence stars, mass-ratios span from XX to 1 and
+periods from XX to YY years.
+- Population of companions and trends (e.g., with metallicity) -> star and
+  planet formation processes
+
+- Expect differences in evolved star pop. from main sequence populations:
+  - mass loss
+  - envelope -> engulfment?
+- But less is known about the population of companions to evolved stars.
+- Current samples limited in dynamical range of chemical properties, location,
+  age (local to sun)
+- Brighter, but typically larger distances, direct imaging not feasible
+- Eclipsing companions good, but low mass companions stars -> hard to find
+  (giants are big)
+- Need large sample of stars with time-domain spectroscopy
 
@@
 However, with few or imprecise radial-velocity measurements, the orbital
-properties of the companion(s) are very poorly constrained.
+properties of the companion(s) are very poorly constrained (e.g.,
+\citealt{Price-Whelan:2017}).
+The vast majority of spectroscopic targets with repeat observations in the
+largest (by number of objects) stellar spectroscopic surveys are often observed
+just a few times with sparse, non-uniform phase coverage.
 Most prior searches for companions using survey RV data have therefore
-restricted their searches to only sources with many epochs, so that the orbital
-solution can be unambiguously determined.
-The vast majority of spectroscopic targets with repeat observations in the
-largest (by number of objects) stellar spectroscopic surveys are in the opposite
-regime: targets are often observed just a few times with sparse, non-uniform
-phase coverage.
-
-% With such data, even if the period or other generic orbital parameters
-% are very poorly constrained, limits on companion mass or pericentric distance
-% of the companion may still XX.
+restricted their searches to only sources with many, high-precision epochs, so
+that the orbital solution can be unambiguously determined (e.g.,
+\citealt{Troup:2016}), or have used simple statistics computed from the data to
+study multiplicity (e.g., ${\rm RV}_{\rm max} - {\rm RV}_{\rm min}$;
+\citealt{Badenes:2017}).
 
@@
     &\approx& \mathcal{Z}_n \, \frac{1}{K_n} \,
-      \sum_k^{K_n} \frac{p(\theta_n^{(k)} \given \alpha)}{p(\theta_n^{(k)} \given \alpha_0)}\\
+      \sum_k^{K_n} \frac{p(\theta_{nk} \given \alpha)}{p(\theta_{nk} \given \alpha_0)}\\
     &=& \frac{\mathcal{Z}_n}{K_n} \,
-      \sum_k^{K_n} \frac{p(y_n^{(k)} \given \alpha)}{p(y_n^{(k)} \given \alpha_0)}
+      \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
 \end{eqnarray}
@@
     p(\{D_n\} \given \alpha) &\propto& \prod_n^N \frac{1}{K_n} \,
-      \sum_k^{K_n} \frac{p(y_n^{(k)} \given \alpha)}{p(y_n^{(k)} \given \alpha_0)}\\
+      \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}\\
     \ln p(\{D_n\} \given \alpha) &\propto& \sum_n^N \left[
-      \ln\left( \sum_k^{K_n} \frac{p(y_n^{(k)} \given \alpha)}{p(y_n^{(k)} \given \alpha_0)} \right)
+      \ln\left( \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
       - \ln K_n\right]\\
     &\propto& \sum_n^N \left[
-      \underset{k}{\rm logsumexp}\left[ \ln{p(y_n^{(k)} \given \alpha)} - \ln{p(y_n^{(k)} \given \alpha_0)} \right]
+      \underset{k}{\rm logsumexp}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
       - \ln K_n\right]

### 2017-12-15 paper/1-catalog/twoface-catalog.tex
intro stab

@@
 
-Nearly all stars form with stellar companions in multiplets or clusters, through
-subsequent dynamical evolution that stars may lose their neighbors.
+Stars typically have companions.
+
 Main sequence stars in the solar neighborhood more often appear in binary or
-multiple star systems than singles (\citealt{ragavhan,tokovinin,todo}).
-- Related to fragmentation in collapsing protostellar / molecular cloud ->
-  multiple stars (e.g., \citealt{Raskutti:2016})
-- Population of stellar binaries and multiplets -> information about star
-  formation, dynamics
+multiple star systems rather than as solitary stars
+(\citealt{ragavhan,tokovinin,todo}).
+This is likely a generic outcome of star formation: turbulent fragmentation in
+collapsing protostellar clouds tends to produce multiplets of stars within
+length-scales comparable to their spheres of influence (e.g.,
+\citealt{Raskutti:2016}).
+Binary and multiple star systems are therefore of great interest in
+astrophysics: the population of stars and their companions encodes information
+about star formation processes (\citealt{todo}), stellar parameters and
+evolution (\citealt{todo}), and the dynamics of multi-body systems
+(\citealt{todo}).
 
 Even without stellar or sub-stellar companions, single stars likely have
-planetary-mass companions (\citealt{todo}.
-- Different formation channel for planets?
-- Sub-stellar companions?
+planetary-mass companions (\citealt{todo}).
+\todo{Planetary systems are interesting because...yadda yadda}
+Differences between the population of stellar companions and that of
+planetary-mass companions likely relate directly to the different formation
+channels for these objects.
 
-- Almost all stars have companions of some mass
-These systems have a large dynamic range of constituent and orbital
+\todo{should we name ``stars with companions'' something like ``stellar
+systems''?}
+Stars with companions have a large dynamic range of constituent and orbital
 characteristics.
-For example, for binary main-sequence stars, mass-ratios span from XX to 1 and
+For example, for binary main sequence stars, mass-ratios span from XX to 1 and
 periods from XX to YY years.
-- Population of companions and trends (e.g., with metallicity) -> star and
+For planetary companions of main sequence stars, \todo{}.
+- Some indication of trends (e.g., with metallicity) -> star and
   planet formation processes
 
+Most of what is known about stellar companions comes from main sequence stars.
+- Some evolved stars have special roles in exotic systems like cataclysmic
+  variables (donor stars)
+- But less is known about the population of companions to evolved stars.
 - Expect differences in evolved star pop. from main sequence populations:
@@
   - envelope -> engulfment?
-- But less is known about the population of companions to evolved stars.
 - Current samples limited in dynamical range of chemical properties, location,

### 2018-01-21 scripts/run_more_samples.py
build in some hacks

@@ -79,9 +79,12 @@ def main(config_file, pool, seed, overwrite=False):
                          max_batch_size=2**24) # MAGIC NUMBER
 
     # Get all stars in this JokerRun that need more prior samples
+    # TODO HACK: because I suck, some got mis-labeled "needs mcmc" - so re-run
+    # on all of those too:
     star_query = session.query(AllStar).join(StarResult, JokerRun, Status)\
                                        .filter(JokerRun.name == run.name)\
-                                       .filter(Status.id == 1)
+                                       .filter(Status.id.in_([1, 2]))
+    #                                    .filter(Status.id == 1)
 
     # Base query to get a StarResult for a given Star so we can update the
     # status, etc.
@@ -99,7 +102,7 @@ def main(config_file, pool, seed, overwrite=False):
 
     count = 0 # how many stars we've processed in this star batch
     batch_size = 16 # MAGIC NUMBER: how many stars to process before committing
-    for star in star_query.all():
+    for star in star_query.limit(4).all():
 
         if result_query.filter(AllStar.apogee_id == star.apogee_id).count() < 1:
             logger.debug('Star {0} has no result object!'
@@ -124,6 +127,11 @@ def main(config_file, pool, seed, overwrite=False):
                 # HACK: data=data, prior_cache_file=run.prior_samples_file,
                 data=data, prior_cache_file=new_path,
                 return_logprobs=True)
+
+            samples, ln_prior = joker.iterative_rejection_sample(
+                data=data, n_requested_samples=run.requested_samples_per_star,
+                # HACK: prior_cache_file=run.prior_samples_file,
+                prior_cache_file=new_path, return_logprobs=True)
 
         except Exception as e:
             logger.warning("\t Failed sampling for star {0} \n Error: {1}"

### 2018-01-21 scripts/run_more_samples.py
split batches better

@@ -67,7 +67,8 @@ def main(config_file, pool, seed, overwrite=False):
     rnd = np.random.RandomState(seed=seed)
     logger.debug("Creating TheJoker instance with {0}, {1}".format(rnd, pool))
     params = run.get_joker_params()
-    joker = TheJoker(params, random_state=rnd, pool=pool)
+    joker = TheJoker(params, random_state=rnd, pool=pool,
+                     n_batches=8 * pool.size) # HACK: magic number
 
     # TODO: a temporary hack for the end-of-2017 apogee-jitter run: we need to
     # make sure a 2nd prior cache exists with even more samples!

### 2018-01-21 scripts/run_more_samples.py
only 1

@@ -103,7 +103,7 @@ def main(config_file, pool, seed, overwrite=False):
 
     count = 0 # how many stars we've processed in this star batch
     batch_size = 16 # MAGIC NUMBER: how many stars to process before committing
-    for star in star_query.limit(4).all():
+    for star in star_query.limit(1).all(): # HACK: REMOVE
 
         if result_query.filter(AllStar.apogee_id == star.apogee_id).count() < 1:
             logger.debug('Star {0} has no result object!'

### 2018-01-21 scripts/run_more_samples.py
abck to all

@@ -103,7 +103,7 @@ def main(config_file, pool, seed, overwrite=False):
 
     count = 0 # how many stars we've processed in this star batch
     batch_size = 16 # MAGIC NUMBER: how many stars to process before committing
-    for star in star_query.limit(1).all(): # HACK: REMOVE
+    for star in star_query.all():
 
         if result_query.filter(AllStar.apogee_id == star.apogee_id).count() < 1:
             logger.debug('Star {0} has no result object!'

### 2018-01-21 paper/1-catalog/twoface-catalog.tex
helpers
minor changes

@@
 Even few observations are useful for detecting companions, though such data are
-difficult to use for constraining the physical properties of companions because
-permitted orbital solutions are highly degenerate.
+difficult to use for constraining physical properties because permitted orbital
+solutions are highly degenerate.
 % Aims
-Here we perform a search for secondary companions of red giant branch (\RGB)
-stars using data from the \apogee\ survey (\DR), which has measured multi-epoch
-($\geq 3$) but sparse radial velocities for $> 80,000$ \RGB\ stars.
+Here we perform a search for secondary companions of stars observed by the
+\apogee\ survey (\DR), which has measured multi-epoch ($\geq 3$) but sparse
+radial velocities for close to 100,000 stars.
+% Methods
+We use a custom-built Monte Carlo sampler (\thejoker) to deliver (often highly
+multi-modal) posterior beliefs about companion orbital parameters.
 In some cases, where the primary-star masses are known, the mass-function
@@
 inferred.
-% Methods
-We use a custom-built Monte Carlo sampler (\thejoker) combined with stellar
-parameters and primary mass estimates to deliver (often highly multi-modal)
-posterior beliefs about companion mass, pericenter distance, and other orbital
-parameters for the stars.
-We compare the companion populations of red clump (\RC) versus \RGB\ stars at
-similar surface gravity and temperature, and study the metallicity dependence
-of companion properties.
 % Results
@@
 \todo{We additionally find trends in XX with chemical abundances.}
+
+We compare the companion populations of red clump (\RC) versus \RGB\ stars at
+similar surface gravity and temperature, and study the metallicity dependence
+of companion properties.
 
@@
 primary orbital parameters $\bs{w} = (P, e, \omega, \phi_0, K, v_0)$ and the
-jitter parameter, $y = \ln s^2$, using {\it The Joker}.
+jitter parameter, $y = \ln s^2$, using \textit{The Joker}.
 To obtain this sampling, we use an interim prior on the jitter parameter
@@
 
-It is a pleasure to thank XX.
+It is a pleasure to thank Keith Hawkins (Columbia), XXXXXXX.
 

### 2018-01-21 scripts/run_apogee.py
mixxing import

@@ -45,6 +45,7 @@ from thejoker.sampler import TheJoker
 import yaml
 
 # Project
+from twoface import unimodal_P
 from twoface.log import log as logger
 from twoface.db import db_connect, get_run
 from twoface.db import JokerRun, AllStar, StarResult, Status

### 2018-01-23 paper/1-catalog/twoface-catalog.tex
more text

@@
 \newcommand{\apogee}{\project{\acronym{APOGEE}}}
-\newcommand{\sdssiii}{\project{\acronym{SDSS-III}}}
+\newcommand{\sdssiv}{\project{\acronym{SDSS-IV}}}
 \newcommand{\thejoker}{\project{The~Joker}}
@@
 
-Stars typically have companions.
-
-Main sequence stars in the solar neighborhood more often appear in binary or
-multiple star systems rather than as solitary stars
-(\citealt{ragavhan,tokovinin,todo}).
+Stars typically have companions. Main sequence stars in the solar neighborhood
+more often appear in binary or multiple star systems rather than as solitary
+stars (\citealt{ragavhan,tokovinin,todo}).
 This is likely a generic outcome of star formation: turbulent fragmentation in
@@
 
-Even without stellar or sub-stellar companions, single stars likely have
-planetary-mass companions (\citealt{todo}).
-\todo{Planetary systems are interesting because...yadda yadda}
-Differences between the population of stellar companions and that of
-planetary-mass companions likely relate directly to the different formation
-channels for these objects.
+% Even without stellar or sub-stellar companions, the vast majority of stars have
+% planetary-mass companions (\citealt{todo}).
+% \todo{Planetary systems are interesting because...blah}
+% Differences between the population of stellar companions and that of
+% planetary-mass companions likely relate directly to the different formation
+% channels for these objects.
 
-\todo{should we name ``stars with companions'' something like ``stellar
-systems''?}
-Stars with companions have a large dynamic range of constituent and orbital
+Most of what is known about stellar companions comes from studies of
+main-sequence (MS) stars.
+MS stars with companions have a large dynamic range of constituent and orbital
 characteristics.
-For example, for binary main sequence stars, mass-ratios span from XX to 1 and
-periods from XX to YY years.
-For planetary companions of main sequence stars, \todo{}.
-- Some indication of trends (e.g., with metallicity) -> star and
-  planet formation processes
+For example, binary stars have mass-ratios that span from $\approx 10^-3$
+\todo{citation because this is made up} to 1, and have periods from days to
+millions of years.
+\todo{Describe distributions over these quantities}
+\todo{multiplicity in general: abundance, numbers, etc.}
+\todo{some other stats from Moe and Di Stefano}
 
-Most of what is known about stellar companions comes from main sequence stars.
-- Some evolved stars have special roles in exotic systems like cataclysmic
-  variables (donor stars)
-- But less is known about the population of companions to evolved stars.
+Some evolved stars have special roles in well-studied exotic multiple systems
+like dwarf novae or cataclysmic variables \todo{are those examples correct?},
+but by-in-large, less is known about the population of companions to evolved
+stars.
+
 - Expect differences in evolved star pop. from main sequence populations:
@@
 - Need large sample of stars with time-domain spectroscopy
+
+% For planetary companions of main sequence stars, \todo{}.
+% - Some indication of trends (e.g., with metallicity) -> star and
+%   planet formation processes
 
@@
 orbital parameters, given any number of radial-velocity measurements.
-Here we use \thejoker\ to generate posterior pdf samples for \todo{which?} stars
-observed by the \apogee\ surveys (see \sectionname~\ref{sec:data};
-\citealt{Majewski:2015}).
+Here we use \thejoker\ to generate posterior pdf samples for stars observed by
+the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2015}).
 
@@
 fit as single-line objects.
+When this constraint is not met, \thejoker\ will in general fail, and a model
+that fits for a mixture of stellar spectra is more appropriate (e.g.,
+\citealt{Kareem}).
 The subset of \RGB\ stars in the ``red clump'' (\RC) are even more powerful as
@@
 With primary-star mass estimates, the binary-orbit fitting will return
-$m_2\,\sin i$ estimates for the secondary, and not just estimates of the
-mass-function.
+$m_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
+of the so-called ``binary mass function.''
 Additionally, the \apogee\ pipelines (\citealt{Garcia-Perez:2016}) and also
@@
 If there are causal relationships between chemical abundances and binary
-companions---as are expected---these should be measurable here.
+companions---as are expected---these should be measurable.
 
@@
 \sectionname~\ref{sec:whatever}), we deliver a catalog of binary-star systems
-from the \apogee\ survey with no cuts or constraints on data quality or volume.
+from the \apogee\ survey \todo{with ... (tease some statistics)}.
 We additionally highlight the subset of this catalog with \RC\ star primaries,
@@
 
-\todo{HOGG: fill in this ish}
+All data used in this work comes from the publicly-available data release 14
+(\DR) of the \apogee\ survey (\citealt{apogeedr14}), a component of the Sloan
+Digital Sky Survey IV (\sdssiv; \citealt{sdss,Blanton:2017}).
+\apogee\ is an infrared spectroscopic survey designed to map \RGB\ stars across
+much of the Milky Way.
+Targets are selected with simple color and brightness cuts, but the survey uses
+fiber-plugged plates with a maximum of \todo{100?} fibers and a $\approx
+1.5~{\rm deg}^2$ field of view, leading to ``pencil-beam''-like sampling of the
+stellar distribution.
+Data taken as part of the \apogee\ survey is reduced with the \todo{APOGEE
+spectral parameter...} (\acronym{ASPCAP}; \citealt{Garcia-Perez:2016}), which
+measures stellar parameters, chemical abundances, and radial velocities for each
+target.
+\todo{talk about visits, how the velocities measured...}
 
-HOGG: Why and how does \apogee\ rock it?
-
-ASPCAP: \citealt{Garcia-Perez:2016}
-SDSSIV: \citealt{Blanton:2017}
-
-Overview of what \apogee\ is, and what data we are using.
-
-How we updated the individual-visit radial velocity measurements.
-
-What we did (if anything) with the missing visits.
+From the \apogee\ \DR\ data products (the \texttt{allStar} and \texttt{allVisit}
+files), we select all stars with $\geq 3$ visits that each pass a set of quality
+cuts.
+For each visit, we require that the following bits are not set in the
+\texttt{STARFLAGS} bitmask: \texttt{TODO}.
+For each star, we require that the following bits are not set in the
+\texttt{ASPCAPFLAGS} bitmask: \texttt{TODO}.
+After these cuts, and the requirement of $\geq 3$ visits for a given star, we
+are left with \todo{XX} visits for \todo{XX} unique stars (unique
+\texttt{APOGEE\_ID}).
+\todo{add numpy code for exact filter?}
 

### 2018-02-01 twoface/plot.py
option to not plot P extreme

@@ -12,7 +12,7 @@ __all__ = ['plot_data_orbits']
 
 def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
                      xlim_choice='wide', n_times=4096, title=None,
-                     ax=None):
+                     ax=None, highlight_P_extrema=True):
     """
     Plot the APOGEE RV data vs. time and orbits computed from The Joker samples.
 
@@ -71,20 +71,18 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
                    plot_kwargs=dict(color='#aaaaaa', alpha=0.25, linewidth=0.5),
                    data_plot_kwargs=dict(zorder=5, elinewidth=1,))
 
-    # Darken the shortest period sample
-    dark_style = dict(color='#333333', alpha=0.5, linewidth=1, zorder=10)
+    if highlight_P_extrema:
+        # Darken the shortest period sample
+        dark_style = dict(color='#333333', alpha=0.5, linewidth=1, zorder=10)
 
-    P_min_samples = samples[samples['P'].argmin()]
-    plot_rv_curves(P_min_samples, t_grid, rv_unit=u.km/u.s, ax=ax,
-                   n_plot=1, plot_kwargs=dark_style)
+        P_min_samples = samples[samples['P'].argmin()]
+        plot_rv_curves(P_min_samples, t_grid, rv_unit=u.km/u.s, ax=ax,
+                       n_plot=1, plot_kwargs=dark_style)
 
-    # Darken the longest period sample
-    P_max_samples = samples[samples['P'].argmax()]
-    plot_rv_curves(P_max_samples, t_grid, rv_unit=u.km/u.s, ax=ax,
-                   n_plot=1, plot_kwargs=dark_style)
-
-    # if now.mjd < t_grid.max():
-    #     ax.axvline(now.mjd, zorder=-10, color='#31a354', alpha=0.4)
+        # Darken the longest period sample
+        P_max_samples = samples[samples['P'].argmax()]
+        plot_rv_curves(P_max_samples, t_grid, rv_unit=u.km/u.s, ax=ax,
+                       n_plot=1, plot_kwargs=dark_style)
 
     ax.set_xlim(t_grid.min(), t_grid.max())
 

### 2018-02-01 paper/1-catalog/twoface-catalog.tex
data

@@
 % \setlength{\parindent}{1.1\baselineskip}
+
+\graphicspath{{figures/}}
 
@@
 
-From the \apogee\ \DR\ data products (the \texttt{allStar} and \texttt{allVisit}
-files), we select all stars with $\geq 3$ visits that each pass a set of quality
-cuts.
-For each visit, we require that the following bits are not set in the
-\texttt{STARFLAGS} bitmask: \texttt{TODO}.
-For each star, we require that the following bits are not set in the
-\texttt{ASPCAPFLAGS} bitmask: \texttt{TODO}.
+We use the primary data products from \apogee\ \DR\ (i.e. the \texttt{allStar}
+and \texttt{allVisit} files) which contain 258475 unique source IDs
+(\texttt{APOGEE\_ID}) and 1054381 unique visits.
+We select all stars with $\geq 3$ visits that each pass a set of quality cuts,
+described below.
+For each visit, we require that the visit velocity uncertainty is $< 100~\kms$
+(\texttt{VRELERR}) and the following bits are not set in the \texttt{STARFLAGS}
+bitmask: \texttt{PERSIST\_HIGH}, \texttt{PERSIST\_JUMP\_POS},
+\texttt{PERSIST\_JUMP\_NEG}, \texttt{VERY\_BRIGHT\_NEIGHBOR}, \texttt{LOW\_SNR}.
+For each star, we require that $0 < \log g < 3.5$ and the following bits are not
+set in the \texttt{ASPCAPFLAGS} bitmask: \texttt{STAR\_BAD}.
 After these cuts, and the requirement of $\geq 3$ visits for a given star, we
@@
 \todo{add numpy code for exact filter?}
+\figurename~\ref{fig:nvisits} shows the number of stars in several bins of
+number of visits that pass the above quality cuts: \todo{XX}\% of the stars in
+\DR\ have $< 8$ visits.
+
+% Notebook: "Numbers after cuts"
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{nvisits.pdf}
+\end{center}
+\caption{%
+TODO: \label{fig:nvisits}
+}
+\end{figure}
 

### 2018-02-06 paper/1-catalog/twoface-catalog.tex
update to aas 62
writing

@@
-\documentclass[modern, letterpaper]{aastex61}
+\documentclass[modern, letterpaper]{aastex62}
 
@@
 
-\newcommand{\nRC}{XX}
-\newcommand{\nRGB}{XX}
-\newcommand{\ntotal}{XX}
-\newcommand{\ncompanions}{XX}
+\newcommand{\nprior}{536,870,912}
+\newcommand{\nposterior}{256}
 
@@
 Digital Sky Survey IV (\sdssiv; \citealt{sdss,Blanton:2017}).
-\apogee\ is an infrared spectroscopic survey designed to map \RGB\ stars across
-much of the Milky Way.
+\apogee\ is designed to map stars across much of the Milky Way by obtaining
+high-resolution ($R \sim 22500$) infrared ($H$-band) spectroscopy of primarily
+\RGB\ stars.
 Targets are selected with simple color and brightness cuts, but the survey uses
-fiber-plugged plates with a maximum of \todo{100?} fibers and a $\approx
-1.5~{\rm deg}^2$ field of view, leading to ``pencil-beam''-like sampling of the
-stellar distribution.
-Data taken as part of the \apogee\ survey is reduced with the \todo{APOGEE
-spectral parameter...} (\acronym{ASPCAP}; \citealt{Garcia-Perez:2016}), which
-measures stellar parameters, chemical abundances, and radial velocities for each
-target.
-\todo{talk about visits, how the velocities measured...}
+fiber-plugged plates with a maximum of 300 fibers per each $\approx 1.5~{\rm
+deg}^2$ field of view, leading to ``pencil-beam''-like sampling of the stellar
+distribution.
+Data taken as part of the \apogee\ survey is reduced with the \apogee\ Stellar
+Parameters and Chemical Abundances Pipeline (\acronym{ASPCAP};
+\citealt{Garcia-Perez:2016}), which measures stellar parameters, chemical
+abundances, and radial velocities for each target.
+In order to meet signal-to-noise ratio requirements, most \apogee\ stars are
+observed multiple times in a series of ``visits,'' typically with at least one
+visit separated by a month or more in order to help identify binary stars.
+\acronym{ASPCAP} iteratively solves for the relative velocities of each visit
+spectrum while determining the optimal co-addition of the visit spectra.
 
@@
 After these cuts, and the requirement of $\geq 3$ visits for a given star, we
-are left with \todo{XX} visits for \todo{XX} unique stars (unique
-\texttt{APOGEE\_ID}).
-\todo{add numpy code for exact filter?}
+are left with \todo{XX} visits for \todo{XX} unique sources.
 \figurename~\ref{fig:nvisits} shows the number of stars in several bins of
@@
 \caption{%
-TODO: \label{fig:nvisits}
+Number of \apogee\ stars in logarithmic bins of number of visits that pass the
+quality cuts described in \sectionname~\ref{sec:data}.
+In total, this work uses \todo{XX} visits and \todo{XX} unique sources.
+\label{fig:nvisits}
 }
@@
 
-\subsection{Orbit fitting and velocity modeling}\label{sec:fitting}
+\subsection{Orbit inference and velocity modeling}\label{sec:fitting}
 
@@
   At present, we ignore the possibility of higher-order multiple systems.
-  \todo{revisit this - do we want to also sample over a long-term trend?}
+  % \todo{revisit this - do we want to also sample over a long-term trend?}
 \item[Kepler] Related to the first point, all velocity variations of the primary
-  are gravitational, and we therefore ignore the possibility of intrinsic
-  variation from, e.g., stellar oscillations.
+  are gravitational, and we therefore ignore the possibility of coherent intrinsic variation from, e.g., stellar oscillations.
 \item[SB1] All spectra are single-lined; that is, we assume that the secondary
@@
   However, there are known double-lined binary stars in the \apogee\ catalog
-  (El-Badry et al., in prep.), and an unknown fraction of \RGB--\RGB\ binaries.
+  \citealt{El-Badry:2017}, and an expected but unknown fraction of \RGB--\RGB\
+  binaries.
 \item[simple noise model] Measurements are unbiased and noise estimates are
-  correct up to an \todo{unknown} extra variance (the ``jitter'').
+  correct up to an unknown extra variance.
   All noise contributions result in Gaussian uncertainties on the individual
   radial-velocity measurements.
+  We often refer to the extra variance as a ``jitter'' parameter because for
+  upper \RGB\ stars the inferred extra variance will be a combination of extra
+  uncertainty and true astrophysical surface jitter (\todo{CITATION NEEDED}).
 \end{description}
@@
 \thejoker\ is a custom-built Monte Carlo sampler designed to produce independent
-posterior samples in orbital parameters, given radial-velocity measurements
-under the assumptions above.
-Our parametrization is similar to the notation in \citet{Murray:2010}:
+posterior samples in Keplerian orbital parameters, given radial-velocity
+measurements under the assumptions listed above.
+Our parametrization of the orbital elements is similar to the notation in
+\citet{Murray:2010}:
 The radial velocity $v$ at time $t$ is given by
 \begin{equation}
-  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, \phi_0)\right) +
+  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, M_0)\right) +
     e\,\cos\omega]
 \end{equation}
-where $\bs{\theta} = (P,e,\phi_0,\omega,K,v_0)$---period, eccentricity, initial
-phase, argument of pericenter, velocity semi-amplitude, Barycentric velocity---
-and the true anomaly, $f$, is a function of the specified parameters (see
-\sectionname~2 of \citealt{Price-Whelan:2017} or \eqname~63 in
-\citealt{Murray:2010}).
+where $\bs{\theta} = (P,e,M_0,\omega,K,v_0)$---period, eccentricity, mean
+anomaly at a reference time $t_0$, argument of pericenter, velocity
+semi-amplitude, Barycentric velocity---and the true anomaly, $f$, is a function
+of time and the specified parameters (see \sectionname~2 of
+\citealt{Price-Whelan:2017} or \eqname~63 in \citealt{Murray:2010}).
 
 \thejoker\ was designed for the extremely multi-modal pdfs expected when the
-number of radial-velocity measurements of a source is small, or the data have
-low signal-to-noise.
+number of radial-velocity measurements of a source is small, or the data are
+sparse (in phase-coverage) or noisy.
 While other Markov Chain Monte Carlo (MCMC) methods have difficulty producing
-independent samples with such data, \thejoker\ succeeds by brute force.
+independent samples with such data, \thejoker\ succeeds by brute force:
 After generating an initial (very large) library of prior samples from an
@@
 proceeds iteratively: since it is easier to accept samples when the data is
-sparse or noisy, far more prior sample draws (and thus likelihood
-evaluations) must occur under very constraining data.
+sparse or noisy, far more prior sample draws (and thus likelihood evaluations)
+must occur under very constraining data.
+
+\subsection{Model selection}
 
@@
 
-\subsection{Infer the jitter}
+\section{Results}
 
+\subsection{Initial inidivual-system posterior samplings}
+\label{sec:initsampling}
+
+For all \todo{XX} \apogee\ stars with $\geq 3$ good visits (see
+\sectionname~\ref{sec:data}), we use \thejoker\ to generate posterior samplings
+for each star under the assumptions listed above (see
+\sectionname~\ref{sec:fitting}).
+We start by generating a library of \nprior\ prior samples generated under the
+prior defined in \citet{Price-Whelan:2017}: uniform or isotropic in angle
+parameters, uniform in log-period over the domain $[1, 32768]~{\rm day}$, and a
+beta distribution over eccentricity with fixed parameters
+(\citealt{Kipping:2013}).
+For the prior over the extra variance or jitter parameter, $s$, we initially use
+a Gaussian over the transformed parameter $y = \ln s^2$ with the mean and
+standard deviation of the initial distribution set to $(\mu_{y,0}, \sigma_{y,0})
+= (10.6, 3)$ in units of ${\rm m}~{\rm s}^{-1}$.
+
+We request \nposterior\ posterior samples for each source, and generate the
+samples using an iterative procedure that adapts to the \todo{informativeness}
+of the data to more efficiently generate posterior samples.
+For sources with very constraining data, \thejoker\ may return fewer than the
+requested number of samples.
+When this happens, we proceed in one of two ways:
+\begin{enumerate}
+    \item \emph{needs MCMC}: If just 1 posterior sample is returned after
+    exhausting the full library of prior samples, or if multiple (but fewer than
+    \nposterior) are returned that all lie within a single mode of the posterior
+    pdf, the orbit-space that is consistent with the data is treated as
+    effectively unimodal. In this case, we use the location of the returned
+    sample, or the median over returned samples to generate a small ball of
+    initial conditions and use the ensemble MCMC sampler (\citealt{Goodman:})
+    implemented in \python\ (\citealt{Foreman-Mackey:}) to continue sampling
+    until we obtain \nposterior\ samples.
+    \item \emph{needs more prior samples}: If more than 1 posterior sample is
+    returned after exhausting the full library of prior samples and the samples
+    lie in multiple modes of the posterior pdf, the star is flagged. In this
+    case, the only way forward is to generate more prior samples and continue
+    running \thejoker\ for these sources to generate more posterior samples.
+\end{description}
+
+A full run of \thejoker\ on all \todo{XX} \apogee\ stars used in this work took
+approximately 300 hours on a compute cluster with 448 cores, with the time
+dominated by sources with many ($\gtrsim 10$) visits.
+
+\subsection{Inferring the extra variance distribution}
+\label{sec:inferjitter}
+
+We use the per-source posterior samplings under the default prior described
+above to infer the parameters of the assumed Gaussian distribution over the
+log-jitter: $(\mu_y, \sigma_y)$.
+This inference serves as a test-case for future work, where we intend to use the
+independent posterior samplings to construct a hierarchical inference over
+companion property distributions, and is also a test of the visit velocity
+uncertainties reported in the \apogee\ data products.
+If the catalog uncertainties are significantly underestimated, we expect the
+inferred log-jitter distribution parameters to tend towards larger values.
+
+To update the log-jitter distribution parameters, we maximize the marginal
+likelihood of these population-level parameters.
+We compute this marginal likelihood using the per-object posterior samples
+re-weighted by the ratio of the value of the hyperprior evaluated at a new
+location over the value of the default prior (see
+\sectionname~\ref{sec:initsampling}).
+This trick has been used in other hierarchical inferences as a way to
+marginalize over the per-object parameters
+(\citealt{Hogg:2008,Foreman-Mackey:...}), and we describe how to compute the
+marginal likelihood in detail below.
+
+\todo{Notation relative to above...}
 For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
-primary orbital parameters $\bs{w} = (P, e, \omega, \phi_0, K, v_0)$ and the
-jitter parameter, $y = \ln s^2$, using \textit{The Joker}.
-To obtain this sampling, we use an interim prior on the jitter parameter
-parametrized by a mean and root-variance $\alpha_0 = (\mu_{y,0}, \sigma_{y,0})$.
+primary orbital parameters $\bs{w} = (P, e, \omega, M_0, K, v_0)$ and the
+jitter parameter, $y = \ln s^2$, using \thejoker.
+To obtain this sampling, we use an interim (Gaussian) prior on the jitter
+parameter parametrized by a mean and standard deviation $\alpha_0 = (\mu_{y,0},
+\sigma_{y,0})$.
 
@@
       \frac{p(\theta_n \given D_n, \alpha_0)}{p(\theta_n \given D_n, \alpha_0)}\\
-    &=& \int \dd \theta_n \, p(D_n \given \theta_n) \, p(\theta_n \given \alpha) \,
-      p(\theta_n \given D_n, \alpha_0) \,
-      \frac{p(D_n \given \alpha_0)}{p(D_n \given \theta_n) \,
-      p(\theta_n \given \alpha_0)}\\
     &=& p(D_n \given \alpha_0) \, \int \dd \theta_n \,
@@
       p(\theta_n \given D_n, \alpha_0)\\
-    &\approx& \mathcal{Z}_n \, \frac{1}{K_n} \,
-      \sum_k^{K_n} \frac{p(\theta_{nk} \given \alpha)}{p(\theta_{nk} \given \alpha_0)}\\
-    &=& \frac{\mathcal{Z}_n}{K_n} \,
+    &\approx& \frac{\mathcal{Z}_n}{K_n} \,
       \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
 \end{eqnarray}
+
+Factor y from theta, so most priors cancel. Up to multiplicative constant all we need. Above just for one source, full population:
 
@@
 
-\section{Results}
+TODO: apply to all stars with $>10$ visits
 
-\subsection{Catalog / posterior samples whatever} \label{sec:full-catalog}
+\subsection{Catalog of posterior samplings for APOGEE DR14 RGB stars}
+\label{sec:fullcatalog}
 
-\todo{APW}
+TODO: re-do exactly \sectionname~\ref{sec:initsampling}, but with $(\mu_{y}, \sigma_{y}) = (9.5, 1.6)$ again in units of ${\rm m}~{\rm s}^{-1}$.
+
+TODO: how many hit each category (flag) after the initial run
 

### 2018-02-07 paper/1-catalog/twoface-catalog.tex
Cleaning up and various style edits

@@
 \newcommand{\nposterior}{256}
+\newcommand{\nstars}{98,758}
+\newcommand{\nvisits}{388,196}
 
@@
 \citealt{Troup:2016}), or have used simple statistics computed from the data to
-study multiplicity (e.g., ${\rm RV}_{\rm max} - {\rm RV}_{\rm min}$;
+study multiplicity (e.g., $\textrm{RV}_\textrm{max} - \textrm{RV}_\textrm{min}$;
 \citealt{Badenes:2017}).
@@
 Targets are selected with simple color and brightness cuts, but the survey uses
-fiber-plugged plates with a maximum of 300 fibers per each $\approx 1.5~{\rm
-deg}^2$ field of view, leading to ``pencil-beam''-like sampling of the stellar
-distribution.
+fiber-plugged plates with a maximum of 300 fibers per each $\approx
+1.5~\textrm{deg}^2$ field of view, leading to ``pencil-beam''-like sampling of
+the stellar distribution.
 Data taken as part of the \apogee\ survey is reduced with the \apogee\ Stellar
@@
 We use the primary data products from \apogee\ \DR\ (i.e. the \texttt{allStar}
-and \texttt{allVisit} files) which contain 258475 unique source IDs
-(\texttt{APOGEE\_ID}) and 1054381 unique visits.
+and \texttt{allVisit} files) which contain 258,475 unique source IDs
+(\texttt{APOGEE\_ID}) and 1,054,381 unique visits.
 We select all stars with $\geq 3$ visits that each pass a set of quality cuts,
@@
 After these cuts, and the requirement of $\geq 3$ visits for a given star, we
-are left with \todo{XX} visits for \todo{XX} unique sources.
+are left with \nvisits\ visits for \nstars\ unique sources.
 \figurename~\ref{fig:nvisits} shows the number of stars in several bins of
@@
 \caption{%
+\todo{Update with new apogee db file}
 Number of \apogee\ stars in logarithmic bins of number of visits that pass the
 quality cuts described in \sectionname~\ref{sec:data}.
-In total, this work uses \todo{XX} visits and \todo{XX} unique sources.
+In total, this work uses \nvisits\ visits and \nstars\ unique sources.
 \label{fig:nvisits}
@@
 
-For all \todo{XX} \apogee\ stars with $\geq 3$ good visits (see
+For all \nstars\ \apogee\ stars with $\geq 3$ good visits (see
 \sectionname~\ref{sec:data}), we use \thejoker\ to generate posterior samplings
@@
 \sectionname~\ref{sec:fitting}).
-We start by generating a library of \nprior\ prior samples generated under the
-prior defined in \citet{Price-Whelan:2017}: uniform or isotropic in angle
-parameters, uniform in log-period over the domain $[1, 32768]~{\rm day}$, and a
-beta distribution over eccentricity with fixed parameters
-(\citealt{Kipping:2013}).
+We start by generating a library of \nprior\ prior samples generated under a
+prior similar to that defined in \citet{Price-Whelan:2017}: uniform or isotropic
+in angle parameters, uniform in log-period over the domain $[1,
+32768]~\textrm{day}$, and a beta distribution over eccentricity with fixed
+parameters (\citealt{Kipping:2013}).
 For the prior over the extra variance or jitter parameter, $s$, we initially use
@@
 standard deviation of the initial distribution set to $(\mu_{y,0}, \sigma_{y,0})
-= (10.6, 3)$ in units of ${\rm m}~{\rm s}^{-1}$.
+= (10.6, 3)$ in units of $\textrm{m}~\textrm{s}^{-1}$.
 
@@
     sample, or the median over returned samples to generate a small ball of
-    initial conditions and use the ensemble MCMC sampler (\citealt{Goodman:})
-    implemented in \python\ (\citealt{Foreman-Mackey:}) to continue sampling
-    until we obtain \nposterior\ samples.
+    initial conditions and use standard MCMC to continue sampling until we obtain \nposterior\ samples. In detail, we use an ensemble sampler
+    (\citealt{Goodman:}) implemented in \python\ (\citealt{Foreman-Mackey:}),
+    run with \todo{XX} walkers for \todo{XX} steps, and downsample the resulting
+    MCMC chains until we have \nposterior\ samples.
     \item \emph{needs more prior samples}: If more than 1 posterior sample is
@@
     lie in multiple modes of the posterior pdf, the star is flagged. In this
-    case, the only way forward is to generate more prior samples and continue
+    case, the only way to proceed is to generate more prior samples and continue
     running \thejoker\ for these sources to generate more posterior samples.
@@
 
-A full run of \thejoker\ on all \todo{XX} \apogee\ stars used in this work took
+A full run of \thejoker\ on all \nstars\ \apogee\ stars used in this work took
 approximately 300 hours on a compute cluster with 448 cores, with the time
@@
 
-We use the per-source posterior samplings under the default prior described
-above to infer the parameters of the assumed Gaussian distribution over the
-log-jitter: $(\mu_y, \sigma_y)$.
+As an initial use of the per-source posterior samplings, we use a hierarchical
+Bayesian model to infer the parameters of the (assumed Gaussian) distribution
+over the log-jitter parameter: $(\mu_y, \sigma_y)$.
 This inference serves as a test-case for future work, where we intend to use the
 independent posterior samplings to construct a hierarchical inference over
-companion property distributions, and is also a test of the visit velocity
-uncertainties reported in the \apogee\ data products.
-If the catalog uncertainties are significantly underestimated, we expect the
-inferred log-jitter distribution parameters to tend towards larger values.
+companion property distributions.
+This is also a test of the visit velocity uncertainties reported in the \apogee\
+data products: If the catalog uncertainties are significantly underestimated, we
+expect the inferred log-jitter distribution parameters to tend towards larger
+values.
 
-To update the log-jitter distribution parameters, we maximize the marginal
-likelihood of these population-level parameters.
+In detail, we maximize the marginal likelihood of the population-level
+parameters $\bs{\alpha} = (\mu_y, \sigma_y)$.
 We compute this marginal likelihood using the per-object posterior samples
-re-weighted by the ratio of the value of the hyperprior evaluated at a new
-location over the value of the default prior (see
+re-weighted by the ratio of the value of the hyperprior evaluated at a given
+vector $\bs{\alpha}$ over the value of the default prior at the previously
+assumed values, $\bs{\alpha}_0 = (\mu_{y,0}, \sigma_{y,0})$ (see
 \sectionname~\ref{sec:initsampling}).
@@
 marginalize over the per-object parameters
-(\citealt{Hogg:2008,Foreman-Mackey:...}), and we describe how to compute the
+(\citealt{Hogg:2008,Foreman-Mackey:...}); we describe how to compute the
 marginal likelihood in detail below.
 
-\todo{Notation relative to above...}
 For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
-primary orbital parameters $\bs{w} = (P, e, \omega, M_0, K, v_0)$ and the
-jitter parameter, $y = \ln s^2$, using \thejoker.
+primary orbital parameters $\bs{\theta} = (P, e, \omega, M_0, K, v_0)$ and the
+jitter parameter, $y = \ln s^2$, using \thejoker; For brevity in expressions
+below, we will use the vector
+\begin{equation}
+    \bs{w} = (\bs{\theta}, y)
+\end{equation}
+to represent the full set of parameters.
 To obtain this sampling, we use an interim (Gaussian) prior on the jitter
-parameter parametrized by a mean and standard deviation $\alpha_0 = (\mu_{y,0},
-\sigma_{y,0})$.
+parameter parametrized by a mean and standard deviation, i.e. $\alpha_0 =
+(\mu_{y,0}, \sigma_{y,0})$ as described above.
+For a given source, the posterior samples in the above parameters are drawn from
+the distribution
+\begin{equation}
+    \bs{w}_k \sim p(\bs{w}_k \given D, \bs{\alpha}_0)
+\end{equation}
+where $D$ represents the data for a given object.
 
-\begin{eqnarray}
-    \bs{\theta}_{k} &=& (\bs{w}, y)_{k}\\
-    \bs{\theta}_{k} &\sim& p(\bs{\theta}_k \given \{D_n\}, \alpha_0)
-\end{eqnarray}
-
-New hyperprior, parameters to infer: $\alpha = (\mu_y, \sigma_y)$.
-
-We want to obtain posterior samples in
-\begin{eqnarray}
-    p(\alpha \given \{D_n\}) &\propto& p(\{D_n\} \given \alpha) \, p(\alpha)
-\end{eqnarray}
-where
-\begin{eqnarray}
-    p(\{D_n\} \given \alpha) &=& \prod_n^N p(D_n \given \alpha) \quad .
-\end{eqnarray}
-
-\begin{eqnarray}
-    p(D_n \given \alpha) &=& \int \dd \theta_n \, p(D_n \given \theta_n) \,
-      p(\theta_n \given \alpha)\\
-    &=& \int \dd \theta_n \, p(D_n \given \theta_n) \, p(\theta_n \given \alpha) \,
-      \frac{p(\theta_n \given D_n, \alpha_0)}{p(\theta_n \given D_n, \alpha_0)}\\
-    &=& p(D_n \given \alpha_0) \, \int \dd \theta_n \,
-      \frac{p(\theta_n \given \alpha)}{p(\theta_n \given \alpha_0)} \,
-      p(\theta_n \given D_n, \alpha_0)\\
-    &\approx& \frac{\mathcal{Z}_n}{K_n} \,
-      \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
-\end{eqnarray}
+We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
+given a new set of hyper-parameters $\bs{\alpha}$
+\begin{equation}
+    p(\{D_n\} \given \bs{\alpha}) = \prod_n^N p(D_n \given \bs{\alpha})
+\end{equation}
+where in the above, we have assumed that this likelihood is separable ( the data
+for each source are independent).
+The per-source marginal likelihood in the above expression is given by
+\begin{align}
+    p(D_n \given \bs{\alpha}) &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \,
+      p(\bs{w}_n \given \bs{\alpha})\\
+    &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \, p(\bs{w}_n \given \bs{\alpha}) \,
+      \frac{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}\\
+    &= p(D_n \given \bs{\alpha}_0) \, \int \dd \bs{w}_n \,
+      \frac{p(\bs{w}_n \given \bs{\alpha})}{p(\bs{w}_n \given \bs{\alpha}_0)} \,
+      p(\bs{w}_n \given D_n, \bs{\alpha}_0) \label{eq:marglike} \quad .
+\end{align}
+Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
+be simplified to a sum over prior value ratios of the $K$ posterior samples for
+each $n$ star
+\begin{equation}
+    \approx \frac{\mathcal{Z}_n}{K_n} \,
+      \sum_k^{K_n} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
+      \quad .
+\end{equation}
 

### 2018-02-07 paper/1-catalog/twoface-catalog.tex
More methods

@@
 \newcommand{\nvisits}{388,196}
-
-% define macros for math
-\newcommand{\hyperpars}{\gamma}
-\newcommand{\pars}{\theta}
 
@@
 % Notebook: "Numbers after cuts"
-\begin{figure}[hp]
+\begin{figure}[h]
 \begin{center}
-\includegraphics[width=\textwidth]{nvisits.pdf}
+\includegraphics[width=0.7\textwidth]{nvisits.pdf}
 \end{center}
@@
 We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
-given a new set of hyper-parameters $\bs{\alpha}$
+given a new set of hyperparameters $\bs{\alpha}$
 \begin{equation}
@@
 Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
-be simplified to a sum over prior value ratios of the $K$ posterior samples for
-each $n$ star
+be simplified to a sum over prior value ratios of the $K$ posterior samples in
+the log-jitter parameter for each $n$ star
 \begin{equation}
-    \approx \frac{\mathcal{Z}_n}{K_n} \,
-      \sum_k^{K_n} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
+    \approx \frac{\mathcal{Z}_n}{K} \,
+      \sum_k^{K} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
+\end{equation}
+where we have canceled the other priors (over $\bs{\theta}$), and all
+normalization constants appear in the constant scale factor $\mathcal{Z}_n$.
+
+The above expresion gives the marginal likelihood of the velocity data for a single source given new hyperparameters $\bs{\alpha}$.
+The full marginal likelihood is then the product of these individual likelihoods
+\begin{align}
+    p(\{D_n\} \given \alpha) &\propto \prod_n^N \frac{1}{K} \,
+      \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
       \quad .
-\end{equation}
+\end{align}
+In practice, we evaluate the log-marginal-likelihood
+\begin{align}
+    \ln p(\{D_n\} \given \alpha) &\propto \sum_n^N \left[
+      \ln\left( \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
+      - \ln K\right]\label{eq:lnlike}\\
+    &\propto \sum_n^N \left[
+      \underset{k}{\textrm{logsumexp}}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
+      - \ln K\right]
+\end{align}
+where $\textrm{logsumexp}$ (the log-sum-exp trick) provides a more stable
+estimate of the sum in \eqname~\ref{eq:lnlike}.
 
-Factor y from theta, so most priors cancel. Up to multiplicative constant all we need. Above just for one source, full population:
+We use 1,825 stars with $>10$ visits and $\log g < 2$ (to avoid upper RGB stars
+that have large intrinsic jitter) and maximize the above likelihood to determine
+better hyperparameters for the log-jitter parameter distribution.
+\figurename~\ref{fig:infer-jitter} shows the distribution corresponding to the
+maximum-likelihood hyperparameters, $\alpha^* = (\mu_y^*, \sigma_y^*) = (9.50,
+1.64)$.
+\todo{discussion about this}
+\todo{some discussion of how this relates to empirical bayes?}
 
-\begin{eqnarray}
-    p(\{D_n\} \given \alpha) &\propto& \prod_n^N \frac{1}{K_n} \,
-      \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}\\
-    \ln p(\{D_n\} \given \alpha) &\propto& \sum_n^N \left[
-      \ln\left( \sum_k^{K_n} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
-      - \ln K_n\right]\\
-    &\propto& \sum_n^N \left[
-      \underset{k}{\rm logsumexp}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
-      - \ln K_n\right]
-\end{eqnarray}
-
-TODO: apply to all stars with $>10$ visits
+% Notebook: "Infer jitter"
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{infer-jitter}
+\end{center}
+\caption{%
+\todo{move notebook to figures path, re-do with "run2" or "run1" samples...}
+Inferred prior over log-jitter parameter ($y = \ln s^2$) using posterior samples
+for 1,825 lower-RGB stars with $>10$ visits.
+Left panel shows the ... Right panel shows the ...
+\label{fig:infer-jitter}
+}
+\end{figure}
 

### 2018-02-07 paper/1-catalog/twoface-catalog.tex
References.

@@
 more often appear in binary or multiple star systems rather than as solitary
-stars (\citealt{ragavhan,tokovinin,todo}).
-This is likely a generic outcome of star formation: turbulent fragmentation in
-collapsing protostellar clouds tends to produce multiplets of stars within
-length-scales comparable to their spheres of influence (e.g.,
+stars (e.g., \citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014}).
+This is likely a generic outcome of star formation: For example, turbulent
+fragmentation in collapsing protostellar clouds can produce stellar multiplets
+within length-scales comparable to their spheres of influence (e.g.,
 \citealt{Raskutti:2016}).
 Binary and multiple star systems are therefore of great interest in
-astrophysics: the population of stars and their companions encodes information
-about star formation processes (\citealt{todo}), stellar parameters and
-evolution (\citealt{todo}), and the dynamics of multi-body systems
-(\citealt{todo}).
+astrophysics: The population of stars and their companions encodes information
+about star formation processes, stellar parameters and evolution, and the
+dynamics of multi-body systems (for recent reviews, see
+\citealt{Duchene:2013,Moe:2017}).
 

### 2018-02-07 scripts/exp2/exp2.py
exp2

@@ -1,0 +1,367 @@
+import os
+from os import path
+import sys
+import time
+
+# Third-party
+from astropy.io import fits
+from astropy.table import Table, join
+from astropy.time import Time
+import astropy.units as u
+import h5py
+import matplotlib as mpl
+import matplotlib.pyplot as plt
+import numpy as np
+import schwimmbad
+
+from thejoker import RVData, JokerParams, TheJoker, JokerSamples
+from thejoker.plot import plot_rv_curves
+from thejoker.log import log as logger
+from thejoker.utils import quantity_to_hdf5
+from twoface.sample_prior import make_prior_cache
+
+# Default root data path containing APOGEE_DR13/14 paths
+APOGEE_DATA_PATH = path.expanduser('~/data/')
+# CACHE_PATH = path.abspath(path.join(path.dirname(path.abspath(__file__)),
+#                           '..', '..', 'cache'))
+CACHE_PATH = path.join(path.dirname(path.abspath(__file__)), 'cache')
+PLOT_PATH = path.join(path.dirname(path.abspath(__file__)), 'plots')
+for P in [CACHE_PATH, PLOT_PATH]:
+    os.makedirs(P, exist_ok=True)
+
+# Columns to read in from the allStar and allVisit files:
+star_columns = ['APOGEE_ID', 'NVISITS',
+                'TEFF', 'TEFF_ERR', 'LOGG', 'LOGG_ERR', 'M_H', 'M_H_ERR']
+visit_columns = ['VISIT_ID', 'APOGEE_ID', 'MJD', 'JD', 'VREL', 'VRELERR',
+                 'VHELIO', 'SNR', 'CHISQ']
+
+def read_table(filename, columns):
+    """
+    Read the specified columns from the specified FITS file and return an
+    Astropy Table object.
+    """
+    tbl = fits.getdata(filename)
+    return Table(tbl.view(tbl.dtype, np.ndarray)[columns])
+
+def load_apogee_dr13_dr14_rg(data_path):
+    if data_path is None:
+        data_path = APOGEE_DATA_PATH
+
+    else:
+        data_path = path.expanduser(data_path)
+
+    allstar_dr13 = read_table(path.join(data_path, 'APOGEE_DR13',
+                                        'allStar-l30e.2.fits'),
+                              star_columns)
+    allvisit_dr13 = read_table(path.join(data_path, 'APOGEE_DR13',
+                                         'allVisit-l30e.2.fits'),
[…]

### 2018-02-07 scripts/exp2/run.sh

@@ -1,0 +1,21 @@
+#!/bin/bash
+#SBATCH -J exp2          # job name
+#SBATCH -o exp2.o%j             # output file name (%j expands to jobID)
+#SBATCH -e exp2.e%j             # error file name (%j expands to jobID)
+#SBATCH -n 224
+#SBATCH -t 12:00:00             # run time (hh:mm:ss) - 1.5 hours
+#SBATCH --mail-user=adrn@princeton.edu
+#SBATCH --mail-type=begin       # email me when the job starts
+#SBATCH --mail-type=end         # email me when the job finishes
+
+cd /tigress/adrianp/projects/twoface/scripts/exp2/
+
+module load openmpi/gcc/1.10.2/64
+
+source activate twoface
+
+date
+
+srun python exp2.py -v --mpi
+
+date

### 2018-02-08 twoface/db/tests/test_database.py
fix test and hide warnings

@@ -37,7 +37,8 @@ class TestDB(object):
         s = self.session
 
         # a target in both test FITS files included in the repo
-        test_target_ID = "4264.2M00000032+5737103"
+        star = s.query(AllStar).filter(AllStar.apogee_id == "2M00000068+5710233").one()
+        test_target_ID = star.target_id
 
         # get star entry and check total num of visits
         star = s.query(AllStar).filter(AllStar.target_id == test_target_ID).one()
@@ -45,7 +46,7 @@ class TestDB(object):
 
         # get a visit and check that it has one star
         visit = s.query(AllVisit).filter(AllVisit.target_id == test_target_ID).limit(1).one()
-        assert len(visit.stars) == 2
+        assert len(visit.stars) == 1
 
     def test_jokerrun_cascade(self):
         """

### 2018-02-20 paper/1-catalog/twoface-catalog.tex
Some mucking about in the intro, and change jitter to excess variance.

@@
 
-\author{Hans-Walter~Rix}
+\author[0000-0003-4996-9069]{Hans-Walter~Rix}
 \affiliation{Max-Planck-Institut f\"ur Astronomie,
@@
 
-% \author{Jason~Cao}
-% \affiliation{Center for Cosmology and Particle Physics,
-%              Department of Physics,
-%              New York University, 726 Broadway,
-%              New York, NY 10003, USA}
+\author{APOGEE Team}
 
@@
 % Context
-Repeat radial-velocity measurements of stars can be used to identify stellar-,
+Repeat radial-velocity measurements of stars can be used to identify stellar,
 sub-stellar, and planetary-mass companions.
-Even few observations are useful for detecting companions, though such data are
-difficult to use for constraining physical properties because permitted orbital
-solutions are highly degenerate.
+Even sparse or few-epoch observations are useful for detecting companions,
+though such data are difficult to use for measuring physical properties of
+individual systems because permitted orbital solutions can be highly degenerate.
 % Aims
-Here we perform a search for secondary companions of stars observed by the
-\apogee\ survey (\DR), which has measured multi-epoch ($\geq 3$) but sparse
-radial velocities for close to 100,000 stars.
+Using a custom-built Monte Carlo sampler (\thejoker) to deliver (often highly
+multi-modal) posterior beliefs about companion orbital parameters, here we
+perform a search for secondary companions of over 100,000 stars observed by the
+\apogee\ survey (\DR) with $\geq 3$ epochs.
 % Methods
-We use a custom-built Monte Carlo sampler (\thejoker) to deliver (often highly
-multi-modal) posterior beliefs about companion orbital parameters.
-In some cases, where the primary-star masses are known, the mass-function
-degeneracy is broken and the secondary-companion masses ($m_2\,\sin i$) can be
-inferred.
+We select stars with probable companions by \todo{how?}.
 % Results
-We provide a catalog of \ncompanions\ companions for which the companion
-properties can be confidently determined, along with posterior samplings for all
-\ntotal\ stars in the parent sample.
-\todo{We find differences between the companions around \RC\ and \RGB\
-stars that are in/consistent with the theoretical prediction that
-stellar companions should not survive the common-envelope \RGB\ phase.}
-\todo{We additionally find trends in XX with chemical abundances.}
-
-We compare the companion populations of red clump (\RC) versus \RGB\ stars at
-similar surface gravity and temperature, and study the metallicity dependence
-of companion properties.
-
+We provide (1) a catalog of \ncompanions\ companions for which the companion
+properties can be confidently determined, (2) a catalog of stars that likely
+have companions, but for  which more observations are needed to uniquely
+determine the orbital properties, and (3) posterior samplings for all \ntotal\
+stars in the parent sample.
+We highlight interesting systems and show the characteristics of stars with
+confidently determined companion properties.
 \end{abstract}
@@
 
-Stars typically have companions. Main sequence stars in the solar neighborhood
-more often appear in binary or multiple star systems rather than as solitary
-stars (e.g., \citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014}).
+Stars typically have companions.
+Main sequence stars in the solar neighborhood more often appear in binary or
+multiple star systems rather than as solitary stars (e.g.,
+\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014}).
 This is likely a generic outcome of star formation: For example, turbulent
@@
 
-% Even without stellar or sub-stellar companions, the vast majority of stars have
-% planetary-mass companions (\citealt{todo}).
-% \todo{Planetary systems are interesting because...blah}
-% Differences between the population of stellar companions and that of
-% planetary-mass companions likely relate directly to the different formation
-% channels for these objects.
-
 Most of what is known about stellar companions comes from studies of
-main-sequence (MS) stars.
+nearby main-sequence (MS) stars.
 MS stars with companions have a large dynamic range of constituent and orbital
 characteristics.
-For example, binary stars have mass-ratios that span from $\approx 10^-3$
-\todo{citation because this is made up} to 1, and have periods from days to
-millions of years.
-\todo{Describe distributions over these quantities}
-\todo{multiplicity in general: abundance, numbers, etc.}
-\todo{some other stats from Moe and Di Stefano}
+For example, binary stars have mass-ratios that span from $\approx 0.03$
+to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
+millions of years (e.g., \citealt{Raghavan:2010}).
+\todo{multiplicity in general: occurrence rates}
 
-Some evolved stars have special roles in well-studied exotic multiple systems
-like dwarf novae or cataclysmic variables \todo{are those examples correct?},
-but by-in-large, less is known about the population of companions to evolved
-stars.
-
-- Expect differences in evolved star pop. from main sequence populations:
-  - mass loss
-  - envelope -> engulfment?
-- Current samples limited in dynamical range of chemical properties, location,
-  age (local to sun)
-- Brighter, but typically larger distances, direct imaging not feasible
-- Eclipsing companions good, but low mass companions stars -> hard to find
-  (giants are big)
-- Need large sample of stars with time-domain spectroscopy
-
-% For planetary companions of main sequence stars, \todo{}.
-% - Some indication of trends (e.g., with metallicity) -> star and
-%   planet formation processes
+Less is known about uninteracting or detached companions to evolved stars.
+Such systems are interesting because they enable studying extreme outcomes or
+conditions of star formation and multiplicity.
+For example, evolved stars are typically much brighter than their MS
+counterparts, which enables characterizing the stellar companion population
+throughout the Galactic disk over a wide range of chemical abundances.
+They also provide a means to study the outcomes of stellar engulfment or
+common-envelope evolution: Do short-period companions exist around evolved
+stars, and do their statistics constrain stellar evolution on the red giant
+branch (RGB)?
+\todo{motivate time-domain spectroscopy}
 
@@
 \citealt{Troup:2016}), or have used simple statistics computed from the data to
-study multiplicity (e.g., $\textrm{RV}_\textrm{max} - \textrm{RV}_\textrm{min}$;
-\citealt{Badenes:2017}).
+study multiplicity (e.g., $\textrm{RV}_\textrm{max}$; \citealt{Badenes:2017}).
 
@@
 Here we use \thejoker\ to generate posterior pdf samples for stars observed by
-the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2015}).
+the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2017}).
 
@@
 All data used in this work comes from the publicly-available data release 14
-(\DR) of the \apogee\ survey (\citealt{apogeedr14}), a component of the Sloan
-Digital Sky Survey IV (\sdssiv; \citealt{sdss,Blanton:2017}).
+(\DR) of the \apogee\ survey (\citealt{Majewski:2017,Abolfathi:2017}), a
+component of the Sloan Digital Sky Survey IV (\sdssiv;
+\citealt{Gunn:2006,Blanton:2017}).
 \apogee\ is designed to map stars across much of the Milky Way by obtaining
-high-resolution ($R \sim 22500$) infrared ($H$-band) spectroscopy of primarily
+high-resolution ($R \sim 22,500$) infrared ($H$-band) spectroscopy of primarily
 \RGB\ stars.
@@
 the stellar distribution.
-Data taken as part of the \apogee\ survey is reduced with the \apogee\ Stellar
-Parameters and Chemical Abundances Pipeline (\acronym{ASPCAP};
-\citealt{Garcia-Perez:2016}), which measures stellar parameters, chemical
-abundances, and radial velocities for each target.
 In order to meet signal-to-noise ratio requirements, most \apogee\ stars are
@@
 visit separated by a month or more in order to help identify binary stars.
-\acronym{ASPCAP} iteratively solves for the relative velocities of each visit
-spectrum while determining the optimal co-addition of the visit spectra.
+
+Data taken as part of the \apogee\ survey is reduced with a multi-step data
+reduction pipeline that ultimately solves for the stellar parameters, chemical
+abundances, and radial velocities for each target
+(\citealt{Nidever:2015}).
+Most relevant for this work, the visit radial velocities (RVs) are determined
+using an iterative scheme: the individual visit spectra are combined using
+initial guesses for the relative RVs into a coadded spectrum, which is then used
+to re-derive the relative visit velocities.
+The stellar parameters---surface gravity, $\log g$, and effective temperature,
+$T_{\textrm{eff}}$---and the chemical abundances are determined from the coadded
+spectrum as a part of the \apogee\ Stellar Parameters and Chemical Abundances
+Pipeline (\acronym{ASPCAP}; \citealt{ASPCAP}).
 
@@
 \texttt{PERSIST\_JUMP\_NEG}, \texttt{VERY\_BRIGHT\_NEIGHBOR}, \texttt{LOW\_SNR}.
-For each star, we require that $0 < \log g < 3.5$ and the following bits are not
+For each star, we require that $0 < \log g < 4$ and the following bits are not
 set in the \texttt{ASPCAPFLAGS} bitmask: \texttt{STAR\_BAD}.
@@
 \DR\ have $< 8$ visits.
+\figurename~\ref{fig:loggteff} ...
 
-% Notebook: "Numbers after cuts"
+% Notebook: "Numbers of visits and sample overview"
 \begin{figure}[h]
@@
 \caption{%
-\todo{Update with new apogee db file}
 Number of \apogee\ stars in logarithmic bins of number of visits that pass the
@@
 \label{fig:nvisits}
+}
+\end{figure}
+
+% Notebook: "Numbers of visits and sample overview"
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{logg_teff_feh.pdf}
+\end{center}
+\caption{%
+\textit{Left panel}: Number of stars in bins of iron abundance,
+$[\textrm{Fe}/\textrm{H}]$, that pass the quality cuts described in
+\sectionname~\ref{sec:data}.
+\textit{Right panel}: Distribution of stars in our sample in stellar parameters
+log-surface-gravity, $\log g$, and effective temperature, $T_{\textrm{eff}}$,
+with points colored by the iron abundance.
+\label{fig:loggteff}
 }
@@
   substantially more luminous than their typical companion.
-  However, there are known double-lined binary stars in the \apogee\ catalog
-  \citealt{El-Badry:2017}, and an expected but unknown fraction of \RGB--\RGB\
-  binaries.
+  However, there are known main-sequence double-lined binary stars in the
+  \apogee\ catalog \citealt{El-Badry:2018}, and an expected but unknown fraction
+  of \RGB--\RGB\ binaries.
 \item[simple noise model] Measurements are unbiased and noise estimates are
-  correct up to an unknown extra variance.
+  correct up to an unknown excess variance.
   All noise contributions result in Gaussian uncertainties on the individual
   radial-velocity measurements.
-  We often refer to the extra variance as a ``jitter'' parameter because for
-  upper \RGB\ stars the inferred extra variance will be a combination of extra
-  uncertainty and true astrophysical surface jitter (\todo{CITATION NEEDED}).
+  For upper \RGB\ stars the inferred excess variance will be a combination of
+  extra systematic uncertainty and true astrophysical surface jitter
+  (\todo{CITATION NEEDED}).
 \end{description}
@@
 
-\subsection{Model selection}
-
-In addition to generating posterior samples for each source, we also compute the
-fully marginalized likelihood (FML) under two other models: (1) a model in which
-the radial velocity of the source is constant, and (2) a model in which the
-radial-velocity variations are purely linear.
-We later use these FML values to define cuts on our full catalog of posterior
-samples to produce a catalog of companions.
-
-\todo{equations?}
-
-Examples of outputs and etc.
+% \subsection{Model selection}
+%
+% In addition to generating posterior samples for each source, we also compute the
+% fully marginalized likelihood (FML) under two other models: (1) a model in which
+% the radial velocity of the source is constant, and (2) a model in which the
+% radial-velocity variations are purely linear.
+% We later use these FML values to define cuts on our full catalog of posterior
+% samples to produce a catalog of companions.
+%
+% \todo{equations?}
+%
+% Examples of outputs and etc.
 
@@
 We start by generating a library of \nprior\ prior samples generated under a
-prior similar to that defined in \citet{Price-Whelan:2017}: uniform or isotropic
-in angle parameters, uniform in log-period over the domain $[1,
-32768]~\textrm{day}$, and a beta distribution over eccentricity with fixed
-parameters (\citealt{Kipping:2013}).
-For the prior over the extra variance or jitter parameter, $s$, we initially use
+prior similar to that defined in \citet{Price-Whelan:2017}:
+\begin{itemize}
+    \item uniform or isotropic in angle parameters,
+    \item uniform in log-period over the domain $[1,32768]~\textrm{day}$,
+    \item a beta distribution over eccentricity with fixed parameters (\citealt{Kipping:2013}).
+\end{itemize}
+For the prior over the excess variance parameter, $s$, we initially use
 a Gaussian over the transformed parameter $y = \ln s^2$ with the mean and
@@
 
-\subsection{Inferring the extra variance distribution}
+\subsection{Inferring the excess variance distribution}
 \label{sec:inferjitter}
@@
 Bayesian model to infer the parameters of the (assumed Gaussian) distribution
-over the log-jitter parameter: $(\mu_y, \sigma_y)$.
+over the log-excess-variance parameter: $(\mu_y, \sigma_y)$.
 This inference serves as a test-case for future work, where we intend to use the
@@
 data products: If the catalog uncertainties are significantly underestimated, we
-expect the inferred log-jitter distribution parameters to tend towards larger
-values.
+expect the inferred log-excess-variance distribution parameters to tend towards
+larger values.
 
@@
 primary orbital parameters $\bs{\theta} = (P, e, \omega, M_0, K, v_0)$ and the
-jitter parameter, $y = \ln s^2$, using \thejoker; For brevity in expressions
-below, we will use the vector
+excess-variance parameter, $y = \ln s^2$, using \thejoker; For brevity in
+expressions below, we will use the vector
 \begin{equation}
@@
 to represent the full set of parameters.
-To obtain this sampling, we use an interim (Gaussian) prior on the jitter
-parameter parametrized by a mean and standard deviation, i.e. $\alpha_0 =
-(\mu_{y,0}, \sigma_{y,0})$ as described above.
+To obtain this sampling, we use an interim (Gaussian) prior on the
+excess-variance parameter parametrized by a mean and standard deviation, i.e.
+$\alpha_0 = (\mu_{y,0}, \sigma_{y,0})$ as described above.
 For a given source, the posterior samples in the above parameters are drawn from
@@
 be simplified to a sum over prior value ratios of the $K$ posterior samples in
-the log-jitter parameter for each $n$ star
+the log-excess-variance parameter for each $n$ star
 \begin{equation}
@@
 that have large intrinsic jitter) and maximize the above likelihood to determine
-better hyperparameters for the log-jitter parameter distribution.
+better hyperparameters for the log-excess-variance parameter distribution.
 \figurename~\ref{fig:infer-jitter} shows the distribution corresponding to the

### 2018-02-20 scripts/run_apogee_continue_mcmc.py
don't give pool to joker

@@ -86,7 +86,7 @@ def main(config_file, pool, seed, overwrite=False):
     rnd = np.random.RandomState(seed=seed)
     logger.debug("Creating TheJoker instance with {0}, {1}".format(rnd, pool))
     params = run.get_joker_params()
-    joker = TheJoker(params, random_state=rnd, pool=pool)
+    joker = TheJoker(params, random_state=rnd)
 
     # Get all stars in this JokerRun that "need mcmc"
     star_query = session.query(AllStar).join(StarResult, JokerRun, Status)\

### 2018-02-20 scripts/run_apogee_continue_mcmc.py
don't redo done ones

@@ -34,6 +34,12 @@ def emcee_worker(task):
     cache_path, results_filename, apogee_id, data, joker = task
     n_walkers = 1024
 
+    chain_path = path.join(cache_path, '{0}.npy'.format(apogee_id))
+
+    if path.exists(chain_path):
+        logger.debug('{0} already done'.format(apogee_id))
+        return
+
     with h5py.File(results_filename, 'r') as f:
         samples0 = JokerSamples.from_hdf5(f[apogee_id])
 
@@ -43,7 +49,7 @@ def emcee_worker(task):
                                                 n_walkers=n_walkers,
                                                 return_sampler=True)
 
-    np.save(path.join(cache_path, '{0}.npy'.format(apogee_id)), sampler.chain)
+    np.save(chain_path, sampler.chain)
 
     model_path = path.join(cache_path, 'model.pickle')
     if not path.exists(model_path):
@@ -103,7 +109,7 @@ def main(config_file, pool, seed, overwrite=False):
 
     tasks = [(cache_path, results_filename, star.apogee_id,
               star.apogeervdata(), joker)
-             for star in star_query.all()][:1]
+             for star in star_query.all()]
     session.close()
 
     for r in pool.map(emcee_worker, tasks):

### 2018-02-20 paper/1-catalog/twoface-catalog.tex
emcee sec and refs

@@
   radial-velocity measurements.
-  For upper \RGB\ stars the inferred excess variance will be a combination of
-  extra systematic uncertainty and true astrophysical surface jitter
-  (\todo{CITATION NEEDED}).
 \end{description}
@@
 \begin{equation}
-  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, M_0)\right) +
+  v(t;\bs{\theta}) = v_0 + K\,[\cos\left(\omega + f(t; e, P, M_0, t_0)\right) +
     e\,\cos\omega]
@@
 \citealt{Price-Whelan:2017} or \eqname~63 in \citealt{Murray:2010}).
+In addition to the orbital parameters listed above, \thejoker\ can also generate
+samples in an ``excess variance'' parameter, $s^2$, that is added to the
+per-visit measurement variances.
+This parameter allows us to test whether the visit RV uncertainties are
+underestimated
+For upper \RGB\ stars the inferred excess variance will be a combination of
+extra systematic uncertainty and true astrophysical surface jitter
+(\todo{CITATION NEEDED}).
 
@@
 
-\section{Results}
-
-\subsection{Initial inidivual-system posterior samplings}
+\subsection{Initial individual-system posterior samplings}
 \label{sec:initsampling}
@@
 \end{itemize}
-For the prior over the excess variance parameter, $s$, we initially use
+For the prior over the excess variance parameter, $s^2$, we initially use
 a Gaussian over the transformed parameter $y = \ln s^2$ with the mean and
@@
 = (10.6, 3)$ in units of $\textrm{m}~\textrm{s}^{-1}$.
+The reference time for each star is set to the minimum visit epoch; $M_0$ then
+becomes the mean anomaly at the first visit observation.
+See \tablename~\ref{tbl:params} for a description of all parameters and priors
+used.
 
-We request \nposterior\ posterior samples for each source, and generate the
-samples using an iterative procedure that adapts to the \todo{informativeness}
-of the data to more efficiently generate posterior samples.
+\begin{table}[h!]
+    \centering
+    \begin{tabular}{ r c l }
+    \hline
+    name & prior & description \\
+    \hline
+    $P$ & $\ln P \sim \mathcal{U}(1, 32768)~\textrm{day}$ & period \\
+    $e$ & $e \sim \textrm{Beta}(0.867, 3.03)$ & eccentricity \\
+    $t_0$ & fixed & reference time \\
+    $M_0$ & $M_0 \sim \mathcal{U}(0, 2\pi)~\textrm{rad}$ & mean anomaly at reference time \\
+    $\omega$ & $\omega \sim \mathcal{U}(0, 2\pi)~\textrm{rad}$ & argument of pericenter \\
+    $s^2$ & $\ln s^2 \sim \mathcal{N}(\mu_y, \sigma_y^2)$ & extra variance added to each visit variance \\
+    $K$ & TODO & velocity semi-amplitude \\
+    $v_0$ & TODO & system barycentric velocity \\
+    \hline
+    \end{tabular}
+    \caption{Summary and description of parameters.
+$\textrm{Beta}(a, b)$ is the beta distribution with \todo{a, b}, $\mathcal{U}(a, b)$ the uniform distribution over the domain $(a, b)$, and $\mathcal{N}(\mu, \sigma^2)$ is the normal distribution with mean $\mu$ and variance $\sigma^2$.}
+    \label{tbl:params}
+\end{table}
+
+We request \nposterior\ posterior samples for each source.
+Depending on the data quality and phase coverage of the visits, \thejoker\ will
+require different numbers of prior samples in order to rejection sample down to
+the requested number of posterior samples: For few-epoch or noisy RV data, many
+prior samples will pass the rejection step, whereas for very precise or
+many-epoch RV data, \thejoker\ may need to go process the full library of prior
+samples.
+We therefore generate the posterior samples using an iterative procedure that
+adaptively predicts how many prior samples to test for each star.
 For sources with very constraining data, \thejoker\ may return fewer than the
-requested number of samples.
+requested number of samples (as few as one sample).
 When this happens, we proceed in one of two ways:
@@
     \nposterior) are returned that all lie within a single mode of the posterior
-    pdf, the orbit-space that is consistent with the data is treated as
-    effectively unimodal. In this case, we use the location of the returned
-    sample, or the median over returned samples to generate a small ball of
-    initial conditions and use standard MCMC to continue sampling until we obtain \nposterior\ samples. In detail, we use an ensemble sampler
-    (\citealt{Goodman:}) implemented in \python\ (\citealt{Foreman-Mackey:}),
-    run with \todo{XX} walkers for \todo{XX} steps, and downsample the resulting
-    MCMC chains until we have \nposterior\ samples.
+    pdf, the posterior pdf over orbital parameters is treated as effectively
+    unimodal.
+    In this case, we use the location of the returned sample, or the median over
+    returned samples to generate a small ball of initial conditions and use
+    standard MCMC to continue sampling until we obtain \nposterior\ samples.
+    In detail, we use an ensemble sampler (\citealt{Goodman:2010}) implemented
+    in \python\ (\citealt{Foreman-Mackey:2013}) for this step.
+    We run with 1024 walkers for 16384 steps, take the final walker positions,
+    and downsample until we have \nposterior\ samples.
+    We compute convergence statistics for each run and report whether the chains
+    appear converged or not.
     \item \emph{needs more prior samples}: If more than 1 posterior sample is
@@
 dominated by sources with many ($\gtrsim 10$) visits.
+\todo{XX} stars were flagged as \emph{needs MCMC}, \todo{YY} were flagged as
+\emph{needs more prior samples}, and the remaining \todo{ZZ} stars completed normally.
+
+\section{Results}
 

### 2018-02-20 paper/1-catalog/twoface-catalog.tex
more refs and text

@@
 
+\subsection{Implementation notes}
+
+\todo{Nontrivial computation - details about running on cluster, TheJoker vs. emcee}
+
 \section{Results}
@@
 As an initial use of the per-source posterior samplings, we use a hierarchical
-Bayesian model to infer the parameters of the (assumed Gaussian) distribution
-over the log-excess-variance parameter: $(\mu_y, \sigma_y)$.
+Bayesian model to infer the parameters of the (assumed Gaussian) prior over the
+log-excess-variance parameter, $(\mu_y, \sigma_y)$ (see
+\tablename~\ref{tbl:params}).
 This inference serves as a test-case for future work, where we intend to use the
 independent posterior samplings to construct a hierarchical inference over
-companion property distributions.
+companion population properties.
 This is also a test of the visit velocity uncertainties reported in the \apogee\
@@
 marginalize over the per-object parameters
-(\citealt{Hogg:2008,Foreman-Mackey:...}); we describe how to compute the
+(\citealt{Hogg:2010,Foreman-Mackey:2014}); we describe how to compute the
 marginal likelihood in detail below.
@@
 1.64)$.
-\todo{discussion about this}
-\todo{some discussion of how this relates to empirical bayes?}
+\todo{Reactions to the inferred parameters?}
+\todo{Discuss how this relates to empirical Bayes?}
 

### 2018-02-21 scripts/run_apogee_continue_mcmc.py
Add plotting functionality to mcmc continue script.
Also added a new plot function for diagnosing MCMC.

@@ -28,6 +28,7 @@ from twoface.log import log as logger
 from twoface.db import db_connect, get_run
 from twoface.db import JokerRun, AllStar, StarResult, Status
 from twoface.config import TWOFACE_CACHE_PATH
+from twoface.plot import plot_mcmc_diagnostic
 
 
 def emcee_worker(task):
@@ -35,26 +36,35 @@ def emcee_worker(task):
     n_walkers = 1024
 
     chain_path = path.join(cache_path, '{0}.npy'.format(apogee_id))
+    plot_path = path.join(cache_path, '{0}.png'.format(apogee_id))
 
-    if path.exists(chain_path):
-        logger.debug('{0} already done'.format(apogee_id))
-        return
+    sampler = None
+    if not path.exists(chain_path):
+        logger.debug('Running MCMC for {0}'.format(apogee_id))
 
-    with h5py.File(results_filename, 'r') as f:
-        samples0 = JokerSamples.from_hdf5(f[apogee_id])
+        with h5py.File(results_filename, 'r') as f:
+            samples0 = JokerSamples.from_hdf5(f[apogee_id])
 
-    model, samples, sampler = joker.mcmc_sample(data, samples0,
-                                                n_burn=0,
-                                                n_steps=16384,
-                                                n_walkers=n_walkers,
-                                                return_sampler=True)
+        model, samples, sampler = joker.mcmc_sample(data, samples0,
+                                                    n_burn=0,
+                                                    n_steps=16384,
+                                                    n_walkers=n_walkers,
+                                                    return_sampler=True)
 
-    np.save(chain_path, sampler.chain)
+        np.save(chain_path, sampler.chain)
 
-    model_path = path.join(cache_path, 'model.pickle')
-    if not path.exists(model_path):
-        with open(model_path, 'wb') as f:
-            pickle.dump(model, f)
+        model_path = path.join(cache_path, 'model.pickle')
+        if not path.exists(model_path):
+            with open(model_path, 'wb') as f:
+                pickle.dump(model, f)
+
+    if not path.exists(plot_path):
+        if sampler is None:
+            chain = np.load(chain_path)
+            
+        fig = plot_mcmc_diagnostic(chain)
+        fig.savefig(plot_path, dpi=250)
+        del fig
 
 
 def main(config_file, pool, seed, overwrite=False):

### 2018-02-21 twoface/plot.py

@@ -1,6 +1,7 @@
 # Third-party
 from astropy.time import Time
 import astropy.units as u
+from astropy.stats import median_absolute_deviation
 import matplotlib.pyplot as plt
 import numpy as np
 
@@ -93,4 +94,37 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
     if title is not None:
         ax.set_title(title)
 
+    return fig
+
+
+def plot_mcmc_diagnostic(chain):
+    """
+    TODO:
+    """
+
+    names = [r'$\ln P$', r'$\sqrt{K}\,\cos M_0$', r'$\sqrt{K}\,\sin M_0$',
+             r'$\sqrt{e}\,\cos \omega$', r'$\sqrt{e}\,\sin \omega$',
+             r'$\ln s^2$', '$v_0$']
+
+    ndim = chain.shape[-1]
+    assert ndim == len(names)
+
+    fig, axes = plt.subplots(ndim, 3, figsize=(12, 16), sharex=True)
+
+    for k in range(ndim):
+        axes[k, 0].set_ylabel(names[k])
+        axes[k, 0].plot(chain[..., k].T, marker='',
+                        drawstyle='steps-mid',
+                        alpha=0.1, rasterized=True)
+        axes[k, 1].plot(np.median(chain[..., k], axis=0),
+                        marker='', drawstyle='steps-mid')
+
+        std = 1.5 * median_absolute_deviation(chain[..., k], axis=0)
+        axes[k, 2].plot(std, marker='', drawstyle='steps-mid')
+
+    axes[0, 0].set_title('walkers')
+    axes[0, 1].set_title('med(walkers)')
+    axes[0, 2].set_title('1.5 MAD(walkers)')
+
+    fig.tight_layout()
     return fig

### 2018-02-21 scripts/initdb.py
add Ness RG table back

@@ -3,18 +3,17 @@ import sys
 
 # Project
 from twoface.log import log as logger
-from twoface.db.init import initialize_db, load_red_clump, load_nessrg
+from twoface.db.init import initialize_db, load_nessrg
+
 
 def main(allVisit_file, allStar_file, rc_file, nessrg_file, **_):
     database_name = 'apogee.sqlite' # TODO: should this be enforced?
 
     initialize_db(allVisit_file, allStar_file, database_name)
 
-    if rc_file is not None:
-        load_red_clump(rc_file, database_name, overwrite=False) # HACK
-
     if nessrg_file is not None:
         load_nessrg(nessrg_file, database_name, overwrite=False) # HACK
+
 
 if __name__ == "__main__":
     from argparse import ArgumentParser

### 2018-02-21 twoface/db/init.py

@@ -12,7 +12,7 @@ from ..config import TWOFACE_CACHE_PATH
 from ..util import Timer
 from ..log import log as logger
 from .connect import db_connect, Base
-from .model import (AllStar, AllVisit, AllVisitToAllStar, Status, RedClump,
+from .model import (AllStar, AllVisit, AllVisitToAllStar, Status, NessRG,
                     StarResult)
 from .query_helpers import paged_query
 
@@ -64,11 +64,12 @@ def initialize_db(allVisit_file, allStar_file, database_file,
     allstar_tbl = fits.getdata(norm(allStar_file))
 
     # Remove bad velocities and flagged bad visits:
-    skip_mask = np.sum(2 ** np.array([9, 12, 13, # PERSIST_HIGH, PERSIST_JUMP_POS, PERSIST_JUMP_NEG
+    # PERSIST_HIGH, PERSIST_JUMP_POS, PERSIST_JUMP_NEG
+    skip_mask = np.sum(2 ** np.array([9, 12, 13,
                                       3, 4])) # VERY_BRIGHT_NEIGHBOR, LOW_SNR
     allvisit_tbl = allvisit_tbl[np.isfinite(allvisit_tbl['VHELIO']) &
                                 np.isfinite(allvisit_tbl['VRELERR']) &
-                                (allvisit_tbl['VRELERR'] < 100.) & # MAGIC NUMBER
+                                (allvisit_tbl['VRELERR'] < 100.) & # MAGIC
                                 ((allvisit_tbl['STARFLAG'] & skip_mask) == 0)]
 
     v_apogee_ids, counts = np.unique(allvisit_tbl['APOGEE_ID'],
@@ -234,7 +235,7 @@ def initialize_db(allVisit_file, allStar_file, database_file,
 
     q = session.query(AllStar).order_by(AllStar.id)
 
-    for i,sub_q in enumerate(paged_query(q, page_size=batch_size)):
+    for i, sub_q in enumerate(paged_query(q, page_size=batch_size)):
         for star in sub_q:
             if len(star.visits) > 0:
                 continue
@@ -416,7 +417,7 @@ def load_nessrg(filename, database_file, overwrite=False, batch_size=4096):
             # Retrieve the parent AllStar record
             try:
                 star = session.query(AllStar).filter(
-                    AllStar.apogee_id == row['2MASS']).one()
+                    AllStar.apogee_id == row['2MASS']).limit(1).one()
[…]

### 2018-02-21 twoface/db/model.py

@@ -20,8 +20,8 @@ from .quantity_type import QuantityTypeClassFactory
 from . import numpy_adapt # just need to execute code
 
 
-__all__ = ['AllStar', 'AllVisit', 'RedClump', 'JokerRun', 'StarResult',
-           'Status', 'AllVisitToAllStar']
+__all__ = ['AllStar', 'AllVisit', 'JokerRun', 'StarResult',
+           'Status', 'AllVisitToAllStar', 'NessRG']
 
 
 class Status(Base):
@@ -508,204 +508,46 @@ class JokerRun(Base):
                            anomaly_tol=anomaly_tol, **jitter_kwargs)
 
 
-# OLD
-
-class RedClump(Base):
-    __tablename__ = 'redclump'
+class NessRG(Base):
+    __tablename__ = 'nessrg'
 
     id = Column(types.Integer, primary_key=True)
 
-    allstar_id = Column("allstar_id", types.Integer,
+    allstar_id = Column('allstar_id', types.Integer,
                         ForeignKey('allstar.id', ondelete='CASCADE'),
                         index=True)
-    star = relationship("AllStar",
-                        backref=backref("red_clump",
-                                        cascade="all, delete-orphan"))
+    star = relationship("AllStar", backref=backref("ness_rg", uselist=False))
 
-    # These are in AllStar
-    # apstar_id = Column("apstar_id", types.String)
-    # aspcap_id = Column("aspcap_id", types.String)
-    # apogee_id = Column("apogee_id", types.String)
-    # telescope = Column("telescope", types.String)
-    # location_id = Column("location_id", types.SmallInteger)
-    # field = Column("field", types.String)
-    # j = Column("j", types.REAL)
-    # j_err = Column("j_err", types.REAL)
-    # h = Column("h", types.REAL)
-    # h_err = Column("h_err", types.REAL)
-    # k = Column("k", types.REAL)
-    # k_err = Column("k_err", types.REAL)
-    # ra = Column("ra", types.Numeric)
-    # dec = Column("dec", types.Numeric)
-    # glon = Column("glon", types.Numeric)
-    # glat = Column("glat", types.Numeric)
[…]

### 2018-02-21 twoface/db/model.py
oops forgot exp

@@ -547,7 +547,9 @@ class NessRG(Base):
         return np.exp(self.lnAge) * u.Gyr
 
     def get_mass_samples(self, size=1):
-        return np.random.normal(self.lnM, self.e_logM, size=size) * u.Msun
+        return np.exp(np.random.normal(self.lnM, self.e_logM,
+                                       size=size)) * u.Msun
 
     def get_age_samples(self, size=1):
-        return np.random.normal(self.lnAge, self.e_logAge, size=size) * u.Gyr
+        return np.exp(np.random.normal(self.lnAge, self.e_logAge,
+                                       size=size)) * u.Gyr

### 2018-02-22 twoface/db/init.py
fix rc tests

@@ -12,11 +12,11 @@ from ..config import TWOFACE_CACHE_PATH
 from ..util import Timer
 from ..log import log as logger
 from .connect import db_connect, Base
-from .model import (AllStar, AllVisit, AllVisitToAllStar, Status, NessRG,
-                    StarResult)
+from .model import AllStar, AllVisit, Status, NessRG
 from .query_helpers import paged_query
 
-__all__ = ['initialize_db', 'load_red_clump', 'load_nessrg']
+__all__ = ['initialize_db', 'load_nessrg']
+
 
 def tblrow_to_dbrow(tblrow, colnames, varchar_cols=[]):
     row_data = dict()
@@ -258,108 +258,6 @@ def initialize_db(allVisit_file, allStar_file, database_file,
     session.commit()
     session.close()
 
-def load_red_clump(filename, database_file, overwrite=False, batch_size=4096):
-    """Load the red clump catalog stars into the database.
-
-    Information about this catalog can be found here:
-    http://www.sdss.org/dr13/data_access/vac/
-
-    Parameters
-    ----------
-    filename : str
-        Full path to APOGEE red clump FITS file.
-    database_file : str
-        Filename (not path) of database file in cache path.
-    overwrite : bool (optional)
-        Overwrite any data already loaded into the database.
-    batch_size : int (optional)
-        How many rows to create before committing.
-    """
-
-    database_path = join(TWOFACE_CACHE_PATH, database_file)
-
-    norm = lambda x: abspath(expanduser(x))
-    rc_tbl = Table.read(norm(filename), format='fits', hdu=1)
-
-    Session, engine = db_connect(database_path)
-    logger.debug("Connected to database at '{}'".format(database_path))
-    session = Session()
-
-    # What columns do we load?
-    rc_skip = ['ID', 'ALLSTAR_ID']
-    rc_colnames = []
-    rc_varchar = []
-    for x in RedClump.__table__.columns:
-        col = str(x).split('.')[1].upper()
-        if col in rc_skip:
-            continue
-
-        if str(x.type) == 'VARCHAR':
-            rc_varchar.append(col)
-
-        rc_colnames.append(col)
-
-    # What APOGEE IDs are already loaded as RC stars?
-    rc_ap_ids = session.query(AllStar.apstar_id).join(RedClump).all()
-    rc_ap_ids = [x[0] for x in rc_ap_ids]
-
-    rcstars = []
-    with Timer() as t:
[…]

### 2018-02-22 twoface/db/tests/test_database.py

@@ -10,7 +10,8 @@ import yaml
 from ...config import TWOFACE_CACHE_PATH
 from ..connect import db_connect
 from ..model import AllStar, AllVisit, StarResult, Status, JokerRun
-from ..init import initialize_db, load_red_clump
+from ..init import initialize_db
+
 
 class TestDB(object):
 
@@ -28,8 +29,6 @@ class TestDB(object):
                       allStar_file=get_pkg_data_filename('test-allStar.fits'),
                       database_file=self.db_path,
                       drop_all=True)
-
-        load_red_clump(get_pkg_data_filename('test-rc.fits'), self.db_path)
 
         self.session = Session()
 

### 2018-02-22 twoface/db/model.py
fast apogee id retrieval

@@ -360,6 +360,10 @@ class AllStar(Base):
                 (((cls.fparam4+8.39) - (cls.fparam5+7.78)) > -0.6) &
                 (((cls.fparam4+8.39) - (cls.fparam5+7.78)) < 0.2))
 
+    def get_apogee_id(self, session, apogee_id):
+        return session.query(AllStar).filter(
+            AllStar.apogee_id == apogee_id).limit(1).one()
+
 
 class AllVisit(Base):
     __tablename__ = 'allvisit'

### 2018-02-22 scripts/run_apogee_continue_mcmc.py
also plot orbits

@@ -28,7 +28,7 @@ from twoface.log import log as logger
 from twoface.db import db_connect, get_run
 from twoface.db import JokerRun, AllStar, StarResult, Status
 from twoface.config import TWOFACE_CACHE_PATH
-from twoface.plot import plot_mcmc_diagnostic
+from twoface.plot import plot_mcmc_diagnostic, plot_data_orbits
 
 
 def emcee_worker(task):
@@ -37,6 +37,7 @@ def emcee_worker(task):
 
     chain_path = path.join(cache_path, '{0}.npy'.format(apogee_id))
     plot_path = path.join(cache_path, '{0}.png'.format(apogee_id))
+    orbits_plot_path = path.join(cache_path, '{0}-orbits.png'.format(apogee_id))
 
     sampler = None
     if not path.exists(chain_path):
@@ -61,9 +62,16 @@ def emcee_worker(task):
     if not path.exists(plot_path):
         if sampler is None:
             chain = np.load(chain_path)
-            
+
         fig = plot_mcmc_diagnostic(chain)
         fig.savefig(plot_path, dpi=250)
+        del fig
+
+        with open(model_path, 'rb') as f:
+            model = pickle.load(f)
+        samples = model.unpack_samples_mcmc(chain[:, -1])
+        fig = plot_data_orbits(data, samples)
+        fig.savefig(orbits_plot_path, dpi=250)
         del fig
 
 

### 2018-02-22 scripts/run_apogee_continue_mcmc.py
derp

@@ -62,6 +62,8 @@ def emcee_worker(task):
     if not path.exists(plot_path):
         if sampler is None:
             chain = np.load(chain_path)
+        else:
+            chain = sampler.chain
 
         fig = plot_mcmc_diagnostic(chain)
         fig.savefig(plot_path, dpi=250)

### 2018-02-22 scripts/run_apogee_continue_mcmc.py
model path

@@ -38,6 +38,7 @@ def emcee_worker(task):
     chain_path = path.join(cache_path, '{0}.npy'.format(apogee_id))
     plot_path = path.join(cache_path, '{0}.png'.format(apogee_id))
     orbits_plot_path = path.join(cache_path, '{0}-orbits.png'.format(apogee_id))
+    model_path = path.join(cache_path, 'model.pickle')
 
     sampler = None
     if not path.exists(chain_path):
@@ -54,7 +55,6 @@ def emcee_worker(task):
 
         np.save(chain_path, sampler.chain)
 
-        model_path = path.join(cache_path, 'model.pickle')
         if not path.exists(model_path):
             with open(model_path, 'wb') as f:
                 pickle.dump(model, f)

### 2018-02-22 scripts/run_apogee_continue_mcmc.py
set t0

@@ -16,6 +16,7 @@ import os
 import pickle
 
 # Third-party
+from astropy.time import Time
 import h5py
 import numpy as np
 from schwimmbad import choose_pool
@@ -72,6 +73,7 @@ def emcee_worker(task):
         with open(model_path, 'rb') as f:
             model = pickle.load(f)
         samples = model.unpack_samples_mcmc(chain[:, -1])
+        samples.t0 = Time(data._t0_bmjd, format='mjd', scale='tcb')
         fig = plot_data_orbits(data, samples)
         fig.savefig(orbits_plot_path, dpi=250)
         del fig

### 2018-02-22 paper/1-catalog/twoface-catalog.tex
trivialities
adjust numbers to reflect new run.
troup figures

@@
 \newcommand{\nposterior}{256}
-\newcommand{\nstars}{98,758}
-\newcommand{\nvisits}{388,196}
+\newcommand{\nstars}{96,231}
+\newcommand{\nvisits}{397,559}
 
@@
 
-\title{Binary companions of red giant stars with APOGEE: \\
-       catalog and companions that survive the common envelope}
+\title{Binary companions of red giant stars with APOGEE DR14}
 
@@
 
-Less is known about uninteracting or detached companions to evolved stars.
+Less is known about non-interacting or detached companions to evolved stars.
 Such systems are interesting because they enable studying extreme outcomes or
@@
 
-\subsection{Differences in companions of RC and RGB stars}
-\label{sec:rc-rgb}
-
-\todo{APW}
-
-\subsection{Trends in RGB companions with chemical abundances}
-\label{sec:rgb-chemistry}
-
-\todo{APW}
-
 \section{Discussion}
@@
 \emph{by name} here.
+
+\subsection{Comparison to previous work} \label{sec:compare-troup}
+
+Stuff.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{unimodal.pdf}
+\end{center}
+\caption{%
+\label{fig:troup-unimodal}
+}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{multimodal.pdf}
+\end{center}
+\caption{%
+\label{fig:troup-multimodal}
+}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{data-changed.pdf}
+\end{center}
+\caption{%
+\label{fig:troup-datachanged}
+}
+\end{figure}
 
@@
 
-\facility{\sdssiii, \apogee}
+\facility{\sdssiv, \apogee}
 

### 2018-02-22 twoface/db/model.py
add classmethod for fast star retrieval

@@ -360,7 +360,8 @@ class AllStar(Base):
                 (((cls.fparam4+8.39) - (cls.fparam5+7.78)) > -0.6) &
                 (((cls.fparam4+8.39) - (cls.fparam5+7.78)) < 0.2))
 
-    def get_apogee_id(self, session, apogee_id):
+    @classmethod
+    def get_apogee_id(cls, session, apogee_id):
         return session.query(AllStar).filter(
             AllStar.apogee_id == apogee_id).limit(1).one()
 

### 2018-02-22 twoface/plot.py
new plot functions

@@ -1,19 +1,22 @@
 # Third-party
-from astropy.time import Time
 import astropy.units as u
 from astropy.stats import median_absolute_deviation
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
 import numpy as np
 
+from thejoker import JokerSamples
 from thejoker.plot import plot_rv_curves
 from thejoker.sampler.likelihood import get_ivar
 
-__all__ = ['plot_data_orbits']
+from .samples_analysis import unimodal_P
+
+__all__ = ['plot_data_orbits', 'plot_mcmc_diagnostic', 'plot_two_panel']
 
 
 def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
-                     xlim_choice='wide', n_times=4096, title=None,
-                     ax=None, highlight_P_extrema=True):
+                     xlim_choice='wide', ylim_fac=1., n_times=4096, title=None,
+                     ax=None, highlight_P_extrema=True, plot_kwargs=None):
     """
     Plot the APOGEE RV data vs. time and orbits computed from The Joker samples.
 
@@ -48,8 +51,6 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
     else:
         fig = ax.figure
 
-    now = Time.now()
-
     if xlim_choice == 'tight': # twice the span of the data
         w = np.ptp(data.t.mjd)
         t_grid = np.linspace(data.t.mjd.min() - w*0.05,
@@ -67,9 +68,15 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
         raise ValueError('Invalid xlim_choice {0}. Can be "wide" or "tight".'
                          .format(xlim_choice))
 
+    if plot_kwargs is None:
+        plot_kwargs = dict()
+
+    plot_kwargs.setdefault('color', '#aaaaaa')
+    # plot_kwargs.setdefault('alpha', 0.25)
+    plot_kwargs.setdefault('linewidth', 0.5)
     plot_rv_curves(samples, t_grid, rv_unit=u.km/u.s, data=data, ax=ax,
                    n_plot=min(len(samples['P']), n_orbits),
-                   plot_kwargs=dict(color='#aaaaaa', alpha=0.25, linewidth=0.5),
+                   plot_kwargs=plot_kwargs,
[…]

### 2018-02-22 twoface/plot.py
fix phase-fold func

@@ -137,7 +137,7 @@ def plot_mcmc_diagnostic(chain):
     return fig
 
 
-def plot_two_panel(star, samples, axes=None, tight=True, title=None,
+def plot_two_panel(data, samples, axes=None, tight=True, title=None,
                    plot_data_orbits_kw=None, scatter_kw=None):
     """Make a two-panel plot with (1) the data and orbit samples, and (2) the
     orbit samples in period-eccentricity space.
@@ -159,8 +159,6 @@ def plot_two_panel(star, samples, axes=None, tight=True, title=None,
 
     if scatter_kw is None:
         scatter_kw = dict()
-
-    data = star.apogeervdata(clean=False)
 
     if axes is None:
         fig = plt.figure(figsize=(12, 4.3))
@@ -200,36 +198,38 @@ def plot_two_panel(star, samples, axes=None, tight=True, title=None,
     return fig
 
 
-def plot_phase_fold(star, samples):
+def plot_phase_fold(data, sample):
     """Plot the data phase-folded at the median period of the input samples,
     and the residuals as a function of phase.
 
     Parameters
     ----------
     star : `~twoface.db.AllStar`
-    samples : `~thejoker.JokerSamples`
+    sample : `~thejoker.JokerSamples`
 
     Returns
     -------
     fig : `matplotlib.Figure`
     """
-    data = star.apogeervdata(clean=False)
 
-    if not unimodal_P(samples, data):
+    if not unimodal_P(sample, data):
         raise ValueError('multi-modal period distribution')
 
+    if len(sample) > 1:
+        raise ValueError('can only pass a single sample to phase-fold at.')
+
     # HACK: hard-set getting the median
-    orbit = samples.median().get_orbit(0)
-    M0 = samples['M0'][0]
-    P = samples['P'][0]
-    s = samples['jitter'][0]
+    orbit = sample.get_orbit(0)
+    M0 = sample['M0'][0]
+    P = sample['P'][0]
+    s = sample['jitter'][0]
     t0 = data.t0 + (P/(2*np.pi)*M0).to(u.day, u.dimensionless_angles())
     phase = data.phase(P=P, t0=t0)
 
     # compute chi^2 of the orbit fit
     residual = data.rv - orbit.radial_velocity(data.t)
     err = np.sqrt(data.stddev**2 + s**2)
-    chisq = np.sum((residual**2 / err**2).decompose())
[…]

### 2018-02-22 twoface/samples_analysis.py

@@ -18,7 +18,7 @@ def unimodal_P(samples, data):
     """
 
     P_samples = samples['P'].to(u.day).value
-    P_min = P_samples.min()
+    P_min = np.min(P_samples)
     T = np.ptp(data.t.mjd)
     delta = 4*P_min**2 / (2*np.pi*T)
 

### 2018-02-22 scripts/run_apogee_continue_mcmc.py
compile results

@@ -16,6 +16,7 @@ import os
 import pickle
 
 # Third-party
+from astropy.stats import median_absolute_deviation
 from astropy.time import Time
 import h5py
 import numpy as np
@@ -110,16 +111,32 @@ def main(config_file, pool, seed, overwrite=False):
                       "you sure you ran `run_apogee.py`?"
                       .format(results_filename))
 
+    # The file to write the MCMC samples to:
+    mcmc_filename = path.join(TWOFACE_CACHE_PATH,
+                              "{0}-mcmc.hdf5".format(run.name))
+
+    if not path.exists(mcmc_filename): # ensure it exists
+        with h5py.File(mcmc_filename, 'w') as f:
+            pass
+
     # Create TheJoker sampler instance with the specified random seed and pool
     rnd = np.random.RandomState(seed=seed)
     logger.debug("Creating TheJoker instance with {0}, {1}".format(rnd, pool))
     params = run.get_joker_params()
+
+    # HACK: test
+    params.jitter = (8.5, 0.9)
     joker = TheJoker(params, random_state=rnd)
 
-    # Get all stars in this JokerRun that "need mcmc"
-    star_query = session.query(AllStar).join(StarResult, JokerRun, Status)\
+    # Get all stars in this JokerRun that "need mcmc" that are not in the MCMC
+    # cache file already
+    with h5py.File(mcmc_filename, 'r') as f:
+        done_ids = list(f.keys())
+
+    base_query = session.query(AllStar).join(StarResult, JokerRun, Status)\
                                        .filter(JokerRun.name == run.name)\
                                        .filter(Status.id == 2)
+    star_query = base_query.filter(~AllStar.apogee_id.in_(done_ids))
 
     n_stars = star_query.count()
     logger.info("{0} stars left to process for run more samples '{1}'"
@@ -138,6 +155,48 @@ def main(config_file, pool, seed, overwrite=False):
         pass
 
     pool.close()
+
+    # Now go through all of the output files and collect them!
+    with open(path.join(cache_path, 'model.pickle')) as f:
+        model = pickle.load(f)
+
+    with h5py.File(mcmc_filename) as f:
+        for star in base_query.all():
+            if star.apogee_id in f:
+                continue
+
+            tmp_file = path.join(cache_path, '{0}.npy'.format(star.apogee_id))
+            chain = np.load(tmp_file)
+
+            g = f.create_group(star.apogee_id)
+
+            try:
+                g2 = g.create_group('chain-stats')
+
[…]

### 2018-02-22 scripts/run_apogee_continue_mcmc.py
more debug [ci skip]

@@ -62,6 +62,8 @@ def emcee_worker(task):
                 pickle.dump(model, f)
 
     if not path.exists(plot_path):
+        logger.debug('Making plots for {0}'.format(apogee_id))
+
         if sampler is None:
             chain = np.load(chain_path)
         else:

### 2018-02-22 scripts/run_apogee_continue_mcmc.py
fix some bugs and more debug

@@ -159,19 +159,21 @@ def main(config_file, pool, seed, overwrite=False):
     pool.close()
 
     # Now go through all of the output files and collect them!
-    with open(path.join(cache_path, 'model.pickle')) as f:
+    with open(path.join(cache_path, 'model.pickle'), 'rb') as f:
         model = pickle.load(f)
 
     with h5py.File(mcmc_filename) as f:
         for star in base_query.all():
             if star.apogee_id in f:
+                logger.debug('Star {0} already in MCMC cache file'.format(star.apogee_id))
                 continue
 
             tmp_file = path.join(cache_path, '{0}.npy'.format(star.apogee_id))
             chain = np.load(tmp_file)
 
             g = f.create_group(star.apogee_id)
-
+            
+            logger.debug('Adding star {0} to MCMC cache'.format(star.apogee_id))
             try:
                 g2 = g.create_group('chain-stats')
 

### 2018-02-22 paper/1-catalog/twoface-catalog.tex
Discussion words

@@
 
-\subsection{Comparison to previous work} \label{sec:compare-troup}
+\subsection{Comparison to other APOGEE companion catalogs}
+\label{sec:compare-troup}
 
-Stuff.
+There are at least two other recent catalogs of stellar systems and companions
+based on \apogee\ data.
+
+One of these studies focused on decomposing spectra of MS stars into mixtures of
+stellar spectra (\citealt{El-Badry:2018}).
+Conceptually, this method works because (a) for two unequal-mass stars,
+unexpected absorption lines will appear superimposed on the brighter star's
+spectrum, and (b) for close to equal-mass stars, the line depths and ratios will
+not be well-matched by a single stellar model.
+Using this technique, they identified thousands of candidate MS binaries and
+trinaries, but did not consider giant stars ($\log g > 4$).
+This sample is therefore complimentary to and non-overlapping with the catalog
+presented in this work.
+
+The other recent catalog searched for stellar and substellar companions to all
+stars in \apogee\ \acronym{DR12} (including the RGB) that passed a series of
+quality cuts, and had $\geq 8$ visits (\citealt{Troup:2016}).
+For each star in the sample, orbits were fit to the visit RVs using a multi-step
+orbit-fitting procedure: it starts by identifying significant periods and a few
+harmonics of those periods, then fits a Keplerian orbit at each of these
+harmonics using least-squares fitting (\citealt{De-Lee:2013}) with a modified
+$\chi^2$ statistic that penalizes fits in which the phase coverage of the data
+is poor.
+This procedure is not guaranteed to provide a unique orbit solution.
+
+Of the 382 companions released as a part of this previous search, only 188 of
+the host stars passed the stellar parameter and quality cuts used to define the
+parent sample in this work (see \sectionname~\ref{sec:data}).
+We have looked at all of the overlapping stars to compare the previously derived
+companion orbital properties to the posterior samplings derived with \thejoker.
+We find that the comparisons fall in three categories:
+(1) the parameters reported in \citet{Troup:2016} agree with the posterior
+samplings, and the period distribution appears unimodal,
+(2) the parameters reported in \citet{Troup:2016} identify one possible mode of
+a likely multi-modal posterior pdf over orbital parameters, and
+(3) the data changed significantly between \apogee\ \acronym{DR12} and \DR, so
+no meaningful comparisons can be made; Roughly 1/3 of the comparison sample
+falls into each class.
+\figurename~\ref{fig:troup-unimodal} shows a few representative cases in which
+the \citet{Troup:2016} orbital parameters (orbit shown as orange line in left
+panels, parameters shown as orange + in right panels) is consistent with the
+orbit samples from \thejoker.
+\figurename~\ref{fig:troup-multimodal} shows a few representative cases in which
+we find that the posterior pdf over orbital parameters is multimodal, and the
+\citet{Troup:2016} orbit identifies one of these modes.
+For completeness, \figurename~\ref{fig:troup-datachanged} shows two instances in
+which the orbital parameters from \citet{Troup:2016} no longer make sense,
+likely because the data changed between data releases.
 
@@
 \caption{%
+TODO:
 \label{fig:troup-unimodal}
@@
 \caption{%
+TODO:
 \label{fig:troup-multimodal}
@@
 \caption{%
+TODO:
 \label{fig:troup-datachanged}

### 2018-02-23 twoface/plot.py
don't need indices

@@ -220,9 +220,9 @@ def plot_phase_fold(data, sample):
 
     # HACK: hard-set getting the median
     orbit = sample.get_orbit(0)
-    M0 = sample['M0'][0]
-    P = sample['P'][0]
-    s = sample['jitter'][0]
+    M0 = sample['M0']
+    P = sample['P']
+    s = sample['jitter']
     t0 = data.t0 + (P/(2*np.pi)*M0).to(u.day, u.dimensionless_angles())
     phase = data.phase(P=P, t0=t0)
 

### 2018-02-23 paper/1-catalog/twoface-catalog.tex
Restructure results to be more to-the-point

@@
 
-\subsection{Initial individual-system posterior samplings}
+\subsection{Individual-system posterior samplings}
 \label{sec:initsampling}
+
+\todo{Explain that this is the procedure, but two full runs described below in results...}
 
@@
 a Gaussian over the transformed parameter $y = \ln s^2$ with the mean and
-standard deviation of the initial distribution set to $(\mu_{y,0}, \sigma_{y,0})
-= (10.6, 3)$ in units of $\textrm{m}~\textrm{s}^{-1}$.
+standard deviation $(\mu_y, \sigma_y)$ indicated where the runs are described
+(see \sectionname~\ref{sec:results}).
 The reference time for each star is set to the minimum visit epoch; $M_0$ then
 becomes the mean anomaly at the first visit observation.
-See \tablename~\ref{tbl:params} for a description of all parameters and priors
+\tablename~\ref{tbl:params} contains descriptions of all parameters and priors
 used.
@@
 requested number of samples (as few as one sample).
-When this happens, we proceed in one of two ways:
-\begin{enumerate}
-    \item \emph{needs MCMC}: If just 1 posterior sample is returned after
-    exhausting the full library of prior samples, or if multiple (but fewer than
-    \nposterior) are returned that all lie within a single mode of the posterior
-    pdf, the posterior pdf over orbital parameters is treated as effectively
-    unimodal.
-    In this case, we use the location of the returned sample, or the median over
-    returned samples to generate a small ball of initial conditions and use
-    standard MCMC to continue sampling until we obtain \nposterior\ samples.
-    In detail, we use an ensemble sampler (\citealt{Goodman:2010}) implemented
+If just 1 posterior sample is returned after exhausting the full library of
+prior samples, or if multiple (but fewer than \nposterior) are returned that all
+lie within a single mode of the posterior pdf, the posterior pdf over orbital
+parameters is treated as effectively unimodal: these stars are flagged ``needs
+MCMC.''
+If more than 1 posterior sample is returned after exhausting the full library of
+prior samples, and the samples lie in multiple modes of the posterior pdf, the
+only way to proceed is to generate more prior samples and continue running
+\thejoker: these stars are flagged ``needs more prior samples.''
+
+\subsubsection{``Needs MCMC'': Following up \thejoker\ with MCMC}
+
+In this case, we use the location of the returned sample, or a randomly chosen
+sample from those returned to generate a small ball of initial conditions and
+use standard MCMC to continue sampling until we obtain \nposterior\ samples.
+
+We use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010}) implemented
 in \python\ (\citealt{Foreman-Mackey:2013}) for this step.
-    We run with 1024 walkers for 16384 steps, take the final walker positions,
-    and downsample until we have \nposterior\ samples.
-    We compute convergence statistics for each run and report whether the chains
-    appear converged or not.
-    \item \emph{needs more prior samples}: If more than 1 posterior sample is
-    returned after exhausting the full library of prior samples and the samples
-    lie in multiple modes of the posterior pdf, the star is flagged. In this
-    case, the only way to proceed is to generate more prior samples and continue
-    running \thejoker\ for these sources to generate more posterior samples.
-\end{description}
+We run with 1024 walkers for 16384 steps, take the final walker positions, and
+downsample until we have \nposterior\ samples.
+\todo{We compute convergence statistics for each run and report whether the
+chains appear converged or not.}
 
-A full run of \thejoker\ on all \nstars\ \apogee\ stars used in this work took
-approximately 300 hours on a compute cluster with 448 cores, with the time
-dominated by sources with many ($\gtrsim 10$) visits.
-\todo{XX} stars were flagged as \emph{needs MCMC}, \todo{YY} were flagged as
-\emph{needs more prior samples}, and the remaining \todo{ZZ} stars completed normally.
+\subsubsection{``Needs more prior samples'': Continuing \thejoker\ sampling}
+
+We generate another equal-sized library of prior samples (a total of
+$2\times\nprior$ samples) and continue sampling.
+\todo{We note that because ... rejection sampling procedure, have to re-do
+the rejection step after computing with full library of samples...}
 
@@
 \section{Results}
+\label{sec:results}
 
@@
 \label{sec:inferjitter}
+
+A full run of \thejoker\ on all \nstars\ \apogee\ stars used in this work took
+approximately 300 hours on a compute cluster with 448 cores, with the time
+dominated by sources with many ($\gtrsim 10$) visits.
+
+For this initial run, we do not follow up on stars that return $<$\nposterior\
+samples.
+
+of the initial distribution set to $(\mu_{y,0}, \sigma_{y,0})
+= (10.6, 3)$ in units of $\textrm{m}~\textrm{s}^{-1}$.
 
@@
 
+\todo{Move math to appendix since we cite relevant papers that show how?}
 In detail, we maximize the marginal likelihood of the population-level
@@
 \todo{Reactions to the inferred parameters?}
-\todo{Discuss how this relates to empirical Bayes?}
 
@@
 
-TODO: re-do exactly \sectionname~\ref{sec:initsampling}, but with $(\mu_{y}, \sigma_{y}) = (9.5, 1.6)$ again in units of ${\rm m}~{\rm s}^{-1}$.
+Using the maximum-likelihood hyperparameters derived from the initial posterior
+samplings (see \sectionname~\ref{sec:initsampling}), we update and fix the
+excess variance prior distribution parameters---$(\mu_{y}, \sigma_{y}) = (9.5,
+1.6)$ in units of $\textrm{m}~\textrm{s}^{-1}$---and re-run \thejoker\ on the
+parent sample of \apogee\ \DR\ stars.
+\todo{Discuss how this relates to empirical Bayes?}
 
-TODO: how many hit each category (flag) after the initial run
+From this run, 91,096 stars completed and successfully returned \nposterior\
+samples using \thejoker\ alone.
+The remaining 5,135 stars did not return \nposterior\ samples: 4,744 were
+flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}.
 
@@
 
-\todo{APW}
+\todo{Thresholding on what now?}
 
-Thresholding on what now?
-
-Catalog.
-
-Some highlights from this catalog.
+\todo{Some highlights from this catalog.}
 
@@
 \software{
-The code used in this project is available from
-\url{https://github.com/adrn/TwoFace} under the MIT open-source
-software license.
-This research utilized the following open-source \python\ packages:
+% The code used in this project is available from
+% \url{https://github.com/adrn/TwoFace} under the MIT open-source
+% software license.
     \package{Astropy} (\citealt{Astropy-Collaboration:2013}),
     % \package{corner} (\citealt{Foreman-Mackey:2016}),
-    % \package{emcee} (\citealt{Foreman-Mackey:2013ascl}),
+    \package{emcee} (\citealt{Foreman-Mackey:2013}),
     \package{IPython} (\citealt{Perez:2007}),

### 2018-02-23 paper/1-catalog/twoface-catalog.tex
Math to appendix

@@
 \subsection{Individual-system posterior samplings}
-\label{sec:initsampling}
+\label{sec:samplings}
 
@@
 
-\todo{Move math to appendix since we cite relevant papers that show how?}
 In detail, we maximize the marginal likelihood of the population-level
-parameters $\bs{\alpha} = (\mu_y, \sigma_y)$.
+parameters $(\mu_y, \sigma_y)$.
 We compute this marginal likelihood using the per-object posterior samples
-re-weighted by the ratio of the value of the hyperprior evaluated at a given
-vector $\bs{\alpha}$ over the value of the default prior at the previously
-assumed values, $\bs{\alpha}_0 = (\mu_{y,0}, \sigma_{y,0})$ (see
-\sectionname~\ref{sec:initsampling}).
+re-weighted by the ratio of the value of the hyperprior evaluated at a given,
+new set of parameters $(\mu_y, \sigma_y)$ over the value of the default prior at
+the previously assumed values, $(\mu_{y,0}, \sigma_{y,0})$ (see above).
 This trick has been used in other hierarchical inferences as a way to
-marginalize over the per-object parameters
+marginalize over the per-object parameters to infer population-level parameters
 (\citealt{Hogg:2010,Foreman-Mackey:2014}); we describe how to compute the
-marginal likelihood in detail below.
-
-For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
-primary orbital parameters $\bs{\theta} = (P, e, \omega, M_0, K, v_0)$ and the
-excess-variance parameter, $y = \ln s^2$, using \thejoker; For brevity in
-expressions below, we will use the vector
-\begin{equation}
-    \bs{w} = (\bs{\theta}, y)
-\end{equation}
-to represent the full set of parameters.
-To obtain this sampling, we use an interim (Gaussian) prior on the
-excess-variance parameter parametrized by a mean and standard deviation, i.e.
-$\alpha_0 = (\mu_{y,0}, \sigma_{y,0})$ as described above.
-For a given source, the posterior samples in the above parameters are drawn from
-the distribution
-\begin{equation}
-    \bs{w}_k \sim p(\bs{w}_k \given D, \bs{\alpha}_0)
-\end{equation}
-where $D$ represents the data for a given object.
-
-We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
-given a new set of hyperparameters $\bs{\alpha}$
-\begin{equation}
-    p(\{D_n\} \given \bs{\alpha}) = \prod_n^N p(D_n \given \bs{\alpha})
-\end{equation}
-where in the above, we have assumed that this likelihood is separable ( the data
-for each source are independent).
-The per-source marginal likelihood in the above expression is given by
-\begin{align}
-    p(D_n \given \bs{\alpha}) &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \,
-      p(\bs{w}_n \given \bs{\alpha})\\
-    &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \, p(\bs{w}_n \given \bs{\alpha}) \,
-      \frac{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}\\
-    &= p(D_n \given \bs{\alpha}_0) \, \int \dd \bs{w}_n \,
-      \frac{p(\bs{w}_n \given \bs{\alpha})}{p(\bs{w}_n \given \bs{\alpha}_0)} \,
-      p(\bs{w}_n \given D_n, \bs{\alpha}_0) \label{eq:marglike} \quad .
-\end{align}
-Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
-be simplified to a sum over prior value ratios of the $K$ posterior samples in
-the log-excess-variance parameter for each $n$ star
-\begin{equation}
-    \approx \frac{\mathcal{Z}_n}{K} \,
-      \sum_k^{K} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
-\end{equation}
-where we have canceled the other priors (over $\bs{\theta}$), and all
-normalization constants appear in the constant scale factor $\mathcal{Z}_n$.
-
-The above expresion gives the marginal likelihood of the velocity data for a single source given new hyperparameters $\bs{\alpha}$.
-The full marginal likelihood is then the product of these individual likelihoods
-\begin{align}
-    p(\{D_n\} \given \alpha) &\propto \prod_n^N \frac{1}{K} \,
-      \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
-      \quad .
-\end{align}
-In practice, we evaluate the log-marginal-likelihood
-\begin{align}
-    \ln p(\{D_n\} \given \alpha) &\propto \sum_n^N \left[
-      \ln\left( \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
-      - \ln K\right]\label{eq:lnlike}\\
-    &\propto \sum_n^N \left[
-      \underset{k}{\textrm{logsumexp}}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
-      - \ln K\right]
-\end{align}
-where $\textrm{logsumexp}$ (the log-sum-exp trick) provides a more stable
-estimate of the sum in \eqname~\ref{eq:lnlike}.
+marginal likelihood in detail in \sectionname~\ref{sec:hierarch}.
 
@@
 
+\appendix
+\section{Hierarchical inference of the excess variance parameter}
+\label{sec:hierarch}
+
+For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
+primary orbital parameters $\bs{\theta} = (P, e, \omega, M_0, K, v_0)$ and the
+excess-variance parameter, $y = \ln s^2$, using \thejoker; For brevity in
+expressions below, we will use the vector
+\begin{equation}
+    \bs{w} = (\bs{\theta}, y)
+\end{equation}
+to represent the full set of parameters.
+To obtain this sampling, we use an interim (Gaussian) prior on the
+excess-variance parameter parametrized by a mean and standard deviation, i.e.
+$\alpha_0 = (\mu_{y,0}, \sigma_{y,0})$ as described above.
+For a given source, the posterior samples in the above parameters are drawn from
+the distribution
+\begin{equation}
+    \bs{w}_k \sim p(\bs{w}_k \given D, \bs{\alpha}_0)
+\end{equation}
+where $D$ represents the data for a given object.
+
+We want to compute the likelihood of all data from all $N$ stars, $\{D_n\}$,
+given a new set of hyperparameters $\bs{\alpha}$
+\begin{equation}
+    p(\{D_n\} \given \bs{\alpha}) = \prod_n^N p(D_n \given \bs{\alpha})
+\end{equation}
+where in the above, we have assumed that this likelihood is separable ( the data
+for each source are independent).
+The per-source marginal likelihood in the above expression is given by
+\begin{align}
+    p(D_n \given \bs{\alpha}) &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \,
+      p(\bs{w}_n \given \bs{\alpha})\\
+    &= \int \dd \bs{w}_n \, p(D_n \given \bs{w}_n) \, p(\bs{w}_n \given \bs{\alpha}) \,
+      \frac{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}{p(\bs{w}_n \given D_n, \bs{\alpha}_0)}\\
+    &= p(D_n \given \bs{\alpha}_0) \, \int \dd \bs{w}_n \,
+      \frac{p(\bs{w}_n \given \bs{\alpha})}{p(\bs{w}_n \given \bs{\alpha}_0)} \,
+      p(\bs{w}_n \given D_n, \bs{\alpha}_0) \label{eq:marglike} \quad .
+\end{align}
+Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
+be simplified to a sum over prior value ratios of the $K$ posterior samples in
+the log-excess-variance parameter for each $n$ star
+\begin{equation}
+    \approx \frac{\mathcal{Z}_n}{K} \,
+      \sum_k^{K} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
+\end{equation}
+where we have canceled the other priors (over $\bs{\theta}$), and all
+normalization constants appear in the constant scale factor $\mathcal{Z}_n$.
+
+The above expresion gives the marginal likelihood of the velocity data for a single source given new hyperparameters $\bs{\alpha}$.
+The full marginal likelihood is then the product of these individual likelihoods
+\begin{align}
+    p(\{D_n\} \given \alpha) &\propto \prod_n^N \frac{1}{K} \,
+      \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
+      \quad .
+\end{align}
+In practice, we evaluate the log-marginal-likelihood
+\begin{align}
+    \ln p(\{D_n\} \given \alpha) &\propto \sum_n^N \left[
+      \ln\left( \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
+      - \ln K\right]\label{eq:lnlike}\\
+    &\propto \sum_n^N \left[
+      \underset{k}{\textrm{logsumexp}}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
+      - \ln K\right]
+\end{align}
+where $\textrm{logsumexp}$ (the log-sum-exp trick) provides a more stable
+estimate of the sum in \eqname~\ref{eq:lnlike}.
+
 \end{document}

### 2018-02-23 scripts/run_apogee_continue_mcmc.py
Compute G-R stat, and mean/std

@@ -33,6 +33,24 @@ from twoface.config import TWOFACE_CACHE_PATH
 from twoface.plot import plot_mcmc_diagnostic, plot_data_orbits
 
 
+def gelman_rubin(chain):
+    """
+    Implementation from http://joergdietrich.github.io/emcee-convergence.html
+    """
+    m, n, *_ = chain.shape
+
+    V = np.var(chain, axis=1, ddof=1) # variance over steps
+    W = np.mean(V, axis=0) # mean variance over walkers
+
+    θb = np.mean(chain, axis=1) # mean over steps
+    θbb = np.mean(θb, axis=0) # mean over walkers
+
+    B = n / (m - 1) * np.sum((θbb - θb)**2, axis=0)
+    var_θ = (n - 1) / n * W + 1 / n * B
+
+    return np.sqrt(var_θ / W)
+
+
 def emcee_worker(task):
     cache_path, results_filename, apogee_id, data, joker = task
     n_walkers = 1024
@@ -170,26 +188,37 @@ def main(config_file, pool, seed, overwrite=False):
 
             tmp_file = path.join(cache_path, '{0}.npy'.format(star.apogee_id))
             chain = np.load(tmp_file)
+            n_walkers, n_steps, n_pars = chain.shape
 
             g = f.create_group(star.apogee_id)
-            
+
             logger.debug('Adding star {0} to MCMC cache'.format(star.apogee_id))
             try:
                 g2 = g.create_group('chain-stats')
 
-                # compute running median, running MAD
+                # compute running median, MAD, mean, stddev
                 all_med = []
                 all_mad = []
+                all_mean = []
+                all_std = []
                 for k in range(chain.shape[-1]):
                     all_med.append(np.median(chain[..., k], axis=0))
                     all_mad.append(median_absolute_deviation(chain[..., k],
                                                              axis=0))
+                    all_mean.append(np.mean(chain[..., k], axis=0))
+                    all_std.append(np.std(chain[..., k], axis=0))
 
                 all_med = np.vstack(all_med).T
                 all_mad = np.vstack(all_mad).T
+                all_mean = np.vstack(all_mean).T
+                all_std = np.vstack(all_std).T
+                Rs = gelman_rubin(chain[:, n_steps//2:])
 
                 g2.create_dataset(name='median', data=all_med)
                 g2.create_dataset(name='MAD', data=all_mad)
+                g2.create_dataset(name='mean', data=all_mean)
[…]

### 2018-02-23 paper/1-catalog/twoface-catalog.tex
moved some figures around text

@@
 
+% Notebook: "Numbers of visits and sample overview"
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{nvisits.pdf}
+\end{center}
+\caption{%
+Number of \apogee\ stars in logarithmic bins of number of visits that pass the
+quality cuts described in \sectionname~\ref{sec:data}.
+In total, this work uses \nvisits\ visits and \nstars\ unique sources.
+\label{fig:nvisits}
+}
+\end{figure}
+
 We use the primary data products from \apogee\ \DR\ (i.e. the \texttt{allStar}
@@
 \figurename~\ref{fig:loggteff} ...
-
-% Notebook: "Numbers of visits and sample overview"
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{nvisits.pdf}
-\end{center}
-\caption{%
-Number of \apogee\ stars in logarithmic bins of number of visits that pass the
-quality cuts described in \sectionname~\ref{sec:data}.
-In total, this work uses \nvisits\ visits and \nstars\ unique sources.
-\label{fig:nvisits}
-}
-\end{figure}
 
@@
 
-\todo{Explain that this is the procedure, but two full runs described below in results...}
+Here we describe the procedure we use to generate per-target posterior samplings
+for the \apogee\ targets.
+We execute the full procedure twice for different goals (as described in
+\sectionname~\ref{sec:results}), and only here outline the key steps in the
+pipeline.
 

### 2018-02-23 scripts/run_apogee_continue_mcmc.py
wrap

@@ -183,7 +183,8 @@ def main(config_file, pool, seed, overwrite=False):
     with h5py.File(mcmc_filename) as f:
         for star in base_query.all():
             if star.apogee_id in f:
-                logger.debug('Star {0} already in MCMC cache file'.format(star.apogee_id))
+                logger.debug('Star {0} already in MCMC cache file'
+                             .format(star.apogee_id))
                 continue
 
             tmp_file = path.join(cache_path, '{0}.npy'.format(star.apogee_id))

### 2018-02-23 scripts/run_apogee_continue_mcmc.py
write f4

@@ -55,7 +55,7 @@ def emcee_worker(task):
                                                     n_walkers=n_walkers,
                                                     return_sampler=True)
 
-        np.save(chain_path, sampler.chain)
+        np.save(chain_path, sampler.chain.astype('f4'))
 
         if not path.exists(model_path):
             with open(model_path, 'wb') as f:

### 2018-02-24 paper/1-catalog/twoface-catalog.tex
Restructure and outline

@@
 
-\section{Results}
-\label{sec:results}
-
-\subsection{Inferring the excess variance distribution}
+\section{Experiment: infer the excess variance distribution}
 \label{sec:inferjitter}
@@
 
-\subsection{Catalog of posterior samplings for APOGEE DR14 RGB stars}
-\label{sec:fullcatalog}
+\section{Results: companion catalogs}
+\label{sec:results}
 
@@
 
-\subsection{A catalog of confident companions} \label{sec:conf-companions}
+\todo{Stuff about full catalog of posterior samples, available online}
+
+\subsection{Stars with confident companions}
+\label{sec:catalog-confident}
 
@@
 
-\todo{Some highlights from this catalog.}
+\todo{Examples of stars with no companions at various num. epochs}
+
+\todo{Examples of stars that meet K cut at various num. epochs}
+
+\subsection{Confident companions with highly constrained orbits}
+\label{sec:catalog-multimodal}
+
+\todo{Many examples of stars that meet K cut but solutions limited to a few
+qualitatively different solutions}
+
+\subsection{Confident companions with effectively uniquely determined orbits}
+\label{sec:catalog-unimodal}
+
+\todo{Many examples of stars that are probably unimodal}
 
@@
 
-Return to our various assumptions and make sure that we discuss them
-\emph{by name} here.
+\todo{Return to our various assumptions and make sure that we discuss them
+\emph{by name} here.}
 
@@
 
+\subsection{Population inference}
+
+Comment on things we could do with these samples hierarchical inference...
+
+We are SB1 only, but should do SB2 as well - no reason to restrict.
+
+\section{Conclusions}
+
+Tables available online, full code open source.\footnote{The code used in this
+project is available from \url{https://github.com/adrn/TwoFace} under the MIT
+open-source software license.}
+
 \acknowledgements
@@
 \software{
-% The code used in this project is available from
-% \url{https://github.com/adrn/TwoFace} under the MIT open-source
-% software license.
     \package{Astropy} (\citealt{Astropy-Collaboration:2013}),
-    % \package{corner} (\citealt{Foreman-Mackey:2016}),
     \package{emcee} (\citealt{Foreman-Mackey:2013}),

### 2018-02-24 twoface/mass.py
some new functions

@@ -3,7 +3,10 @@
 # Standard library
 
 # Third-party
+from astropy.constants import G
+import astropy.units as u
 import numpy as np
+from scipy.optimize import root
 
 __all__ = ['CM_NM_to_CNM', 'CM_NM_to_CN', 'get_martig_vec', 'm2_func']
 
@@ -76,3 +79,14 @@ def get_martig_vec(Teff, logg, M_H, C_M, N_M):
 
 def m2_func(m2, m1, sini, mf):
     return (m2*sini)**3 / (m1 + m2)**2 - mf
+
+
+@u.quantity_input(m1=u.Msun, mf=u.Msun)
+def get_m2_min(m1, mf):
+    mf = mf.to(m1.unit)
+    res = root(m2_func, x0=10., args=(m1.value, 1., mf.value))
+    return res.x[0] * m1.unit
+
+
+def stellar_radius(star, mass):
+    return np.sqrt(G*mass / (10**star.logg*u.cm/u.s**2)).to(u.Rsun)

### 2018-02-25 paper/1-catalog/twoface-catalog.tex
More TODO items
fix unclosed todo
fix two undefined

@@
 \newcommand{\nvisits}{397,559}
+\newcommand{\ncompanions}{xx}
 
@@
 have companions, but for  which more observations are needed to uniquely
-determine the orbital properties, and (3) posterior samplings for all \ntotal\
+determine the orbital properties, and (3) posterior samplings for all \nstars\
 stars in the parent sample.
@@
 
+\todo{Period - eccentricity plot? Tidal circularization (See "Debug MCMC
+continue")}
+
+\todo{Stars with Martig or Ness masses - m2min}
+
 \section{Discussion}

### 2018-02-25 paper/1-catalog/twoface-catalog.tex
title tweak; abstract work

@@
 
-\title{Binary companions of red giant stars with APOGEE DR14}
+\title{Binary companions of red giant stars in \apogee\ \DR}
 
@@
 sub-stellar, and planetary-mass companions.
-Even sparse or few-epoch observations are useful for detecting companions,
-though such data are difficult to use for measuring physical properties of
-individual systems because permitted orbital solutions can be highly degenerate.
+Even a very small number of observation epochs can be useful for detecting companions,
+though such data can be difficult to use for characterizing
+individual systems:
+There can be multiple qualitatively different orbital solutions that fit the data.
+For these reasons, we custom-built a Monte Carlo sampler (\thejoker)
+that delivers reliable (and often highly
+multi-modal) posterior samplings for companion orbital parameters given sparse
+radial-velocity data.
 % Aims
-Using a custom-built Monte Carlo sampler (\thejoker) to deliver (often highly
-multi-modal) posterior beliefs about companion orbital parameters, here we
-perform a search for secondary companions of over 100,000 stars observed by the
-\apogee\ survey (\DR) with $\geq 3$ epochs.
+Here we
+perform a search for secondary companions of over 100,000 red-giant
+stars observed in the
+\apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
 % Methods
-We select stars with probable companions by \todo{how?}.
+We select stars with probable companions by making a cut on our posterior belief about
+the amplitude of the stellar radial-velocity variation induced by the orbit \todo{do we?}.
 % Results
-We provide (1) a catalog of \ncompanions\ companions for which the companion
-properties can be confidently determined, (2) a catalog of stars that likely
-have companions, but for  which more observations are needed to uniquely
-determine the orbital properties, and (3) posterior samplings for all \nstars\
+We provide (1)~a catalog of \ncompanions\ companions for which the stellar companion
+properties can be confidently determined in a posterior sense,
+(2)~a catalog of stars that likely
+have companions, but for which (in most cases) more observations
+would be needed to uniquely
+determine the orbital properties, and (3)~posterior samplings for all orbital-companion
+parameters for all \nstars\
 stars in the parent sample.

### 2018-02-25 paper/1-catalog/twoface-catalog.tex
numbers

@@
 % Results
-We provide (1)~a catalog of \ncompanions\ companions for which the stellar companion
+We provide (1)~a catalog of \todo{\ncompanions} companions for which the stellar companion
 properties can be confidently determined in a posterior sense,
-(2)~a catalog of stars that likely
+(2)~a catalog of stars \todo{YYY} stars that likely
 have companions, but for which (in most cases) more observations

### 2018-02-26 paper/1-catalog/twoface-catalog.tex
MCMC section details

@@
 requested number of samples (as few as one sample).
-If just 1 posterior sample is returned after exhausting the full library of
+When this occurs, we continue sampling either using standard MCMC, or by
+increasing the size of the prior cache and continuing rejection sampling with
+\thejoker.
+
+\subsubsection{``Needs MCMC'': Following up \thejoker\ with MCMC}
+
+If just one posterior sample is returned after exhausting the full library of
 prior samples, or if multiple (but fewer than \nposterior) are returned that all
@@
 MCMC.''
-If more than 1 posterior sample is returned after exhausting the full library of
-prior samples, and the samples lie in multiple modes of the posterior pdf, the
-only way to proceed is to generate more prior samples and continue running
-\thejoker: these stars are flagged ``needs more prior samples.''
-
-\subsubsection{``Needs MCMC'': Following up \thejoker\ with MCMC}
-
-In this case, we use the location of the returned sample, or a randomly chosen
-sample from those returned to generate a small ball of initial conditions and
-use standard MCMC to continue sampling until we obtain \nposterior\ samples.
+In this case, we use the location of the returned sample (if only one is
+returned), or a randomly chosen sample from those returned (if multiple samples
+are returned within one mode) to generate a small Gaussian ball of initial
+conditions and use standard MCMC to continue sampling until we obtain
+\nposterior\ samples.
 
 We use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010}) implemented
-in \python\ (\citealt{Foreman-Mackey:2013}) for this step.
-We run with 1024 walkers for 16384 steps, take the final walker positions, and
-downsample until we have \nposterior\ samples.
-\todo{We compute convergence statistics for each run and report whether the
-chains appear converged or not.}
+in \python\ (\package{emcee}; \citealt{Foreman-Mackey:2013}) to perform the
+samplings.
+We transform the standard Keplerian orbital parameters to a safer
+parametrization, $(\ln P, \sqrt{K}\,\cos M_0, \sqrt{K}\,\sin M_0, \sqrt{e}\,\cos
+\omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, to sample in.
+This reparametrization is safer and more efficient for sampling with
+\package{emcee}, which expects parameters to be components of a vector so that
+linear operations can be applied (see, e.g., \citealt{Hogg:2017}); the angle
+variables $(\omega, M_0)$ don't meet this requirement in the standard
+parametrization.
+We use the same prior pdfs as in \thejoker\ when running MCMC (see
+\tablename~\ref{tbl:params}).
+
+For each star that is flagged ``needs MCMC,'' we run \package{emcee} with 1024
+walkers for 16384 steps, take the final walker positions, and downsample at
+random until we have \nposterior\ samples.
+We compute the Gelman--Rubin convergence statistic, $\hat{R}_j$,
+(\citealt{Gelman:1992}) for each parameter $j$ and include these values in the
+catalogs below when standard MCMC is run.
+We also provide a binary flag, ``converged,'' for each sampling continued with
+MCMC that is set to true if:
+\begin{equation}
+\underset{j}{\textrm{mean}}\left(\hat{R}_j\right) < 1.1 \quad .
+\end{equation}
 
@@
 
-We generate another equal-sized library of prior samples (a total of
-$2\times\nprior$ samples) and continue sampling.
+If more than one posterior sample is returned after exhausting the full library
+of prior samples, and the samples lie in multiple modes of the posterior pdf,
+the only way to proceed is to generate more prior samples and continue running
+\thejoker: these stars are flagged ``needs more prior samples.''
+In this case, we generate another equal-sized library of prior samples (a total
+of $2\times\nprior$ samples) and continue rejection sampling.
 \todo{We note that because ... rejection sampling procedure, have to re-do

### 2018-02-26 paper/1-catalog/twoface-catalog.tex
More words in 3.2

@@
 In this case, we generate another equal-sized library of prior samples (a total
-of $2\times\nprior$ samples) and continue rejection sampling.
-\todo{We note that because ... rejection sampling procedure, have to re-do
-the rejection step after computing with full library of samples...}
+of $2\times\nprior$ samples) and re-do the rejection sampling.
+We note that because of the way the rejection sampling step is done, this is not
+equivalent to concatenating the results from a second, independent run of
+\thejoker: the log-likelihood values for all of the prior samples must be used.
+If at the end of this second run the target still has fewer than \nposterior\
+samples, the sampling is flagged as ``incomplete.''
 
@@
 \label{sec:inferjitter}
+
+\todo{Fix this bit from copy-pasting}
 

### 2018-02-27 paper/1-catalog/twoface-catalog.tex
Some BS

@@
 Using the maximum-likelihood hyperparameters derived from the initial posterior
-samplings (see \sectionname~\ref{sec:initsampling}), we update and fix the
+samplings (see \sectionname~\ref{sec:inferjitter}), we update and fix the
 excess variance prior distribution parameters---$(\mu_{y}, \sigma_{y}) = (9.5,
-1.6)$ in units of $\textrm{m}~\textrm{s}^{-1}$---and re-run \thejoker\ on the
+1.6)$, in units of $\textrm{m}~\textrm{s}^{-1}$---and re-run \thejoker\ on the
 parent sample of \apogee\ \DR\ stars.
@@
 The remaining 5,135 stars did not return \nposterior\ samples: 4,744 were
-flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}.
+flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}; We
+continue generating samples for these stars using the methodology explained
+above (see \sectionname~\ref{sec:samplings}).
+In the end, \todo{XX} stars have completed samplings with \nposterior\ samples,
+and, after re-running \thejoker\ with twice as many prior samples, \todo{XX}
+stars still have fewer samples that are multi-modal in period.
 
-\todo{Stuff about full catalog of posterior samples, available online}
+The full catalog of posterior samples for all \apogee\ stars in our parent
+sample is available online.
+\todo{Structure of table, where to download, make a webpage with some
+information?}
 
@@
 \label{sec:catalog-confident}
+
+We don't expect there to be a sharp transition in orbital parameters between
+stars with companions and stars without companions: There is a continuum of
+companion properties.
+For example, the companion masses can be brown dwarf-like, or at long periods,
+or at high inclination, all of which will make the velocity semi-amplitude, $K$,
+small for a given system.
+From the radial velocity data alone, there is therefore no way to select a
+complete sample of stars with and without companions.
 

### 2018-02-27 paper/1-catalog/twoface-catalog.tex
Priors on K v0

@@
 \section{Introduction} \label{sec:intro}
-
-\todo{Intro is missing many citations}
 
@@
     $s^2$ & $\ln s^2 \sim \mathcal{N}(\mu_y, \sigma_y^2)$ & extra variance added to each visit variance \\
-    $K$ & TODO & velocity semi-amplitude \\
-    $v_0$ & TODO & system barycentric velocity \\
+    $K$ & $\mathcal{N}(0, \sigma_{v}^2)~\kms$ & velocity semi-amplitude \\
+    $v_0$ & $\mathcal{N}(0, \sigma_{v}^2)~\kms$ & system barycentric velocity \\
     \hline
     \end{tabular}
-    \caption{Summary and description of parameters.
-$\textrm{Beta}(a, b)$ is the beta distribution with \todo{a, b}, $\mathcal{U}(a, b)$ the uniform distribution over the domain $(a, b)$, and $\mathcal{N}(\mu, \sigma^2)$ is the normal distribution with mean $\mu$ and variance $\sigma^2$.}
+    \caption{Summary and description of parameters. $\textrm{Beta}(a, b)$ is the
+    beta distribution with shape parameters $(a, b)$, $\mathcal{U}(a, b)$ the
+    uniform distribution over the domain $(a, b)$, and $\mathcal{N}(\mu,
+    \sigma^2)$ is the normal distribution with mean $\mu$ and variance
+    $\sigma^2$.
+    For the systemic velocity and semi-amplitude, we use a broad Gaussian prior
+    that is formally inconsistent between \thejoker\ and follow-up MCMC sampling
+    (see \sectionname~\ref{sec:mcmc}): in \thejoker\ we assume Gaussian priors
+    with $\sigma_v$ much larger than the measurement uncertainty so they can be
+    neglected ($\sigma_v \approx \infty$), whereas when running MCMC we fix
+    $\sigma_v = 10^3~\kms$.}
     \label{tbl:params}
@@
 \subsubsection{``Needs MCMC'': Following up \thejoker\ with MCMC}
+\label{sec:mcmc}
 

### 2018-02-27 paper/1-catalog/twoface-catalog.tex
Add symbols

@@
     $\sigma^2$.
-    For the systemic velocity and semi-amplitude, we use a broad Gaussian prior
-    that is formally inconsistent between \thejoker\ and follow-up MCMC sampling
-    (see \sectionname~\ref{sec:mcmc}): in \thejoker\ we assume Gaussian priors
-    with $\sigma_v$ much larger than the measurement uncertainty so they can be
-    neglected ($\sigma_v \approx \infty$), whereas when running MCMC we fix
-    $\sigma_v = 10^3~\kms$.}
+    For the systemic velocity and semi-amplitude, $(v_0, K)$, we use a broad
+    Gaussian prior that is formally inconsistent between \thejoker\ and
+    follow-up MCMC sampling (see \sectionname~\ref{sec:mcmc}): in \thejoker\ we
+    assume Gaussian priors with $\sigma_v$ much larger than the measurement
+    uncertainty so they can be neglected ($\sigma_v \approx \infty$), whereas
+    when running MCMC we fix $\sigma_v = 10^3~\kms$.}
     \label{tbl:params}

### 2018-02-27 paper/1-catalog/twoface-catalog.tex
Caption for infer jitter

@@
 
-\begin{table}[h!]
+\begin{table}[h]
     \centering
-    \begin{tabular}{ r c l }
+    \begin{tabular}{ r l l }
     \hline
@@
 
-\todo{Fix this bit from copy-pasting}
-
-A full run of \thejoker\ on all \nstars\ \apogee\ stars used in this work took
-approximately 300 hours on a compute cluster with 448 cores, with the time
-dominated by sources with many ($\gtrsim 10$) visits.
-
-For this initial run, we do not follow up on stars that return $<$\nposterior\
-samples.
-
-of the initial distribution set to $(\mu_{y,0}, \sigma_{y,0})
-= (10.6, 3)$ in units of $\textrm{m}~\textrm{s}^{-1}$.
-
 As an initial use of the per-source posterior samplings, we use a hierarchical
@@
 
+We execute a full run of \thejoker\ on all \nstars\ \apogee\ stars in our parent
+sample using initial values for the excess variance distribution parameters
+chosen so that $\sqrt{e^{\mu_{y,0}}} \approx 200~\mps$: $(\mu_{y,0},
+\sigma_{y,0}) = (10.6, 3)$ in units of $\mps$.
+This run took approximately 300 hours on a compute cluster with 448 cores, with
+the time dominated by sources with many ($\gtrsim 10$) visits.  For this initial
+run, we do not follow up on stars that return $<$\nposterior\ samples.
+
 We use 1,825 stars with $>10$ visits and $\log g < 2$ (to avoid upper RGB stars
@@
 1.64)$.
-\todo{Reactions to the inferred parameters?}
+These values are consistent with the estimated systematic floor of the visit
+velocity uncertainties of $\approx 100$--$200~\mps$ estimated from stars
+observed multiple times and on multiple plates (\citealt{Nidever:2015}).
+
+However, there are a number of caveats to keep in mind about this estimate of
+the excess variance distribution.
+First, we do not remove triple or other multiple systems: For any individual
+system, the radial velocity variations induced by other massive bodies will lead
+to larger preferred values for the excess variance parameter.
+We don't expect there to be a large number of triple systems in our sample, but
+this is still an important consideration for future efforts.
+Second, we are sensitive to outliers and very non-Gaussian systematic error
+distributions.
+In fact, the residuals we find away from our best-fit two-body orbital solutions
+do not appear strictly Gaussian and have tails towards larger systematic errors.
+\figurename~\ref{fig:residuals} shows the distribution of normalized residuals,
+$R_{nk}$, for each visit: the $k$ visit velocities for each $n$ star, $v_{nk}$,
+minus the predicted radial velocity from the best-fitting sample returned by
+\thejoker, $\hat{v}^*_{nk}$, normalized by the excess-variance-included
+uncertainty, $\sigma_{nk}^* = \sqrt{\sigma_{nk}^2 + \hat{s}_{n}^2}$, where
+$\hat{s}_{n}$ is computed from the excess-variance parameter of the best-fitting
+sample:
+\begin{equation}
+    R_{nk} = \frac{v_{nk} - \hat{v}^*_{nk}}{\sqrt{\sigma_{nk}^2 +
+    \hat{s}_{n}^2}} \quad .
+\end{equation}
 
@@
 \caption{%
-\todo{move notebook to figures path, re-do with "run2" or "run1" samples...}
-Inferred prior over log-jitter parameter ($y = \ln s^2$) using posterior samples
-for 1,825 lower-RGB stars with $>10$ visits.
-Left panel shows the ... Right panel shows the ...
+Inferred prior over excess variance parameter ($y = \ln s^2$) using posterior
+samples for 1,825 lower-RGB stars with $>10$ visits.
+Left and right panels show the distribution corresponding to the
+maximum-likelihood parameters $(\mu_y, \sigma_y) = (9.50, 1.64)$ in log and
+linear, respectively.
 \label{fig:infer-jitter}
@@
 \end{figure}
+
+% Notebook: "Residuals-vs-metadata"
+% \begin{figure}[h]
+% \begin{center}
+% \includegraphics[width=\textwidth]{residuals}
+% \end{center}
+% \caption{%
+% \todo{stuff}
+% \label{fig:residuals}
+% }
+% \end{figure}
 
@@
 samplings (see \sectionname~\ref{sec:inferjitter}), we update and fix the
-excess variance prior distribution parameters---$(\mu_{y}, \sigma_{y}) = (9.5,
-1.6)$, in units of $\textrm{m}~\textrm{s}^{-1}$---and re-run \thejoker\ on the
+excess variance prior distribution parameters---$(\mu_{y}, \sigma_{y}) = (9.50,
+1.64)$, in units of $\textrm{m}~\textrm{s}^{-1}$---and re-run \thejoker\ on the
 parent sample of \apogee\ \DR\ stars.

### 2018-02-27 paper/1-catalog/twoface-catalog.tex
Residuals figure

@@
 distributions.
-In fact, the residuals we find away from our best-fit two-body orbital solutions
-do not appear strictly Gaussian and have tails towards larger systematic errors.
-\figurename~\ref{fig:residuals} shows the distribution of normalized residuals,
-$R_{nk}$, for each visit: the $k$ visit velocities for each $n$ star, $v_{nk}$,
-minus the predicted radial velocity from the best-fitting sample returned by
-\thejoker, $\hat{v}^*_{nk}$, normalized by the excess-variance-included
-uncertainty, $\sigma_{nk}^* = \sqrt{\sigma_{nk}^2 + \hat{s}_{n}^2}$, where
-$\hat{s}_{n}$ is computed from the excess-variance parameter of the best-fitting
-sample:
-\begin{equation}
-    R_{nk} = \frac{v_{nk} - \hat{v}^*_{nk}}{\sqrt{\sigma_{nk}^2 +
-    \hat{s}_{n}^2}} \quad .
-\end{equation}
+We later assess the non-Gaussianity of the visit velocity uncertainties by
+looking at the visit velocity residuals away from the orbit samples produced by
+\thejoker\ using the updated excess variance distribution (see
+\sectionname~\ref{sec:discuss-assumptions}).
 
@@
 \end{figure}
-
-% Notebook: "Residuals-vs-metadata"
-% \begin{figure}[h]
-% \begin{center}
-% \includegraphics[width=\textwidth]{residuals}
-% \end{center}
-% \caption{%
-% \todo{stuff}
-% \label{fig:residuals}
-% }
-% \end{figure}
 
@@
 
+\subsection{Assumptions}
+\label{sec:discuss-assumptions}
+
 \todo{Return to our various assumptions and make sure that we discuss them
 \emph{by name} here.}
+
+We find that the residuals away from our best-fit two-body orbital solutions
+appear close to Gaussian, indicating that the \emph{simple noise model}
+assumption may be sufficient.
+\figurename~\ref{fig:residuals} shows the distribution of normalized residuals,
+$R_{nk}$, for each visit: the $k$ visit velocities for each $n$ star, $v_{nk}$,
+minus the predicted radial velocity from the best-fitting sample returned by
+\thejoker, $\hat{v}^*_{nk}$, normalized by the excess-variance-included
+uncertainty, $\sigma_{nk}^* = \sqrt{\sigma_{nk}^2 + \hat{s}_{n}^2}$, where
+$\hat{s}_{n}$ is computed from the excess-variance parameter of the best-fitting
+sample:
+\begin{equation}
+    R_{nk} = \frac{v_{nk} - \hat{v}^*_{nk}}{\sqrt{\sigma_{nk}^2 +
+    \hat{s}_{n}^2}} \quad . \label{eq:normresid}
+\end{equation}
+
+% Notebook: "Residuals-vs-metadata"
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{residuals}
+\end{center}
+\caption{%
+Histogram shows the distribution of normalized visit residuals (see
+\eqname~\ref{eq:normresid}) for all \nvisits\ used in the final run of
+\thejoker\ (i.e. with updated excess variance distribution parameters inferred
+in \sectionname~\ref{sec:inferjitter}).
+The orange curve shows that expected for Gaussian uncertainties.
+The distribution appears mostly Gaussian over a large range of residual values,
+with slightly more populated tails and evidence of a few catastrophic outliers.
+\label{fig:residuals}
+}
+\end{figure}
 

### 2018-02-27 paper/1-catalog/twoface-catalog.tex
Move some text

@@
   However, there are known main-sequence double-lined binary stars in the
-  \apogee\ catalog \citealt{El-Badry:2018}, and an expected but unknown fraction
-  of \RGB--\RGB\ binaries.
+  \apogee\ catalog (\citealt{El-Badry:2018}), and an expected but unknown
+  fraction of \RGB--\RGB\ binaries.
 \item[simple noise model] Measurements are unbiased and noise estimates are
@@
 
-\subsection{Implementation notes}
+\subsection{Control sample}
+\label{sec:control-sample}
 
-\todo{Nontrivial computation - details about running on cluster, TheJoker vs. emcee}
+\todo{describe how we make it}
+
+% \subsection{Implementation notes}
+%
+% \todo{Nontrivial computation - details about running on cluster, TheJoker vs. emcee}
 
@@
 parent sample of \apogee\ \DR\ stars.
-\todo{Discuss how this relates to empirical Bayes?}
+This methodology, in which we estimate a prior from the data and then fix it, is
+typically referred to as ``empirical Bayes''; In this case, it is an
+approximation to doing a hierarchical inference of the individual system orbital
+parameters and the excess variance distribution hyperparameters simultaneously.
 
@@
 The remaining 5,135 stars did not return \nposterior\ samples: 4,744 were
-flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}; We
-continue generating samples for these stars using the methodology explained
+flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}.
+We continue generating samples for these stars using the methodology explained
 above (see \sectionname~\ref{sec:samplings}).
@@
 
-\subsection{Stars with confident companions}
+\subsection{Stars with companions}
 \label{sec:catalog-confident}
@@
 companion properties.
-For example, the companion masses can be brown dwarf-like, or at long periods,
-or at high inclination, all of which will make the velocity semi-amplitude, $K$,
-small for a given system.
-From the radial velocity data alone, there is therefore no way to select a
-complete sample of stars with and without companions.
+For example, the companion masses can be low mass, or at long periods, or at
+high inclination, all of which will make the velocity semi-amplitude, $K$, small
+for a given system.
+From the radial velocity data alone, there is therefore no way to define a
+simple cut to select a complete sample of stars with companions.
+Nevertheless, it is possible to define a cut that selects stars with
+high-confidence companions while similarly minimizing an estimated
+false-positive rate.
 
-\todo{Thresholding on what now?}
+\todo{False-positive estimate using the null control sample.}
+
+We select a sample of stars that confidently have companions using percentiles
+computed from the posterior samples in the velocity semi-amplitude parameter,
+$K$.
+\figurename~\ref{fig:} shows distributions of percentiles in $\ln K$ computed
+for all stars with posterior samplings from \thejoker.
+\todo{percentiles figure}
 
@@
 \emph{by name} here.}
+
+\todo{\emph{no multiplets} punt to future work}
+
+\todo{\emph{Kepler} evidence for coherent oscillations at the upper giant branch? companions that don't make sense}
+
+\todo{\emph{SB1} comment that there are some companions with masses large enough that they should be evolved, and therefore should contribute to the light!
+where are the companions?
+neutron stars / BHs?}
 

### 2018-02-28 scripts/run_null_control.py
use rnd for seed

@@ -115,7 +115,7 @@ def main(config_file, pool, seed, overwrite=False):
         y = rnd.normal(params.jitter[0], params.jitter[1])
         s = np.exp(0.5 * y) * params._jitter_unit
         std = np.sqrt(s**2 + orig_data.stddev**2).to(orig_data.rv.unit).value
-        new_rv = np.random.normal(np.mean(orig_data.rv).value, std)
+        new_rv = rnd.normal(np.mean(orig_data.rv).value, std)
         data = APOGEERVData(t=orig_data.t, rv=new_rv * orig_data.rv.unit,
                             stddev=orig_data.stddev)
 

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
Words

@@
 \newcommand{\nvisits}{397,559}
-\newcommand{\ncompanions}{xx}
+\newcommand{\ncontrol}{16,384}
+\newcommand{\nunimodal}{XX}
+\newcommand{\nmultimodal}{XX}
 
@@
 % Methods
-We select stars with probable companions by making a cut on our posterior belief about
-the amplitude of the stellar radial-velocity variation induced by the orbit \todo{do we?}.
+We select stars with probable companions by making a cut on our posterior belief
+about the amplitude of the stellar radial-velocity variation induced by the
+orbit.
 % Results
-We provide (1)~a catalog of \todo{\ncompanions} companions for which the stellar companion
-properties can be confidently determined in a posterior sense,
-(2)~a catalog of stars \todo{YYY} stars that likely
-have companions, but for which (in most cases) more observations
-would be needed to uniquely
-determine the orbital properties, and (3)~posterior samplings for all orbital-companion
-parameters for all \nstars\
-stars in the parent sample.
+We provide (1)~a catalog of \nunimodal\ companions for which the stellar
+companion properties can be confidently determined in a posterior sense, (2)~a
+catalog of stars \nmultimodal\ stars that likely have companions, but for which
+(in most cases) more observations would be needed to uniquely determine the
+orbital properties, and (3)~posterior samplings for all orbital-companion
+parameters for all \nstars\ stars in the parent sample.
 We highlight interesting systems and show the characteristics of stars with
@@
 
-\subsection{Control sample}
+\subsection{Null control sample}
 \label{sec:control-sample}
 
-\todo{describe how we make it}
+We construct a control sample of simulated data with no RV variability to assess
+our false-positive rates in the selections below (see
+\sectionname~\ref{sec:results}).
+We randomly pick \ncontrol\ stars from the parent sample used in this work and
+replace the visit velocity measurements with simulated data.
+For each star $n$, we randomly sample an excess variance parameter value from
+the prior, $s_n$.
+For each visit $k$, we then sample a new velocity $v_{nk}$ by drawing from a
+Gaussian with mean equal to the mean of the real data visit velocities,
+$\bar{v}_n$, and variance equal to the sum of the visit variance (uncertainty),
+$\sigma_{nk}$, and the excess variance,
+\begin{equation}
+    v_{nk} \sim \mathcal{N}(\bar{v}_n, \sigma_{nk}^2 + s_n^2) \quad .
+\end{equation}
+When we save the control data, we only store the visit uncertainty and
+``forget'' the fact that the simulated data is generated with excess variance.
 
@@
 
+We also run on the null control sample (see \sectionname~\ref{sec:control}) with
+the same parameters.
+\todo{summary of this run - do any stars get flagged as needs MCMC/etc?}
+
 \subsection{Stars with companions}
@@
 Nevertheless, it is possible to define a cut that selects stars with
-high-confidence companions while similarly minimizing an estimated
-false-positive rate.
-
-\todo{False-positive estimate using the null control sample.}
+high-confidence companions.
 
@@
 $K$.
-\figurename~\ref{fig:} shows distributions of percentiles in $\ln K$ computed
-for all stars with posterior samplings from \thejoker.
-\todo{percentiles figure}
+\figurename~\ref{fig:lnK-percentiles}, left panel, shows distributions of
+percentiles in $\ln K$ computed for all stars with posterior samplings from
+\thejoker.
+\figurename~\ref{fig:lnK-percentiles}, right panel, shows the same for posterior
+samplings for the null control sample.
+We define a cut in the \todo{XX} percentile of the $\ln K$ samples such that
+\todo{XX false-positives...} (vertical line in
+\figurename~\ref{fig:lnK-percentiles}); \nmultimodal\ stars pass this cut.
 
-\todo{Examples of stars with no companions at various num. epochs}
+% % Notebook: "TODO"
+% \begin{figure}[h]
+% \begin{center}
+% \includegraphics[width=\textwidth]{lnK-percentiles}
+% \end{center}
+% \caption{%
+% TODO
+% \label{fig:lnK-percentiles}
+% }
+% \end{figure}
 
-\todo{Examples of stars that meet K cut at various num. epochs}
+\tablename~\ref{tbl:} contains a subset of the \apogee\ \DR\ \texttt{allStar}
+catalog information for the stars that pass the above cut and therefore likely
+have companions; For brevity below, we refer to this sample as the ``high-$K$''
+stars, and the complimentary sample as the ``low-$K$'' stars.
+\figurename~\ref{fig:highK} shows eight examples of high-$K$ stars, i.e. stars
+that likely have companions, with different numbers of visits.
+\figurename~\ref{fig:lowK} is the same for eight low-$K$ stars, i.e. stars that
+either have no companions, low-mass companions, or companions at long periods or
+high inclination.
+
+The majority of the stars in the high-$K$ sample have poorly constrained orbital
+parameters because the posterior pdfs are very multimodal.
+
+% % Notebook: "TODO"
+% Should also include period-eccentricity panels with samples
+% \begin{figure}[h]
+% \begin{center}
+% \includegraphics[width=\textwidth]{highK}
+% \end{center}
+% \caption{%
+% TODO
+% \label{fig:highK}
+% }
+% \end{figure}
+
+% % Notebook: "TODO"
+% \begin{figure}[h]
+% \begin{center}
+% \includegraphics[width=\textwidth]{lowK}
+% \end{center}
+% \caption{%
+% TODO
+% \label{fig:lowK}
+% }
+% \end{figure}
 

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
Discussion of assumption

@@
 
-\subsection{Orbit inference and velocity modeling}\label{sec:fitting}
+\subsection{Orbit inference and velocity modeling}
+\label{sec:fitting}
 
@@
 We execute the full procedure twice for different goals (as described in
-\sectionname~\ref{sec:results}), and only here outline the key steps in the
+\sectionname~\ref{sec:catalogs}), and only here outline the key steps in the
 pipeline.
@@
 standard deviation $(\mu_y, \sigma_y)$ indicated where the runs are described
-(see \sectionname~\ref{sec:results}).
+(see \sectionname~\ref{sec:catalogs}).
 The reference time for each star is set to the minimum visit epoch; $M_0$ then
@@
 our false-positive rates in the selections below (see
-\sectionname~\ref{sec:results}).
+\sectionname~\ref{sec:catalogs}).
 We randomly pick \ncontrol\ stars from the parent sample used in this work and
@@
 
-\section{Results: companion catalogs}
-\label{sec:results}
+\section{Companion catalogs}
+\label{sec:catalogs}
 
@@
 
+\subsection{Confident companions with effectively uniquely determined orbits}
+\label{sec:catalog-unimodal}
+
+\section{Results}
+\label{sec:results}
+
+\todo{Period-logg plot - lack of short-period companions for red clump stars?}
+
+\todo{The companions that don't make sense / can't be}
+
+\todo{Period-eccentricity plot - tidal circularization}
+
+\todo{Companion mass and mass-ratio distribution for stars with Ness / Martig masses}
+
 \section{Discussion}
@@
 
-\todo{Return to our various assumptions and make sure that we discuss them
-\emph{by name} here.}
-
-\todo{\emph{no multiplets} punt to future work}
-
-\todo{\emph{Kepler} evidence for coherent oscillations at the upper giant branch? companions that don't make sense}
-
-\todo{\emph{SB1} comment that there are some companions with masses large enough that they should be evolved, and therefore should contribute to the light!
-where are the companions?
-neutron stars / BHs?}
-
-We find that the residuals away from our best-fit two-body orbital solutions
-appear close to Gaussian, indicating that the \emph{simple noise model}
+It is important to keep in mind that the companion catalogs and results
+presented in this \documentname\ depend on the assumptions laid out in
+\sectionname~\ref{sec:fitting}.
+We have only considered radial velocity modulations from a single massive
+companion---\emph{no multiplets}; This is a fundamental limitation of our search
+and methodology.
+However, stars with multiple companions will likely be included in our companion
+catalog anyway, only with two-body orbital solutions that either don't fit the
+data well or pick out the shortest period orbit.
+
+Related to the above, we assume that all RV variations are
+gravitational---\emph{Kepler}.
+We see tentative evidence for coherent oscillations at the upper giant branch
+(see \figurename~\ref{fig:todo}), which we presently treat as excess variance or
+intrinsic jitter.
+If these oscillations are indeed from asteroseismic modes, \todo{why is this
+interesting?}.
+
+We assume that one star in each two-body system overwhelmingly dominates the
+luminosity and therefore spectrum of each system---\emph{SB1}.
+As seen in \figurename~\ref{fig:todo}, there are some companions with minimum
+masses comparable to or consistent with being larger than the masses of the
+observed star.
+These systems are either (a) nearly edge-on MS--RGB binaries (i.e. consistent
+with our assumption), (b) SB2 systems where the \apogee\ pipeline failed to flag
+the source as having broad lines or a bad fit, or (c) RGB--stellar remnant
+systems with a black hole or neutron star companion.
+A subset of these systems that look like they fall in category (c) are being
+followed-up to obtain further RV measurements to test whether any companions are
+black holes (Geha et al., in prep.).
+
+Finally, we have assumed that the visit velocity error distribution is Gaussian,
+and that the visit velocity uncertainties could be
+under-estimated---\emph{simple noise model}.
+We can test this assumption using posterior samples from \thejoker\ by computing
+the residuals away from our best-fit two-body orbital solutions.
+We find that the normalized residuals appear to be very close to Gaussian over a
+large dynamic range of normalized residual values, indicating that this
 assumption may be sufficient.

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
Err small edits

@@
 for a given system.
-From the radial velocity data alone, there is therefore no way to define a
+Therefore, from the radial velocity data alone, there is no way to define a
 simple cut to select a complete sample of stars with companions.
-Nevertheless, it is possible to define a cut that selects stars with
+It is nevertheless possible to define a cut that selects stars with
 high-confidence companions.

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
catalogs

@@
 parameters because the posterior pdfs are very multimodal.
+The high-$K$ sample includes all stars in the catalogs described in the next two
+subsections, however these two catalogs are mutually exclusive with one another.
 
@@
 
-\todo{Many examples of stars that meet K cut but solutions limited to a few
-qualitatively different solutions}
+A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
+multimodal posterior distributions over orbital parameters that are limited to
+just a few qualitatively different solutions.
+For these stars, one or a few more well-timed RV measurements would likely lead
+to uniquely-determined orbital parameters.
+\todo{How exactly do we select these stars?}
+
+\figurename~\ref{fig:todo} shows a few examples of stars that meet these
+criteria.
+Note that in these cases, there are certain times at which a future observation
+would be far more informative: For example, \todo{top panel, time X vs. time Y}.
+
+% % Notebook: "TODO"
+% \begin{figure}[h]
+% \begin{center}
+% \includegraphics[width=\textwidth]{lowK}
+% \end{center}
+% \caption{%
+% TODO
+% \label{fig:lowK}
+% }
+% \end{figure}
+
+\todo{Describe catalog contents: Also include M2_min and primary mass when
+possible?}
 
@@
 
+\todo{Should figures be phase-folded?}
+
+\todo{Also include M2_min and primary mass when possible?}
+
+\section{Results}
+\label{sec:results}
+
+\todo{Period-logg plot - lack of short-period companions for red clump stars?}
+
+\todo{The companions that don't make sense / can't be}
+
 \todo{Period - eccentricity plot? Tidal circularization (See "Debug MCMC
 continue")}
-
-\todo{Stars with Martig or Ness masses - m2min}
-
-\subsection{Confident companions with effectively uniquely determined orbits}
-\label{sec:catalog-unimodal}
-
-\section{Results}
-\label{sec:results}
-
-\todo{Period-logg plot - lack of short-period companions for red clump stars?}
-
-\todo{The companions that don't make sense / can't be}
-
-\todo{Period-eccentricity plot - tidal circularization}
 

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
results subsections

@@
 
-\todo{Describe catalog contents: Also include M2_min and primary mass when
+\todo{Describe catalog contents: Also include M2min and primary mass when
 possible?}
@@
 
-\todo{Also include M2_min and primary mass when possible?}
+\todo{Also include M2min and primary mass when possible?}
 
@@
 
+\subsection{Red clump stars lack short-period companions}
+
 \todo{Period-logg plot - lack of short-period companions for red clump stars?}
 
-\todo{The companions that don't make sense / can't be}
+\subsection{Evidence for tidal circularization at $P\sim 10~\textrm{day}$}
 
@@
 
+\subsection{Companion masses and mass ratios}
+
 \todo{Companion mass and mass-ratio distribution for stars with Ness / Martig masses}
+
+\subsection{Impossible companions}
+
+\todo{The companions that don't make sense / can't be}
 

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
Fewer short-period companions

@@
 
+Red clump (\RC) stars are horizontal branch stars and have therefore already
+climbed the RGB, where the surface of the star can grow to be several AU in
+size.
+\RC\ stars are therefore expected to less frequently have companions at short
+periods ($P<XX$~day) because the companions would have been engulfed during the
+RGB phase.
+\figurename~\ref{fig:todo} hints that there is an apparent dearth of
+short-period companions around $\log g \sim 2.5$ (the approximate location of
+the red clump).
+Another way to look at this is to compare the fraction of subgiant stars with
+short-period companions to the same for stars near the \RC.
+
 \subsection{Evidence for tidal circularization at $P\sim 10~\textrm{day}$}
@@
 
-\subsection{Impossible companions}
+\subsection{Impossible companions: coherent oscillations?}
 
@@
 
-Tables available online, full code open source.\footnote{The code used in this
+\todo{catalogs available online}, full code open source.\footnote{The code used in this
 project is available from \url{https://github.com/adrn/TwoFace} under the MIT

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
Placeholder for fractions
more todos
Abstract wrapping

@@
 sub-stellar, and planetary-mass companions.
-Even a very small number of observation epochs can be useful for detecting companions,
-though such data can be difficult to use for characterizing
+Even a very small number of observation epochs can be useful for detecting
+companions, though such data can be difficult to use for characterizing
 individual systems:
-There can be multiple qualitatively different orbital solutions that fit the data.
-For these reasons, we custom-built a Monte Carlo sampler (\thejoker)
-that delivers reliable (and often highly
-multi-modal) posterior samplings for companion orbital parameters given sparse
-radial-velocity data.
+There can be multiple qualitatively different orbital solutions that fit the
+data.
+For these reasons, we custom-built a Monte Carlo sampler (\thejoker) that
+delivers reliable (and often highly multi-modal) posterior samplings for
+companion orbital parameters given sparse radial-velocity data.
 % Aims
-Here we
-perform a search for secondary companions of over 100,000 red-giant
-stars observed in the
-\apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
+Here we perform a search for secondary companions of \nstars\ red-giant stars
+observed in the \apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
 % Methods
@@
 companion properties can be confidently determined in a posterior sense, (2)~a
-catalog of stars \nmultimodal\ stars that likely have companions, but for which
+catalog of \nmultimodal\ stars that likely have companions, but for which
 (in most cases) more observations would be needed to uniquely determine the
 orbital properties, and (3)~posterior samplings for all orbital-companion
-parameters for all \nstars\ stars in the parent sample.
+parameters for all stars in the parent sample.
 We highlight interesting systems and show the characteristics of stars with
@@
 
+\todo{To assess this, take boxes in log g-Teff}
+We find that, empirically, XX\% of the subgiants and YY\% of the \RC\ stars have
+companions with $XX < P < XX$~day.
+\todo{Assuming selection function doesn't preferentially bias...}
+
 \subsection{Evidence for tidal circularization at $P\sim 10~\textrm{day}$}
+
+\todo{Expectation from theory? What's spectial at 10 days?}
+
+\figurename~\ref{fig:todo} shows the orbital period and eccentricity...
 

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
Conc

@@
 
-\todo{catalogs available online}, full code open source.\footnote{The code used in this
-project is available from \url{https://github.com/adrn/TwoFace} under the MIT
-open-source software license.}
+All companion catalogs described in this work (\sectionname~\ref{sec:catalogs})
+are available online at \todo{XX}.
+The source code for this project is open source and available from
+\url{https://github.com/adrn/TwoFace} under the MIT open source software
+license.
 
@@
 
-It is a pleasure to thank Keith Hawkins (Columbia), XXXXXXX.
+It is a pleasure to thank Keith Hawkins (Columbia), \todo{who else?}.
 
@@
     \package{scipy} (\url{https://www.scipy.org/}),
-    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}).
+    \package{schwimmbad} (\todo{JOSS paper}),
+    \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}),
+    \package{thejoker} (\todo{thejoker zenodo?}),
+    \package{twobody} (\todo{twobody zenodo?}).
 }
@@
 \facility{\sdssiv, \apogee}
+
+\clearpage
 

### 2018-02-28 paper/1-catalog/twoface-catalog.tex
details about the control sample run

@@
 the same parameters.
-\todo{summary of this run - do any stars get flagged as needs MCMC/etc?}
+From the control sample run, 13 ``stars'' are flagged as \emph{needs MCMC}, and
+1023 are flagged as ``needs more prior samples'';
+For these stars, we don't continue sampling with \thejoker\ or MCMC and only use
+the posterior samples returned from the \thejoker.
+% See: Sample K cuts.ipynb
 

### 2018-03-01 paper/1-catalog/twoface-catalog.tex
With control sample

@@
 \newcommand{\ncontrol}{16,384}
+\newcommand{\nhighK}{4,510}
 \newcommand{\nunimodal}{XX}
@@
 companion properties can be confidently determined in a posterior sense, (2)~a
-catalog of \nmultimodal\ stars that likely have companions, but for which
+catalog of \nhighK\ stars that likely have companions, but for which
 (in most cases) more observations would be needed to uniquely determine the
@@
 $K$.
-\figurename~\ref{fig:lnK-percentiles}, left panel, shows distributions of
-percentiles in $\ln K$ computed for all stars with posterior samplings from
-\thejoker.
-\figurename~\ref{fig:lnK-percentiles}, right panel, shows the same for posterior
-samplings for the null control sample.
-We define a cut in the \todo{XX} percentile of the $\ln K$ samples such that
-\todo{XX false-positives...} (vertical line in
-\figurename~\ref{fig:lnK-percentiles}); \nmultimodal\ stars pass this cut.
-
-% % Notebook: "TODO"
-% \begin{figure}[h]
-% \begin{center}
-% \includegraphics[width=\textwidth]{lnK-percentiles}
-% \end{center}
-% \caption{%
-% TODO
-% \label{fig:lnK-percentiles}
-% }
-% \end{figure}
+\figurename~\ref{fig:lnK-percentiles}, shows the distribution of 1st percentiles
+of $\ln K$ computed for all stars with posterior samplings from \thejoker\
+(filled, blue histogram), along with the same for posterior samplings for the
+null control sample (solid line, black).
+To select stars with companions, we use a cut in the 1st percentile of the $\ln
+K$ samples such that $<1\%$ of the control sample is selected, i.e. our
+estimated false positive rate is $<1\%$.
+We use a threshold of $\ln K = 0$ to meet this constraint (vertical, dashed line
+in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this cut.
+
+% Notebook: "Sample K cuts.ipynb"
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{lnK-percentiles}
+\end{center}
+\caption{%
+TODO
+\label{fig:lnK-percentiles}
+}
+\end{figure}
 

### 2018-03-01 paper/1-catalog/twoface-catalog.tex
Caption

@@
 (filled, blue histogram), along with the same for posterior samplings for the
-null control sample (solid line, black).
+null control sample (solid, dark line).
 To select stars with companions, we use a cut in the 1st percentile of the $\ln
@@
 \caption{%
-TODO
+Distribution of 1st percentiles in $\ln K$ (in units of \kms) for the parent
+\apogee\ sample (blue, filled) and for the control sample (solid, dark line).
+The vertical line (dashed, orange) indicates our adopted cut to select high-$K$
+stars that likely have companions; $<1\%$ of the control sample falls above this
+cut.
 \label{fig:lnK-percentiles}

### 2018-03-02 twoface/db/model.py
Added a boolean column to label whether in high-K sample

@@ -48,6 +48,8 @@ class StarResult(Base):
                        ForeignKey('status.id'),
                        default=0, index=True)
 
+    high_K = Column('high_K', types.Boolean)
+
     star = relationship("AllStar")
     jokerrun = relationship("JokerRun",
                             backref=backref("results",

### 2018-03-02 twoface/plot.py
fix bug with xspan

@@ -51,8 +51,8 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
     else:
         fig = ax.figure
 
+    w = np.ptp(data.t.mjd)
     if xlim_choice == 'tight': # twice the span of the data
-        w = np.ptp(data.t.mjd)
         t_grid = np.linspace(data.t.mjd.min() - w*0.05,
                              data.t.mjd.max() + w*1.05,
                              n_times)
@@ -61,8 +61,9 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
         t_min = data.t.mjd.min()
         t_max = max(data.t.mjd.min() + 2*samples['P'].max().value,
                     data.t.mjd.max())
-        span = t_max - t_min
-        t_grid = np.linspace(t_min - 0.05*span, t_max + 0.05*span, 2048)
+        t_grid = np.linspace(t_min - 0.05*w,
+                             t_max + 0.05*w,
+                             n_times)
 
     elif xlim_choice == 'data': # span of the data
         t_grid = np.linspace(data.t.mjd.min() - w*0.05,

### 2018-03-02 paper/1-catalog/twoface-catalog.tex
high-K figures

@@
 stars, and the complimentary sample as the ``low-$K$'' stars.
-\figurename~\ref{fig:highK} shows eight examples of high-$K$ stars, i.e. stars
-that likely have companions, with different numbers of visits.
-\figurename~\ref{fig:lowK} is the same for eight low-$K$ stars, i.e. stars that
-either have no companions, low-mass companions, or companions at long periods or
-high inclination.
+\figurename s~\ref{fig:highK-0} and \ref{fig:highK-1} show eight examples of
+high-$K$ stars, i.e. stars that likely have companions, with different numbers
+of visits, $N$, indicated on the panels.
+\figurename s~\ref{fig:lowK-0} and \ref{fig:lowK-1} is the same for eight
+low-$K$ stars, i.e. stars that either have no companions, low-mass companions,
+or companions at long periods or high inclination.
 
@@
 
-% % Notebook: "TODO"
-% Should also include period-eccentricity panels with samples
-% \begin{figure}[h]
-% \begin{center}
-% \includegraphics[width=\textwidth]{highK}
-% \end{center}
-% \caption{%
-% TODO
-% \label{fig:highK}
-% }
-% \end{figure}
-
-% % Notebook: "TODO"
-% \begin{figure}[h]
-% \begin{center}
-% \includegraphics[width=\textwidth]{lowK}
-% \end{center}
-% \caption{%
-% TODO
-% \label{fig:lowK}
-% }
-% \end{figure}
+% Notebook: "HighK"
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-0}
+\end{center}
+\caption{%
+TODO
+\label{fig:highK-0}
+}
+\end{figure}
+
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-1}
+\end{center}
+\caption{%
+TODO
+\label{fig:highK-1}
+}
+\end{figure}
+
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{lowK-0}
+\end{center}
+\caption{%
+TODO
+\label{fig:lowK-0}
+}
+\end{figure}
+
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{lowK-1}
+\end{center}
+\caption{%
+TODO
+\label{fig:lowK-1}
+}
+\end{figure}
 

### 2018-03-02 paper/1-catalog/twoface-catalog.tex
Figure captions

@@
 \caption{%
-TODO
+Examples of stars in our high-$K$ sample, i.e. stars that likely have companions
+(see \sectionname~\ref{sec:catalog-confident}), with different numbers of
+visits.
+Left panels show the data (black markers, error bars are the visit velocity
+uncertainties) with 128 random orbits from the \nposterior\ posterior samples
+under-plotted (grey lines); The \texttt{APOGEE\_ID} of each target and the
+number of visits that pass our quality cuts, $N$, are indicated on each panel.
+Right panels show the \nposterior\ posterior samples visualized in
+period--eccentricity space.
+In most cases, despite confidently having companions based on the RV amplitude,
+the permitted orbit fits are highly multimodal.
 \label{fig:highK-0}
@@
 \caption{%
-TODO
+Continuation of \figurename~\ref{fig:highK-0}.
 \label{fig:highK-1}
@@
 \caption{%
-TODO
+Same as \figurename~\ref{fig:highK-0}, but for stars in the low-$K$ sample, i.e.
+stars that have undetected or no companions.
 \label{fig:lowK-0}
@@
 \caption{%
-TODO
+Continuation of \figurename~\ref{fig:lowK-0}.
 \label{fig:lowK-1}

### 2018-03-02 twoface/samples_analysis.py
Utility to get the max likelihood sample

@@ -2,7 +2,7 @@
 import astropy.units as u
 import numpy as np
 
-__all__ = ['unimodal_P']
+__all__ = ['unimodal_P', 'max_likelihood_sample']
 
 
 def unimodal_P(samples, data):
@@ -23,3 +23,23 @@ def unimodal_P(samples, data):
     delta = 4*P_min**2 / (2*np.pi*T)
 
     return np.std(P_samples) < delta
+
+
+def max_likelihood_sample(data, samples):
+    """Return the maximum-likelihood sample.
+
+    Parameters
+    ----------
+    data : `~thejoker.RVData`
+    samples : `~thejoker.JokerSamples`
+
+    """
+    chisqs = np.zeros(len(samples))
+
+    for i in range(len(samples)):
+        orbit = samples.get_orbit(i)
+        residual = data.rv - orbit.radial_velocity(data.t)
+        err = np.sqrt(data.stddev**2 + samples['jitter'][i]**2)
+        chisqs[i] = np.sum((residual**2 / err**2).decompose())
+
+    return samples[np.argmin(chisqs)]

### 2018-03-02 twoface/plot.py
change default orbit color to blue

@@ -5,14 +5,15 @@ import matplotlib.pyplot as plt
 from matplotlib.gridspec import GridSpec
 import numpy as np
 
-from thejoker import JokerSamples
 from thejoker.plot import plot_rv_curves
 from thejoker.sampler.likelihood import get_ivar
 
-from .samples_analysis import unimodal_P
+from .samples_analysis import unimodal_P, max_likelihood_sample
 
 __all__ = ['plot_data_orbits', 'plot_mcmc_diagnostic', 'plot_two_panel']
 
+# TODO: customize this in TheJoker?
+_RV_LBL = 'RV [{0:latex_inline}]'
 
 def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
                      xlim_choice='data', ylim_fac=1., n_times=4096, title=None,
@@ -77,7 +78,7 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
     if plot_kwargs is None:
         plot_kwargs = dict()
 
-    plot_kwargs.setdefault('color', '#aaaaaa')
+    plot_kwargs.setdefault('color', 'tab:blue')
     # plot_kwargs.setdefault('alpha', 0.25)
     plot_kwargs.setdefault('linewidth', 0.5)
     plot_rv_curves(samples, t_grid, rv_unit=u.km/u.s, data=data, ax=ax,
@@ -204,58 +205,63 @@ def plot_two_panel(data, samples, axes=None, tight=True, title=None,
     return fig
 
 
-# def plot_phase_fold(data, period, ax=None):
-#     """
-#     TODO:
-#
-#     Parameters
-#     ----------
-#     star : `~twoface.db.AllStar`
-#     sample : `~thejoker.JokerSamples`
-#
-#     Returns
-#     -------
-#     fig : `matplotlib.Figure`
-#     """
-#
-#     if not unimodal_P(sample, data):
-#         raise ValueError('multi-modal period distribution')
-#
-#     if len(sample) > 1:
-#         raise ValueError('can only pass a single sample to phase-fold at.')
-#
-#     # HACK: hard-set getting the median
-#     orbit = sample.get_orbit(0)
-#     M0 = sample['M0']
-#     P = sample['P']
-#     s = sample['jitter']
-#     t0 = data.t0 + (P/(2*np.pi)*M0).to(u.day, u.dimensionless_angles())
-#     phase = data.phase(P=P, t0=t0)
-#
-#     # compute chi^2 of the orbit fit
[…]

### 2018-03-02 paper/1-catalog/twoface-catalog.tex
unimodal figure and caption

@@
 uncertainties) with 128 random orbits from the \nposterior\ posterior samples
-under-plotted (grey lines); The \texttt{APOGEE\_ID} of each target and the
+under-plotted (lines, blue); The \texttt{APOGEE\_ID} of each target and the
 number of visits that pass our quality cuts, $N$, are indicated on each panel.
@@
 
-% % Notebook: "TODO"
-% \begin{figure}[h]
+% % Notebook: TODO
+% \begin{figure}[hp]
 % \begin{center}
-% \includegraphics[width=\textwidth]{lowK}
+% \includegraphics[width=\textwidth]{highK-unimodal}
 % \end{center}
@@
 % TODO
-% \label{fig:lowK}
+% \label{fig:highK-unimodal}
 % }
@@
 \todo{Also include M2min and primary mass when possible?}
+
+% Notebook: HighK-unimodal
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-unimodal}
+\end{center}
+\caption{%
+Examples of high-$K$ stars with effectively unimodal posterior samplings.
+Each panel shows the data (black markers), phase-folded at the period of the
+posterior sample with maximum posterior probability.
+Visit velocity uncertainties are shown as black error bars (these are typically
+smaller than or equal to the size of the markers), and the inferred jitter as
+grey, capped error bars.
+Line (blue) shows the orbit compute from the posterior sample with maximum
+posterior probability.
+\label{fig:highK-unimodal}
+}
+\end{figure}
 

### 2018-03-02 paper/1-catalog/twoface-catalog.tex
P-e figure

@@
 
-\todo{Many examples of stars that are probably unimodal}
-
-\todo{Should figures be phase-folded?}
-
 \todo{Also include M2min and primary mass when possible?}
@@
 \caption{%
-Examples of high-$K$ stars with effectively unimodal posterior samplings.
+Examples of high-$K$ stars with effectively unimodal posterior samplings and
+converged MCMC samplings.
 Each panel shows the data (black markers), phase-folded at the period of the
@@
 
-\todo{Expectation from theory? What's spectial at 10 days?}
-
-\figurename~\ref{fig:todo} shows the orbital period and eccentricity...
+\todo{Expectation from theory? What's special at 10 days?}
+
+\figurename~\ref{fig:unimodalPe} shows the orbital period and eccentricity...
 
@@
 continue")}
+
+% Notebook: HighK-unimodal
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{unimodal-P-e}
+\end{center}
+\caption{%
+TODO:
+\label{fig:unimodalPe}
+}
+\end{figure}
 

### 2018-03-02 twoface/plot.py
tweaking style

@@ -244,14 +244,15 @@ def plot_phase_fold(data, sample, ax=None, label=True,
     rv_unit = u.km/u.s
     ax.errorbar(phase, data.rv.to(rv_unit).value,
                 data.stddev.to(rv_unit).value,
-                linestyle='none', marker='o', color='k', markersize=5)
+                linestyle='none', marker='o', color='k', markersize=5,
+                zorder=10)
 
     if jitter_errorbar:
         ax.errorbar(phase, data.rv.to(rv_unit).value,
                     np.sqrt(data.stddev**2 +
                             sample['jitter']**2).to(rv_unit).value,
                     linestyle='none', marker='', elinewidth=0., color='#aaaaaa',
-                    alpha=0.9, capsize=3, capthick=1)
+                    alpha=0.9, capsize=3, capthick=1, zorder=9)
 
     phase_grid = np.linspace(0, 1, 1024)
     ax.plot(phase_grid, orbit.radial_velocity(t0 + phase_grid*P),

### 2018-03-03 twoface/mass.py
mf and m2_min improvements

@@ -3,12 +3,16 @@
 # Standard library
 
 # Third-party
+from astropy.utils.misc import isiterable
 from astropy.constants import G
 import astropy.units as u
 import numpy as np
 from scipy.optimize import root
 
-__all__ = ['CM_NM_to_CNM', 'CM_NM_to_CN', 'get_martig_vec', 'm2_func']
+# Project
+from .log import log as logger
+
+__all__ = ['CM_NM_to_CNM', 'CM_NM_to_CN', 'get_martig_vec', 'mf', 'm2_func']
 
 # For abundance transformations:
 log_eC = 8.39
@@ -77,6 +81,11 @@ def get_martig_vec(Teff, logg, M_H, C_M, N_M):
     return vec
 
 
+def mf(P, K, e):
+    mf_circ = P * K**3 / (2*np.pi * G)
+    return (mf_circ * (1 - e**2)**1.5).to(u.Msun)
+
+
 def m2_func(m2, m1, sini, mf):
     return (m2*sini)**3 / (m1 + m2)**2 - mf
 
@@ -84,8 +93,20 @@ def m2_func(m2, m1, sini, mf):
 @u.quantity_input(m1=u.Msun, mf=u.Msun)
 def get_m2_min(m1, mf):
     mf = mf.to(m1.unit)
-    res = root(m2_func, x0=10., args=(m1.value, 1., mf.value))
-    return res.x[0] * m1.unit
+    if isiterable(m1) and isiterable(mf):
+        m2s = []
+        for x, y in zip(m1, mf):
+            try:
+                res = root(m2_func, x0=10., args=(x.value, 1., y.value))
+                m2s.append(res.x[0])
+            except Exception as e:
+                logger.warning('Failed to compute m2_min for sample: {0}'.format(str(e)))
+                m2s.append(np.nan)
+        return m2s * m1.unit
+
+    else:
+        res = root(m2_func, x0=10., args=(m1.value, 1., mf.value))
+        return res.x[0] * m1.unit
 
 
 def stellar_radius(star, mass):

### 2018-03-04 twoface/mass.py
Two more funcs

@@ -12,7 +12,8 @@ from scipy.optimize import root
 # Project
 from .log import log as logger
 
-__all__ = ['CM_NM_to_CNM', 'CM_NM_to_CN', 'get_martig_vec', 'mf', 'm2_func']
+__all__ = ['CM_NM_to_CNM', 'CM_NM_to_CN', 'get_martig_vec', 'mf', 'm2_func',
+           'get_m2_min', 'stellar_radius', 'period_at_surface']
 
 # For abundance transformations:
 log_eC = 8.39
@@ -111,3 +112,8 @@ def get_m2_min(m1, mf):
 
 def stellar_radius(star, mass):
     return np.sqrt(G*mass / (10**star.logg*u.cm/u.s**2)).to(u.Rsun)
+
+
+def period_at_surface(M1, logg, e, M2=0*u.Msun):
+    P = 2*np.pi * G**(1/4.) * (M1+M2) / (M1**(3/4.)) * (10**logg*u.cm/u.s**2)**(-3/4) * (1-e)**(-3/2)
+    return P.to(u.day)

### 2018-03-04 paper/1-catalog/twoface-catalog.tex
Implementation notes
comment change
Description of table

@@
 
+% How do we run TheJoker on all of APOGEE? Run on a cluster, send batches of samples to processors because individual likelihood calls are fast
+%
+% How do we do MCMC? Similar issues as above, but less control because not inherently embarassingly parallel
+
 \section{Experiment: infer the excess variance distribution}
@@
 in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this cut.
+For brevity below, we refer to this sample as the ``high-$K$'' stars, and the
+complimentary sample as the ``low-$K$'' stars.
 
@@
 
-\tablename~\ref{tbl:} contains a subset of the \apogee\ \DR\ \texttt{allStar}
-catalog information for the stars that pass the above cut and therefore likely
-have companions; For brevity below, we refer to this sample as the ``high-$K$''
-stars, and the complimentary sample as the ``low-$K$'' stars.
+\begin{table}
+    \centering
+    \begin{tabular}{c | c}
+    \texttt{APOGEE\_ID} & \texttt{lnK\_per\_1} \\
+    \hline
+    2M00000002+7417074 & -2.028 \\
+    2M00000068+5710233 & -2.600 \\
+    2M00000222+5625359 &  2.134 \\
+    2M00000446+5854329 & -7.199 \\
+    2M00000535+1504343 & -4.457 \\
+    2M00000546+6152107 & -4.388 \\
+    2M00000797+6436119 & -0.774 \\
+    2M00000866+7122144 & -3.313 \\
+    ... & ...
+    \end{tabular}
+    \caption{This table contains the \texttt{APOGEE\_ID} for all \nstars\ stars
+    in the parent sample used in this work, for which we have posterior
+    samplings in orbital parameters.
+    The other column contains the 1st percentile computed over the $\ln K$
+    samples for each source, \texttt{lnK\_per\_1}:
+    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
+    defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.
+    }
+    \label{tbl:lnK-per}
+\end{table}
+
+\tablename~\ref{tbl:lnK-per} contains a list of all stars in the parent sample
+and the value of the 1st percentile over the $\ln K$ samples for each star.
 \figurename s~\ref{fig:highK-0} and \ref{fig:highK-1} show eight examples of
@@
 
+\todo{How do we actually select these?}
+
 A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
@@
 
-% Notebook: HighK-unimodal
+% Notebook: HighK-unimodal-catalog
 \begin{figure}[hp]
@@
 % Notebook: HighK-unimodal
-\begin{figure}[hp]
-\begin{center}
-\includegraphics[width=\textwidth]{unimodal-P-e}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{unimodal-P-e}
 \end{center}
@@
 \section{Discussion}
+
+\todo{comment on short period shit, eccentricities}
 

### 2018-03-04 paper/1-catalog/twoface-catalog.tex
MAP not median

@@
 \newcommand{\RGB}{\acronym{RGB}}
+\newcommand{\quantile}[2]{\ensuremath{\textrm{Q}_{#1}}\left(#2\right)}
 
@@
 \newcommand{\nhighK}{4,510}
-\newcommand{\nunimodal}{XX}
 \newcommand{\nmultimodal}{XX}
+\newcommand{\nunimodal}{319}
 
@@
 
-We also run on the null control sample (see \sectionname~\ref{sec:control}) with
-the same parameters.
+We also run on the null control sample (see
+\sectionname~\ref{sec:control-sample}) with the same parameters.
 From the control sample run, 13 ``stars'' are flagged as \emph{needs MCMC}, and
 1023 are flagged as ``needs more prior samples'';
-For these stars, we don't continue sampling with \thejoker\ or MCMC and only use
-the posterior samples returned from the \thejoker.
+For these control sample stars, we don't continue sampling with \thejoker\ or
+MCMC and only use the ($<$\nposterior) posterior samples returned from the
+\thejoker.
 % See: Sample K cuts.ipynb
@@
 companion properties.
-For example, the companion masses can be low mass, or at long periods, or at
-high inclination, all of which will make the velocity semi-amplitude, $K$, small
-for a given system.
+For example, the companions can be low mass, or at long periods, or at high
+inclination, all of which will make the velocity semi-amplitude, $K$, small for
+a given system.
 Therefore, from the radial velocity data alone, there is no way to define a
@@
 We select a sample of stars that confidently have companions using percentiles
-computed from the posterior samples in the velocity semi-amplitude parameter,
-$K$.
+computed from the log of the posterior samples in the velocity semi-amplitude
+parameter, $\ln K$.
 \figurename~\ref{fig:lnK-percentiles}, shows the distribution of 1st percentiles
@@
     \end{tabular}
-    \caption{This table contains the \texttt{APOGEE\_ID} for all \nstars\ stars
+    \caption{This table contains \texttt{APOGEE\_ID}'s for all \nstars\ stars
     in the parent sample used in this work, for which we have posterior
@@
 of visits, $N$, indicated on the panels.
-\figurename s~\ref{fig:lowK-0} and \ref{fig:lowK-1} is the same for eight
+\figurename s~\ref{fig:lowK-0} and \ref{fig:lowK-1} shows the same for eight
 low-$K$ stars, i.e. stars that either have no companions, low-mass companions,
 or companions at long periods or high inclination.
-
 The majority of the stars in the high-$K$ sample have poorly constrained orbital
 parameters because the posterior pdfs are very multimodal.
-The high-$K$ sample includes all stars in the catalogs described in the next two
-subsections, however these two catalogs are mutually exclusive with one another.
+The high-$K$ sample selected from \tablename~\ref{tbl:lnK-per} includes all
+stars in the catalogs described in the next two subsections, however the
+following two catalogs are mutually exclusive.
 
@@
 multimodal posterior distributions over orbital parameters that are limited to
-just a few qualitatively different solutions.
+a few qualitatively different solutions.
 For these stars, one or a few more well-timed RV measurements would likely lead
 to uniquely-determined orbital parameters.
-\todo{How exactly do we select these stars?}
-
 \figurename~\ref{fig:todo} shows a few examples of stars that meet these
 criteria.
-Note that in these cases, there are certain times at which a future observation
-would be far more informative: For example, \todo{top panel, time X vs. time Y}.
+These stars are prime candidates for follow-up efforts.
+In many of these cases, there are certain times at which a future observation
+would be far more informative.
+Visually, these times correspond to regions of time when the predicted RVs have
+large variance as computed for orbits computed from the posterior samplings.
+For example, \todo{top panel, time X vs. time Y, node vs. large variance}.
+
+\tablename~\ref{tbl:todo} contains ...
+\todo{Describe catalog contents: Also include M2min and primary mass when
+possible?}
 
@@
 
-\todo{Describe catalog contents: Also include M2min and primary mass when
-possible?}
-
 \subsection{Confident companions with effectively uniquely determined orbits}
@@
 
-\todo{Also include M2min and primary mass when possible?}
+Starting from the high-$K$ sample, we select stars that have posterior period
+samples that fall within a single mode.
+We define a period resolution for each star $\Delta =
+\left[4\,P_{\textrm{min}}^2 / (2\pi \, T) \right]$, where $P_{\textrm{min}}$ is
+the minimum period sample value, and $T$ is the epoch span of the data for that
+star.
+We consider a sampling to be unimodal when $P_{\textrm{max}} - P_{\textrm{min}}
+< \Delta$ (see \sectionname~2 of \citealt{Price-Whelan:2017}).
+Unlike the multi-modal stars, the posterior pdfs for these stars can be
+approximated using point-estimates and standard deviations of their respective
+samplings.
+
+\tablename~\ref{tbl:highK-unimodal} contains ...
+
+\input{tables/unimodal-schema}
 
@@
 
+\section{Using data products...}
+\label{sec:code-demo}
+
+\todo{Brief example of how to use twobody and thejoker to visualize orbits, with
+link to a tutorial notebook on github}
+
 \end{document}

### 2018-03-04 paper/1-catalog/twoface-catalog.tex
Describe table

@@
 We consider a sampling to be unimodal when $P_{\textrm{max}} - P_{\textrm{min}}
-< \Delta$ (see \sectionname~2 of \citealt{Price-Whelan:2017}).
+< \Delta$ (see \sectionname~2 of \citealt{Price-Whelan:2017}); \nunimodal\ stars
+pass this cut.
 Unlike the multi-modal stars, the posterior pdfs for these stars can be
@@
 samplings.
-
-\tablename~\ref{tbl:highK-unimodal} contains ...
+We report maximum \textit{a posteriori} sample values for the orbital
+parameters, along with other computed quantities for all \nunimodal\ stars in
+this high-$K$, unimodal sample.
+For stars with measured primary masses, $M_1$, from other work
+(\citealt{Ness:2015}), we compute the minimum companion mass, $M_{2,
+\textrm{min}}$ and include both masses in this catalog.
+\tablename~\ref{tbl:highK-unimodal} contains descriptions and units for all columns in the high-$K$, unimodal sample catalog.
 

### 2018-03-04 paper/1-catalog/twoface-catalog.tex
Clarify

@@
 samplings.
-We report maximum \textit{a posteriori} sample values for the orbital
+We report maximum \textit{a posteriori} (MAP) sample values for the orbital
 parameters, along with other computed quantities for all \nunimodal\ stars in
@@
 \textrm{min}}$ and include both masses in this catalog.
-\tablename~\ref{tbl:highK-unimodal} contains descriptions and units for all columns in the high-$K$, unimodal sample catalog.
+We additionally join with the \apogee\ \DR\ \texttt{allStar} catalog and provide
+all columns from this catalog for convenience.
+\tablename~\ref{tbl:highK-unimodal} contains descriptions and units for all
+columns in the high-$K$, unimodal sample catalog, available online.
+\todo{where?}
 

### 2018-03-04 twoface/samples_analysis.py
add func to compute MAP

@@ -1,8 +1,9 @@
 # Third-party
 import astropy.units as u
 import numpy as np
+from thejoker.sampler.mcmc import TheJokerMCMCModel
 
-__all__ = ['unimodal_P', 'max_likelihood_sample']
+__all__ = ['unimodal_P', 'max_likelihood_sample', 'MAP_sample']
 
 
 def unimodal_P(samples, data):
@@ -22,7 +23,7 @@ def unimodal_P(samples, data):
     T = np.ptp(data.t.mjd)
     delta = 4*P_min**2 / (2*np.pi*T)
 
-    return np.std(P_samples) < delta
+    return np.ptp(P_samples) < delta
 
 
 def max_likelihood_sample(data, samples):
@@ -43,3 +44,24 @@ def max_likelihood_sample(data, samples):
         chisqs[i] = np.sum((residual**2 / err**2).decompose())
 
     return samples[np.argmin(chisqs)]
+
+
+def MAP_sample(data, samples, joker_params):
+    """Return the maximum a posteriori sample.
+
+    Parameters
+    ----------
+    data : `~thejoker.RVData`
+    samples : `~thejoker.JokerSamples`
+    joker_params : `~thejoker.JokerParams`
+
+    """
+    model = TheJokerMCMCModel(joker_params, data)
+
+    ln_ps = np.zeros(len(samples))
+
+    mcmc_p = np.squeeze(model.pack_samples_mcmc(samples))
+    for i in range(len(samples)):
+        ln_ps[i] = model.ln_posterior(mcmc_p[i])
+
+    return samples[np.argmax(ln_ps)]

### 2018-03-04 twoface/plot.py
relative to t0 option

@@ -17,7 +17,8 @@ _RV_LBL = 'RV [{0:latex_inline}]'
 
 def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
                      xlim_choice='data', ylim_fac=1., n_times=4096, title=None,
-                     ax=None, highlight_P_extrema=True, plot_kwargs=None):
+                     ax=None, highlight_P_extrema=True, plot_kwargs=None,
+                     relative_to_t0=False):
     """
     Plot the APOGEE RV data vs. time and orbits computed from The Joker samples.
 
@@ -84,7 +85,8 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
     plot_rv_curves(samples, t_grid, rv_unit=u.km/u.s, data=data, ax=ax,
                    n_plot=min(len(samples['P']), n_orbits),
                    plot_kwargs=plot_kwargs,
-                   data_plot_kwargs=dict(zorder=5, elinewidth=1,))
+                   data_plot_kwargs=dict(zorder=5, elinewidth=1,),
+                   relative_to_t0=relative_to_t0)
 
     if highlight_P_extrema:
         # Darken the shortest period sample
@@ -92,14 +94,20 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
 
         P_min_samples = samples[samples['P'].argmin()]
         plot_rv_curves(P_min_samples, t_grid, rv_unit=u.km/u.s, ax=ax,
-                       n_plot=1, plot_kwargs=dark_style)
+                       n_plot=1, plot_kwargs=dark_style,
+                       relative_to_t0=relative_to_t0)
 
         # Darken the longest period sample
         P_max_samples = samples[samples['P'].argmax()]
         plot_rv_curves(P_max_samples, t_grid, rv_unit=u.km/u.s, ax=ax,
-                       n_plot=1, plot_kwargs=dark_style)
+                       n_plot=1, plot_kwargs=dark_style,
+                       relative_to_t0=relative_to_t0)
 
-    ax.set_xlim(t_grid.min(), t_grid.max())
+    if relative_to_t0:
+        ax.set_xlim(t_grid.min() - data.t0.mjd,
+                    t_grid.max() - data.t0.mjd)
+    else:
+        ax.set_xlim(t_grid.min(), t_grid.max())
 
     _rv = data.rv.to(u.km/u.s).value
     h = np.ptp(_rv)

### 2018-03-04 paper/1-catalog/twoface-catalog.tex
table and figure update

@@
 \RC\ stars are therefore expected to less frequently have companions at short
-periods ($P<XX$~day) because the companions would have been engulfed during the
-RGB phase.
+periods ($P<XX$~day) because short period companions would have been
+common-envelope during the RGB phase.
 \figurename~\ref{fig:todo} hints that there is an apparent dearth of

### 2018-03-06 paper/1-catalog/twoface-catalog.tex
TODOs

@@
 \caption{%
-Examples of high-$K$ stars with effectively unimodal posterior samplings and
-converged MCMC samplings.
+Randomly-selected examples of high-$K$ stars with effectively unimodal posterior
+samplings and converged MCMC samplings.
 Each panel shows the data (black markers), phase-folded at the period of the
@@
 \todo{Period-logg plot - lack of short-period companions for red clump stars?}
+
+\todo{also look at trends with Teff?}
+
+\todo{Cite troup, who also noted this}
 
@@
 
+\todo{Mathieu et al. 2004}
+
 \figurename~\ref{fig:unimodalPe} shows the orbital period and eccentricity...
@@
 
+\todo{eccentricity distribution w/ prior over-plotted}
+
 % Notebook: HighK-unimodal
@@
 
-\subsection{Impossible companions: coherent oscillations?}
-
-\todo{The companions that don't make sense / can't be}
+\todo{Interesting companions w/ large M2min, WD/NS/BH companions?}
+
+\todo{Plot: M1 vs. M2min}
+
+\todo{Plot: M2min vs. a2sini}
+
+\subsection{Eclipsing}
+
+\todo{Stars with Kepler light curves}
 
@@
 \end{figure}
+
+\subsection{Impossible companions: coherent oscillations?}
+
+\todo{The companions that don't make sense / can't be}
 
@@
 Comment on things we could do with these samples hierarchical inference...
+
+We don't need to know the orbits of each individual system: that's the power in
+this.
 

### 2018-03-07 twoface/db/model.py
semi-fix bug with j mag

@@ -93,7 +93,7 @@ class AllStar(Base):
     telescope = Column("telescope", types.String)
     location_id = Column("location_id", types.SmallInteger)
     field = Column("field", types.String)
-    j = Column("j", types.REAL)
+    # j = Column("j", types.REAL)
     j_err = Column("j_err", types.REAL)
     h = Column("h", types.REAL)
     h_err = Column("h_err", types.REAL)
@@ -281,12 +281,12 @@ class AllStar(Base):
     # param_cov = Column("param_cov", postgresql.ARRAY(types.REAL))
     _fp_str = "fparam{i} = Column('fparam{i}', types.REAL)"
     _cov_str = "fparam_cov{i}{j} = Column('fparam_cov{i}{j}', types.REAL)"
-    for i in range(9):
-        exec(_fp_str.format(i=i))
-        for j in range(9):
-            if j < i:
+    for _i in range(9):
+        exec(_fp_str.format(i=_i))
+        for _j in range(9):
+            if _j < _i:
                 continue
-            exec(_cov_str.format(i=i, j=j))
+            exec(_cov_str.format(i=_i, j=_j))
 
     def __repr__(self):
         return ("<ApogeeStar(id='{0}', apogee_id='{1}', {2} results)>"

### 2018-03-07 twoface/plot.py
default don't highlight

@@ -15,6 +15,7 @@ __all__ = ['plot_data_orbits', 'plot_mcmc_diagnostic', 'plot_two_panel']
 # TODO: customize this in TheJoker?
 _RV_LBL = 'RV [{0:latex_inline}]'
 
+
 def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
                      xlim_choice='data', ylim_fac=1., n_times=4096, title=None,
                      ax=None, highlight_P_extrema=True, plot_kwargs=None,
@@ -187,8 +188,8 @@ def plot_two_panel(data, samples, axes=None, tight=True, title=None,
         fig = ax1.figure
 
     # orbits
-    plot_data_orbits_kw['xlim_choice'] = plot_data_orbits_kw.get(
-        'xlim_choice', 'tight')
+    plot_data_orbits_kw.setdefault('xlim_choice', 'tight')
+    plot_data_orbits_kw.setdefault('highlight_P_extrema', False)
     fig = plot_data_orbits(data, samples, ax=ax1,
                            **plot_data_orbits_kw)
     if title is not None:
@@ -200,12 +201,12 @@ def plot_two_panel(data, samples, axes=None, tight=True, title=None,
     scatter_kw.setdefault('linewidth', 0.)
     ax2.scatter(samples['P'], samples['e'], **scatter_kw)
 
-    ax2.set_xscale('log', basex=2)
+    ax2.set_xscale('log', basex=10)
     ax2.set_xlim(0.5, 2**15)
     ax2.set_ylim(-0.02, 1.02)
     ax2.set_xlabel('period, $P$ [day]')
     ax2.set_ylabel('eccentricity, $e$')
-    ax2.xaxis.set_ticks(2**np.arange(0, 15+1, 3))
+    ax2.xaxis.set_ticks(10**np.arange(0, 4+1, 1))
 
     if tight:
         fig.tight_layout()

### 2018-03-07 twoface/mass.py
stability

@@ -99,15 +99,22 @@ def get_m2_min(m1, mf):
         for x, y in zip(m1, mf):
             try:
                 res = root(m2_func, x0=10., args=(x.value, 1., y.value))
+                if not res.success:
+                    raise RuntimeError('Unsuccessful')
                 m2s.append(res.x[0])
             except Exception as e:
-                logger.warning('Failed to compute m2_min for sample: {0}'.format(str(e)))
+                logger.warning('Failed to compute m2_min for sample: {0}'
+                               .format(str(e)))
                 m2s.append(np.nan)
         return m2s * m1.unit
 
     else:
         res = root(m2_func, x0=10., args=(m1.value, 1., mf.value))
-        return res.x[0] * m1.unit
+
+        if res.success:
+            return res.x[0] * m1.unit
+        else:
+            return np.nan * m1.unit
 
 
 def stellar_radius(star, mass):

### 2018-03-07 twoface/samples_analysis.py

@@ -60,7 +60,7 @@ def MAP_sample(data, samples, joker_params):
 
     ln_ps = np.zeros(len(samples))
 
-    mcmc_p = np.squeeze(model.pack_samples_mcmc(samples))
+    mcmc_p = model.pack_samples_mcmc(samples)        
     for i in range(len(samples)):
         ln_ps[i] = model.ln_posterior(mcmc_p[i])
 

### 2018-03-07 twoface/ext/__init__.py
read hat LCs

@@ -1,0 +1,1 @@
+from .hatlc import read_and_filter_sqlitecurve

### 2018-03-07 twoface/ext/hatlc.py

@@ -1,0 +1,1518 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+# From: https://github.com/waqasbhatti/astrobase/
+
+'''hatlc.py - Waqas Bhatti (wbhatti@astro.princeton.edu) - Jan 2016
+License: MIT - see LICENSE for the full text.
+
+This contains functions to read HAT sqlite ("sqlitecurves") and CSV light curves
+generated by the new HAT data server.
+
+The most probably useful functions in this module are:
+
+read_csvlc(lcfile):
+    This reads the HAT data server producd CSV light curve into an lcdict.
+
+    lcfile is the HAT gzipped CSV LC (with a .hatlc.csv.gz extension)
+
+
+read_and_filter_sqlitecurve(lcfile, columns=None, sqlfilters=None,
+                            raiseonfail=False, forcerecompress=False):
+    This reads the sqlitecurve and optionally filters it, returns an lcdict.
+
+    Returns columns requested in columns. If None, then returns all columns
+    present in the latest columnlist in the lightcurve. See COLUMNDEFS for the
+    full list of HAT LC columns.
+
+    If sqlfilters is not None, it must be a list of text sql filters that apply
+    to the columns in the lightcurve.
+
+    This returns an lcdict with an added 'lcfiltersql' key that indicates what
+    the parsed SQL filter string was.
+
+    If forcerecompress = True, will recompress the un-gzipped sqlitecurve even
+    if the gzipped form exists on disk already.
+
+
+Two other functions that might be useful:
+
+normalize_lcdict(lcdict, timecol='rjd', magcols='all', mingap=4.0,
+                 normto='sdssr', debugmode=False):
+
+    This normalizes magnitude columns (specified in the magcols keyword
+    argument) in an lcdict obtained from reading a HAT light curve. This
+    normalization is done by finding 'timegroups' in each magnitude column,
+    assuming that these belong to different 'eras' separated by a specified gap
+    in the mingap keyword argument, and thus may be offset vertically from one
+    another. Measurements within a timegroup are normalized to zero using the
+    meidan magnitude of the timegroup. Once all timegroups have been processed
+    this way, the whole time series is then re-normalized to the specified value
+    in the normto keyword argument.
+
+
+normalize_lcdict_byinst(lcdict, magcols='all', normto='sdssr', debugmode=False)
+
[…]

### 2018-03-07 paper/1-catalog/twoface-catalog.tex
fix troup names and tightness
new table

@@
 % ----------
-% - write a first draft of the introduction
-% - list our assumptions for detection & characterization (TheJoker)
-% - Check abstract numbers (~80,000 RG stars with 3 or more good visits?)
+% - Make sure we don't use $K$ as a number of samples - that's $M$ - because $K$
+%   is the velocity semi-amplitude!
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{lnK-percentiles}
+\includegraphics[width=0.8\textwidth]{lnK-percentiles}
 \end{center}
@@
 
-\subsection{Confident companions with highly constrained orbits}
-\label{sec:catalog-multimodal}
-
-\todo{How do we actually select these?}
-
-A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
-multimodal posterior distributions over orbital parameters that are limited to
-a few qualitatively different solutions.
-For these stars, one or a few more well-timed RV measurements would likely lead
-to uniquely-determined orbital parameters.
-\figurename~\ref{fig:todo} shows a few examples of stars that meet these
-criteria.
-These stars are prime candidates for follow-up efforts.
-In many of these cases, there are certain times at which a future observation
-would be far more informative.
-Visually, these times correspond to regions of time when the predicted RVs have
-large variance as computed for orbits computed from the posterior samplings.
-For example, \todo{top panel, time X vs. time Y, node vs. large variance}.
-
-\tablename~\ref{tbl:todo} contains ...
-\todo{Describe catalog contents: Also include M2min and primary mass when
-possible?}
-
-% % Notebook: TODO
-% \begin{figure}[hp]
-% \begin{center}
-% \includegraphics[width=\textwidth]{highK-unimodal}
-% \end{center}
-% \caption{%
-% TODO
-% \label{fig:highK-unimodal}
-% }
-% \end{figure}
-
 \subsection{Confident companions with effectively uniquely determined orbits}
@@
 star.
-We consider a sampling to be unimodal when $P_{\textrm{max}} - P_{\textrm{min}}
-< \Delta$ (see \sectionname~2 of \citealt{Price-Whelan:2017}); \nunimodal\ stars
+We consider a sampling to be unimodal when
+\begin{equation}
+    P_{\textrm{max}} - P_{\textrm{min}} < \Delta \label{eq:unimodal}
+\end{equation}
+(see \sectionname~2 of \citealt{Price-Whelan:2017}); \nunimodal\ stars
 pass this cut.
@@
 \end{figure}
+
+\subsection{Confident companions with highly constrained orbits}
+\label{sec:catalog-multimodal}
+
+A subset of the high-$K$ sample (\sectionname~\ref{sec:catalog-confident}) have
+multimodal posterior distributions over orbital parameters that are limited to
+a few qualitatively different solutions.
+For these stars, one or a few more RV measurements would likely lead to
+uniquely-determined orbital parameters; These stars are prime candidates for
+follow-up efforts.
+As examples of such cases, we include an additional catalog of systems that have
+effectively bimodal posterior samplings in orbital period.
+We identify these systems using $k$-means clustering (\citealt{Lloyd:1982}) of
+the posterior samples.
+In detail, for all stars in the high-$K$ sample, we compute $\ln P$ for all
+period samples, then use $k$-means clustering with $k=2$ as implemented in the
+\texttt{scikit-learn} package (\citealt{Pedregosa:2011}) to identify two
+clusters of samples.
+For each cluster, we then ask whether the samples assigned to that cluster are
+unimodal (\eqname~\ref{eq:unimodal}): If the samples in each respective cluster
+are unimodal, we call the sampling ``bimodal.''
+
+\figurename~\ref{fig:highK-bimodal} shows a few examples of stars that meet
+this criterion.
+In many of these cases, there are certain times at which a future observation
+would be far more informative.
+Visually, those times correspond to regions of time when the predicted RVs have
+large variance based on the posterior samples at hand.
+% For example, \todo{top panel, time X vs. time Y, node vs. large variance}.
+
+\tablename~\ref{fig:highK-bimodal} contains ...
+\todo{Describe catalog contents: Also include M2min and primary mass when
+possible?}
+
+% Notebook: HighK-multimodal
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-bimodal}
+\end{center}
+\caption{%
+TODO
+\label{fig:highK-bimodal}
+}
+\end{figure}
+
+\begin{table}
+    \centering
+    \begin{tabular}{c | c | c | c}
+    \texttt{APOGEE\_ID} & \texttt{P} & \texttt{e} & \texttt{K} \\
+    & $\mathrm{d}$ & & $\mathrm{km\,s^{-1}}$ \\
+    \hline
+    2M21362657-0017579 & 14.639 & 0.1171 & 30.68 \\
+    2M21362657-0017579 & 2.3049 & 0.0026 & 28.69 \\
+    2M03180303-0004215 & 35.553 & 0.1302 & 4.541 \\
+    2M03180303-0004215 & 107.22 & 0.3192 & 8.792 \\
+    ... & ... & ...
+    \end{tabular}
+    \caption{\todo{copy pasta}This table contains \texttt{APOGEE\_ID}'s for all \nstars\ stars
+    in the parent sample used in this work, for which we have posterior
+    samplings in orbital parameters.
+    The other column contains the 1st percentile computed over the $\ln K$
+    samples for each source, \texttt{lnK\_per\_1}:
+    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
+    defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.
+    }
+    \label{tbl:lnK-per}
+\end{table}
+
+\todo{We could instead select stars with log(Pmax) - log(Pmin) < 1, or some other metric to select things that have constrained periods but are multimodal}
 
@@
 \todo{The companions that don't make sense / can't be}
+
+\todo{Semi-periodic variables, etc.}
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{unimodal.pdf}
+\includegraphics[width=\textwidth]{troup-unimodal.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{multimodal.pdf}
+\includegraphics[width=\textwidth]{troup-multimodal.pdf}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\textwidth]{data-changed.pdf}
+\includegraphics[width=\textwidth]{troup-data-changed.pdf}
 \end{center}

### 2018-03-07 twoface/sample_prior.py
whitespace

@@ -3,6 +3,7 @@ import h5py
 import numpy as np
 from thejoker.sampler import pack_prior_samples
 
+
 def make_prior_cache(filename, thejoker, N, max_batch_size=2**24):
     """
 

### 2018-03-08 paper/1-catalog/twoface-catalog.tex
update figure titles
Big figures to end and table caption

@@
 \newcommand{\nhighK}{4,510}
-\newcommand{\nmultimodal}{XX}
+\newcommand{\nbimodal}{113}
 \newcommand{\nunimodal}{319}
@@
     2M00000446+5854329 & -7.199 \\
-    2M00000535+1504343 & -4.457 \\
-    2M00000546+6152107 & -4.388 \\
-    2M00000797+6436119 & -0.774 \\
-    2M00000866+7122144 & -3.313 \\
     ... & ...
@@
 
-% Notebook: "HighK"
-\begin{figure}[hp]
-\begin{center}
-\includegraphics[width=\textwidth]{highK-0}
-\end{center}
-\caption{%
-Examples of stars in our high-$K$ sample, i.e. stars that likely have companions
-(see \sectionname~\ref{sec:catalog-confident}), with different numbers of
-visits.
-Left panels show the data (black markers, error bars are the visit velocity
-uncertainties) with 128 random orbits from the \nposterior\ posterior samples
-under-plotted (lines, blue); The \texttt{APOGEE\_ID} of each target and the
-number of visits that pass our quality cuts, $N$, are indicated on each panel.
-Right panels show the \nposterior\ posterior samples visualized in
-period--eccentricity space.
-In most cases, despite confidently having companions based on the RV amplitude,
-the permitted orbit fits are highly multimodal.
-\label{fig:highK-0}
-}
-\end{figure}
-
-\begin{figure}[hp]
-\begin{center}
-\includegraphics[width=\textwidth]{highK-1}
-\end{center}
-\caption{%
-Continuation of \figurename~\ref{fig:highK-0}.
-\label{fig:highK-1}
-}
-\end{figure}
-
-\begin{figure}[hp]
-\begin{center}
-\includegraphics[width=\textwidth]{lowK-0}
-\end{center}
-\caption{%
-Same as \figurename~\ref{fig:highK-0}, but for stars in the low-$K$ sample, i.e.
-stars that have undetected or no companions.
-\label{fig:lowK-0}
-}
-\end{figure}
-
-\begin{figure}[hp]
-\begin{center}
-\includegraphics[width=\textwidth]{lowK-1}
-\end{center}
-\caption{%
-Continuation of \figurename~\ref{fig:lowK-0}.
-\label{fig:lowK-1}
-}
-\end{figure}
-
-\subsection{Confident companions with effectively uniquely determined orbits}
+\subsection{Confident companions with uniquely-determined orbits}
 \label{sec:catalog-unimodal}
@@
 columns in the high-$K$, unimodal sample catalog, available online.
-\todo{where?}
+\todo{Link to a webpage with the catalogs and info on how to use with
+TwoBody/TheJoker}
 
 \input{tables/unimodal-schema}
-
-% Notebook: HighK-unimodal-catalog
-\begin{figure}[hp]
-\begin{center}
-\includegraphics[width=\textwidth]{highK-unimodal}
-\end{center}
-\caption{%
-Randomly-selected examples of high-$K$ stars with effectively unimodal posterior
-samplings and converged MCMC samplings.
-Each panel shows the data (black markers), phase-folded at the period of the
-posterior sample with maximum posterior probability.
-Visit velocity uncertainties are shown as black error bars (these are typically
-smaller than or equal to the size of the markers), and the inferred jitter as
-grey, capped error bars.
-Line (blue) shows the orbit compute from the posterior sample with maximum
-posterior probability.
-\label{fig:highK-unimodal}
-}
-\end{figure}
 
@@
 We identify these systems using $k$-means clustering (\citealt{Lloyd:1982}) of
-the posterior samples.
+the posterior samples in period.
 In detail, for all stars in the high-$K$ sample, we compute $\ln P$ for all
@@
 clusters of samples.
+We initialize the cluster positions at either end of the range of possible
+periods.
 For each cluster, we then ask whether the samples assigned to that cluster are
-unimodal (\eqname~\ref{eq:unimodal}): If the samples in each respective cluster
-are unimodal, we call the sampling ``bimodal.''
-
-\figurename~\ref{fig:highK-bimodal} shows a few examples of stars that meet
+unimodal (\eqname~\ref{eq:unimodal}).
+If the samples in each respective cluster are unimodal, we call the sampling
+``bimodal''; \nbimodal\ systems are identified as bimodal.
+
+\figurename~\ref{fig:highK-bimodal} shows a few examples of systems that meet
 this criterion.
@@
 large variance based on the posterior samples at hand.
-% For example, \todo{top panel, time X vs. time Y, node vs. large variance}.
-
-\tablename~\ref{fig:highK-bimodal} contains ...
-\todo{Describe catalog contents: Also include M2min and primary mass when
-possible?}
-
-% Notebook: HighK-multimodal
-\begin{figure}[hp]
-\begin{center}
-\includegraphics[width=\textwidth]{highK-bimodal}
-\end{center}
-\caption{%
-TODO
-\label{fig:highK-bimodal}
-}
-\end{figure}
+For example, in the bottom left panel of \figurename~\ref{fig:highK-bimodal}, an
+observation at $t = 60$ would rule out one of the two possible orbits.
+
+\tablename~\ref{tbl:highK-bimodal} contains a list of all \apogee\ targets with
+bimodal samplings identified in this way.
+The table contains two rows for each source, with the period, eccentricity, and
+velocity semi-amplitude values from the MAP sample from each period mode.
 
@@
     \end{tabular}
-    \caption{\todo{copy pasta}This table contains \texttt{APOGEE\_ID}'s for all \nstars\ stars
-    in the parent sample used in this work, for which we have posterior
-    samplings in orbital parameters.
-    The other column contains the 1st percentile computed over the $\ln K$
-    samples for each source, \texttt{lnK\_per\_1}:
-    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
-    defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.
-    }
-    \label{tbl:lnK-per}
+    \caption{This table contains \texttt{APOGEE\_ID}'s and limited orbital
+    parameter information for all \nbimodal\ stars identified as having bimodal
+    posterior samplings in orbital period.
+    Each source is listed twice, with MAP values of period, \texttt{P},
+    eccentricity, \texttt{e}, and velocity semi-amplitude, \texttt{K}, from each
+    period mode of the posterior sampling for the source.
+    }
+    \label{tbl:highK-bimodal}
 \end{table}
@@
     \package{numpy} (\citealt{Van-der-Walt:2011}),
+    \package{scikit-learn} (\citealt{Pedregosa:2011}),
     \package{scipy} (\url{https://www.scipy.org/}),
@@
 
+\section{Example data and orbit samples}
+
+% Notebook: "HighK"
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-0}
+\end{center}
+\caption{%
+Examples of stars in our high-$K$ sample, i.e. stars that likely have companions
+(see \sectionname~\ref{sec:catalog-confident}), with different numbers of
+visits.
+Left panels show the data (black markers, error bars are the visit velocity
+uncertainties) with 128 random orbits from the \nposterior\ posterior samples
+under-plotted (lines, blue); The \texttt{APOGEE\_ID} of each target and the
+number of visits that pass our quality cuts, $N$, are indicated on each panel.
+Right panels show the \nposterior\ posterior samples visualized in
+period--eccentricity space.
+In most cases, despite confidently having companions based on the RV amplitude,
+the permitted orbit fits are highly multimodal.
+\label{fig:highK-0}
+}
+\end{figure}
+
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-1}
+\end{center}
+\caption{%
+Continuation of \figurename~\ref{fig:highK-0}.
+\label{fig:highK-1}
+}
+\end{figure}
+
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{lowK-0}
+\end{center}
+\caption{%
+Same as \figurename~\ref{fig:highK-0}, but for stars in the low-$K$ sample, i.e.
+stars that have undetected or no companions.
+\label{fig:lowK-0}
+}
+\end{figure}
+
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{lowK-1}
+\end{center}
+\caption{%
+Continuation of \figurename~\ref{fig:lowK-0}.
+\label{fig:lowK-1}
+}
+\end{figure}
+
+% Notebook: HighK-unimodal-catalog
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-unimodal}
+\end{center}
+\caption{%
+Randomly-selected examples of high-$K$ stars with effectively unimodal posterior
+samplings and converged MCMC samplings.
+Each panel shows the data (black markers), phase-folded at the period of the
+posterior sample with maximum posterior probability.
+Visit velocity uncertainties are shown as black error bars (these are typically
+smaller than or equal to the size of the markers), and the inferred jitter as
+grey, capped error bars.
+Line (blue) shows the orbit compute from the posterior sample with maximum
+posterior probability.
+\label{fig:highK-unimodal}
+}
+\end{figure}
+
+% Notebook: HighK-multimodal
+\begin{figure}[hp]
+\begin{center}
+\includegraphics[width=\textwidth]{highK-bimodal}
+\end{center}
+\caption{%
+TODO
+\label{fig:highK-bimodal}
+}
+\end{figure}
+
 \end{document}

### 2018-03-08 paper/1-catalog/twoface-catalog.tex
title and change control to comparison

@@
 
-\title{Binary companions of red giant stars in \apogee\ \DR}
+\title{Binary companions of red giant stars in \apogee\ \DR: \\
+       Search method and catalog of $>$4,500 systems}
 
@@
 
-\subsection{Null control sample}
+\subsection{Null comparison sample}
 \label{sec:control-sample}
 
-We construct a control sample of simulated data with no RV variability to assess
-our false-positive rates in the selections below (see
+We construct a comparison sample of simulated data with no RV variability to
+assess our false-positive rates in the selections below (see
 \sectionname~\ref{sec:catalogs}).
@@
 \end{equation}
-When we save the control data, we only store the visit uncertainty and
+When we save the comparison data, we only store the visit uncertainty and
 ``forget'' the fact that the simulated data is generated with excess variance.
@@
 
-We also run on the null control sample (see
+We also run on the null comparison sample (see
 \sectionname~\ref{sec:control-sample}) with the same parameters.
-From the control sample run, 13 ``stars'' are flagged as \emph{needs MCMC}, and
-1023 are flagged as ``needs more prior samples'';
-For these control sample stars, we don't continue sampling with \thejoker\ or
+From the comparison sample run, 13 ``stars'' are flagged as \emph{needs MCMC},
+and 1023 are flagged as ``needs more prior samples'';
+For these comparison sample stars, we don't continue sampling with \thejoker\ or
 MCMC and only use the ($<$\nposterior) posterior samples returned from the
@@
 (filled, blue histogram), along with the same for posterior samplings for the
-null control sample (solid, dark line).
+null comparison sample (solid, dark line).
 To select stars with companions, we use a cut in the 1st percentile of the $\ln
-K$ samples such that $<1\%$ of the control sample is selected, i.e. our
+K$ samples such that $<1\%$ of the comparison sample is selected, i.e. our
 estimated false positive rate is $<1\%$.
@@
 Distribution of 1st percentiles in $\ln K$ (in units of \kms) for the parent
-\apogee\ sample (blue, filled) and for the control sample (solid, dark line).
+\apogee\ sample (blue, filled) and for the comparison sample (solid, dark line).
 The vertical line (dashed, orange) indicates our adopted cut to select high-$K$
-stars that likely have companions; $<1\%$ of the control sample falls above this
-cut.
+stars that likely have companions; $<1\%$ of the comparison sample falls above
+this cut.
 \label{fig:lnK-percentiles}
@@
 
+% TODO:
+% Notebook: "Results"
+% \begin{figure}[h]
+% \begin{center}
+% \includegraphics[width=\textwidth]{infer-jitter}
+% \end{center}
+% \caption{%
+% Inferred prior over excess variance parameter ($y = \ln s^2$) using posterior
+% samples for 1,825 lower-RGB stars with $>10$ visits.
+% Left and right panels show the distribution corresponding to the
+% maximum-likelihood parameters $(\mu_y, \sigma_y) = (9.50, 1.64)$ in log and
+% linear, respectively.
+% \label{fig:infer-jitter}
+% }
+% \end{figure}
+
 \subsection{Red clump stars lack short-period companions}

### 2018-03-08 paper/1-catalog/twoface-catalog.tex
bimodal table

@@
     \centering
-    \begin{tabular}{c | c | c | c}
-    \texttt{APOGEE\_ID} & \texttt{P} & \texttt{e} & \texttt{K} \\
-    & $\mathrm{d}$ & & $\mathrm{km\,s^{-1}}$ \\
+    \begin{tabular}{c | c | c | c | c | c | c}
+    \texttt{APOGEE\_ID} & \texttt{P} & \texttt{e} & \texttt{K} &
+        \texttt{M1} & \texttt{M2\_min} & \texttt{clean\_flag} \\
+    & $\mathrm{d}$ & & $\mathrm{km\,s^{-1}}$ &
+        $\mathrm{M_{\odot}}$ & $\mathrm{M_{\odot}}$ &  \\
     \hline
-    2M21362657-0017579 & 14.639 & 0.1171 & 30.68 \\
-    2M21362657-0017579 & 2.3049 & 0.0026 & 28.69 \\
-    2M03180303-0004215 & 35.553 & 0.1302 & 4.541 \\
-    2M03180303-0004215 & 107.22 & 0.3192 & 8.792 \\
-    ... & ... & ...
+    2M21362657-0017579 & 14.639 & 0.1171 & 30.68 & -- & -- & 0 \\
+    2M21362657-0017579 & 2.3049 & 0.0026 & 28.69 & -- & -- & 0 \\
+    2M03180303-0004215 & 35.553 & 0.1302 & 4.541 & 1.20 & 0.20 & 0 \\
+    2M03180303-0004215 & 107.22 & 0.3192 & 8.792 & 1.20 & 0.13 & 0 \\
+    ... & ... & ... \\
+    \multicolumn{6}{c}{\textit{(210 rows)}}
     \end{tabular}
@@
     period mode of the posterior sampling for the source.
+    When available, this table also includes point-mass estimates of the primary
+    mass, \texttt{M1}, from \citet{Ness:2015}, and minimum companion masses,
+    \texttt{M2\_min}, computed for each period mode.
     }
@@
 
-% TODO:
+\subsection{Companion masses and mass ratios}
+
+\todo{Companion mass and mass-ratio distribution for stars with Ness / Martig masses}
+
+\todo{Interesting companions w/ large M2min, WD/NS/BH companions?}
+
+\todo{Plot: M1 vs. M2min}
+
+\todo{Plot: M2min vs. a2sini}
+
 % Notebook: "Results"
-% \begin{figure}[h]
-% \begin{center}
-% \includegraphics[width=\textwidth]{infer-jitter}
-% \end{center}
-% \caption{%
-% Inferred prior over excess variance parameter ($y = \ln s^2$) using posterior
-% samples for 1,825 lower-RGB stars with $>10$ visits.
-% Left and right panels show the distribution corresponding to the
-% maximum-likelihood parameters $(\mu_y, \sigma_y) = (9.50, 1.64)$ in log and
-% linear, respectively.
-% \label{fig:infer-jitter}
-% }
-% \end{figure}
-
-\subsection{Red clump stars lack short-period companions}
-
-\todo{Period-logg plot - lack of short-period companions for red clump stars?}
-
-\todo{also look at trends with Teff?}
-
-\todo{Cite troup, who also noted this}
-
-Red clump (\RC) stars are horizontal branch stars and have therefore already
-climbed the RGB, where the surface of the star can grow to be several AU in
-size.
-\RC\ stars are therefore expected to less frequently have companions at short
-periods ($P<XX$~day) because short period companions would have been
-common-envelope during the RGB phase.
-\figurename~\ref{fig:todo} hints that there is an apparent dearth of
-short-period companions around $\log g \sim 2.5$ (the approximate location of
-the red clump).
-Another way to look at this is to compare the fraction of subgiant stars with
-short-period companions to the same for stars near the \RC.
-
-\todo{To assess this, take boxes in log g-Teff}
-We find that, empirically, XX\% of the subgiants and YY\% of the \RC\ stars have
-companions with $XX < P < XX$~day.
-\todo{Assuming selection function doesn't preferentially bias...}
-
-\subsection{Evidence for tidal circularization at $P\sim 10~\textrm{day}$}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{mass}
+\end{center}
+\caption{%
+TODO
+\label{fig:mass}
+}
+\end{figure}
+
+\subsection{Evidence for tidal circularization}
 
@@
 \end{figure}
-
-\subsection{Companion masses and mass ratios}
-
-\todo{Companion mass and mass-ratio distribution for stars with Ness / Martig masses}
-
-\todo{Interesting companions w/ large M2min, WD/NS/BH companions?}
-
-\todo{Plot: M1 vs. M2min}
-
-\todo{Plot: M2min vs. a2sini}
 

### 2018-03-08 twoface/mass.py
debug not warn

@@ -103,8 +103,8 @@ def get_m2_min(m1, mf):
                     raise RuntimeError('Unsuccessful')
                 m2s.append(res.x[0])
             except Exception as e:
-                logger.warning('Failed to compute m2_min for sample: {0}'
-                               .format(str(e)))
+                logger.debug('Failed to compute m2_min for sample: {0}'
+                             .format(str(e)))
                 m2s.append(np.nan)
         return m2s * m1.unit
 

### 2018-03-08 paper/1-catalog/twoface-catalog.tex
some tweaks

@@
 
+We have visually inspected the inferred orbits for all systems in this sample
+and have flagged systems with questionable or invalid fits.
+The value of the flag is: 0, when the orbits look reasonable, 1, when the orbits
+look reasonable but the value of the inferred excess variance is large, and 2,
+when the orbits are clearly poor fits.
+Many of the stars flagged as ``2'' look like they may be triple systems, as they
+tend to have radial velocity variations over two distinct timescales that are
+never well-fit by a single Keplerian orbit.
+This flag is included in the catalog (\tablename~\ref{tbl:highK-unimodal}) as
+the column \texttt{clean\_flag}; To select a clean sample of companions that
+have been successfully vetted by-eye, select only stars with
+\texttt{clean\_flag == 0}.
+
 \input{tables/unimodal-schema}
@@
 For each cluster, we then ask whether the samples assigned to that cluster are
-unimodal (\eqname~\ref{eq:unimodal}).
+unimodal (\eqname~\ref{eq:catalog-unimodal}).
 If the samples in each respective cluster are unimodal, we call the sampling
@@
 velocity semi-amplitude values from the MAP sample from each period mode.
+When available, this table also includes an estimate of the primary mass, $M_1$,
+from \citet{Ness:2015} and an estimate of the minimum companion mass, $M_{2,
+\textrm{min}}$.
+
+We again visually inspect all systems in this sample and assign a flag based on
+the apparent quality of the inferred orbits (see
+\sectionname~\ref{sec:unimodal}).
+Again, to select a clean sample of companions from this catalog that have been
+successfully vetted by-eye, select only stars with \texttt{clean\_flag == 0}.
 
@@
 
-\todo{We could instead select stars with log(Pmax) - log(Pmin) < 1, or some other metric to select things that have constrained periods but are multimodal}
+% Note: We could instead select stars with log(Pmax) - log(Pmin) < 1, or some
+% other metric to select things that have constrained periods but are
+% multimodal. This would get us more stars, but may be harder to interpret...
 
@@
 
+Here we highlight a few interesting results from visualizing the properties of
+companions in the unimodal and bimodal samples defined above (see \sectionname
+s~\ref{sec:catalog-unimodal} and \ref{sec:catalog-bimodal}).
+In all figures, we only plot companions with inferred orbits that pass our
+visual inspection (\texttt{clean\_flag == 0}).
+We additionally restrict to stars with $\log g > 2$ to avoid upper RGB stars
+that may have coherent oscillations that mimic RV modulations from companions.
+
 \subsection{Companion masses and mass ratios}
 
-\todo{Companion mass and mass-ratio distribution for stars with Ness / Martig masses}
-
-\todo{Interesting companions w/ large M2min, WD/NS/BH companions?}
-
-\todo{Plot: M1 vs. M2min}
-
-\todo{Plot: M2min vs. a2sini}
+Most of the companions we find have minimum masses $M_{2, \textrm{min}} <
+1~\Msun$.
+\figurename~\ref{fig:mass}, left panel shows primary and companion mass
+estimates for systems with unimodal (black markers) and bimodal (blue lines)
+period samplings using primary masses from \citet{Ness:2015}.
+Upper dashed line shows the $M_{2, \textrm{min}} = M_1$ line, and lower dashed
+line shows the Hydrogen-burning limit, $M_{2, \textrm{min}} = 0.08~\msun$.
+For the systems with bimodal samplings, we compute the minimum companion mass
+for each period mode and connect these estimates with a line.
+Three interesting systems are immediately obvious: two systems with bimodal
+period samplings in which both period modes put the minimum companion mass below
+the Hydrogen-burning limit, and one system in which $M_{2, \textrm{min}} > M_1$.
 
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{mass}
+\includegraphics[width=\textwidth]{mass}
 \end{center}
@@
 
-\subsection{Evidence for tidal circularization}
+\subsubsection{2M04582049+2232000}
+% BD candidate
+
+\subsubsection{2M20130687+2044173}
+% BD candidate
+
+\subsubsection{2M01231070+1801407}
+% NS/BH candidate
+
+\subsection{Tidal circularization}
 

### 2018-03-09 paper/1-catalog/twoface-catalog.tex
some words and captions

@@
 
-\begin{table}
-    \centering
-    \begin{tabular}{c | c}
-    \texttt{APOGEE\_ID} & \texttt{lnK\_per\_1} \\
-    \hline
-    2M00000002+7417074 & -2.028 \\
-    2M00000068+5710233 & -2.600 \\
-    2M00000222+5625359 &  2.134 \\
-    2M00000446+5854329 & -7.199 \\
-    ... & ...
-    \end{tabular}
-    \caption{This table contains \texttt{APOGEE\_ID}'s for all \nstars\ stars
-    in the parent sample used in this work, for which we have posterior
-    samplings in orbital parameters.
-    The other column contains the 1st percentile computed over the $\ln K$
-    samples for each source, \texttt{lnK\_per\_1}:
-    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
-    defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.
-    }
-    \label{tbl:lnK-per}
-\end{table}
-
 \tablename~\ref{tbl:lnK-per} contains a list of all stars in the parent sample
@@
 
-\input{tables/unimodal-schema}
-
 \subsection{Confident companions with highly constrained orbits}
@@
 
-\begin{table}
-    \centering
-    \begin{tabular}{c | c | c | c | c | c | c}
-    \texttt{APOGEE\_ID} & \texttt{P} & \texttt{e} & \texttt{K} &
-        \texttt{M1} & \texttt{M2\_min} & \texttt{clean\_flag} \\
-    & $\mathrm{d}$ & & $\mathrm{km\,s^{-1}}$ &
-        $\mathrm{M_{\odot}}$ & $\mathrm{M_{\odot}}$ &  \\
-    \hline
-    2M21362657-0017579 & 14.639 & 0.1171 & 30.68 & -- & -- & 0 \\
-    2M21362657-0017579 & 2.3049 & 0.0026 & 28.69 & -- & -- & 0 \\
-    2M03180303-0004215 & 35.553 & 0.1302 & 4.541 & 1.20 & 0.20 & 0 \\
-    2M03180303-0004215 & 107.22 & 0.3192 & 8.792 & 1.20 & 0.13 & 0 \\
-    ... & ... & ... \\
-    \multicolumn{6}{c}{\textit{(210 rows)}}
-    \end{tabular}
-    \caption{This table contains \texttt{APOGEE\_ID}'s and limited orbital
-    parameter information for all \nbimodal\ stars identified as having bimodal
-    posterior samplings in orbital period.
-    Each source is listed twice, with MAP values of period, \texttt{P},
-    eccentricity, \texttt{e}, and velocity semi-amplitude, \texttt{K}, from each
-    period mode of the posterior sampling for the source.
-    When available, this table also includes point-mass estimates of the primary
-    mass, \texttt{M1}, from \citet{Ness:2015}, and minimum companion masses,
-    \texttt{M2\_min}, computed for each period mode.
-    }
-    \label{tbl:highK-bimodal}
-\end{table}
-
 % Note: We could instead select stars with log(Pmax) - log(Pmin) < 1, or some
@@
 Most of the companions we find have minimum masses $M_{2, \textrm{min}} <
-1~\Msun$.
+1~\msun$.
 \figurename~\ref{fig:mass}, left panel shows primary and companion mass
 estimates for systems with unimodal (black markers) and bimodal (blue lines)
-period samplings using primary masses from \citet{Ness:2015}.
-Upper dashed line shows the $M_{2, \textrm{min}} = M_1$ line, and lower dashed
+period samplings, again using primary masses from \citet{Ness:2015}.
+Upper dashed line shows the $M_{2, \textrm{min}} = M_1$ curve, and lower dashed
 line shows the Hydrogen-burning limit, $M_{2, \textrm{min}} = 0.08~\msun$.
@@
 for each period mode and connect these estimates with a line.
-Three interesting systems are immediately obvious: two systems with bimodal
-period samplings in which both period modes put the minimum companion mass below
-the Hydrogen-burning limit, and one system in which $M_{2, \textrm{min}} > M_1$.
+
+\figurename~\ref{fig:mass}, right panel shows the ratio of the primary stellar
+radius to the projected separation of the two bodies, $R_1 / (a\,\sin i)$, and
+the minimum mass ratio, $q_{\textrm{min}} = M_{2, \textrm{min}} / M_1$, for the
+same samples.
+Again, black large markers indicate systems from the unimodal sample, and blue
+markers and lines connect the two estimates for systems with bimodal period
+samplings.
+Upper dashed line in this panel indicates the Roche radius, computed using an approximate functional form (\citealt{Eggleton:1983})
+\begin{equation}
+    \frac{R}{a} = \frac{0.49\,q^{-2/3}}{0.6\,q^{-2/3} +
+        \ln\left(1 + q^{-1/3}\right)} \quad .
+\end{equation}
+All of these systems are generally consistent with being detached binaries.
+
+In \figurename~\ref{fig:mass}, three interesting systems are immediately
+obvious: two systems with bimodal period samplings in which both period modes
+put the minimum companion mass below the Hydrogen-burning limit, and one system
+in which $M_{2, \textrm{min}} > M_1$; We discuss these systems below
 
@@
 \caption{%
-TODO
+\textit{Left panel:} Estimated minimum companion mass and primary mass for
+systems in the high-$K$, unimodal sample (black markers) and bimodal sample
+(blue markers and lines).
+Lines connecting markers (blue) show estimates from each period mode for each
+system with bimodal period samplings.
+Upper, dashed line shows the equal-mass curve, and lower dashed line shows the
+Hydrogen-burning limit.
+\textit{Right panel:} Ratio of primary stellar radius, $R_1$, to projected
+separation of the two bodies, $a\,\sin i$, and minimum mass ratio,
+$q_{\textrm{min}}$, again for systems in the high-$K$, unimodal sample
+(black markers) and bimodal sample (blue markers and lines).
 \label{fig:mass}
@@
 
-\subsubsection{2M04582049+2232000}
+\subsubsection{2M04582049+2232000 -- brown dwarf companion candidate}
 % BD candidate
 
-\subsubsection{2M20130687+2044173}
+\todo{What to show?}
+
+\subsubsection{2M20130687+2044173 -- brown dwarf companion candidate}
 % BD candidate
 
-\subsubsection{2M01231070+1801407}
+\todo{What to show?}
+
+\subsubsection{2M01231070+1801407 -- black hole companion candidate}
 % NS/BH candidate
 
+\todo{What to show?}
+
 \subsection{Tidal circularization}
@@
 
-\figurename~\ref{fig:unimodalPe} shows the orbital period and eccentricity...
-
-\todo{Period - eccentricity plot? Tidal circularization (See "Debug MCMC
-continue")}
-
-\todo{eccentricity distribution w/ prior over-plotted}
-
-% Notebook: HighK-unimodal
+\figurename~\ref{fig:PeK}, left panel, shows the normalized orbital period and
+eccentricity for all systems in the high-$K$, unimodal sample, and, right panel,
+the RV semi-amplitude.
+We normalize the inferred periods by the orbital period of a companion at the
+surface of the primary, $P_{\textrm{surface}}$:
+To do this for all systems in this sample, we assume that all primary stars have
+masses equal to the median mass over all stars with prior mass measurements in
+this sample, $M_1 = 1.36~\msun$, and assume a companion mass also equal to the
+median over minimum companion masses $M_{2, \textrm{min}} = 0.5~\msun$.
+We then use the surface gravity, $\log g$, of the primary, and eccentricity to
+compute
+\begin{align}
+    P_{\textrm{surface}} &= \frac{2\pi \, G^{1/4}}{(1-e)^{-3/2}} \,
+        \left(\frac{M_1 + M_2}{M_1^{3/4}}\right) \,
+        10^{-3/4\,\log g} \\
+    &= XX ...
+\end{align}
+where we use $M_{2, \textrm{min}}$ as $M_2$.
+There is a clear signature of circularization just under $P/P_{\textrm{surface}}
+\approx 10$.
+\todo{why is this novel with a large sample of giant stars}
+
+% Notebook: Results
 \begin{figure}[h]
 \begin{center}
-\includegraphics[width=0.7\textwidth]{unimodal-P-e}
+\includegraphics[width=\textwidth]{P-e-K}
 \end{center}
@@
 TODO:
-\label{fig:unimodalPe}
-}
-\end{figure}
-
-\subsection{Eclipsing}
-
-\todo{Stars with Kepler light curves}
+\label{fig:PeK}
+}
+\end{figure}
+
+% Notebook: Results
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{ecc-dist}
+\end{center}
+\caption{%
+TODO:
+\label{fig:ecc}
+}
+\end{figure}
+
+% \subsection{Eclipsing}
+%
+% \todo{Stars with Kepler light curves}
 
@@
 
-\todo{comment on short period shit, eccentricities}
+\todo{comment on short period shit at low logg, eccentricities}
 
@@
 
-\section{Example data and orbit samples}
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{c | c}
+    \multicolumn{2}{c}{\textbf{Parent sample of \apogee\ \DR\ stars}}\\
+    \hline
+    \texttt{APOGEE\_ID} & \texttt{lnK\_per\_1} \\
+    \hline
+    2M00000002+7417074 & -2.028 \\
+    2M00000068+5710233 & -2.600 \\
+    2M00000222+5625359 &  2.134 \\
+    2M00000446+5854329 & -7.199 \\
+    ... & ... \\
+    \hline
+    \multicolumn{2}{c}{\textit{(\nstars\ rows)}}
+    \end{tabular}
+    \caption{This table contains \texttt{APOGEE\_ID}'s for all \nstars\ stars
+    in the parent sample used in this work, for which we have posterior
+    samplings in orbital parameters.
+    The other column contains the 1st percentile computed over the $\ln K$
+    samples for each source, \texttt{lnK\_per\_1}:
+    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
+    defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.
+    }
+    \label{tbl:lnK-per}
+\end{table}
+
+\input{tables/unimodal-schema}
+
+\begin{table}[ht]
+    \centering
+    \begin{tabular}{c | c | c | c | c | c | c}
+    \multicolumn{7}{c}{\textbf{High-$K$, bimodal systems}}\\
+    \hline
+    \texttt{APOGEE\_ID} & \texttt{P} & \texttt{e} & \texttt{K} &
+        \texttt{M1} & \texttt{M2\_min} & \texttt{clean\_flag} \\
+    & $\mathrm{d}$ & & $\mathrm{km\,s^{-1}}$ &
+        $\mathrm{M_{\odot}}$ & $\mathrm{M_{\odot}}$ &  \\
+    \hline
+    2M21362657-0017579 & 14.639 & 0.1171 & 30.68 & -- & -- & 0 \\
+    2M21362657-0017579 & 2.3049 & 0.0026 & 28.69 & -- & -- & 0 \\
+    2M03180303-0004215 & 35.553 & 0.1302 & 4.541 & 1.20 & 0.20 & 0 \\
+    2M03180303-0004215 & 107.22 & 0.3192 & 8.792 & 1.20 & 0.13 & 0 \\
+    ... & ... & ... \\
+    \hline
+    \multicolumn{6}{c}{\textit{(210 rows)}}
+    \end{tabular}
+    \caption{This table contains \texttt{APOGEE\_ID}'s and limited orbital
+    parameter information for all \nbimodal\ stars identified as having bimodal
+    posterior samplings in orbital period.
+    Each source is listed twice, with MAP values of period, \texttt{P},
+    eccentricity, \texttt{e}, and velocity semi-amplitude, \texttt{K}, from each
+    period mode of the posterior sampling for the source.
+    When available, this table also includes point-mass estimates of the primary
+    mass, \texttt{M1}, from \citet{Ness:2015}, and minimum companion masses,
+    \texttt{M2\_min}, computed for each period mode.
+    }
+    \label{tbl:highK-bimodal}
+\end{table}
+
+\clearpage
 

### 2018-03-10 paper/1-catalog/twoface-catalog.tex
Captions and intro

@@
 %   is the velocity semi-amplitude!
+% - Figure out what to say about upper RGB stars / coherent oscillations!
 
@@
 \newcommand{\DR}{\acronym{DR14}}
+\newcommand{\DRtw}{\acronym{DR12}}
 \newcommand{\RC}{\acronym{RC}}
@@
 nearby main-sequence (MS) stars.
+\todo{occurrence rates}
 MS stars with companions have a large dynamic range of constituent and orbital
@@
 millions of years (e.g., \citealt{Raghavan:2010}).
-\todo{multiplicity in general: occurrence rates}
-
-Less is known about non-interacting or detached companions to evolved stars.
-Such systems are interesting because they enable studying extreme outcomes or
-conditions of star formation and multiplicity.
-For example, evolved stars are typically much brighter than their MS
-counterparts, which enables characterizing the stellar companion population
-throughout the Galactic disk over a wide range of chemical abundances.
-They also provide a means to study the outcomes of stellar engulfment or
-common-envelope evolution: Do short-period companions exist around evolved
-stars, and do their statistics constrain stellar evolution on the red giant
-branch (RGB)?
-\todo{motivate time-domain spectroscopy}
-
+Less is known about population properties of non-interacting or detached
+companions to evolved stars.
+
+Large spectroscopic surveys with repeat observations offer a novel way to study
+binary star and companion populations.
 Time-domain radial-velocity measurements of stars contain information about
-massive companions: even with two successive observations of a single star, a
+massive companions: Even with two successive observations of a single star, a
 difference in the measured radial velocities implies the existence of at least
@@
 \figurename~\ref{fig:nvisits} shows the number of stars in several bins of
-number of visits that pass the above quality cuts: \todo{XX}\% of the stars in
-\DR\ have $< 8$ visits.
-\figurename~\ref{fig:loggteff} ...
+number of visits that pass the above quality cuts: \todo{92}\% of the stars in
+our parent sample have $< 8$ visits.
+\figurename~\ref{fig:loggteff} shows the stellar parameters of all \nstars\
+stars in the parent sample used in this work.
 
@@
   hierarchical so that the period of the inner binary is typically much shorter
-  than the orbital period of the outer body (\todo{CITATION NEEDED}).
+  than the orbital period of the outer body (e.g., \citealt{Tokovinin:2018}).
   At present, we ignore the possibility of higher-order multiple systems.
@@
 never well-fit by a single Keplerian orbit.
+The stars flagged as ``1'' tend to be either upper RGB stars, where
+astrophysical surface jitter can lead to RV modulations, or SB2 systems, where
+absorption lines from the secondary confuse the RV pipeline and lead to strange
+RV signals.
 This flag is included in the catalog (\tablename~\ref{tbl:highK-unimodal}) as
@@
 the apparent quality of the inferred orbits (see
-\sectionname~\ref{sec:unimodal}).
+\sectionname~\ref{sec:catalog-unimodal}).
 Again, to select a clean sample of companions from this catalog that have been
@@
 companions in the unimodal and bimodal samples defined above (see \sectionname
-s~\ref{sec:catalog-unimodal} and \ref{sec:catalog-bimodal}).
+s~\ref{sec:catalog-unimodal} and \ref{sec:catalog-multimodal}).
 In all figures, we only plot companions with inferred orbits that pass our
@@
 obvious: two systems with bimodal period samplings in which both period modes
-put the minimum companion mass below the Hydrogen-burning limit, and one system
-in which $M_{2, \textrm{min}} > M_1$; We discuss these systems below
+put the minimum companion mass below the Hydrogen-burning limit (brown dwarf
+candidates), and one system in which $M_{2, \textrm{min}} > M_1$ (a neutron star
+or black hole candidate).
+These systems are being followed-up and will be studied in future work.
 
@@
 
-\subsubsection{2M04582049+2232000 -- brown dwarf companion candidate}
-% BD candidate
-
-\todo{What to show?}
-
-\subsubsection{2M20130687+2044173 -- brown dwarf companion candidate}
-% BD candidate
-
-\todo{What to show?}
-
-\subsubsection{2M01231070+1801407 -- black hole companion candidate}
-% NS/BH candidate
-
-\todo{What to show?}
-
 \subsection{Tidal circularization}
 
-\todo{Expectation from theory? What's special at 10 days?}
-
-\todo{Mathieu et al. 2004}
-
-\figurename~\ref{fig:PeK}, left panel, shows the normalized orbital period and
-eccentricity for all systems in the high-$K$, unimodal sample, and, right panel,
-the RV semi-amplitude.
-We normalize the inferred periods by the orbital period of a companion at the
-surface of the primary, $P_{\textrm{surface}}$:
+From studies of binary star systems in open clusters, it is clear that
+short-period binaries tend to be more circular (have smaller eccentricities) as
+compared to longer-period binaries of the same age (e.g.,
+\citealt{Mathieu:2005}).
+Within a given population of binaries, this manifests as an apparently steep
+transition from a distribution of eccentricities to mainly circularized orbits
+at some characteristic period that depends on the age and evolutionary state of
+the population (see, e.g., \figurename~5 in \citealt{Mathieu:2005})---i.e. above
+this critical value of the period (or log-period), systems have some
+distribution of eccentricities, and below the eccentricities are close to zero.
+Many studies have measured this ``circularization period,'' predominantly for
+main-sequence binaries (e.g., \citealt{Meibom:2006, Kjurkchieva:2017}), and
+have found that it tends to be between 5--20 days, depending on age (e.g.,
+\citealt{Mathieu:1988}).
+For giant stars, the circularization period appears to be longer, closer to
+$\approx 100$ days (\citealt{Mayor:1984}).
+
+Theories for the tidal circularization process generally rely on tidal
+dissipation via eddy viscosity in the convective regions of stars to explain
+this (e.g., \citealt{Zahn:1989}).
+The circularization timescale and period should therefore depend on the size of
+the convective region, and the stellar parameters of the primary star.
+RGB stars are therefore excellent test-cases for these theories, since
+they are fully-convective:
+By comparing to (small) samples of giant star binaries in open clusters, it
+seems that the circularization timescales predicted by eddy viscosity
+dissipation at least qualitatively scale to giant stars
+(\citealt{Verbunt:1995}).
+
+It is less clear what to expect for red clump (\RC) stars:
+RC stars are horizontal branch stars and have therefore already climbed the RGB.
+Short period, circularized companions in place before or while on the RGB could
+be engulfed, or will at least have been common-envelope with the evolving star.
+However, it is difficult to select samples of \RC\ stars to study this process
+because first-ascent RGB stars will have similar colors and stellar parameters.
+\citet{Verbunt:1995} argue that companion orbits may be a way to separate the
+two overlapping populations: eccentric, short-period orbits around stars near
+the red-clump should be first-ascent giants.
+
+The samples produced in this work provide excellent datasets to study tidal
+circularization of companions to giant stars across a range of ages and
+metallicities
+However, our samples consist of a mix of subgiants, giants, and RC stars which
+have distinct distributions in mass and stellar radius.
+To visualize the distribution of systems in period--eccentricity space, we have
+therefore computed normalized periods, where we normalize by the orbital period
+of a typical companion on a circular orbit at the surface of the primary,
+$P_{\textrm{surface}}$.
 To do this for all systems in this sample, we assume that all primary stars have
@@
         \left(\frac{M_1 + M_2}{M_1^{3/4}}\right) \,
-        10^{-3/4\,\log g} \\
-    &= XX ...
+        10^{-3/4\,\log g}
 \end{align}
 where we use $M_{2, \textrm{min}}$ as $M_2$.
+
+\figurename~\ref{fig:PeK}, left panel, shows this normalized orbital period and
+eccentricity for all systems in the high-$K$, unimodal sample, and, right panel,
+the RV semi-amplitude.
 There is a clear signature of circularization just under $P/P_{\textrm{surface}}
 \approx 10$.
-\todo{why is this novel with a large sample of giant stars}
+\todo{what else can or should we say?}
 
@@
 \caption{%
-TODO:
+\textit{Left panel:} Orbital period and eccentricity for all high-$K$ unimodal
+systems with $\log g > 2$, with period values normalized by the orbital period
+of a hypothetical companion at the surface of each primary star,
+$P_{\textrm{surface}}$, assuming $M_1 = 1.36~\msun$ and $M_2 =0.5~\msun$.
+Stars in the \apogee\ \DR\ red clump catalog are outlined in red.
+The sharp transition from eccentric systems to almost all circular orbits at
+$P/P_\textrm{surface} \approx 7$ is likely an outcome of tidal circularization.
+\textit{Right panel:} Normalized period and inferred radial velocity amplitude
+for the same systems.
+Points below $P/P_\textrm{surface} = 1$ are likely systems where the primary and
+companion mass assumptions used to compute $P_\textrm{surface}$ are invalid.
 \label{fig:PeK}
@@
 % Notebook: Results
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{ecc-dist}
-\end{center}
-\caption{%
-TODO:
-\label{fig:ecc}
-}
-\end{figure}
+% \begin{figure}[h]
+% \begin{center}
+% \includegraphics[width=0.7\textwidth]{ecc-dist}
+% \end{center}
+% \caption{%
+% TODO:
+% \label{fig:ecc}
+% }
+% \end{figure}
 
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{residuals}
+\includegraphics[width=0.6\textwidth]{residuals}
 \end{center}
@@
 \end{figure}
-
-\subsection{Impossible companions: coherent oscillations?}
-
-\todo{The companions that don't make sense / can't be}
-
-\todo{Semi-periodic variables, etc.}
 
@@
 The other recent catalog searched for stellar and substellar companions to all
-stars in \apogee\ \acronym{DR12} (including the RGB) that passed a series of
-quality cuts, and had $\geq 8$ visits (\citealt{Troup:2016}).
+stars in \apogee\ \DRtw\ (including the RGB) that passed a series of quality
+cuts, and had $\geq 8$ visits (\citealt{Troup:2016}).
 For each star in the sample, orbits were fit to the visit RVs using a multi-step
@@
 a likely multi-modal posterior pdf over orbital parameters, and
-(3) the data changed significantly between \apogee\ \acronym{DR12} and \DR, so
+(3) the data changed significantly between \apogee\ \DRtw\ and \DR, so
 no meaningful comparisons can be made; Roughly 1/3 of the comparison sample
@@
 
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{troup-unimodal.pdf}
-\end{center}
-\caption{%
-TODO:
-\label{fig:troup-unimodal}
-}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{troup-multimodal.pdf}
-\end{center}
-\caption{%
-TODO:
-\label{fig:troup-multimodal}
-}
-\end{figure}
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{troup-data-changed.pdf}
-\end{center}
-\caption{%
-TODO:
-\label{fig:troup-datachanged}
-}
-\end{figure}
-
 \subsection{Population inference}
 
-Comment on things we could do with these samples hierarchical inference...
+\todo{Comment on things we could do with these samples hierarchical
+inference...}
 
@@
     \package{scipy} (\url{https://www.scipy.org/}),
-    \package{schwimmbad} (\todo{JOSS paper}),
+    \package{schwimmbad} (\citealt{Price-Whelan:2017a}),
     \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}),
-    \package{thejoker} (\todo{thejoker zenodo?}),
-    \package{twobody} (\todo{twobody zenodo?}).
+    \package{thejoker} (\citealt{Price-Whelan:2017b}),
+    \package{twobody} (\todo{twobody zenodo}).
 }
@@
 
+% Notebook: Compare Troup
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{troup-unimodal.pdf}
+\end{center}
+\caption{%
+Four examples of stars that also appear in the \citet{Troup:2016} companion
+catalog where our orbit samplings are unimodal, and agree well with the previous
+orbit fits.
+Left panels show the data from \apogee\ \DR\ (black markers), the orbit fit
+(orange, thick line; \citealt{Troup:2016}), and 128 orbits computed from
+posterior samples generated in this work (blue, thinner lines).
+Right panels show the period and eccentricity of the orbit fit (orange, +
+marker; \citealt{Troup:2016}) and the posterior samples from this work (blue
+markers).
+\label{fig:troup-unimodal}
+}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{troup-multimodal.pdf}
+\end{center}
+\caption{%
+Same as \figurename~\ref{fig:troup-unimodal}, but for four stars where we find a
+multi-modal posterior sampling.
+In this cases, the orbit fit from \citet{Troup:2016} generally identify one of
+the possible period modes, but multiple orbit solutions are typically allowed.
+\label{fig:troup-multimodal}
+}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{troup-data-changed.pdf}
+\end{center}
+\caption{%
+Same as \figurename~\ref{fig:troup-unimodal} and
+\figurename~\ref{fig:troup-multimodal}, but for two stars in which no meaningful
+comparison can be made because the visit velocities changed significantly
+between \DRtw\ and \DR.
+\label{fig:troup-datachanged}
+}
+\end{figure}
+
 \end{document}

### 2018-03-17 paper/1-catalog/twoface-catalog.tex
Address HWR comments - remove tidal circ bit

@@
 \newcommand{\quantile}[2]{\ensuremath{\textrm{Q}_{#1}}\left(#2\right)}
+\newcommand{\pdf}{\textit{pdf}}
 
@@
 % Context
-Repeat radial-velocity measurements of stars can be used to identify stellar,
-sub-stellar, and planetary-mass companions.
-Even a very small number of observation epochs can be useful for detecting
-companions, though such data can be difficult to use for characterizing
-individual systems:
-There can be multiple qualitatively different orbital solutions that fit the
-data.
+Multi-epoch radial velocity measurements of stars can be used to identify
+stellar, sub-stellar, and planetary-mass companions.
+Even a small number of observation epochs can be useful for detecting
+companions, though there can be multiple qualitatively different orbital
+solutions that fit the data.
 For these reasons, we custom-built a Monte Carlo sampler (\thejoker) that
@@
 % Aims
-Here we perform a search for secondary companions of \nstars\ red-giant stars
-observed in the \apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
+Here we perform a search for companions to \nstars\ red-giant stars observed in
+the \apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
 % Methods
@@
 We provide (1)~a catalog of \nunimodal\ companions for which the stellar
-companion properties can be confidently determined in a posterior sense, (2)~a
-catalog of \nhighK\ stars that likely have companions, but for which
-(in most cases) more observations would be needed to uniquely determine the
-orbital properties, and (3)~posterior samplings for all orbital-companion
-parameters for all stars in the parent sample.
+companion properties can be confidently determined, (2)~a catalog of \nhighK\
+stars that likely have companions, but for which more observations would be
+needed to uniquely determine the orbital properties, and (3)~posterior samplings
+for the full orbital parameters for all stars in the parent sample.
 We highlight interesting systems and show the characteristics of stars with
@@
 multiple star systems rather than as solitary stars (e.g.,
-\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014}).
-This is likely a generic outcome of star formation: For example, turbulent
+\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014,Moe:2017}).
+This is likely a generic outcome of star formation, as, for example turbulent
 fragmentation in collapsing protostellar clouds can produce stellar multiplets
@@
 However, with few or imprecise radial-velocity measurements, the orbital
-properties of the companion(s) are very poorly constrained (e.g.,
+properties of the companion(s) may be poorly constrained (e.g.,
 \citealt{Price-Whelan:2017}).
 The vast majority of spectroscopic targets with repeat observations in the
-largest (by number of objects) stellar spectroscopic surveys are often observed
-just a few times with sparse, non-uniform phase coverage.
+largest (by number of objects) stellar spectroscopic surveys are observed just a
+few times with sparse, non-uniform phase coverage.
 Most prior searches for companions using survey RV data have therefore
@@
 
-If there are only a few radial-velocity measurements made per star, and the
-companion spectrum is not observed, any measured radial-velocity values will be
-consistent with many different combinations of (primary) orbital parameters
-(period, amplitude, eccentricity, etc.).
+If there are only a few radial-velocity epochs per star, and if the companion
+spectrum is not observed, the data will be consistent with many different
+combinations of (primary) orbital parameters (period, amplitude, eccentricity,
+etc.).
 To identify companions to the typical star observed in a spectroscopic survey,
@@
 In general, the likelihood function---and the posterior probability distribution
-function (pdf) under any reasonable prior pdf---will be highly multimodal, and
+function (\pdf) under any reasonable prior \pdf---will be highly multimodal, and
 many of the modes will have comparable integrated probability density.
@@
 
-We have solved this problem previously, though with limitations (to be discussed
-more below), with \thejoker\ (\citealt{Price-Whelan:2017}).
-\thejoker\ is a Monte Carlo rejection sampler that is computationally expensive
-but probabilistically righteous:
-It delivers independent posterior pdf samples for single-companion binary
+We have solved this problem previously with \thejoker\
+(\citealt{Price-Whelan:2017}), a Monte Carlo rejection sampler that is
+computationally expensive but probabilistically righteous:
+It delivers independent posterior \pdf samples for single-companion binary
 orbital parameters, given any number of radial-velocity measurements.
-Here we use \thejoker\ to generate posterior pdf samples for stars observed by
+Here we use \thejoker\ to generate posterior \pdf samples for stars observed by
 the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2017}).
@@
 ideal for the study of single-line binary systems.
-For one, because they are so luminous, they are unlikely (in general) to have
-equally-bright companions, and their spectra are therefore well-approximated or
-fit as single-line objects.
+In general, they are unlikely to have equally-bright companions, and their
+spectra are therefore fit as single-line objects.
 When this constraint is not met, \thejoker\ will in general fail, and a model
 that fits for a mixture of stellar spectra is more appropriate (e.g.,
-\citealt{Kareem}).
-The subset of \RGB\ stars in the ``red clump'' (\RC) are even more powerful as
+\citealt{El-Badry:2018}).
+The subset of giant stars in the ``red clump'' (\RC) are even more powerful as
 they are standard candles and have masses that can be estimated using
 spectroscopy (using dredged-up elements; \citealt{Martig:2016,Ness:2016}).
-With primary-star mass estimates, the binary-orbit fitting will return
+With mass estimates for the primary star, the binary-orbit fitting will return
 $m_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
@@
 
-By making cuts on this library of posterior pdf samples (described in detail in
+By making cuts on this library of posterior \pdf samples (described in detail in
 \sectionname~\ref{sec:whatever}), we deliver a catalog of binary-star systems
@@
 
-Data taken as part of the \apogee\ survey is reduced with a multi-step data
+Data taken as part of the \apogee\ survey are reduced with a multi-step data
 reduction pipeline that ultimately solves for the stellar parameters, chemical
@@
 \caption{%
-Number of \apogee\ stars in logarithmic bins of number of visits that pass the
-quality cuts described in \sectionname~\ref{sec:data}.
+Number of \apogee\ \DR\ stars in logarithmic bins of number of visits that pass
+the quality cuts described in \sectionname~\ref{sec:data}.
 In total, this work uses \nvisits\ visits and \nstars\ unique sources.
@@
 \figurename~\ref{fig:nvisits} shows the number of stars in several bins of
-number of visits that pass the above quality cuts: \todo{92}\% of the stars in
+number of visits that pass the above quality cuts: 92\% of the stars in
 our parent sample have $< 8$ visits.
@@
 with a single companion.
-This sampling is performed under a relatively uninformative prior pdf, and the
-resulting posterior samplings are used to discover and characterize individual
-binary-star systems and generate a catalog of companions.
-We use \thejoker\ to perform the posterior samplings; we briefly describe the
-algorithm below, but a full description is given in previous work
-(\citealt{Price-Whelan:2017}).
-In this \sectionname, we describe the assumptions and method used to generate
-individual-system samplings.
-
-We perform our individual system fits---that is, make our posterior
-samplings---under the following assumptions:
+This sampling is performed with \thejoker\ (\citealt{Price-Whelan:2017}) under a
+relatively uninformative prior \pdf, and the resulting posterior samplings are
+used to discover and characterize individual binary-star systems and generate a
+catalog of companions.
+We now describe the assumptions and method used to generate samplings for
+individual systems:
 \begin{description}
@@
   % \todo{revisit this - do we want to also sample over a long-term trend?}
-\item[Kepler] Related to the first point, all velocity variations of the primary
-  are gravitational, and we therefore ignore the possibility of coherent intrinsic variation from, e.g., stellar oscillations.
+\item[Kepler] All velocity variations of the primary
+  are gravitational, and we therefore ignore the possibility of coherent
+  intrinsic variation from, e.g., stellar oscillations.
 \item[SB1] All spectra are single-lined; that is, we assume that the secondary
@@
 anomaly at a reference time $t_0$, argument of pericenter, velocity
-semi-amplitude, Barycentric velocity---and the true anomaly, $f$, is a function
+semi-amplitude, barycentric velocity---and the true anomaly, $f$, is a function
 of time and the specified parameters (see \sectionname~2 of
@@
 
-\thejoker\ was designed for the extremely multi-modal pdfs expected when the
+\thejoker\ was designed for the extremely multi-modal \pdf s expected when the
 number of radial-velocity measurements of a source is small, or the data are
@@
 After generating an initial (very large) library of prior samples from an
-assumed prior pdf (see below), the (typically multi-modal) likelihood is
+assumed prior \pdf (see below), the (typically multi-modal) likelihood is
 evaluated at each sample and used to rejection sample.
@@
 
-Here we describe the procedure we use to generate per-target posterior samplings
-for the \apogee\ targets.
+Here we describe the specifics of generating posterior samplings for all of the
+\apogee\ targets in our parent sample.
 We execute the full procedure twice for different goals (as described in
@@
     \item uniform in log-period over the domain $[1,32768]~\textrm{day}$,
-    \item a beta distribution over eccentricity with fixed parameters (\citealt{Kipping:2013}).
+    \item a beta distribution over eccentricity (using parameters from
+    \citealt{Kipping:2013}).
 \end{itemize}
-For the prior over the excess variance parameter, $s^2$, we initially use
-a Gaussian over the transformed parameter $y = \ln s^2$ with the mean and
-standard deviation $(\mu_y, \sigma_y)$ indicated where the runs are described
-(see \sectionname~\ref{sec:catalogs}).
+For the excess variance parameter, $s^2$, we initially use a Gaussian over the
+transformed parameter $y = \ln s^2$ with the mean and standard deviation
+$(\mu_y, \sigma_y)$ indicated where the runs are described (see
+\sectionname~\ref{sec:catalogs}).
 The reference time for each star is set to the minimum visit epoch; $M_0$ then
@@
 prior samples, or if multiple (but fewer than \nposterior) are returned that all
-lie within a single mode of the posterior pdf, the posterior pdf over orbital
+lie within a single mode of the posterior \pdf, the posterior \pdf over orbital
 parameters is treated as effectively unimodal: these stars are flagged ``needs
@@
 parametrization, $(\ln P, \sqrt{K}\,\cos M_0, \sqrt{K}\,\sin M_0, \sqrt{e}\,\cos
-\omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, to sample in.
+\omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, for MCMC sampling.
 This reparametrization is safer and more efficient for sampling with
@@
 parametrization.
-We use the same prior pdfs as in \thejoker\ when running MCMC (see
+We use the same prior \pdf s as in \thejoker\ when running MCMC (see
 \tablename~\ref{tbl:params}).
@@
 
-\subsubsection{``Needs more prior samples'': Continuing \thejoker\ sampling}
+\subsubsection{``Needs more prior samples'': Extending \thejoker\ sampling}
 
 If more than one posterior sample is returned after exhausting the full library
-of prior samples, and the samples lie in multiple modes of the posterior pdf,
+of prior samples, and the samples lie in multiple modes of the posterior \pdf,
 the only way to proceed is to generate more prior samples and continue running
@@
 
+\todo{Rix suggests this needs 1-2 more sentences to ease the transition: "It comes as an unexpected abrupt turn in the logical flow."}
 As an initial use of the per-source posterior samplings, we use a hierarchical
@@
 (\citealt{Hogg:2010,Foreman-Mackey:2014}); we describe how to compute the
-marginal likelihood in detail in \sectionname~\ref{sec:hierarch}.
+marginal likelihood in detail in Appendix~\ref{sec:hierarch}.
 
@@
 
-Using the maximum-likelihood hyperparameters derived from the initial posterior
+Using the most likely hyperparameters derived from the initial posterior
 samplings (see \sectionname~\ref{sec:inferjitter}), we update and fix the
@@
 parent sample of \apogee\ \DR\ stars.
-This methodology, in which we estimate a prior from the data and then fix it, is
-typically referred to as ``empirical Bayes''; In this case, it is an
-approximation to doing a hierarchical inference of the individual system orbital
-parameters and the excess variance distribution hyperparameters simultaneously.
+This approach of estimating a prior \pdf\ from the data and then fixing the
+parameters is typically referred to as ``empirical Bayes''; in this case, it is
+an approximation to doing a hierarchical inference of the individual system
+orbital parameters and the excess variance distribution hyperparameters
+simultaneously.
 
@@
 flagged as \emph{needs more prior samples}, 391 as \emph{needs MCMC}.
-We continue generating samples for these stars using the methodology explained
-above (see \sectionname~\ref{sec:samplings}).
+We then proceed to generate \nposterior\ samples for these two subsets using the
+methodology explained above (see \sectionname~\ref{sec:samplings}).
 In the end, \todo{XX} stars have completed samplings with \nposterior\ samples,
@@
 
-We don't expect there to be a sharp transition in orbital parameters between
-stars with companions and stars without companions: There is a continuum of
-companion properties.
+We don't expect a sharp transition in inferred orbital parameters between stars
+with companions and stars without companions: There is a continuum of companion
+properties.
 For example, the companions can be low mass, or at long periods, or at high
@@
 a given system.
-Therefore, from the radial velocity data alone, there is no way to define a
-simple cut to select a complete sample of stars with companions.
+Therefore there is no simple cut on radial velocity data alone that would select
+a complete sample of stars with companions.
 It is nevertheless possible to define a cut that selects stars with
@@
 The majority of the stars in the high-$K$ sample have poorly constrained orbital
-parameters because the posterior pdfs are very multimodal.
+parameters because the posterior \pdf s are very multimodal.
 The high-$K$ sample selected from \tablename~\ref{tbl:lnK-per} includes all
@@
 
-\subsection{Confident companions with uniquely-determined orbits}
+\subsection{Companions with uniquely-determined orbits}
 \label{sec:catalog-unimodal}
@@
 pass this cut.
-Unlike the multi-modal stars, the posterior pdfs for these stars can be
+Unlike the multi-modal stars, the posterior \pdf s for these stars can be
 approximated using point-estimates and standard deviations of their respective
@@
 
-\subsection{Confident companions with highly constrained orbits}
+\subsection{Companions with highly constrained orbits}
 \label{sec:catalog-multimodal}
@@
 estimates for systems with unimodal (black markers) and bimodal (blue lines)
-period samplings, again using primary masses from \citet{Ness:2015}.
+period samplings for the subset of /\nunimodal\ unimodal and /\nbimodal\ bimodal
+systems with primary mass estimates, again using primary masses from
+\citet{Ness:2015}.
 Upper dashed line shows the $M_{2, \textrm{min}} = M_1$ curve, and lower dashed
@@
 samplings.
-Upper dashed line in this panel indicates the Roche radius, computed using an approximate functional form (\citealt{Eggleton:1983})
+Upper dashed line in this panel indicates the Roche radius, computed using an
+approximate functional form (\citealt{Eggleton:1983})
 \begin{equation}
     \frac{R}{a} = \frac{0.49\,q^{-2/3}}{0.6\,q^{-2/3} +
-        \ln\left(1 + q^{-1/3}\right)} \quad .
+        \ln\left(1 + q^{-1/3}\right)}
 \end{equation}
+where $q = \M_2 / M_1$.
 All of these systems are generally consistent with being detached binaries.
@@
 \caption{%
-\textit{Left panel:} Estimated minimum companion mass and primary mass for
-systems in the high-$K$, unimodal sample (black markers) and bimodal sample
-(blue markers and lines).
+\textit{Left panel:} Estimated minimum companion mass and primary mass for a
+subset of the systems in the high-$K$, unimodal converged sample (31 systems;
+black circles), unimodal unconverged sample (38 systems; grey circles), and
+bimodal sample (25 systems; blue squares and lines) with $\log g > 2$ and
+previously measured primary masses (\citealt{Ness:2015}).
 Lines connecting markers (blue) show estimates from each period mode for each
@@
 separation of the two bodies, $a\,\sin i$, and minimum mass ratio,
-$q_{\textrm{min}}$, again for systems in the high-$K$, unimodal sample
-(black markers) and bimodal sample (blue markers and lines).
+$q_{\textrm{min}}$, for the same systems and markers in left panel.
 \label{fig:mass}
@@
 \end{figure}
-
-\subsection{Tidal circularization}
-
-From studies of binary star systems in open clusters, it is clear that
-short-period binaries tend to be more circular (have smaller eccentricities) as
-compared to longer-period binaries of the same age (e.g.,
-\citealt{Mathieu:2005}).
-Within a given population of binaries, this manifests as an apparently steep
-transition from a distribution of eccentricities to mainly circularized orbits
-at some characteristic period that depends on the age and evolutionary state of
-the population (see, e.g., \figurename~5 in \citealt{Mathieu:2005})---i.e. above
-this critical value of the period (or log-period), systems have some
-distribution of eccentricities, and below the eccentricities are close to zero.
-Many studies have measured this ``circularization period,'' predominantly for
-main-sequence binaries (e.g., \citealt{Meibom:2006, Kjurkchieva:2017}), and
-have found that it tends to be between 5--20 days, depending on age (e.g.,
-\citealt{Mathieu:1988}).
-For giant stars, the circularization period appears to be longer, closer to
-$\approx 100$ days (\citealt{Mayor:1984}).
-
-Theories for the tidal circularization process generally rely on tidal
-dissipation via eddy viscosity in the convective regions of stars to explain
-this (e.g., \citealt{Zahn:1989}).
-The circularization timescale and period should therefore depend on the size of
-the convective region, and the stellar parameters of the primary star.
-RGB stars are therefore excellent test-cases for these theories, since
-they are fully-convective:
-By comparing to (small) samples of giant star binaries in open clusters, it
-seems that the circularization timescales predicted by eddy viscosity
-dissipation at least qualitatively scale to giant stars
-(\citealt{Verbunt:1995}).
-
-It is less clear what to expect for red clump (\RC) stars:
-RC stars are horizontal branch stars and have therefore already climbed the RGB.
-Short period, circularized companions in place before or while on the RGB could
-be engulfed, or will at least have been common-envelope with the evolving star.
-However, it is difficult to select samples of \RC\ stars to study this process
-because first-ascent RGB stars will have similar colors and stellar parameters.
-\citet{Verbunt:1995} argue that companion orbits may be a way to separate the
-two overlapping populations: eccentric, short-period orbits around stars near
-the red-clump should be first-ascent giants.
-
-The samples produced in this work provide excellent datasets to study tidal
-circularization of companions to giant stars across a range of ages and
-metallicities
-However, our samples consist of a mix of subgiants, giants, and RC stars which
-have distinct distributions in mass and stellar radius.
-To visualize the distribution of systems in period--eccentricity space, we have
-therefore computed normalized periods, where we normalize by the orbital period
-of a typical companion on a circular orbit at the surface of the primary,
-$P_{\textrm{surface}}$.
-To do this for all systems in this sample, we assume that all primary stars have
-masses equal to the median mass over all stars with prior mass measurements in
-this sample, $M_1 = 1.36~\msun$, and assume a companion mass also equal to the
-median over minimum companion masses $M_{2, \textrm{min}} = 0.5~\msun$.
-We then use the surface gravity, $\log g$, of the primary, and eccentricity to
-compute
-\begin{align}
-    P_{\textrm{surface}} &= \frac{2\pi \, G^{1/4}}{(1-e)^{-3/2}} \,
-        \left(\frac{M_1 + M_2}{M_1^{3/4}}\right) \,
-        10^{-3/4\,\log g}
-\end{align}
-where we use $M_{2, \textrm{min}}$ as $M_2$.
-
-\figurename~\ref{fig:PeK}, left panel, shows this normalized orbital period and
-eccentricity for all systems in the high-$K$, unimodal sample, and, right panel,
-the RV semi-amplitude.
-There is a clear signature of circularization just under $P/P_{\textrm{surface}}
-\approx 10$.
-\todo{what else can or should we say?}
-
-% Notebook: Results
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=\textwidth]{P-e-K}
-\end{center}
-\caption{%
-\textit{Left panel:} Orbital period and eccentricity for all high-$K$ unimodal
-systems with $\log g > 2$, with period values normalized by the orbital period
-of a hypothetical companion at the surface of each primary star,
-$P_{\textrm{surface}}$, assuming $M_1 = 1.36~\msun$ and $M_2 =0.5~\msun$.
-Stars in the \apogee\ \DR\ red clump catalog are outlined in red.
-The sharp transition from eccentric systems to almost all circular orbits at
-$P/P_\textrm{surface} \approx 7$ is likely an outcome of tidal circularization.
-\textit{Right panel:} Normalized period and inferred radial velocity amplitude
-for the same systems.
-Points below $P/P_\textrm{surface} = 1$ are likely systems where the primary and
-companion mass assumptions used to compute $P_\textrm{surface}$ are invalid.
-\label{fig:PeK}
-}
-\end{figure}
-
-% Notebook: Results
-% \begin{figure}[h]
-% \begin{center}
-% \includegraphics[width=0.7\textwidth]{ecc-dist}
-% \end{center}
-% \caption{%
-% TODO:
-% \label{fig:ecc}
-% }
-% \end{figure}
 
@@
 (2) the parameters reported in \citet{Troup:2016} identify one possible mode of
-a likely multi-modal posterior pdf over orbital parameters, and
+a likely multi-modal posterior \pdf over orbital parameters, and
 (3) the data changed significantly between \apogee\ \DRtw\ and \DR, so
@@
 \figurename~\ref{fig:troup-multimodal} shows a few representative cases in which
-we find that the posterior pdf over orbital parameters is multimodal, and the
+we find that the posterior \pdf over orbital parameters is multimodal, and the
 \citet{Troup:2016} orbit identifies one of these modes.
@@
 
-\clearpage
-
 \bibliographystyle{aasjournal}
 \bibliography{refs}
+
+\clearpage
 

### 2018-03-19 twoface/mass.py
asini function

@@ -13,7 +13,8 @@ from scipy.optimize import root
 from .log import log as logger
 
 __all__ = ['CM_NM_to_CNM', 'CM_NM_to_CN', 'get_martig_vec', 'mf', 'm2_func',
-           'get_m2_min', 'stellar_radius', 'period_at_surface']
+           'get_m2_min', 'stellar_radius', 'period_at_surface', 'asini',
+           'a2sini']
 
 # For abundance transformations:
 log_eC = 8.39
@@ -117,10 +118,18 @@ def get_m2_min(m1, mf):
             return np.nan * m1.unit
 
 
-def stellar_radius(star, mass):
-    return np.sqrt(G*mass / (10**star.logg*u.cm/u.s**2)).to(u.Rsun)
+def stellar_radius(logg, mass):
+    return np.sqrt(G*mass / (10**logg*u.cm/u.s**2)).to(u.Rsun)
 
 
 def period_at_surface(M1, logg, e, M2=0*u.Msun):
     P = 2*np.pi * G**(1/4.) * (M1+M2) / (M1**(3/4.)) * (10**logg*u.cm/u.s**2)**(-3/4) * (1-e)**(-3/2)
     return P.to(u.day)
+
+
+def asini(P, e, K, m1, m2):
+    return (P/(2*np.pi) * np.sqrt(1-e**2) * K * (m1+m2) / m2).to(u.au)
+
+
+def a2sini(P, e, K, m1, m2):
+    return (P/(2*np.pi) * (m1/m2) * np.sqrt(1-e**2) * K).to(u.au)

### 2018-03-19 paper/1-catalog/twoface-catalog.tex
Placeholder for upper RGB
TODOs
new figures and shite
new figure
logg teff

@@
 \newcommand{\pdf}{\textit{pdf}}
+\newcommand{\logg}{\ensuremath{\log g}}
+\newcommand{\Teff}{\ensuremath{T_{\textrm{eff}}}}
 
@@
 to re-derive the relative visit velocities.
-The stellar parameters---surface gravity, $\log g$, and effective temperature,
-$T_{\textrm{eff}}$---and the chemical abundances are determined from the coadded
+The stellar parameters---surface gravity, \logg, and effective temperature,
+\Teff---and the chemical abundances are determined from the coadded
 spectrum as a part of the \apogee\ Stellar Parameters and Chemical Abundances
@@
 \texttt{PERSIST\_JUMP\_NEG}, \texttt{VERY\_BRIGHT\_NEIGHBOR}, \texttt{LOW\_SNR}.
-For each star, we require that $0 < \log g < 4$ and the following bits are not
+For each star, we require that $0 < \logg < 4$ and the following bits are not
 set in the \texttt{ASPCAPFLAGS} bitmask: \texttt{STAR\_BAD}.
@@
 \textit{Right panel}: Distribution of stars in our sample in stellar parameters
-log-surface-gravity, $\log g$, and effective temperature, $T_{\textrm{eff}}$,
-with points colored by the iron abundance.
+log-surface-gravity, \logg, and effective temperature, \Teff, with points
+colored by the iron abundance.
 \label{fig:loggteff}
@@
 
-We use 1,825 stars with $>10$ visits and $\log g < 2$ (to avoid upper RGB stars
+We use 1,825 stars with $>10$ visits and $\logg > 2$ (to avoid upper RGB stars
 that have large intrinsic jitter) and maximize the above likelihood to determine
@@
 visual inspection (\texttt{clean\_flag == 0}).
-We additionally restrict to stars with $\log g > 2$ to avoid upper RGB stars
+We additionally restrict to stars with $\logg > 2$ to avoid upper RGB stars
 that may have coherent oscillations that mimic RV modulations from companions.
+
+\todo{here:} \figurename~\ref{fig:logg-teff-p} ...
+
+% Notebook: "Results"
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\textwidth]{mass}
+\end{center}
+\caption{%
+\textit{Left panel:}
+\textit{Right panel:}
+\label{fig:logg-teff-p}
+}
+\end{figure}
 
@@
 black circles), unimodal unconverged sample (38 systems; grey circles), and
-bimodal sample (25 systems; blue squares and lines) with $\log g > 2$ and
+bimodal sample (25 systems; blue squares and lines) with $\logg > 2$ and
 previously measured primary masses (\citealt{Ness:2015}).
@@
 
+\subsection{Coherent...}
+
+\todo{Coherent oscillations?}
+
 % \subsection{Eclipsing}
@@
 \todo{comment on short period shit at low logg, eccentricities}
+
+\todo{Discuss tidal circ and mention worth pursuing. Don't say we are going to
+do this. Emphasize that the data contain information about those things.}
+
+\todo{High-mass companions}
 
@@
 Using this technique, they identified thousands of candidate MS binaries and
-trinaries, but did not consider giant stars ($\log g > 4$).
+trinaries, but did not consider giant stars ($\logg > 4$).
 This sample is therefore complimentary to and non-overlapping with the catalog

### 2018-03-20 paper/1-catalog/twoface-catalog.tex
Figure

@@
 visual inspection (\texttt{clean\_flag == 0}).
-We additionally restrict to stars with $\logg > 2$ to avoid upper RGB stars
-that may have coherent oscillations that mimic RV modulations from companions.
-
-\todo{here:} \figurename~\ref{fig:logg-teff-p} ...
+
+\todo{here:} \figurename~\ref{fig:logg-teff-p-e} ...
 
@@
 \begin{center}
-\includegraphics[width=\textwidth]{mass}
-\end{center}
-\caption{%
-\textit{Left panel:}
-\textit{Right panel:}
-\label{fig:logg-teff-p}
+\includegraphics[width=\textwidth]{P-Teff-logg-e}
+\end{center}
+\caption{%
+\textit{Top:} Inferred orbital period and eccentricity for all systems with unimodal posterior samplings from \thejoker.
+\textit{Bottom left:}
+\textit{Bottom Right:}
+\label{fig:logg-teff-p-e}
 }
@@
 for each period mode and connect these estimates with a line.
+Here we restrict to stars with $\logg > 2$ to avoid upper RGB stars that may
+have surface oscillations that mimic RV modulations from companions.
 

### 2018-03-20 paper/1-catalog/twoface-catalog.tex
change cut in lnK
Edits and cleaning

@@
 \newcommand{\ncontrol}{16,384}
-\newcommand{\nhighK}{4,510}
-\newcommand{\nbimodal}{113}
-\newcommand{\nunimodal}{319}
+\newcommand{\lnKcut}{-0.2}
+\newcommand{\nhighK}{4,898}
+\newcommand{\nbimodal}{106}
+\newcommand{\nunimodal}{322}
 
@@
 
-\title{Binary companions of red giant stars in \apogee\ \DR: \\
-       Search method and catalog of $>$4,500 systems}
+\title{Binary companions of evolved stars in \apogee\ \DR: \\
+       Search method and catalog of $\sim$5,000 companions}
 
@@
 for the full orbital parameters for all stars in the parent sample.
-We highlight interesting systems and show the characteristics of stars with
+We highlight interesting systems and show the characteristics of systems with
 confidently determined companion properties.
@@
 nearby main-sequence (MS) stars.
-\todo{occurrence rates}
+Nearly 50\% of MS stars in the solar neighborhood have companions
+(e.g., \citealt{Tokovinin:2014}).
 MS stars with companions have a large dynamic range of constituent and orbital
@@
 computationally expensive but probabilistically righteous:
-It delivers independent posterior \pdf samples for single-companion binary
+It delivers independent posterior \pdf\ samples for single-companion binary
 orbital parameters, given any number of radial-velocity measurements.
-Here we use \thejoker\ to generate posterior \pdf samples for stars observed by
+Here we use \thejoker\ to generate posterior \pdf\ samples for stars observed by
 the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2017}).
 
-The \apogee\ surveys primarily target red-giant-branch (\RGB) stars, which are
-ideal for the study of single-line binary systems.
+The \apogee\ surveys primarily target red-giant-branch (\RGB) and evolved stars,
+which are ideal for the study of single-line binary systems.
 In general, they are unlikely to have equally-bright companions, and their
@@
 With mass estimates for the primary star, the binary-orbit fitting will return
-$m_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
+$M_2\,\sin i$ (minimum mass) estimates for the secondary, and not just estimates
 of the so-called ``binary mass function.''
@@
 
-By making cuts on this library of posterior \pdf samples (described in detail in
-\sectionname~\ref{sec:whatever}), we deliver a catalog of binary-star systems
-from the \apogee\ survey \todo{with ... (tease some statistics)}.
-We additionally highlight the subset of this catalog with \RC\ star primaries,
-for which we can compute companion mass and pericentric distance.
+By making cuts on this library of posterior \pdf\ samples (described in detail
+in \sectionname~\ref{sec:catalogs}), we deliver catalogs of binary-star systems
+from the \apogee\ survey, and show the bulk properties of these systems.
 
@@
 1.5~\textrm{deg}^2$ field of view, leading to ``pencil-beam''-like sampling of
-the stellar distribution.
+the Galactic stellar distribution.
 In order to meet signal-to-noise ratio requirements, most \apogee\ stars are
@@
 extra systematic uncertainty and true astrophysical surface jitter
-(\todo{CITATION NEEDED}).
-
-\thejoker\ was designed for the extremely multi-modal \pdf s expected when the
+(e.g., \cite{Hekker:2008}).
+
+\thejoker\ was designed for the extremely multi-modal \pdf 's expected when the
 number of radial-velocity measurements of a source is small, or the data are
@@
 After generating an initial (very large) library of prior samples from an
-assumed prior \pdf (see below), the (typically multi-modal) likelihood is
+assumed prior \pdf\ (see below), the (typically multi-modal) likelihood is
 evaluated at each sample and used to rejection sample.
@@
 \end{itemize}
-For the excess variance parameter, $s^2$, we initially use a Gaussian over the
-transformed parameter $y = \ln s^2$ with the mean and standard deviation
-$(\mu_y, \sigma_y)$ indicated where the runs are described (see
-\sectionname~\ref{sec:catalogs}).
+For the excess variance parameter, $s^2$, we use a Gaussian over the transformed
+parameter $y = \ln s^2$ with the mean and standard deviation $(\mu_y, \sigma_y)$
+indicated where the runs are described (see \sectionname~\ref{sec:catalogs}).
 The reference time for each star is set to the minimum visit epoch; $M_0$ then
@@
 prior samples will pass the rejection step, whereas for very precise or
-many-epoch RV data, \thejoker\ may need to go process the full library of prior
+many-epoch RV data, \thejoker\ may need to process the full library of prior
 samples.
@@
 prior samples, or if multiple (but fewer than \nposterior) are returned that all
-lie within a single mode of the posterior \pdf, the posterior \pdf over orbital
+lie within a single mode of the posterior \pdf, the posterior \pdf\ over orbital
 parameters is treated as effectively unimodal: these stars are flagged ``needs
@@
 
-We use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010}) implemented
-in \python\ (\package{emcee}; \citealt{Foreman-Mackey:2013}) to perform the
-samplings.
+In detail, we use an ensemble MCMC sampling algorithm (\citealt{Goodman:2010})
+implemented in \python\ (\package{emcee}; \citealt{Foreman-Mackey:2013}) to
+perform the samplings.
 We transform the standard Keplerian orbital parameters to a safer
@@
 parametrization.
-We use the same prior \pdf s as in \thejoker\ when running MCMC (see
+We use the same prior \pdf 's as in \thejoker\ when running MCMC (see
 \tablename~\ref{tbl:params}).
@@
 samples, the sampling is flagged as ``incomplete.''
+% TODO: this isn't done yet in the catalogs!
 
@@
 When we save the comparison data, we only store the visit uncertainty and
-``forget'' the fact that the simulated data is generated with excess variance.
+``forget'' the fact that the simulated data is generated with excess variance
+(to be inferrred with \thejoker).
 
@@
 
-\todo{Rix suggests this needs 1-2 more sentences to ease the transition: "It comes as an unexpected abrupt turn in the logical flow."}
-As an initial use of the per-source posterior samplings, we use a hierarchical
+% \todo{Rix suggests this needs 1-2 more sentences to ease the transition: "It comes as an unexpected abrupt turn in the logical flow."}
+With posterior samplings for all \apogee\ \DR\ systems in hand, and as an
+initial use of the per-source posterior samplings, we use a hierarchical
 Bayesian model to infer the parameters of the (assumed Gaussian) prior over the
@@
 To select stars with companions, we use a cut in the 1st percentile of the $\ln
-K$ samples such that $<1\%$ of the comparison sample is selected, i.e. our
-estimated false positive rate is $<1\%$.
-We use a threshold of $\ln K = 0$ to meet this constraint (vertical, dashed line
-in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this cut.
+K$ samples such that $1\%$ of the comparison sample is selected, i.e. our
+estimated false positive rate is $\approx 1\%$.
+We use a threshold of $\ln K = \lnKcut$ to meet this constraint (vertical,
+dashed line in \figurename~\ref{fig:lnK-percentiles}); \nhighK\ stars pass this
+cut.
 For brevity below, we refer to this sample as the ``high-$K$'' stars, and the
@@
 The vertical line (dashed, orange) indicates our adopted cut to select high-$K$
-stars that likely have companions; $<1\%$ of the comparison sample falls above
-this cut.
+stars that likely have companions at $\ln K = \lnKcut$; $1\%$ of the comparison
+sample falls above this cut.
 \label{fig:lnK-percentiles}
@@
 The majority of the stars in the high-$K$ sample have poorly constrained orbital
-parameters because the posterior \pdf s are very multimodal.
+parameters because the posterior \pdf 's are very multimodal.
 The high-$K$ sample selected from \tablename~\ref{tbl:lnK-per} includes all
@@
 pass this cut.
-Unlike the multi-modal stars, the posterior \pdf s for these stars can be
+Unlike the multi-modal stars, the posterior \pdf 's for these stars can be
 approximated using point-estimates and standard deviations of their respective
@@
 
-We have visually inspected the inferred orbits for all systems in this sample
-and have flagged systems with questionable or invalid fits.
+We have also visually inspected the inferred orbits for all systems in this
+sample and have flagged systems with questionable or invalid fits.
 The value of the flag is: 0, when the orbits look reasonable, 1, when the orbits
@@
 For each cluster, we then ask whether the samples assigned to that cluster are
-unimodal (\eqname~\ref{eq:catalog-unimodal}).
+unimodal (\eqname~\ref{eq:unimodal}).
 If the samples in each respective cluster are unimodal, we call the sampling
@@
 For example, in the bottom left panel of \figurename~\ref{fig:highK-bimodal}, an
-observation at $t = 60$ would rule out one of the two possible orbits.
+observation at $t = 60$ would rule out one of the two possible  classes of
+orbits.
 
@@
 from \citet{Ness:2015} and an estimate of the minimum companion mass, $M_{2,
-\textrm{min}}$.
+\textrm{min}}$ for each mode.
 
@@
 s~\ref{sec:catalog-unimodal} and \ref{sec:catalog-multimodal}).
-In all figures, we only plot companions with inferred orbits that pass our
+In all figures below, we only plot companions with inferred orbits that pass our
 visual inspection (\texttt{clean\_flag == 0}).
 
-\todo{here:} \figurename~\ref{fig:logg-teff-p-e} ...
+\subsection{Systems with unimodal and bimodal posterior samplings}
+
+\figurename~\ref{fig:logg-teff-p-e} shows the companion and stellar properties
+of the unimodal and bimodal samples: upper left panel shows inferred period and
+eccentricity of the binary orbit, lower left panel shows inferred period and
+surface gravity of the primary star, lower right panel shows the primary star
+stellar parameters.
+
+The period-eccentricity plot (upper left panel in
+\figurename~\ref{fig:logg-teff-p-e}) shows a clear signature of tidal
+circularization: Such plots for main sequence stars typically show a more
+distinct circularization period, below which the majority of stars have
+eccentricities close to zero.
+Here, we see that close to $P\approx 10~\textrm{day}$ the majority of companion
+orbits appear to be circular, but between 10--100~days there appears to be a
+gradual decrease in eccentricity rather than a sharp transition.
+This is likely because the primary stars in this sample have a large range in
+surface gravities and therefore a large range in stellar radii (see lower right
+panel).
+
+In the lower left panel of \figurename~\ref{fig:logg-teff-p-e}, the diagonal
+(thick) line shows the orbital period of a hypothetical massless companion at
+the surface of a $1.35~\msun$ primary star (the median of our sample with
+measured masses) with the given surface gravity.
+The vertical line in this panel shows the same for a $1.35~\msun$ primary with
+$\logg=0$, i.e. the orbital period at the surface of such a star close to the
+tip of the RGB.
+The horizontal line in this panel shows the approximate upper bound of the red
+clump, above which most stars should be fully convective.
+Interestingly, there is a dearth of short-period systems for stars above the red
+clump ($\logg \lesssim 2.3$) in the triangle defined by the three qualitative
+lines, suggestive of possible companion engulfment as the primary star envelope
+encases the companion.
+
+The lower right panel of \figurename~\ref{fig:logg-teff-p-e} shows the stellar
+parameters of all primary stars in the unimodal and bimodal samples.
+Relative to \figurename~\ref{fig:loggteff}, upper RGB stars appear to be
+under-represented in these companion catalogs, another tentative signature of
+engulfment or depletion of companions on shorter-period orbits.
 
@@
 \caption{%
-\textit{Top:} Inferred orbital period and eccentricity for all systems with unimodal posterior samplings from \thejoker.
-\textit{Bottom left:}
-\textit{Bottom Right:}
+\textit{Top:} Inferred orbital period and eccentricity for all systems with
+unimodal and bimodal posterior samplings from \thejoker.
+\textit{Bottom left:} Inferred orbital period and primary star surface gravity,
+\logg.
+Diagonal line indicates the orbital period of a hypothetical massless companion
+at the surface of a $1.35~\msun$ primary star with the given \logg, vertical
+line shows the same for $\logg = 0$, and horizontal line shows the rough upper
+boundary of the red clump.
+\textit{Bottom Right:} Stellar parameters for all primary stars in the unimodal
+and bimodal samples.
 \label{fig:logg-teff-p-e}
@@
 \figurename~\ref{fig:mass}, left panel shows primary and companion mass
-estimates for systems with unimodal (black markers) and bimodal (blue lines)
-period samplings for the subset of /\nunimodal\ unimodal and /\nbimodal\ bimodal
-systems with primary mass estimates, again using primary masses from
-\citet{Ness:2015}.
+estimates for systems with unimodal (circle markers) and bimodal (square markes,
+blue lines) period samplings for the subset of 69/\nunimodal\ unimodal and
+25/\nbimodal\ bimodal systems with primary mass estimates, again using primary
+masses from \citet{Ness:2015}.
 Upper dashed line shows the $M_{2, \textrm{min}} = M_1$ curve, and lower dashed
@@
 the minimum mass ratio, $q_{\textrm{min}} = M_{2, \textrm{min}} / M_1$, for the
-same samples.
-Again, black large markers indicate systems from the unimodal sample, and blue
-markers and lines connect the two estimates for systems with bimodal period
-samplings.
+same systems.
+Again, circle (black, grey) markers indicate systems from the unimodal sample,
+and square (blue) markers and lines connect the two estimates for systems with
+bimodal period samplings.
 Upper dashed line in this panel indicates the Roche radius, computed using an
@@
 \end{equation}
-where $q = \M_2 / M_1$.
-All of these systems are generally consistent with being detached binaries.
-
-In \figurename~\ref{fig:mass}, three interesting systems are immediately
-obvious: two systems with bimodal period samplings in which both period modes
-put the minimum companion mass below the Hydrogen-burning limit (brown dwarf
-candidates), and one system in which $M_{2, \textrm{min}} > M_1$ (a neutron star
-or black hole candidate).
-These systems are being followed-up and will be studied in future work.
+where $q = M_2 / M_1$.
+All of these systems are consistent with being detached binaries, with the
+exception of one notable system (discussed below).
+
+In \figurename~\ref{fig:mass}, a few interesting systems are immediately
+obvious: From the left panel, two systems have bimodal period samplings in which
+both period modes put the minimum companion mass below the Hydrogen-burning
+limit (brown dwarf candidates), and two systems have $M_{2, \textrm{min}} > M_1$
+(neutron star or black hole candidates).
+In the right panel, one system appears right on the limit of being an
+interacting binary ($q_{\textrm{min}} \approx 0.5$), but \todo{does it have
+x-ray or UV data?}.
+These systems are prime candidates for spectroscopic follow-up.
 
@@
 
-\subsection{Coherent...}
-
-\todo{Coherent oscillations?}
-
-% \subsection{Eclipsing}
-%
-% \todo{Stars with Kepler light curves}
-
 \section{Discussion}
 
-\todo{comment on short period shit at low logg, eccentricities}
-
-\todo{Discuss tidal circ and mention worth pursuing. Don't say we are going to
-do this. Emphasize that the data contain information about those things.}
-
-\todo{High-mass companions}
+\subsection{Impossible companions and the upper RGB}
+
+Paradoxically, there exist short period ($\lesssim 100~\textrm{day}$) companions
+at low surface gravities ($\logg \lesssim 1$), obvious in the lower-left panel
+of \figurename~\ref{fig:logg-teff-p-e}: These companions would orbit within the
+surface of their primary stars.
+Each of these systems appears fine from the \apogee\ data quality flags, but
+could nonetheless have incorrect stellar parameters.
+As a test, we would ideally be able to compare the spectroscopic stellar
+parameters to an independent determination from, e.g., asteroseismology.
+Unfortunately, none of these short-period, low-\logg\ stars in the unimodal or
+bimodal samples appear in the \acronym{APOKASC} (\citealt{Serenelli:2017})
+catalog, which provides asteroseismic stellar parameters for a few hundred
+\apogee\ dwarf and giant stars.
+We therefore instead cross-match all high-$K$ stars with $\logg < 1$ and $99\%$
+of their period samples below $100~\textrm{day}$ to the \acronym{APOKASC}
+catalog: One star appears in both samples 2M19024490+4419523.
+In \apogee\ \DR, this star has $\logg = 0.57$, but asteroseismic $\logg > 4$
+(e.g., \texttt{HUBER\_LOGG}), consistent with being an M dwarf.
+Many of these systems could therefore be low-mass dwarf stars with incorrect
+stellar parameters in \apogee\ \DR.
+
+Another possibility is that \thejoker\ interprets semi-coherent surface
+oscillations (from asteroseismic modes) with sparse time coverage as orbital RV
+variations (see, e.g., \citealt{Hekker:2008}).
+However, for RGB stars with $\logg < 1$, these modes would likely have
+frequencies between $\nu_{\textrm{max}} \approx 0.5$--$5~\mu\textrm{Hz}$
+(\citealt{Garcia:2018}), corresponding to periods between $P \approx
+20$--$2~\textrm{day}$ and amplitudes between $\approx
+20$--$200~\textrm{m}~\textrm{s}^{-1}$ (\citealt{Huber:2011,Huber:2017}).
+Except for the lowest \logg\ stars, these amplitudes wouldn't pass our cut on
+$K$ (\sectionname~\ref{sec:catalog-confident}).
+
+These systems would need precise, long-term photometric follow-up to obtain
+asteroseismic parameters to confirm or explain their existence.
 
@@
 gravitational---\emph{Kepler}.
-We see tentative evidence for coherent oscillations at the upper giant branch
-(see \figurename~\ref{fig:todo}), which we presently treat as excess variance or
-intrinsic jitter.
-If these oscillations are indeed from asteroseismic modes, \todo{why is this
-interesting?}.
+We see tentative evidence for surface oscillations at the upper giant branch
+(see \figurename~\ref{fig:P-Teff-logg-e}, lower left), which we presently treat
+as excess variance or intrinsic jitter.
+Further observations are needed to determine whether these ``systems'' have
+incorrect stellar parameters, or indeed show surface oscillations related to
+asteroseismic modes.
 
@@
 luminosity and therefore spectrum of each system---\emph{SB1}.
-As seen in \figurename~\ref{fig:todo}, there are some companions with minimum
+As seen in \figurename~\ref{fig:mass}, there are some companions with minimum
 masses comparable to or consistent with being larger than the masses of the
@@
 followed-up to obtain further RV measurements to test whether any companions are
-black holes (Geha et al., in prep.).
+black holes.
+However, \thejoker\ can and should be extended to support sampling for SB2
+systems with just one additional parameter (typically the mass ratio, or ratio
+of velocity semi-amplitudes).
 
@@
 
-Of the 382 companions released as a part of this previous search, only 188 of
-the host stars passed the stellar parameter and quality cuts used to define the
+Of the 382 companions released as a part of the previous search, only 188 of the
+host stars passed the stellar parameter and quality cuts used to define the
 parent sample in this work (see \sectionname~\ref{sec:data}).
@@
 (2) the parameters reported in \citet{Troup:2016} identify one possible mode of
-a likely multi-modal posterior \pdf over orbital parameters, and
+a likely multi-modal posterior \pdf\ over orbital parameters, and
 (3) the data changed significantly between \apogee\ \DRtw\ and \DR, so
@@
 \figurename~\ref{fig:troup-multimodal} shows a few representative cases in which
-we find that the posterior \pdf over orbital parameters is multimodal, and the
+we find that the posterior \pdf\ over orbital parameters is multimodal, and the
 \citet{Troup:2016} orbit identifies one of these modes.
@@
 
-\todo{Comment on things we could do with these samples hierarchical
-inference...}
-
-We don't need to know the orbits of each individual system: that's the power in
-this.
-
-We are SB1 only, but should do SB2 as well - no reason to restrict.
+The main motivation for this work was to produce posterior samplings for all
+\apogee\ \DR\ stars to be used in a population inference.
+To use these samplings for population or hierarchical inference, the orbits of
+each individual system don't have to be unimodal and thus \emph{all} \nstars\
+samplings can be used in conjunction without well-determined orbital parameters
+for the majority of the stars.
+These samplings will be useful for constraining (through hierarchical inference)
+the period and eccentricity distributions of evolved stars, and the occurrence
+rates of companions to evolved stars to test for signatures of engulfment.
 
@@
 
+We have selected a catalog of nearly 5,000 stellar systems with companions by
+making cuts on posterior beliefs about the amplitude of orbital radial velocity
+variations for often low-epoch, sparsely sampled radial velocity data.
+We provide posterior samplings over orbital parameters for all \nstars\ in the
+parent sample of \apogee\ \DR\ systems used in this work, along with several
+sub-catalogs with better constrained orbital information:
+\begin{itemize}
+    \item A catalog of \nunimodal\ systems with unimodal posterior samplings,
+    and therefore uniquely determined orbits, of which 227 are newly discovered
+    binary star systems.
+    \item A catalog of \nbimodal\ systems with highly constrained posterior
+    samplings that, in period, span two distinct period modes.
+    For these systems, one or a few more radial velocity measurements would
+    uniquely determine their orbits.
+    79 of these systems are newly discovered binary star systems.
+    \item A catalog of \nhighK\ systems with radial velocity variations
+    consistent with having a companion, but which need further radial velocity
+    measurements to better constrain the orbital properties.
+\end{itemize}
 All companion catalogs described in this work (\sectionname~\ref{sec:catalogs})
 are available online at \todo{XX}.
+
 The source code for this project is open source and available from
@@
 
-It is a pleasure to thank Keith Hawkins (Columbia), \todo{who else?}.
+It is a pleasure to thank Keith Hawkins (Columbia), Yuan-Sen Ting (Princeton),
+\todo{who else?}.
 

### 2018-03-20 paper/1-catalog/twoface-catalog.tex
sdss ack

@@
 
+Funding for the Sloan Digital Sky Survey IV has been provided by the Alfred P.
+Sloan Foundation, the U.S.
+Department of Energy Office of Science, and the Participating Institutions.
+SDSS-IV acknowledges
+support and resources from the Center for High-Performance Computing at
+the University of Utah. The SDSS web site is www.sdss.org.
+
+SDSS-IV is managed by the Astrophysical Research Consortium for the
+Participating Institutions of the SDSS Collaboration including the
+Brazilian Participation Group, the Carnegie Institution for Science,
+Carnegie Mellon University, the Chilean Participation Group, the French Participation Group, Harvard-Smithsonian Center for Astrophysics,
+Instituto de Astrof\'isica de Canarias, The Johns Hopkins University,
+Kavli Institute for the Physics and Mathematics of the Universe (IPMU) /
+University of Tokyo, Lawrence Berkeley National Laboratory,
+Leibniz Institut f\"ur Astrophysik Potsdam (AIP),
+Max-Planck-Institut f\"ur Astronomie (MPIA Heidelberg),
+Max-Planck-Institut f\"ur Astrophysik (MPA Garching),
+Max-Planck-Institut f\"ur Extraterrestrische Physik (MPE),
+National Astronomical Observatories of China, New Mexico State University,
+New York University, University of Notre Dame,
+Observat\'ario Nacional / MCTI, The Ohio State University,
+Pennsylvania State University, Shanghai Astronomical Observatory,
+United Kingdom Participation Group,
+Universidad Nacional Aut\'onoma de M\'exico, University of Arizona,
+University of Colorado Boulder, University of Oxford, University of Portsmouth,
+University of Utah, University of Virginia, University of Washington, University of Wisconsin,
+Vanderbilt University, and Yale University.
+
 The authors are pleased to acknowledge that the work reported on in this

### 2018-04-02 twoface/mass.py
more plotting options

@@ -123,7 +123,8 @@ def stellar_radius(logg, mass):
 
 
 def period_at_surface(M1, logg, e, M2=0*u.Msun):
-    P = 2*np.pi * G**(1/4.) * (M1+M2) / (M1**(3/4.)) * (10**logg*u.cm/u.s**2)**(-3/4) * (1-e)**(-3/2)
+    R1 = np.sqrt(G*M1 / (10**logg * u.cm/u.s**2))
+    P = 2*np.pi * (G*(M1+M2) / R1**3)**(-1/2) * (1-e)**(-3/2)
     return P.to(u.day)
 
 

### 2018-04-02 twoface/plot.py

@@ -18,8 +18,7 @@ _RV_LBL = 'RV [{0:latex_inline}]'
 
 def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
                      xlim_choice='data', ylim_fac=1., n_times=4096, title=None,
-                     ax=None, highlight_P_extrema=True, plot_kwargs=None,
-                     relative_to_t0=False):
+                     ax=None, highlight_P_extrema=True, plot_kwargs=None, data_plot_kwargs=None, relative_to_t0=False):
     """
     Plot the APOGEE RV data vs. time and orbits computed from The Joker samples.
 
@@ -81,12 +80,18 @@ def plot_data_orbits(data, samples, n_orbits=128, jitter=None,
         plot_kwargs = dict()
 
     plot_kwargs.setdefault('color', 'tab:blue')
-    # plot_kwargs.setdefault('alpha', 0.25)
     plot_kwargs.setdefault('linewidth', 0.5)
+
+    if data_plot_kwargs is None:
+        data_plot_kwargs = dict()
+
+    data_plot_kwargs.setdefault('zorder', 5)
+    data_plot_kwargs.setdefault('elinewidth', 1)
+    
     plot_rv_curves(samples, t_grid, rv_unit=u.km/u.s, data=data, ax=ax,
                    n_plot=min(len(samples['P']), n_orbits),
                    plot_kwargs=plot_kwargs,
-                   data_plot_kwargs=dict(zorder=5, elinewidth=1,),
+                   data_plot_kwargs=data_plot_kwargs,
                    relative_to_t0=relative_to_t0)
 
     if highlight_P_extrema:
@@ -349,5 +354,5 @@ def plot_phase_fold_residual(data, sample, axes=None, label=True,
 
     axes[0].set_xlim(0, 1)
     axes[1].set_xlim(0, 1)
-    
+
     return fig

### 2018-04-02 paper/1-catalog/twoface-catalog.tex
refs and update unimodal table
Coauthors

@@
 
-\author{APOGEE Team}
+% APOGEE RV var co-authors
+\author{Nathan De Lee}
+\affiliation{Department of Physics and Astronomy, Vanderbilt University, VU Station 1807, Nashville, TN 37235, USA}
+
+\author{Steven R. Majewski}
+\affiliation{Department of Astronomy, University of Virginia, Charlottesville, VA 22904-4325, USA}
+
+\author{David L. Nidever}
+\affiliation{Department of Physics, Montana State University, P.O. Box
+173840, Bozeman, MT 59717-3840}
+\affiliation{National Optical Astronomy Observatory, 950 North Cherry
+Ave, Tucson, AZ 85719}
+
+\author{Nicholas Troup}
+\affiliation{}
+
+% Other APOGEE co-authors:
+\author{Jos\'e G. Fern\'andez-Trincado}
+\affiliation{Departamento de Astronom\'ia, Universidad de Concepci\'on, Casilla 160-C, Concepci\'on, Chile}
+
+\author{Domingo An\'ibal Garc\'ia-Hern\'andez}
+\affiliation{Instituto de Astrof\'isica de Canarias, 38205 La Laguna, Tenerife, Spain}
+
+\author{Pen\'elope Longa-Pe\~na}
+\affiliation{Unidad de Astronom\'ia, Universidad de Antofagasta, Avenida Angamos 601, Antofagasta 1270300, Chile}
+
+\author{Christian Nitschelm}
+\affiliation{Unidad de Astronom\'ia, Universidad de Antofagasta, Avenida Angamos 601, Antofagasta 1270300, Chile}
+
+\author{Olga Zamora}
+\affiliation{Instituto de Astrof\'isica de Canarias, 38205 La Laguna, Tenerife, Spain}
 
@@
 
-It is a pleasure to thank Keith Hawkins (Columbia), Yuan-Sen Ting (Princeton),
-\todo{who else?}.
+It is a pleasure to thank
+Kevin Covey (WWU),
+Marina Kounkel (WWU),
+Keith Hawkins (Columbia),
+Melissa Ness (Columbia/Flatiron),
+David Spergel (Princeton/Flatiron),
+and Yuan-Sen Ting (Princeton).
 

### 2018-04-02 paper/1-catalog/twoface-catalog.tex
Rix comments

@@
 stellar, sub-stellar, and planetary-mass companions.
-Even a small number of observation epochs can be useful for detecting
+Even a small number of observation epochs can be informative about
 companions, though there can be multiple qualitatively different orbital
 solutions that fit the data.
-For these reasons, we custom-built a Monte Carlo sampler (\thejoker) that
-delivers reliable (and often highly multi-modal) posterior samplings for
-companion orbital parameters given sparse radial-velocity data.
+We have custom-built a Monte Carlo sampler (\thejoker) that delivers reliable
+(and often highly multi-modal) posterior samplings for companion orbital
+parameters given sparse radial-velocity data.
 % Aims
-Here we perform a search for companions to \nstars\ red-giant stars observed in
-the \apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
+Here we use \thejoker\ to perform a search for companions to \nstars\ red-giant
+stars observed in the \apogee\ survey (\DR) with $\geq 3$ spectroscopic epochs.
 % Methods
@@
 companion properties can be confidently determined, (2)~a catalog of \nhighK\
-stars that likely have companions, but for which more observations would be
-needed to uniquely determine the orbital properties, and (3)~posterior samplings
-for the full orbital parameters for all stars in the parent sample.
+stars that likely have companions, but would require more observations to
+uniquely determine the orbital properties, and (3)~posterior samplings for the
+full orbital parameters for all stars in the parent sample.
 We highlight interesting systems and show the characteristics of systems with
@@
 \section{Introduction} \label{sec:intro}
-
-Stars typically have companions.
-Main sequence stars in the solar neighborhood more often appear in binary or
-multiple star systems rather than as solitary stars (e.g.,
-\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014,Moe:2017}).
-This is likely a generic outcome of star formation, as, for example turbulent
-fragmentation in collapsing protostellar clouds can produce stellar multiplets
-within length-scales comparable to their spheres of influence (e.g.,
-\citealt{Raskutti:2016}).
-Binary and multiple star systems are therefore of great interest in
-astrophysics: The population of stars and their companions encodes information
-about star formation processes, stellar parameters and evolution, and the
-dynamics of multi-body systems (for recent reviews, see
-\citealt{Duchene:2013,Moe:2017}).
-
-Most of what is known about stellar companions comes from studies of
-nearby main-sequence (MS) stars.
-Nearly 50\% of MS stars in the solar neighborhood have companions
-(e.g., \citealt{Tokovinin:2014}).
-MS stars with companions have a large dynamic range of constituent and orbital
-characteristics.
-For example, binary stars have mass-ratios that span from $\approx 0.03$
-to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
-millions of years (e.g., \citealt{Raghavan:2010}).
-Less is known about population properties of non-interacting or detached
-companions to evolved stars.
 
@@
 
-We have solved this problem previously with \thejoker\
-(\citealt{Price-Whelan:2017}), a Monte Carlo rejection sampler that is
-computationally expensive but probabilistically righteous:
+We have solved the problem of deriving comprehensive multi-modal \pdf\ samplings
+previously with \thejoker\ (\citealt{Price-Whelan:2017}), a Monte Carlo
+rejection sampler that is computationally expensive but probabilistically
+righteous:
 It delivers independent posterior \pdf\ samples for single-companion binary
@@
 
-The \apogee\ surveys primarily target red-giant-branch (\RGB) and evolved stars,
-which are ideal for the study of single-line binary systems.
-In general, they are unlikely to have equally-bright companions, and their
+Stars typically have companions.
+Main sequence stars in the solar neighborhood more often appear in binary or
+multiple star systems rather than as solitary stars (e.g.,
+\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014,Moe:2017}).
+This is likely a generic outcome of star formation, as, for example turbulent
+fragmentation in collapsing protostellar clouds can produce stellar multiplets
+within length-scales comparable to their spheres of influence (e.g.,
+\citealt{Raskutti:2016}).
+Binary and multiple star systems are therefore of great interest in
+astrophysics: The population of stars and their companions encodes information
+about star formation processes, stellar parameters and evolution, and the
+dynamics of multi-body systems (for recent reviews, see
+\citealt{Duchene:2013,Moe:2017}).
+
+Most of what is known about stellar companions comes from studies of
+nearby main-sequence (MS) stars.
+Nearly 50\% of MS stars in the solar neighborhood have companions
+(e.g., \citealt{Tokovinin:2014}).
+MS stars with companions have a large dynamic range of constituent and orbital
+characteristics.
+For example, binary stars have mass-ratios that span from $\approx 0.03$
+to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
+millions of years (e.g., \citealt{Raghavan:2010}).
+Less is known about population properties of non-interacting or detached
+companions to evolved stars.
+
+The \apogee\ surveys primarily target red-giant-branch (\RGB) and other evolved
+stars (e.g., red clump giants, RC), which are ideal for the study of single-line
+binary systems.
+In general, they are unlikely to have comparably-bright companions, and their
 spectra are therefore fit as single-line objects.
@@
 \citealt{El-Badry:2018}).
-The subset of giant stars in the ``red clump'' (\RC) are even more powerful as
-they are standard candles and have masses that can be estimated using
-spectroscopy (using dredged-up elements; \citealt{Martig:2016,Ness:2016}).
+The subset of RC stars are even more powerful as they are standard candles and
+have masses that can be estimated using spectroscopy (using dredged-up elements;
+\citealt{Martig:2016,Ness:2016}).
 With mass estimates for the primary star, the binary-orbit fitting will return
@@
 
-All data used in this work comes from the publicly-available data release 14
+All data used in this work come from the publicly-available data release 14
 (\DR) of the \apogee\ survey (\citealt{Majewski:2017,Abolfathi:2017}), a
@@
 \begin{description}
-\item[no multiplets] All radial-velocity variations of the primary star are
-  induced by a single companion.
+\item[no multiple companions] All radial-velocity variations of the primary
+  star are induced by a single companion.
   This is motivated by the idea that triple star systems are usually
@@
 indicated where the runs are described (see \sectionname~\ref{sec:catalogs}).
-The reference time for each star is set to the minimum visit epoch; $M_0$ then
+The reference time for each star is set to the first visit epoch; $M_0$ then
 becomes the mean anomaly at the first visit observation.
@@
 Depending on the data quality and phase coverage of the visits, \thejoker\ will
-require different numbers of prior samples in order to rejection sample down to
+require different numbers of prior samples in order to rejection-sample down to
 the requested number of posterior samples: For few-epoch or noisy RV data, many
@@
 to larger preferred values for the excess variance parameter.
-We don't expect there to be a large number of triple systems in our sample, but
+We do not expect there to be a large number of triple systems in our sample, but
 this is still an important consideration for future efforts.
@@
 
-We don't expect a sharp transition in inferred orbital parameters between stars
-with companions and stars without companions: There is a continuum of companion
-properties.
+We do not expect a sharp transition in inferred orbital parameters between stars
+with and without companions: There is a continuum of companion properties.
 For example, the companions can be low mass, or at long periods, or at high
@@
 a given system.
-Therefore there is no simple cut on radial velocity data alone that would select
-a complete sample of stars with companions.
+Therefore, there is no simple cut on radial velocity data alone that would
+select a complete sample of stars with companions.
 It is nevertheless possible to define a cut that selects stars with
@@
 For these stars, one or a few more RV measurements would likely lead to
-uniquely-determined orbital parameters; These stars are prime candidates for
+uniquely-determined orbital parameters, making these stars prime candidates for
 follow-up efforts.
@@
 
-Paradoxically, there exist short period ($\lesssim 100~\textrm{day}$) companions
-at low surface gravities ($\logg \lesssim 1$), obvious in the lower-left panel
-of \figurename~\ref{fig:logg-teff-p-e}: These companions would orbit within the
-surface of their primary stars.
+Paradoxically, there appear to exist short period ($\lesssim 100~\textrm{day}$)
+companions at low surface gravities ($\logg \lesssim 1$), obvious in the
+lower-left panel of \figurename~\ref{fig:logg-teff-p-e}: These companions would
+orbit within the surface of their primary stars.
 Each of these systems appears fine from the \apogee\ data quality flags, but
@@
 We have only considered radial velocity modulations from a single massive
-companion---\emph{no multiplets}; This is a fundamental limitation of our search
-and methodology.
+companion---\emph{no multiple companions}; This is a fundamental limitation of
+our search and methodology.
 However, stars with multiple companions will likely be included in our companion
@@
 
-Of the 382 companions released as a part of the previous search, only 188 of the
-host stars passed the stellar parameter and quality cuts used to define the
-parent sample in this work (see \sectionname~\ref{sec:data}).
+Of the 382 companions released as a part of the \citet{Troup:2016} search, only
+188 of the host stars passed the stellar parameter and quality cuts used to
+define the parent sample in this work (see \sectionname~\ref{sec:data}).
 We have looked at all of the overlapping stars to compare the previously derived
@@
 a likely multi-modal posterior \pdf\ over orbital parameters, and
-(3) the data changed significantly between \apogee\ \DRtw\ and \DR, so
-no meaningful comparisons can be made; Roughly 1/3 of the comparison sample
-falls into each class.
+(3) the radial velocity data used in \citet{Troup:2016} changed significantly
+between \apogee\ \DRtw\ and \DR, so no meaningful comparisons can be made;
+roughly 1/3 of the comparison sample falls into each class.
 \figurename~\ref{fig:troup-unimodal} shows a few representative cases in which
@@
     \item A catalog of \nunimodal\ systems with unimodal posterior samplings,
-    and therefore uniquely determined orbits, of which 227 are newly discovered
+    and therefore uniquely determined orbits, of which 225 are newly discovered
     binary star systems.
@@
     uniquely determine their orbits.
-    79 of these systems are newly discovered binary star systems.
+    90 of these systems are newly discovered binary star systems.
     \item A catalog of \nhighK\ systems with radial velocity variations
@@
 and Yuan-Sen Ting (Princeton).
+
+HWR acknowledges support by Sonderforschungsbereich SFB 881 (A3) of the German
+Research Foundation (DFG), and from the European Research Council under the
+European Union’s Seventh Framework Programme (FP 7) ERC Grant Agreement
+n. [321035].
 

### 2018-04-02 paper/1-catalog/twoface-catalog.tex
galex note

@@
 \newcommand{\nbimodal}{106}
-\newcommand{\nunimodal}{322}
+\newcommand{\nunimodal}{320}
 
@@
 
-Large spectroscopic surveys with repeat observations offer a novel way to study
-binary star and companion populations.
 Time-domain radial-velocity measurements of stars contain information about
-massive companions: Even with two successive observations of a single star, a
-difference in the measured radial velocities implies the existence of at least
-one companion.
+massive companions.
+Even with two successive observations of a single star, a difference in the
+measured radial velocities implies the existence of at least one companion.
 However, with few or imprecise radial-velocity measurements, the orbital
@@
 the \apogee\ survey (see \sectionname~\ref{sec:data}; \citealt{Majewski:2017}).
-
-Stars typically have companions.
-Main sequence stars in the solar neighborhood more often appear in binary or
-multiple star systems rather than as solitary stars (e.g.,
-\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014,Moe:2017}).
-This is likely a generic outcome of star formation, as, for example turbulent
-fragmentation in collapsing protostellar clouds can produce stellar multiplets
-within length-scales comparable to their spheres of influence (e.g.,
-\citealt{Raskutti:2016}).
-Binary and multiple star systems are therefore of great interest in
-astrophysics: The population of stars and their companions encodes information
-about star formation processes, stellar parameters and evolution, and the
-dynamics of multi-body systems (for recent reviews, see
-\citealt{Duchene:2013,Moe:2017}).
-
-Most of what is known about stellar companions comes from studies of
-nearby main-sequence (MS) stars.
-Nearly 50\% of MS stars in the solar neighborhood have companions
-(e.g., \citealt{Tokovinin:2014}).
-MS stars with companions have a large dynamic range of constituent and orbital
-characteristics.
-For example, binary stars have mass-ratios that span from $\approx 0.03$
-to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
-millions of years (e.g., \citealt{Raghavan:2010}).
-Less is known about population properties of non-interacting or detached
-companions to evolved stars.
 
@@
 from the \apogee\ survey, and show the bulk properties of these systems.
+
+Binary and multiple star systems are of great interest in astrophysics: The
+population of stars and their companions encodes information about star
+formation processes, stellar parameters and evolution, and the dynamics of
+multi-body systems (for recent reviews, see \citealt{Duchene:2013,Moe:2017}).
+Most of what is known about stellar companions comes from studies of nearby
+main-sequence (MS) stars (e.g.,
+\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014,Moe:2017}).
+Nearly 50\% of MS stars in the solar neighborhood have companions
+(e.g., \citealt{Tokovinin:2014}).
+MS stars with companions have a large dynamic range of constituent and orbital
+characteristics.
+For example, binary stars have mass-ratios that span from $\approx 0.03$
+to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
+millions of years (e.g., \citealt{Raghavan:2010}).
+Less is known about population properties of non-interacting or detached
+companions to evolved stars.
+The catalogs and methodology presented in this work are a first step towards
+performing population inferences of binary star systems with evolved star
+members.
 
@@
 methodology explained above (see \sectionname~\ref{sec:samplings}).
-In the end, \todo{XX} stars have completed samplings with \nposterior\ samples,
-and, after re-running \thejoker\ with twice as many prior samples, \todo{XX}
-stars still have fewer samples that are multi-modal in period.
-
+% In the end, \todo{XX} stars have completed samplings with \nposterior\ samples,
+% and, after re-running \thejoker\ with twice as many prior samples, \todo{XX}
+% stars still have fewer samples that are multi-modal in period.
 The full catalog of posterior samples for all \apogee\ stars in our parent
@@
 In the right panel, one system appears right on the limit of being an
-interacting binary ($q_{\textrm{min}} \approx 0.5$), but \todo{does it have
-x-ray or UV data?}.
+interacting binary ($q_{\textrm{min}} \approx 0.5$), but has no significant UV
+flux (\acronym{GALEX} DR5; \citealt{Bianchi:2011}).
 These systems are prime candidates for spectroscopic follow-up.

### 2018-04-03 paper/1-catalog/twoface-catalog.tex
Links to website

@@
 The full catalog of posterior samples for all \apogee\ stars in our parent
-sample is available online.
-\todo{Structure of table, where to download, make a webpage with some
-information?}
+sample is available online.\footnote{\url{http://adrian.pw/twoface.html}}
 
@@
 \tablename~\ref{tbl:highK-unimodal} contains descriptions and units for all
-columns in the high-$K$, unimodal sample catalog, available online.
-\todo{Link to a webpage with the catalogs and info on how to use with
-TwoBody/TheJoker}
+columns in the high-$K$, unimodal sample catalog, available online.\footnote{\url{http://adrian.pw/twoface.html}}
 
@@
 All companion catalogs described in this work (\sectionname~\ref{sec:catalogs})
-are available online at \todo{XX}.
+are available online.\footnote{\url{http://adrian.pw/twoface.html}}
 
@@
 
-\section{Using data products...}
+\section{Data products}
 \label{sec:code-demo}
 
-\todo{Brief example of how to use twobody and thejoker to visualize orbits, with
-link to a tutorial notebook on github}
+% \todo{Brief example of how to use twobody and thejoker to visualize orbits, with
+% link to a tutorial notebook on github}
 
@@
 \caption{%
-TODO
+Randomly-selected examples of high-$K$ stars with bimodal posterior samplings.
+Same as \figurename~\ref{fig:highK-unimodal}, but for bimodal posterior
+samplings.
 \label{fig:highK-bimodal}

### 2018-04-03 paper/1-catalog/twoface-catalog.tex
Marla

@@
 Kevin Covey (WWU),
+Marla Geha (Yale),
 Marina Kounkel (WWU),

### 2018-04-06 paper/1-catalog/twoface-catalog.tex
Affil
M not K

@@
 \documentclass[modern, letterpaper]{aastex62}
-
-% to-do list
-% ----------
-% - Make sure we don't use $K$ as a number of samples - that's $M$ - because $K$
-%   is the velocity semi-amplitude!
-% - Figure out what to say about upper RGB stars / coherent oscillations!
 
@@
 
-% other notes
-% -----------
-% - Binary engulfment reference: https://arxiv.org/pdf/1002.2216.pdf
-
 \include{gitstuff}
@@
 \author{Nathan De Lee}
-\affiliation{Department of Physics and Astronomy, Vanderbilt University, VU Station 1807, Nashville, TN 37235, USA}
+\affiliation{Department of Physics, Geology, and Engineering Technology, Northern Kentucky Univerity, Highland Heights, KY 41099}
+\affiliation{Department of Physics & Astronomy, Vanderbilt University, Nashville, TN 37235}
 
@@
 \author{Nicholas Troup}
-\affiliation{}
+\affiliation{Department of Physics, Salisbury University, Salisbury, MD 21801}
 
@@
 full orbital parameters for all stars in the parent sample.
-We highlight interesting systems and show the characteristics of systems with
-confidently determined companion properties.
+We show the characteristics of systems with confidently determined companion
+properties and highlight interesting systems with candidate compact object
+companions.
 \end{abstract}
@@
 We see tentative evidence for surface oscillations at the upper giant branch
-(see \figurename~\ref{fig:P-Teff-logg-e}, lower left), which we presently treat
+(see \figurename~\ref{fig:logg-teff-p-e}, lower left), which we presently treat
 as excess variance or intrinsic jitter.
@@
     \package{sqlalchemy} (\url{https://www.sqlalchemy.org/}),
-    \package{thejoker} (\citealt{Price-Whelan:2017b}),
-    \package{twobody} (\todo{twobody zenodo}).
+    \package{thejoker} (\citealt{Price-Whelan:2017b}).
+    % \package{twobody} (\todo{twobody zenodo}).
 }
@@
 
-For each $n$ of $N$ RGB stars in APOGEE, we obtain $K$ posterior samples over
+For each $n$ of $N$ RGB stars in APOGEE, we obtain $M$ posterior samples over
 primary orbital parameters $\bs{\theta} = (P, e, \omega, M_0, K, v_0)$ and the
@@
 \begin{equation}
-    \bs{w}_k \sim p(\bs{w}_k \given D, \bs{\alpha}_0)
+    \bs{w}_m \sim p(\bs{w}_m \given D, \bs{\alpha}_0)
 \end{equation}
@@
 Using the Monte Carlo integration approximation, \eqname~\ref{eq:marglike} can
-be simplified to a sum over prior value ratios of the $K$ posterior samples in
+be simplified to a sum over prior value ratios of the $M$ posterior samples in
 the log-excess-variance parameter for each $n$ star
 \begin{equation}
-    \approx \frac{\mathcal{Z}_n}{K} \,
-      \sum_k^{K} \frac{p(y_{nk} \given \bs{\alpha})}{p(y_{nk} \given \bs{\alpha}_0)}
+    \approx \frac{\mathcal{Z}_n}{M} \,
+      \sum_m^{M} \frac{p(y_{nm} \given \bs{\alpha})}{p(y_{nm} \given \bs{\alpha}_0)}
 \end{equation}
@@
 \begin{align}
-    p(\{D_n\} \given \alpha) &\propto \prod_n^N \frac{1}{K} \,
-      \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)}
+    p(\{D_n\} \given \alpha) &\propto \prod_n^N \frac{1}{M} \,
+      \sum_m^{M} \frac{p(y_{nm} \given \alpha)}{p(y_{nm} \given \alpha_0)}
       \quad .
@@
     \ln p(\{D_n\} \given \alpha) &\propto \sum_n^N \left[
-      \ln\left( \sum_k^{K} \frac{p(y_{nk} \given \alpha)}{p(y_{nk} \given \alpha_0)} \right)
-      - \ln K\right]\label{eq:lnlike}\\
+      \ln\left( \sum_m^{M} \frac{p(y_{nm} \given \alpha)}{p(y_{nm} \given \alpha_0)} \right)
+      - \ln M\right]\label{eq:lnlike}\\
     &\propto \sum_n^N \left[
-      \underset{k}{\textrm{logsumexp}}\left[ \ln{p(y_{nk} \given \alpha)} - \ln{p(y_{nk} \given \alpha_0)} \right]
-      - \ln K\right]
+      \underset{k}{\textrm{logsumexp}}\left[ \ln{p(y_{nm} \given \alpha)} - \ln{p(y_{nm} \given \alpha_0)} \right]
+      - \ln M\right]
 \end{align}

### 2018-04-09 twoface/samples_analysis.py
option to return index of MAP sample

@@ -46,7 +46,7 @@ def max_likelihood_sample(data, samples):
     return samples[np.argmin(chisqs)]
 
 
-def MAP_sample(data, samples, joker_params):
+def MAP_sample(data, samples, joker_params, return_index=False):
     """Return the maximum a posteriori sample.
 
     Parameters
@@ -60,8 +60,12 @@ def MAP_sample(data, samples, joker_params):
 
     ln_ps = np.zeros(len(samples))
 
-    mcmc_p = model.pack_samples_mcmc(samples)        
+    mcmc_p = model.pack_samples_mcmc(samples)
     for i in range(len(samples)):
         ln_ps[i] = model.ln_posterior(mcmc_p[i])
 
-    return samples[np.argmax(ln_ps)]
+    if return_index:
+        idx = np.argmax(ln_ps)
+        return samples[idx], idx
+    else:
+        return samples[np.argmax(ln_ps)]

### 2018-04-09 paper/1-catalog/twoface-catalog.tex
ampersand
moar authors

@@
 \affiliation{Department of Physics, Geology, and Engineering Technology, Northern Kentucky Univerity, Highland Heights, KY 41099}
-\affiliation{Department of Physics & Astronomy, Vanderbilt University, Nashville, TN 37235}
+\affiliation{Department of Physics \& Astronomy, Vanderbilt University, Nashville, TN 37235}
 
@@
 
-\author{Domingo An\'ibal Garc\'ia-Hern\'andez}
-\affiliation{Instituto de Astrof\'isica de Canarias, 38205 La Laguna, Tenerife, Spain}
+\author{Domingo A. Garc\'ia-Hern\'andez}
+\affiliation{Instituto de Astrof\'isica de Canarias (IAC), E-38205 La Laguna, Tenerife, Spain}
+\affiliation{Universidad de La Laguna (ULL), Departamento de Astrof\'isica, E-38206 La Laguna, Tenerife, Spain}
 
@@
 
+\author{Jennifer Sobeck}
+\affiliation{Department of Astronomy, University of Washington, Box 351580,
+Seattle, WA 98195, USA}
+
 \author{Olga Zamora}
-\affiliation{Instituto de Astrof\'isica de Canarias, 38205 La Laguna, Tenerife, Spain}
+\affiliation{Instituto de Astrof\'isica de Canarias (IAC), E-38205 La Laguna, Tenerife, Spain}
+\affiliation{Universidad de La Laguna (ULL), Departamento de Astrof\'isica, E-38206 La Laguna, Tenerife, Spain}
 
@@
 
+DWH was partially supported by the NSF (grant AST-1517237).
 HWR acknowledges support by Sonderforschungsbereich SFB 881 (A3) of the German
@@
 n. [321035].
+DAGH and OZ acknowledge support provided by the Spanish Ministry of Economy and
+Competitiveness (MINECO) under grant AYA-2017-88254-P.
 

### 2018-04-09 paper/2-population/twoface-pop.tex
rename tex
titles
author shit
intro copy pasta

--- 2017-08-05 paper/main.tex
@@
 % - Reserve () for function arguments; use [] or {} for outer shit.
-% - Always prior pdf or posterior pdf, never prior or posterior (that's your arse).
-% - Use \sectionname not Section, \figname not Figure, \documentname not Article or Paper or paper.
-% - Make sure that two assumptions in the two assumption lists only have the same name if they really are the same assumption: These are proper names!
+% - Always prior pdf or posterior pdf, never prior or posterior (that's your
+%   arse).
+% - Use \sectionname not Section, \figname not Figure, \documentname not Article
+%   or Paper or paper.
+% - Make sure that two assumptions in the two assumption lists only have the
+%   same name if they really are the same assumption: These are proper names!
 % - Hyphenate binary-star when it is an adjective, not when it is a noun!
 % - Where there are defined math symbols (like \pars), use them!
+
+% other notes
+% - Binary engulfment reference: https://arxiv.org/pdf/1002.2216.pdf
 
@@
 
-\title{Binary companions of red clump stars}
+\title{Binary companions of red clump stars II: \\
+       population inference}
 
-\author{Adrian~M.~Price-Whelan}
+\author[0000-0003-0872-7098]{Adrian~M.~Price-Whelan}
 \affiliation{Department of Astrophysical Sciences,
@@
 \section{Introduction} \label{sec:intro}
+
+Stars typically have companions.
+Main sequence stars in the solar neighborhood more often appear in binary or
+multiple star systems rather than as solitary stars (e.g.,
+\citealt{Duquennoy:1991,Raghavan:2010,Tokovinin:2014,Moe:2017}).
+This is likely a generic outcome of star formation, as, for example turbulent
+fragmentation in collapsing protostellar clouds can produce stellar multiplets
+within length-scales comparable to their spheres of influence (e.g.,
+\citealt{Raskutti:2016}).
+Binary and multiple star systems are therefore of great interest in
+astrophysics: The population of stars and their companions encodes information
+about star formation processes, stellar parameters and evolution, and the
+dynamics of multi-body systems (for recent reviews, see
+\citealt{Duchene:2013,Moe:2017}).
+
+Most of what is known about stellar companions comes from studies of
+nearby main-sequence (MS) stars.
+Nearly 50\% of MS stars in the solar neighborhood have companions
+(e.g., \citealt{Tokovinin:2014}).
+MS stars with companions have a large dynamic range of constituent and orbital
+characteristics.
+For example, binary stars have mass-ratios that span from $\approx 0.03$
+to 1 (e.g., \citealt{Kraus:2008}), and have periods from days to
+millions of years (e.g., \citealt{Raghavan:2010}).
+Less is known about population properties of non-interacting or detached
+companions to evolved stars.
 

### 2018-04-12 paper/1-catalog/twoface-catalog.tex
PL comments

@@
 spectrum is not observed, the data will be consistent with many different
-combinations of (primary) orbital parameters (period, amplitude, eccentricity,
+combinations of primary orbital parameters (period, amplitude, eccentricity,
 etc.).
@@
 n. [321035].
+PL was partly funded by "Programa de Iniciaci\'on en Investigaci\'on –
+Universidad de Antofagasta."
 DAGH and OZ acknowledge support provided by the Spanish Ministry of Economy and

### 2018-04-24 paper/1-catalog/twoface-catalog.tex
Fix MCMC parametrization
Address referee comments

@@
 companions---as are expected---these should be measurable.
-
 By making cuts on this library of posterior \pdf\ samples (described in detail
@@
 assumed prior \pdf\ (see below), the (typically multi-modal) likelihood is
-evaluated at each sample and used to rejection sample.
+evaluated at each sample and used to reject all prior samples above the
+likelihood surface (the rejection sampling step).
 In practice, given a number of requested samples for each star, the sampling
@@
 perform the samplings.
+\package{emcee} uses an ensemble of ``walkers'' (separate Markov chains) that
+naturally adapt to the geometry of the parameter-space by generating proposal
+steps based on the locations of other walkers (i.e. the ensemble chains are not
+independent).
 We transform the standard Keplerian orbital parameters to a safer
-parametrization, $(\ln P, \sqrt{K}\,\cos M_0, \sqrt{K}\,\sin M_0, \sqrt{e}\,\cos
-\omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, for MCMC sampling.
+parametrization, $(\ln P, \sqrt{K}\,\cos (M_0 + \omega), \sqrt{K}\,\sin (M_0 +
+\omega), \sqrt{e}\,\cos \omega, \sqrt{e}\,\cos \omega, \ln s^2, v_0)$, for MCMC
+sampling.
 This reparametrization is safer and more efficient for sampling with

### 2018-04-24 paper/1-catalog/twoface-catalog.tex
colon capitalization

@@
 righteous:
-It delivers independent posterior \pdf\ samples for single-companion binary
+it delivers independent posterior \pdf\ samples for single-companion binary
 orbital parameters, given any number of radial-velocity measurements.
@@
 
-Binary and multiple star systems are of great interest in astrophysics: The
+Binary and multiple star systems are of great interest in astrophysics: the
 population of stars and their companions encodes information about star
@@
 \citet{Murray:2010}:
-The radial velocity $v$ at time $t$ is given by
+the radial velocity $v$ at time $t$ is given by
 \begin{equation}
@@
 independent samples with such data, \thejoker\ succeeds by brute force:
-After generating an initial (very large) library of prior samples from an
+after generating an initial (very large) library of prior samples from an
 assumed prior \pdf\ (see below), the (typically multi-modal) likelihood is
@@
 require different numbers of prior samples in order to rejection-sample down to
-the requested number of posterior samples: For few-epoch or noisy RV data, many
+the requested number of posterior samples: for few-epoch or noisy RV data, many
 prior samples will pass the rejection step, whereas for very precise or
@@
 This is also a test of the visit velocity uncertainties reported in the \apogee\
-data products: If the catalog uncertainties are significantly underestimated, we
+data products: if the catalog uncertainties are significantly underestimated, we
 expect the inferred log-excess-variance distribution parameters to tend towards
@@
 the excess variance distribution.
-First, we do not remove triple or other multiple systems: For any individual
+First, we do not remove triple or other multiple systems: for any individual
 system, the radial velocity variations induced by other massive bodies will lead
@@
 We do not expect a sharp transition in inferred orbital parameters between stars
-with and without companions: There is a continuum of companion properties.
+with and without companions: there is a continuum of companion properties.
 For example, the companions can be low mass, or at long periods, or at high
@@
 \figurename~\ref{fig:logg-teff-p-e}) shows a clear signature of tidal
-circularization: Such plots for main sequence stars typically show a more
+circularization: such plots for main sequence stars typically show a more
 distinct circularization period, below which the majority of stars have
@@
 In \figurename~\ref{fig:mass}, a few interesting systems are immediately
-obvious: From the left panel, two systems have bimodal period samplings in which
+obvious: from the left panel, two systems have bimodal period samplings in which
 both period modes put the minimum companion mass below the Hydrogen-burning
@@
 companions at low surface gravities ($\logg \lesssim 1$), obvious in the
-lower-left panel of \figurename~\ref{fig:logg-teff-p-e}: These companions would
+lower-left panel of \figurename~\ref{fig:logg-teff-p-e}: these companions would
 orbit within the surface of their primary stars.
@@
 of their period samples below $100~\textrm{day}$ to the \acronym{APOKASC}
-catalog: One star appears in both samples 2M19024490+4419523.
+catalog: one star appears in both samples 2M19024490+4419523.
 In \apogee\ \DR, this star has $\logg = 0.57$, but asteroseismic $\logg > 4$
@@
     samples for each source, \texttt{lnK\_per\_1}:
-    The ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
+    the ``high-$K$'' sample (see \sectionname~\ref{sec:catalog-confident}) is
     defined using this column, by selecting \texttt{lnK\_per\_1 > 0}.

### 2018-04-24 paper/1-catalog/twoface-catalog.tex
thanks
git hash
Ref tidal circ paper

@@
 surface gravities and therefore a large range in stellar radii (see lower right
-panel).
+panel); this result is explored in more detail in companion work
+(\citealt{Circularization:2018}).
 
@@
 license.
+This paper version was generated with Git commit \texttt{\githash\,(\gitdate)}.
 
@@
 and Yuan-Sen Ting (Princeton).
+We thank the anonymous referee for constructive comments that improved this
+manuscript.
 
